### 2021-06-12 docs/elsarticle/elsarticle-template.tex
paper

\documentclass[review]{elsarticle}

\usepackage{lineno,hyperref}
\modulolinenumbers[5]

\journal{CIRP Journal of Manufacturing Science and Technology}

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
%\bibliographystyle{model1-num-names}

%% Numbered without titles
%\bibliographystyle{model1a-num-names}

%% Harvard
%\bibliographystyle{model2-names.bst}\biboptions{authoryear}

%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}

%% APA style
%\bibliographystyle{model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frontmatter}

\title{An Intuitive Spatial AR System for Augmenting Planar Scenes in Industrial Settings} %\title{Elsevier \LaTeX\ template\tnoteref{mytitlenote}}
%\tnotetext[mytitlenote]{Fully documented templates are available in the elsarticle package on \href{http://www.ctan.org/tex-archive/macros/latex/contrib/elsarticle}{CTAN}.}

%% Group authors per affiliation:
\author{Michael Horn\'{a}\v{c}ek\corref{mycorrespondingauthor}}
\ead{michael.hornacek@tuwien.ac.at}
\address{Human Centered Cyber Physical Production and Assembly Systems, Institute for Management Sciences, TU Vienna, Austria}

\begin{abstract}
Augmented reality (AR) promises to enable use cases in industrial settings that include the embedding of work steps directly into the scene, potentially reducing or altogether obviating the need for workers to refer to instructions in paper form or on a screen. In turn, spatial AR is a form of AR whereby the augmentation of the scene is carried out using a projector, with the advantage or rending the augmentation visible to all onlookers simultaneously without calling for each to wear AR glasses. However, care must be taken to distort the images to be projected in a manner that they appear undistorted to the viewer, since the geometry of the scene as it relates to the geometry of the projector plays a role in how the pixels of the projector’s image plane map to points in the scene. For planar scene geometry (such as a floor, wall, or table), this can be done in a cumbersome manual process called keystone correction, often using software bundled with the projector. We propose a system that produces the same effect by placing the desired augmentations intuitively in accordance with an image of the scene acquired by a camera. We achieve this by distorting the image to be projected using a plane-induced homography computed to produce the effect of projecting the image not from the actual projector pose, but in accordance with a \textit{virtual} projector pose (i) facing directly downwards to the scene plane and (ii) rotated to place the image plane of the virtual projector in line with that of the camera.  
\end{abstract}

\begin{keyword}
Spatial augmented reality (SAR) \sep Industry 4.0
\end{keyword}

\end{frontmatter}

\linenumbers

\section{Introduction}

A central challenge of spatial AR, however,  the geometry of the scene as it relates to the geometry of the projector plays a role in how the pixels of the projector’s image plane map to points in the scene. This is why, unless special care is taken,

While pointing a projector at a scene and projecting an image is on its own not a challenge, doing so in a manner such that the projected content appear to be naturally part of the scene is more tricky. The intuition for the general problem we face can be drawn from using a projector to project an image to a flat wall: unless the projector faces the wall frontally, the bounds of a projected rectangular image will not appear rectangular, but more generally as a trapezoid (i.e., the image will appear distorted). The cause for this effect is that the geometry of the scene as it relates to the geometry of the projector plays a role in how the pixels of the projector’s image plane map to points in the scene.

The central objective we set for ourselves is to enable correcting for distortion caused by projecting to arbitrary scene geometry in an automatic fashion. This is to be done in a manner that lends itself to practical application in a real-world setting where the scene to be augmented—e.g., an object being gradually assembled and its placement relative to the projector—does not remain static. Finally, the solution is ideally to rely on the existing hardware setup—i.e., projector with steerable mirror system and stereo camera—mounted in the Pilotfabrik  of TU Vienna.

\paragraph{Usage} Once the package is properly installed, you can use the document class \emph{elsarticle} to create a manuscript. Please make sure that your manuscript follows the guidelines in the Guide for Authors of the relevant journal. It is not necessary to typeset your manuscript in exactly the same way as an article, unless you are submitting to a camera-ready copy (CRC) journal.

\paragraph{Functionality} The Elsevier article class is based on the standard article class and supports almost all of the functionality of that class. In addition, it features commands and options to format the
\begin{itemize}
\item document style
\item baselineskip
\item front matter
\item keywords and MSC codes
\item theorems, definitions and proofs
\item lables of enumerations
\item citation style and labeling.
\end{itemize}

\section{Front matter}

The author names and affiliations could be formatted in two ways:
\begin{enumerate}[(1)]
\item Group the authors per affiliation.
\item Use footnotes to indicate the affiliations.
\end{enumerate}
See the front matter of this document for examples. You are recommended to conform your choice to the journal you are submitting to.

\section{Bibliography styles}

There are various bibliography styles available. You can select the style of your choice in the preamble of this document. These styles are Elsevier styles based on standard styles like Harvard and Vancouver. Please use Bib\TeX\ to generate your bibliography and include DOIs whenever available.

Here are two sample references: \cite{Feynman1963118,Dirac1953888}.

\section*{References}

\bibliography{mybibfile}

\end{document}

### 2021-06-12 src/apps/calibrateProj/calibrateProj.cpp

@@ -3,10 +3,10 @@
  *   michael.hornacek@gmail.com
  *   IMW-CPS TU Vienna, Austria
  *
- *   calibrateProj <boardSqSize> <boardDimsX> <boardDimsY> <circlesDimsX> <circlesDimsY> <outDir> <numIms> <imDir> <cam0Path> <circlesImPath> <verticalNegOffset> [<visImIdx>] [<visIm>]
+ *   calibrateProj <boardSqSize> <boardDimsX> <boardDimsY> <circlesDimsX> <circlesDimsY> <outDir> <numIms> <cam0ImDir> <cam1ImDir> <cam0Path> <cam1Path> <circlesImPath> <verticalNegOffset> [<visImIdx>] [<visIm>]
  *
  *   Example invocation:
- *   calibrateProj 0.0565 4 6 4 11 C:\Users\micha\Desktop\spatial-ar\in_out\calibrateProj\out 14 C:\Users\micha\Desktop\spatial-ar\in_out\splitZed\projCalib\outLeft C:\Users\micha\Desktop\spatial-ar\in_out\calibrateCam\out\cam_0.yml C:\Users\micha\Desktop\spatial-ar\in_out\calibrateProj\acircles_pattern_960x600.png 1.75 3 C:\Users\micha\Desktop\spatial-ar\in_out\applyHomography\holodeck.png
+ *   calibrateProj 0.0565 4 6 4 11 C:\Users\micha\Desktop\spatial-ar\in_out\calibrateProj\out 14 C:\Users\micha\Desktop\spatial-ar\in_out\splitZed\projCalib\outLeft C:\Users\micha\Desktop\spatial-ar\in_out\splitZed\projCalib\outRight C:\Users\micha\Desktop\spatial-ar\in_out\calibrateCam\out\cam_0.yml C:\Users\micha\Desktop\spatial-ar\in_out\calibrateCam\out\cam_1.yml C:\Users\micha\Desktop\spatial-ar\in_out\calibrateProj\acircles_pattern_960x600.png 1.75 3 C:\Users\micha\Desktop\spatial-ar\in_out\applyHomography\holodeck.png
  */
 
 
@@ -111,8 +111,10 @@ static const char* keys =
     "{@circlesDimsY | | ...}"
     "{@outDir | | ...}"
     "{@numIms | | ...}"
-    "{@imDir | | ...}"
+    "{@cam0ImDir | | ...}"
+    "{@cam1ImDir | | ...}"
[…]

### 2021-06-12 src/cpslib/libCalib/CalibPattern.cpp

@@ -92,15 +92,14 @@ void CalibPattern::findCirclesImPts(cv::Mat& im, cv::Size circlesPatternSize,
 
     cv::cvtColor(imUndistort, imGrayUndistort, cv::COLOR_BGR2GRAY);
     cv::bitwise_not(imGrayUndistort, imGrayUndistortInvert);
-    // cv::threshold(candidateIm, candidateImThresh, 100, 255, THRESH_BINARY);
 
     outImSize = cv::Size(im.cols, im.rows);
 
     std::vector<cv::Point2f> candidateCirclesImPts;
     if (findCirclesGrid(imGrayUndistortInvert, circlesPatternSize, candidateCirclesImPts, cv::CALIB_CB_ASYMMETRIC_GRID))
     {
-        //cv::cornerSubPix(imGrayUndistort, candidateCirclesImPts, cv::Size(5, 5), cv::Size(-1, -1),
-        //    TermCriteria(TermCriteria::EPS + TermCriteria::COUNT, 40, 0.001));
+        cv::cornerSubPix(imGrayUndistort, candidateCirclesImPts, cv::Size(5, 5), cv::Size(-1, -1),
+            cv::TermCriteria(cv::TermCriteria::EPS + cv::TermCriteria::COUNT, 40, 0.001));
 
         drawChessboardCorners(outImVis, circlesPatternSize, candidateCirclesImPts, true);
         outCirclesImPts.push_back(candidateCirclesImPts);
@@ -137,3 +136,4 @@ void CalibPattern::findChessboardAndCirclesImPts(std::string inDir, int numIms, 
         cv::waitKey(100);
     }
 }
+

### 2021-06-12 docs/elsarticle/elsarticle-template.tex
paper

@@
 Augmented reality (AR) promises to enable use cases in industrial settings that include the embedding of work steps directly into the scene, potentially reducing or altogether obviating the need for workers to refer to instructions in paper form or on a screen.
-In turn, spatial AR is a form of AR whereby the augmentation of the scene is carried out using a projector, with the advantage or rending the augmentation visible to all onlookers simultaneously without calling for each to wear AR glasses.
+In turn, spatial AR is a form of AR whereby the augmentation of the scene is carried out using a projector, with the advantage of rending the augmentation visible to all onlookers simultaneously without calling for each to wear AR glasses.
 However, care must be taken to distort the images to be projected in a manner that they appear undistorted to the viewer, since the geometry of the scene as it relates to the geometry of the projector plays a role in how the pixels of the projector’s image plane map to points in the scene.

### 2021-06-13 docs/elsarticle/elsarticle-template.tex
paper

@@
 \usepackage{lineno,hyperref}
+\usepackage{algorithm}
+\usepackage{algorithmic}
+\usepackage{amsmath,amssymb}
+\usepackage{graphicx}
+\usepackage{subfig}
+\usepackage{tabularx}
+\usepackage[usenames,dvipsnames]{xcolor}
+\usepackage{colortbl}
 \modulolinenumbers[5]
@@
 
-\title{An Intuitive Spatial AR System for Augmenting Planar Scenes in Industrial Settings} %\title{Elsevier \LaTeX\ template\tnoteref{mytitlenote}}
+\title{A Spatial AR System for Wide-area Axis-aligned Augmentation of Planar Scenes in Industrial Settings} %\title{Elsevier \LaTeX\ template\tnoteref{mytitlenote}}
 %\tnotetext[mytitlenote]{Fully documented templates are available in the elsarticle package on \href{http://www.ctan.org/tex-archive/macros/latex/contrib/elsarticle}{CTAN}.}
@@
 %% Group authors per affiliation:
-\author{Michael Horn\'{a}\v{c}ek\corref{mycorrespondingauthor}}
-\ead{michael.hornacek@tuwien.ac.at}
+\author{Michael Horn\'{a}\v{c}ek}
+\author{Hans K\"{u}ffner-McCauley}
+\author{Sebastian Schlund}
 \address{Human Centered Cyber Physical Production and Assembly Systems, Institute for Management Sciences, TU Vienna, Austria}
@@
 Augmented reality (AR) promises to enable use cases in industrial settings that include the embedding of work steps directly into the scene, potentially reducing or altogether obviating the need for workers to refer to instructions in paper form or on a screen.
-In turn, spatial AR is a form of AR whereby the augmentation of the scene is carried out using a projector, with the advantage of rending the augmentation visible to all onlookers simultaneously without calling for each to wear AR glasses.
-However, care must be taken to distort the images to be projected in a manner that they appear undistorted to the viewer, since the geometry of the scene as it relates to the geometry of the projector plays a role in how the pixels of the projector’s image plane map to points in the scene.
+In turn, spatial AR is a form of AR whereby the augmentation of the scene is carried out using a projector, with the advantage of rendering the augmentation visible to all onlookers simultaneously without calling for each to wear AR glasses.
+However, care must be taken to distort the images to be projected in a manner that they appear undistorted to the viewer, since the geometry of the scene as it relates to the geometry of the projector plays a role in how the pixels of the projector's image plane map to points in the scene.
 For planar scene geometry (such as a floor, wall, or table), this can be done in a cumbersome manual process called keystone correction, often using software bundled with the projector.
-We propose a system that produces the same effect by placing the desired augmentations intuitively in accordance with an image of the scene acquired by a camera.
-We achieve this by distorting the image to be projected using a plane-induced homography computed to produce the effect of projecting the image not from the actual projector pose, but in accordance with a \textit{virtual} projector pose (i) facing directly downwards to the scene plane and (ii) rotated to place the image plane of the virtual projector in line with that of the camera.
+
+We propose a system that produces the effect of keystone correction analytically, intuitively placing the desired augmentations in a manner aligned with the axes of an image of the scene acquired by a camera.
+Moreover, our system is able to handle a projector equipped with a steerable mirror, enabling wide-area factory floor augmentation.
 \end{abstract}
@@
 \begin{keyword}
-Spatial augmented reality (SAR) \sep Industry 4.0
+Spatial augmented reality (SAR) \sep Industry 4.0 \sep Pilotfabrik
 \end{keyword}
@@
 
-\section{Introduction}
+\section{Introduction}\label{sec:intro}
 
-A central challenge of spatial AR, however,  the geometry of the scene as it relates to the geometry of the projector plays a role in how the pixels of the projector’s image plane map to points in the scene.
-This is why, unless special care is taken,
+Augmented reality (AR) promises to enable use cases in industrial settings that include the embedding of work steps directly into the scene, potentially reducing or altogether doing away with the need for workers to refer to instructions in paper form or on a screen.
+Typically, AR works by embedding the augmentation in an image of the scene acquired from the viewpoint of a single individual, with the resulting augmented image in turn displayed using some form of AR glasses.
+Reliance on AR glasses has two adverse consequences: (i) AR glasses must be worn by each individual wishing to view the augmentation, and (ii) such glasses---in some cases taking the form of a helmet in order to house multiple sensors in support of accurately tracking the viewpoint of the viewer relative to the scene---can be obtrusive.
+Spatial AR is a form of augmented reality carried out not by embedding the augmentation in an image of the scene as with AR glasses, but by projection to the scene itself, thus eliminating both aforementioned problems.
+Yet considering for the moment a planar surface to be augmented, unless the projector faces the surface frontally, the bounds of a projected rectangular image will not appear rectangular; more generally, they will instead appear trapezoidal (i.e., the image will appear distorted).
+Such distortions can be eliminated by carrying out a cumbersome manual process called keystone correction, often using software bundled with the projector.
 
-While pointing a projector at a scene and projecting an image is on its own not a challenge, doing so in a manner such that the projected content appear to be naturally part of the scene is more tricky.
-The intuition for the general problem we face can be drawn from using a projector to project an image to a flat wall: unless the projector faces the wall frontally, the bounds of a projected rectangular image will not appear rectangular, but more generally as a trapezoid (i.e., the image will appear distorted).
-The cause for this effect is that the geometry of the scene as it relates to the geometry of the projector plays a role in how the pixels of the projector’s image plane map to points in the scene.
+Using the $X$- and $Y$-axes of an image of the scene as a proxy, our contribution is to propose a system that produces the effect of keystone correction analytically, and in a manner aligning the axes of the augmentation with those of the proxy image.
+We achieve this by distorting the image to be projected using a plane-induced homography computed to produce the effect of projecting the image not from the actual projector viewpoint, but in accordance with the viewpoint of a \textit{virtual} projector (i) facing directly downwards to the scene plane and (ii) rotated to place the axes of the image plane of the virtual projector in line with those of the camera.
+This enables intuitive placement of augmentations in the scene, and eliminates the need for manual keystone correction.
+Moreover, our system is able to handle a projector equipped with a steerable mirror without need for explicitly modeling the action of the steerable mirror on the projector.
 
-The central objective we set for ourselves is to enable correcting for distortion caused by projecting to arbitrary scene geometry in an automatic fashion.
-This is to be done in a manner that lends itself to practical application in a real-world setting where the scene to be augmented—e.g., an object being gradually assembled and its placement relative to the projector—does not remain static.
-Finally, the solution is ideally to rely on the existing hardware setup—i.e., projector with steerable mirror system and stereo camera—mounted in the Pilotfabrik  of TU Vienna.
+\subsection{Related Work}
 
-\paragraph{Usage} Once the package is properly installed, you can use the document class \emph{elsarticle} to create a manuscript.
-Please make sure that your manuscript follows the guidelines in the Guide for Authors of the relevant journal.
-It is not necessary to typeset your manuscript in exactly the same way as an article, unless you are submitting to a camera-ready copy (CRC) journal.
+\section{Hardware Setup}
 
-\paragraph{Functionality} The Elsevier article class is based on the standard article class and supports almost all of the functionality of that class.
-In addition, it features commands and options to format the
-\begin{itemize}
-\item document style
-\item baselineskip
-\item front matter
-\item keywords and MSC codes
-\item theorems, definitions and proofs
-\item lables of enumerations
-\item citation style and labeling.
-\end{itemize}
+The hardware setup employed in this work comprised a Panasonic XXX projector with a steerable mirror system manufactured by XX.
+In addition, we used a Zed 2 stereo camera manufactured by Stereolabs, yet relied only on the left view.
+The setup was mounted on the ceiling of the Pilotfabrik, a collaborative space for research on Industry 4.0 topics situated in Vienna, Austria.
+The floorspace used for our experiments measured dimensions of X~m~$\times$~Y~m; the projector was mounted at approximately the center of this space, at a height of ca.\ XX m.
 
-\section{Front matter}
+\section{Approach}
 
-The author names and affiliations could be formatted in two ways:
-\begin{enumerate}[(1)]
-\item Group the authors per affiliation.
-\item Use footnotes to indicate the affiliations.
-\end{enumerate}
-See the front matter of this document for examples. You are recommended to conform your choice to the journal you are submitting to.
+The problem of correcting for distortions as outlined in Section~\ref{sec:intro} requires knowledge of the manner in which the respective rays through the pixels of the projector's image plane fan out into the scene, and the geometry of the scene itself within at least the projector’s field of view, as illustrated in Figure 1.
+This is because the scene point `illuminated' by a pixel in the projector's image plane is given by intersecting its corresponding ray with the geometry of the scene surface.
+To model this interaction calls for a one-time camera-projector calibration, which includes recovery of the scene plane.
+Next, the relative camera-projector-scene plane geometry is used to compute a plane-induced homography.
 
-\section{Bibliography styles}
+The hardware setup employed
 
-There are various bibliography styles available.
-You can select the style of your choice in the preamble of this document.
-These styles are Elsevier styles based on standard styles like Harvard and Vancouver.
-Please use Bib\TeX\ to generate your bibliography and include DOIs whenever available.
+\subsection{Camera-projector Calibration}
+
+\paragraph{Camera 2D-3D correspondences}
+
+\paragraph{Projector 2D-3D correspondences}
+
+\subsection{Distortion}
+
+\paragraph{Virtual projector}
+
+\paragraph{Plane-induced homography} Let $\mathtt{K}_\text{proj}$ express the $3~\times~3$ calibration matrix of the projector and $(\mathtt{R}', \mathbf{t}') \in SE(3)$ the rigid body transformation that transforms points from the coordinate frame of the projector to that of the virtual projector.
+
+\begin{equation}
+\mathtt{H} = \mathtt{K}_\text{proj}\left(\mathtt{R}' + \frac{1}{Z^{} \mathbf{n}^\top \mathbf{p}^{}}\mathbf{t}'\mathbf{n}^\top\right)\mathtt{K}_\text{proj}^{-1},
+\label{homgen}
+\end{equation}
+where ()
+
+\section{Evaluation}
+
+\section{Conclusion}
 
 Here are two sample references: \cite{Feynman1963118,Dirac1953888}.
-
-\section*{References}
 

### 2021-06-14 docs/elsarticle/elsarticle-template.tex
paper

@@
 \begin{abstract}
-Augmented reality (AR) promises to enable use cases in industrial settings that include the embedding of work steps directly into the scene, potentially reducing or altogether obviating the need for workers to refer to instructions in paper form or on a screen.
+Augmented reality (AR) promises to enable use cases in industrial settings that include the embedding of work step instructions directly into the scene, potentially reducing or altogether obviating the need for workers to refer to instructions in paper form or on a screen.
 In turn, spatial AR is a form of AR whereby the augmentation of the scene is carried out using a projector, with the advantage of rendering the augmentation visible to all onlookers simultaneously without calling for each to wear AR glasses.
@@
 
-Augmented reality (AR) promises to enable use cases in industrial settings that include the embedding of work steps directly into the scene, potentially reducing or altogether doing away with the need for workers to refer to instructions in paper form or on a screen.
+Augmented reality (AR) promises to enable use cases in industrial settings that include the embedding of work step instructions directly into the scene, potentially reducing or altogether doing away with the need for workers to refer to instructions in paper form or on a screen.
 Typically, AR works by embedding the augmentation in an image of the scene acquired from the viewpoint of a single individual, with the resulting augmented image in turn displayed using some form of AR glasses.
@@
 
-The hardware setup employed in this work comprised a Panasonic XXX projector with a steerable mirror system manufactured by XX.
+The hardware setup employed in this work comprised a Panasonic PT-RZ660BE projector with a steerable mirror system manufactured by Dynamic Projection Institute.
 In addition, we used a Zed 2 stereo camera manufactured by Stereolabs, yet relied only on the left view.
-The setup was mounted on the ceiling of the Pilotfabrik, a collaborative space for research on Industry 4.0 topics situated in Vienna, Austria.
+The setup was mounted on the ceiling of the Pilotfabrik\footnote{\url{https://www.pilotfabrik.at/}} of TU Vienna, a collaborative space for research on Industry 4.0 topics situated in Vienna, Austria.
 The floorspace used for our experiments measured dimensions of X~m~$\times$~Y~m; the projector was mounted at approximately the center of this space, at a height of ca.\ XX m.
@@
 
-\paragraph{Camera 2D-3D correspondences}
+\paragraph{Camera calibration}
 
-\paragraph{Projector 2D-3D correspondences}
+\paragraph{Projector calibration}
 
-\subsection{Distortion}
+The pose of the projector is given in the coordinate frame of the camera.
+
+\subsection{Correcting for Distortion}
 
@@
 \begin{equation}
-\mathtt{H} = \mathtt{K}_\text{proj}\left(\mathtt{R}' + \frac{1}{Z^{} \mathbf{n}^\top \mathbf{p}^{}}\mathbf{t}'\mathbf{n}^\top\right)\mathtt{K}_\text{proj}^{-1},
+\mathtt{H} = \mathtt{K}_\text{proj}\left(\mathtt{R}' - \frac{\mathbf{t}'\mathbf{n}^\top}{d}\right)\mathtt{K}_\text{proj}^{-1},
 \label{homgen}
 \end{equation}
-where ()
+where $(\mathbf{n}^\top, -d)^\top$ gives the scene plane and $d = \mathbf{n}^\top\mathbf{X}$, for any point $\mathbf{X}$ in the plane.
 

### 2021-06-18 paper/elsarticle/elsarticle-template.tex
moved

--- 2021-06-14 docs/elsarticle/elsarticle-template.tex
@@
 This enables intuitive placement of augmentations in the scene, and eliminates the need for manual keystone correction.
-Moreover, our system is able to handle a projector equipped with a steerable mirror without need for explicitly modeling the action of the steerable mirror on the projector.
+Moreover, a consequence of our approach is that our system is able to handle a projector equipped with a steerable mirror without need for explicitly modeling the action of the steerable mirror on the projector.
 
 \subsection{Related Work}
+
+todo - maybe Majesa would like to help?
 
@@
 
-The problem of correcting for distortions as outlined in Section~\ref{sec:intro} requires knowledge of the manner in which the respective rays through the pixels of the projector's image plane fan out into the scene, and the geometry of the scene itself within at least the projector’s field of view, as illustrated in Figure 1.
+The problem of correcting for distortions as outlined in Section~\ref{sec:intro} requires knowledge of the manner in which the respective rays through the pixels of the projector's image plane fan out into the scene (i.e., we must `calibrate' the projector) and the geometry of the scene itself (i.e., we must recover the scene plane) within at least the projector's field of view.
 This is because the scene point `illuminated' by a pixel in the projector's image plane is given by intersecting its corresponding ray with the geometry of the scene surface.
-To model this interaction calls for a one-time camera-projector calibration, which includes recovery of the scene plane.
-Next, the relative camera-projector-scene plane geometry is used to compute a plane-induced homography.
+To model this interaction calls for (i) a one-time projector calibration, which in our approach calls for additionally calibrating a camera overlooking the scene and includes recovery of the scene plane as a convenient side effect.
+Next, (ii) the relative camera-projector-scene plane geometry is used to compute a plane-induced homography that distorts the image to be projected in a manner that it appear undistorted to the viewer, and placed in alignment with the axes of a proxy image of the scene.
+These to points are treated in Sections~\ref{sec:approach:geometry} and \ref{sec:approach:homography}, respectively.
 
-The hardware setup employed
+\subsection{Recovering Geometry}\label{sec:approach:geometry}
 
-\subsection{Camera-projector Calibration}
+The `output' of projector (or camera) calibration is a mathematical model that determines, for each pixel xproj in the projector's (or camera's) image plane πproj, the ray from the projector’s center of projection Cproj through xproj, along which the pixel is projected to the scene.
+Inversely, the model also determines the projection $\mathbf{x}'_\text{proj}$ of any point $\mathbf{X}'_\text{proj}$ in the scene to the projector's (or camera's) image plane $\mathbf{\pi}_\text{proj}$, obtained by intersecting $\mathbf{\pi}_\text{proj}$ with the ray from $\mathbf{C}_\text{proj}$ through $\mathbf{X}'_\text{proj}$.
+Such a model is said to function in accordance with the principle of central projection.
+
+Notably, central projection is the same principle that underlies the modeling of conventional cameras.
+A camera modeled in accordance with the principle of central projection is called a pinhole camera.
+A pinhole camera can be expressed in terms of a $3~\times~3$ camera calibration matrix $\mathtt{K}$,  derived from the camera's focal length and principal point.
+In practice, a pinhole camera model works well in modeling how the respective rays through the pixels of the image plane fan out into the scene, on the condition that images to which the model is applied have been corrected for lens distortions caused by the camera’s lens system.
+The `output' of camera calibration is accordingly not only the matrix $\mathtt{K}$, but the associated lens distortions model coefficients as well.
+The lens distortion model coefficients enable applying a lens distortion model to undistort any image acquired by the camera.
+
+In practice, calibrating a camera relies on (i) establishing 2D-3D correspondences between pixels in the camera's image plane and corresponding points in the scene, and on (ii) using those correspondences as input to an optimization procedure that outputs the calibration matrix $\mathtt{K}$ and the associated distortion model coefficients.
+Calibration of a projector can be carried out in precisely the same manner insofar as step (ii) is concerned; the major difference in projector calibration relative to camera calibration concerns the manner in which 2D-3D correspondences are identified, i.e., between pixels in the image plane of the projector and the corresponding points in the scene.
 
 \paragraph{Camera calibration}
+
+\paragraph{Scene plane recovery}
 
@@
 
-\subsection{Correcting for Distortion}
+\subsection{Correcting for Distortion}\label{sec:approach:homography}
 
@@
 
+todo - maybe Hans would like to help?
+
 \section{Conclusion}
-
-Here are two sample references: \cite{Feynman1963118,Dirac1953888}.
 

### 2021-06-20 paper/elsarticle/elsarticle-template.tex
paper

--- 2021-06-14 docs/elsarticle/elsarticle-template.tex
@@
 Augmented reality (AR) promises to enable use cases in industrial settings that include the embedding of work step instructions directly into the scene, potentially reducing or altogether obviating the need for workers to refer to instructions in paper form or on a screen.
-In turn, spatial AR is a form of AR whereby the augmentation of the scene is carried out using a projector, with the advantage of rendering the augmentation visible to all onlookers simultaneously without calling for each to wear AR glasses.
+In turn, \textit{spatial} AR is a form of AR whereby the augmentation of the scene is carried out using a projector, with the advantage of rendering the augmentation visible to all onlookers simultaneously without calling for each to wear AR glasses.
 However, care must be taken to distort the images to be projected in a manner that they appear undistorted to the viewer, since the geometry of the scene as it relates to the geometry of the projector plays a role in how the pixels of the projector's image plane map to points in the scene.
@@
 
-We propose a system that produces the effect of keystone correction analytically, intuitively placing the desired augmentations in a manner aligned with the axes of an image of the scene acquired by a camera.
-Moreover, our system is able to handle a projector equipped with a steerable mirror, enabling wide-area factory floor augmentation.
+We propose a system that produces the effect of keystone correction analytically, and that intuitively places the desired augmentations in a manner aligned with the axes of an image of the scene acquired by a camera.
+Moreover, our system is able to handle a projector equipped with a steerable mirror, enabling wide-area factory floor augmentation exceeding the projector's own immediate field of view.
 \end{abstract}
@@
 This enables intuitive placement of augmentations in the scene, and eliminates the need for manual keystone correction.
-Moreover, our system is able to handle a projector equipped with a steerable mirror without need for explicitly modeling the action of the steerable mirror on the projector.
+Moreover, a consequence of our approach is that our system is able to handle a projector equipped with a steerable mirror (without need for explicitly modeling the action of the steerable mirror on the projector), thereby enabling wide-area applications.
 
 \subsection{Related Work}
+
+todo - maybe Majesa would like to help? Sebastian's inputs would be valuable for more general background story.
 
@@
 
-The problem of correcting for distortions as outlined in Section~\ref{sec:intro} requires knowledge of the manner in which the respective rays through the pixels of the projector's image plane fan out into the scene, and the geometry of the scene itself within at least the projector’s field of view, as illustrated in Figure 1.
+Correcting for projective distortions of the sort outlined in Section~\ref{sec:intro} requires knowledge of the manner in which the respective rays through the pixels of the projector's image plane fan out into the scene (i.e., we must `calibrate' the projector) and the geometry of the scene itself (i.e., we must recover the scene plane relative to the projector) within at least the projector's field of view.
 This is because the scene point `illuminated' by a pixel in the projector's image plane is given by intersecting its corresponding ray with the geometry of the scene surface.
-To model this interaction calls for a one-time camera-projector calibration, which includes recovery of the scene plane.
-Next, the relative camera-projector-scene plane geometry is used to compute a plane-induced homography.
+To model this interaction calls for (i) a one-time projector calibration, which in our approach calls for additionally calibrating a camera overlooking the scene and includes recovery of the scene plane as a convenient side effect.
+Next, (ii) the relative camera-projector-scene plane geometry is used to compute a plane-induced homography that distorts the image to be projected in a manner that it appear undistorted to the viewer, and placed in alignment with the axes of a proxy image of the scene.
+These two points are treated in Sections~\ref{sec:approach:geometry} and \ref{sec:approach:homography}, respectively.
 
-The hardware setup employed
+\subsection{Recovering Geometry}\label{sec:approach:geometry}
 
-\subsection{Camera-projector Calibration}
+\begin{figure}
+    \centering
+		\subfloat[\centering Circles pattern image, in image plane of projector (detected 2D circles pattern center points overlain).]{{\includegraphics[height=3.25cm]{images/circles.png} }}
+    \qquad
+    \subfloat[\centering Projector calibration image (one of several), in image plane of camera (detected 2D projected circles pattern center points and chessboard corners overlain).]{{\includegraphics[height=3.25cm]{images/detections.png}}}
+		\caption{Recovering 2D positions in support of projector calibration.
+(a) 2D positions of the 2D-3D correspondences to be used for calibrating the projector are obtained by detecting---in the image plane of the projector---the circle centers in the circles pattern image projected by the projector to each of the target locations in the scene plane.
+(b) For each such target location, an image is acquired from the viewpoint of the camera and the circle centers of the projected circles pattern are detected.
+A chessboard pattern to be used for recovering the local scene plane is placed near the projected pattern, whose corners are likewise detected.
+Since the projected circles pattern lies in this plane, we are in a subsequent step able to rely on that plane to derive the 3D positions corresponding to the 2D projected circles pattern center points detected in the image.} %At this stage, the chessboard is used to recover the local scene plane for each target location.
+    \label{fig:example}
+\end{figure}
 
-\paragraph{Camera calibration}
+\begin{figure}
+		\centerline{\includegraphics[scale=.5]{images/2d3d.png}}
+		\caption{Scene plane via camera resection with respect to chessboard correspondences (blue); 3D circles pattern points via intersection of back-projections of detected circles pattern points with scene plane (red).}
+    \label{fig:example}
+\end{figure}
 
-\paragraph{Projector calibration}
+Calibration of a projector (or camera) in the sense we employ the term here\footnote{We are referring to a geometric calibration; not, e.g., to a color calibration.} entails that one become able to project a scene point~$\mathbf{X}$ to its corresponding pixel~$\mathbf{x}$ in the projector's (or camera's) image plane, or compute the `back-projection' of $\mathbf{x}$, the ray from the projector's (or camera's) center of projection~$\mathbf{C}$ through $\mathbf{x}$ along which $\mathbf{X}$ must lie.
+Such a calibration can be expressed in terms of (i) a $3~\times~3$ calibration matrix~$\mathtt{K}$ derived from the projector's (or camera's) focal length and principal point, and (ii) the coefficients of a lens distortion model to account for radial or tangential distortions caused by the lens system.
 
-The pose of the projector is given in the coordinate frame of the camera.
+In practice, calibrating a camera relies on (i) establishing 2D-3D correspondences between pixels in the camera's image plane and corresponding points in the scene, and on (ii) using those correspondences as input to an optimization procedure that outputs the calibration matrix $\mathtt{K}$, the associated lens distortion model coefficients, and, for each calibration image, the pose (i.e., position and orientation) of the camera relative to the 3D points.
+To calibrate a camera, a calibration surface such as a chessboard pattern is used to identify the correspondences.
+Calibration of a projector can be carried out in precisely the same manner insofar as step (ii) is concerned; the major difference in projector calibration relative to camera calibration concerns the manner in which 2D-3D correspondences are identified, i.e., between pixels in the image plane of the projector and the corresponding points in the scene.
+A convenient consequences of the approach we take to identifying the 3D points of the 2D-3D correspondences needed for projector calibration is, for each target location, recovery of the pose (i.e., position and orientation) of (i) the projector and of (ii) the scene plane, both relative to the coordinate frame of the camera.
 
-\subsection{Correcting for Distortion}
+\paragraph{Camera calibration} The recovery of 2D-3D correspondences in support of camera calibration is carried out, conventionally, by relying on a planar calibration surface of known scale to automatically identify correspondences between the 3D points on the calibration surface and their 2D correspondences in the image plane.
+The classical calibration surface is a chessboard pattern.
+The 3D corner points of the chessboard are obtained \textit{a priori} in a coordinate system defined in the plane of the chessboard\footnote{E.g., $(0,0,0), (1.5,0,0), (3,0,0), \dots, (9,7.5,0)$ for a chessboard with $7~\times{}~6$ corners, with each square of length and width of 1.5 unit, respectively.
+Note that the units of the chessboard's 3D points give the units of the camera calibration, and---owing to how our projector calibration relies on the camera calibration---of the projector calibration as well.}, requiring knowledge only of the dimensions of the chessboard pattern and of the length of a side of a chessboard square.
+The corresponding 2D points are obtained, in the same order, using a specialized algorithm.
+A set of calibration images is acquired, each with the calibration pattern visible in a different part of the image plane, and such that the center and all corners and edges of the image plane are covered.
+2D-3D correspondences are then recovered for each calibration image, and the resulting list is passed on as input to an optimization procedure that relies on bundle adjustment to yield the camera calibration matrix~$\mathtt{K}_\text{cam}$ and the associated lens distortion model coefficients.
+
+\paragraph{Projector calibration} As in the case of camera calibration, projector calibration relies on 2D-3D correspondences, yet in this case we obtain them by on projecting a pattern of circles.
+We rely on an algorithm to detect the circle pattern center points in the circles pattern image in the image plane of the projector, gibing the 2D positions of our 2D-3D correspondences.
+We project the circles pattern image to each of the target locations, yielding a projector calibration image for each.
+Given a calibration image, we can detect the centers of the \textit{projected} circles pattern; if we know the scene plane, then given such a 2D circle center $\mathbf{x}$, its 3D correspondence is obtained by intersecting the back-projection of $\mathbf{x}$ with the scene plane.
+Since the algorithm that yields 2D circle centers does so in a consistent ordering, we thus obtain the 2D-3D correspondences required for projector calibration, yielding the projector calibration matrix~$\mathtt{K}_\text{proj}$.
+
+We recover the scene plane via camera resection by applying a PnP algorithm to the with respect to 2D-3D correspondences of a chessboard pattern.
+Note that this step is separate from camera calibration, yet could well be carried out using the same pattern used in the camera calibration step.\footnote{The critical point is that the pattern should ideally be coplanar with the local scene plane, meaning its height above the scene plane should not exceed a few millimeters.} While a single image of such a chessboard pattern placed on the floor could be sufficient if the floor is even, we place a chessboard pattern in close proximity to the projected circles pattern in each projector calibration image in order to recover the scene plane locally to each target location in order to account for the possibility of an uneven floor.
+
+The pose (i.e., position and orientation) of the projector, for each projector calibration image, is provided alongside $\mathtt{K}_\text{proj}$ by the aforementioned optimization procedure based on bundle adjustment.
+Note that for a fixed projector with steerable mirror, given a projector calibration image, the recovered projector's pose is the pose the projector would have to have to project to the given target location \textit{in the absence of the mirror}.
+
+\subsection{Correcting for Projective Distortion}\label{sec:approach:homography}
+
+If the projector is calibrated and the scene plane is known in the coordinate frame of the projector, a `virtual' projector (with the same calibration $\mathtt{K}$ and lens distortion coefficients) can be placed elsewhere relative to the scene plane.
+The projection of an image from the virtual projector to the scene plane can in turn be projected to the image plane of the original projector;
 

### 2021-06-20 src/apps/calibrateProj/calibrateProj.cpp

@@ -187,17 +187,65 @@ void display()
             if (showVirtual)
                 pointCloud2Circles->display(pointSize, 0, 1, 0);
             else
+            {
                 pointCloudCircles->display(pointSize, 1, 0, 0);
+
+                std::vector<cv::Point3f> points = pointCloudCircles->getPoints();
+                std::vector<cv::Point3f> points2 = pointCloud->getPoints();
+
+                float f = cams[0].getf();
+                float CCDWidth_half_mm = 20.0 * 0.01 * 0.5;
+
+                for (int ptIdx = 0; ptIdx < points.size(); ptIdx++)
+                {
+                    glBegin(GL_POINTS);
+                        glPointSize(0.5 * pointSize);
+                        glColor3f(1.0, 0.0, 0.0);
+                        glVertex3f(CCDWidth_half_mm * points[ptIdx].x / points[ptIdx].z, CCDWidth_half_mm * points[ptIdx].y / points[ptIdx].z, CCDWidth_half_mm);
+                    glEnd();
+
+                    glBegin(GL_LINES);
+                        glColor3f(0.9, 0.9, 0.9);
+                        glVertex3f(0, 0, 0);
+                        glVertex3f(points[ptIdx].x, points[ptIdx].y, points[ptIdx].z);
+                    glEnd();
+
+                    glBegin(GL_POINTS);
+                        glPointSize(0.5 * pointSize);
+                        glColor3f(1.0, 0.0, 0.0);
+                        glVertex3f(CCDWidth_half_mm * points[ptIdx].x / points[ptIdx].z, CCDWidth_half_mm * points[ptIdx].y / points[ptIdx].z, CCDWidth_half_mm);
+                    glEnd();
+                }
+                   
+                for (int ptIdx = 0; ptIdx < points2.size(); ptIdx++)
+                {
+                    glBegin(GL_POINTS);
+                        glPointSize(0.5 * pointSize);
+                        glColor3f(0.0, 0.0, 1.0);
+                        glVertex3f(CCDWidth_half_mm * points2[ptIdx].x / points2[ptIdx].z, CCDWidth_half_mm * points2[ptIdx].y / points2[ptIdx].z, CCDWidth_half_mm);
+                    glEnd();
+                }
+            }
         }
 
-        planeVis->display(2.5, pointSize);
+        planeVis->display(1.5, pointSize);
 
         float offset = 0.02;
         for (int i = 0; i < cams.size(); i++)
[…]

### 2021-06-20 src/cpslib/libCalib/CalibPattern.cpp

@@ -128,7 +128,7 @@ void CalibPattern::findChessboardAndCirclesImPts(std::string inDir, int numIms, 
             applyIntrinsics, K, distCoeffs);
 
         std::stringstream ss2;
-        ss2 << "projector calibration image (50% resized)";
+        ss2 << "projector calibration image " << ss1.str() << " (50% resized)";
 
         cv::resize(imVis, imVis, cv::Size(imVis.cols * 0.5, imVis.rows * 0.5));
 

### 2021-06-22 paper/elsarticle/elsarticle-template.tex
paper

--- 2021-06-14 docs/elsarticle/elsarticle-template.tex
@@
 \begin{abstract}
-Augmented reality (AR) promises to enable use cases in industrial settings that include the embedding of work step instructions directly into the scene, potentially reducing or altogether obviating the need for workers to refer to instructions in paper form or on a screen.
-In turn, spatial AR is a form of AR whereby the augmentation of the scene is carried out using a projector, with the advantage of rendering the augmentation visible to all onlookers simultaneously without calling for each to wear AR glasses.
+Augmented reality (AR) promises to enable use cases in industrial settings that include the embedding of assembly instructions directly into the scene, potentially reducing or altogether obviating the need for workers to refer to instructions in paper form or on a screen.
+In turn, \textit{spatial} AR is a form of AR whereby the augmentation of the scene is carried out using a projector, with the advantage of rendering the augmentation visible to all onlookers simultaneously without calling for each to wear some form of head mounted display.
 However, care must be taken to distort the images to be projected in a manner that they appear undistorted to the viewer, since the geometry of the scene as it relates to the geometry of the projector plays a role in how the pixels of the projector's image plane map to points in the scene.
@@
 
-Augmented reality (AR) promises to enable use cases in industrial settings that include the embedding of work step instructions directly into the scene, potentially reducing or altogether doing away with the need for workers to refer to instructions in paper form or on a screen.
-Typically, AR works by embedding the augmentation in an image of the scene acquired from the viewpoint of a single individual, with the resulting augmented image in turn displayed using some form of AR glasses.
-Reliance on AR glasses has two adverse consequences: (i) AR glasses must be worn by each individual wishing to view the augmentation, and (ii) such glasses---in some cases taking the form of a helmet in order to house multiple sensors in support of accurately tracking the viewpoint of the viewer relative to the scene---can be obtrusive.
-Spatial AR is a form of augmented reality carried out not by embedding the augmentation in an image of the scene as with AR glasses, but by projection to the scene itself, thus eliminating both aforementioned problems.
+Augmented reality (AR) \cite{van2010survey,zhou2008trends} promises to enable use cases in industrial settings that include the embedding of assembly instructions directly into the scene \cite{schlund2018moglichkeiten,masood2019augmented,gattullo2019towards}, potentially reducing or altogether doing away with the need for workers to refer to instructions in paper form or on a screen.
+Typically, AR works by embedding the augmentation in an image of the scene acquired from the viewpoint of a single individual, with the resulting augmented image in turn displayed using some form of head mounted display.
+Reliance on head mounted displays has two adverse consequences: (i) a head mounted display must be worn by each individual wishing to view the augmentation, and (ii) such glasses---in some cases taking the form of a helmet in order to house multiple sensors in support of accurately tracking the viewpoint of the viewer relative to the scene---can be obtrusive.
+Spatial AR is a form of augmented reality carried out not by embedding the augmentation in an image of the scene as with head mounted displays, but by projection to the scene itself \cite{bimber2019spatial}, thus eliminating both aforementioned problems.
 Yet considering for the moment a planar surface to be augmented, unless the projector faces the surface frontally, the bounds of a projected rectangular image will not appear rectangular; more generally, they will instead appear trapezoidal (i.e., the image will appear distorted).
@@
 
-The problem of correcting for distortions as outlined in Section~\ref{sec:intro} requires knowledge of the manner in which the respective rays through the pixels of the projector's image plane fan out into the scene, and the geometry of the scene itself within at least the projector’s field of view, as illustrated in Figure 1.
+Correcting for projective distortions of the sort outlined in Section~\ref{sec:intro} calls for knowledge of the manner in which the respective rays through the pixels of the projector's image plane fan out into the scene (i.e., we must `calibrate' the projector) and the geometry of the scene itself (i.e., we must recover the scene plane relative to the projector) within at least the projector's field of view.
 This is because the scene point `illuminated' by a pixel in the projector's image plane is given by intersecting its corresponding ray with the geometry of the scene surface.
-To model this interaction calls for a one-time camera-projector calibration, which includes recovery of the scene plane.
-Next, the relative camera-projector-scene plane geometry is used to compute a plane-induced homography.
+To model this interaction, we (i) carry out a one-time projector calibration, which in our approach calls for additionally calibrating a camera overlooking the scene and includes recovery of the scene plane as a convenient side effect.
+Next, we use the relative camera-projector-scene plane geometry to (ii) compute a plane-induced homography that distorts the image to be projected in a manner that it appear undistorted to the viewer, and placed in alignment with the axes of a proxy image of the scene.
+These two points are treated in Sections~\ref{sec:approach:geometry} and \ref{sec:approach:homography}, respectively.
 
-The hardware setup employed
+\subsection{Recovering Geometry}\label{sec:approach:geometry}
 
-\subsection{Camera-projector Calibration}
+\begin{figure}
+    \centering
+		\subfloat[\centering Circles pattern image, in image plane of projector (detected 2D circles pattern center points overlain).]{{\includegraphics[height=3.25cm]{images/circles.png} }}
+    \qquad
+    \subfloat[\centering Projector calibration image (one for each target location), in image plane of camera (detected 2D projected circles pattern center points and chessboard corners overlain).]{{\includegraphics[height=3.25cm]{images/detections.png}}}
+		\caption{Recovering 2D positions in support of projector calibration.
+(a) 2D positions of the 2D-3D correspondences to be used for calibrating the projector are obtained by detecting---in the image plane of the projector---the circle centers in the circles pattern image projected by the projector to each of the target locations in the scene plane.
+(b) For each such target location, an image is acquired from the viewpoint of the camera and the circle centers of the projected circles pattern are detected.
+A chessboard pattern to be used for recovering the local scene plane is placed near the projected pattern, whose corners are likewise detected.} %At this stage, the chessboard is used to recover the local scene plane for each target location.
+Since the projected circles pattern lies in this plane, we are in a subsequent step able to rely on that plane to derive the 3D positions corresponding to the 2D projected circles pattern center points detected in the image.
+    \label{fig:2d}
+\end{figure}
 
-\paragraph{Camera calibration}
+\begin{figure}
+		\centerline{\includegraphics[scale=.35]{images/2d3d.png}}
+		\caption{Scene plane (gray) obtained via spatial resection with respect to 2D-3D correspondences obtained using a chessboard pattern (blue); 3D circles pattern points obtained by intersection with the scene plane of back-projections (likewise gray) of circles pattern center points detected in the image plane (red).
+Note that the points in the figure are the points recovered for the projector calibration image in Figure~\ref{fig:2d}(b), acquired by the camera (frustum in black, with up vector in red).}
+    \label{fig:3d}
+\end{figure}
 
-\paragraph{Projector calibration}
+Calibration of a projector (or camera) in the sense we employ the term here\footnote{We are referring to a geometric calibration; not, e.g., to a color calibration.} entails that one become able to project a scene point~$\mathbf{X}$ to its corresponding pixel~$\mathbf{x}$ in the projector's (or camera's) image plane, or compute the `back-projection' of $\mathbf{x}$, i.e., the ray from the projector's (or camera's) center of projection~$\mathbf{C}$ through $\mathbf{x}$ along which $\mathbf{X}$ must lie.
+Such a calibration can be expressed in terms of (i) a $3~\times{}~3$ calibration matrix~$\mathtt{K}$ derived from the projector's (or camera's) focal length and principal point \cite{Hartley2004}, and (ii) the coefficients of a lens distortion model to correct for radial or tangential distortions caused by the lens system \cite{duane1971close}.
 
-The pose of the projector is given in the coordinate frame of the camera.
+In practice, calibrating a camera relies on (i) establishing 2D-3D correspondences between pixels in the camera's image plane and corresponding points in the scene, and on (ii) using those correspondences as input to an optimization procedure that relies on bundle adjustment \cite{triggs1999bundle} to output the calibration matrix $\mathtt{K}$, the associated lens distortion model coefficients, and, for each calibration image, the pose (i.e., position and orientation) of the camera relative to the 3D points \cite{Hartley2004,zhang2000flexible}.
+To calibrate a camera, a calibration surface such as a chessboard pattern is used to identify the correspondences.
+Calibration of a projector can be carried out in precisely the same manner insofar as step (ii) is concerned; the major difference in projector calibration relative to camera calibration concerns the manner in which 2D-3D correspondences are identified, i.e., between pixels in the image plane of the projector and the corresponding points in the scene.
+A convenient consequences of the approach we take to identifying the 3D points of the 2D-3D correspondences needed for projector calibration is, for each target location, recovery of the pose (i.e., position and orientation) of (i) the projector and of (ii) the scene plane, both relative to the coordinate frame of the camera.
 
-\subsection{Correcting for Distortion}
+\paragraph{Camera calibration} The recovery of 2D-3D correspondences in support of camera calibration is carried out by relying on a planar calibration surface of known scale to automatically identify correspondences between the 3D points on the calibration surface and their 2D correspondences in the image plane.
+The classical calibration surface is a chessboard pattern.
+The 3D corner points of the chessboard are obtained \textit{a priori} in a coordinate system defined in the plane of the chessboard\footnote{E.g., $(0,0,0), (1.5,0,0), (3,0,0), \dots, (9,7.5,0)$ for a chessboard with $7~\times{}~6$ corners, with each square of length and width of 1.5 unit, respectively.
+Note that the units of the chessboard's 3D points give the units of the camera calibration, and---owing to how our projector calibration relies on the camera calibration---of the projector calibration as well.}, requiring knowledge only of the dimensions of the chessboard pattern and of the length of a side of a chessboard square.
+The corresponding 2D points are obtained, in the same order, using a specialized algorithm.
+A set of calibration images is acquired, each with the calibration pattern visible in a different part of the image plane, and such that the center and all corners and edges of the image plane are covered.
+2D-3D correspondences are then recovered for each calibration image, and the resulting list is passed on as input to an optimization procedure that relies on bundle adjustment to yield the camera calibration matrix~$\mathtt{K}_\text{cam}$ and the associated lens distortion model coefficients.
+
+\paragraph{Projector calibration} As in the case of camera calibration, projector calibration relies on 2D-3D correspondences, yet we obtain them in this case by projecting a pattern of circles.
+We rely on an algorithm to detect the circle pattern center points in the circles pattern image in the image plane of the projector, giving the 2D positions of our 2D-3D correspondences.
+We project the circles pattern image to each of the target locations, and use the camera to acquire a projector calibration image for each.
+Given a projector calibration image, we again detect the centers of the \textit{projected} circles pattern; given the scene plane (a point to which we shall return in the paragraph that follows), then given such a 2D circle center $\mathbf{x}$, its 3D correspondence is obtained by intersecting the back-projection of $\mathbf{x}$ with the scene plane.
+Since the algorithm that yields 2D circle centers does so in a consistent ordering, we thus obtain the 2D-3D correspondences between the projector's image plane and the scene required for projector calibration, yielding the projector calibration matrix~$\mathtt{K}_\text{proj}$.
+
+We recover the scene plane via spatial resection by applying a PnP algorithm \cite{terzakis2020consistently} to the with respect to 2D-3D correspondences of a chessboard pattern.
+Note that this step is separate from camera calibration, yet could well be carried out using the same pattern used in the camera calibration step.\footnote{The critical point is that the pattern should ideally be coplanar with the local scene plane, meaning its height above the scene plane should not exceed a few millimeters.} While a single image of such a chessboard pattern placed on the floor could be sufficient if the floor is even, we place a chessboard pattern in close proximity to the projected circles pattern in each projector calibration image in order to recover the scene plane locally to each target location, in order to account for the possibility of an uneven floor.
+
+The pose (i.e., position and orientation) of the projector, for each projector calibration image, is provided alongside $\mathtt{K}_\text{proj}$ by the aforementioned optimization.
+Note that for a fixed projector with steerable mirror, given a projector calibration image, the recovered projector's pose is the pose the projector would have to have had to project to the given target location \textit{in the absence of the mirror}.
+
+\subsection{Correcting for Projective Distortion}\label{sec:approach:homography}
+
+If the projector is calibrated and the scene plane is known in the coordinate frame of the projector, a `virtual' projector (with the same calibration $\mathtt{K}$ and lens distortion coefficients) can be placed elsewhere relative to the scene plane.
+The projection of an image from the virtual projector to the scene plane can in turn be projected to the image plane of the original projector;
 

### 2021-06-22 paper/elsarticle/elsarticle-template.tex
paper

@@
 Augmented reality (AR) promises to enable use cases in industrial settings that include the embedding of assembly instructions directly into the scene, potentially reducing or altogether obviating the need for workers to refer to instructions in paper form or on a screen.
-In turn, \textit{spatial} AR is a form of AR whereby the augmentation of the scene is carried out using a projector, with the advantage of rendering the augmentation visible to all onlookers simultaneously without calling for each to wear some form of head mounted display.
+In turn, \textit{spatial} AR is a form of AR whereby the augmentation of the scene is carried out using a projector, with the advantage of rendering the augmentation visible to all onlookers simultaneously without calling for each to wear some form of head-mounted display.
 However, care must be taken to distort the images to be projected in a manner that they appear undistorted to the viewer, since the geometry of the scene as it relates to the geometry of the projector plays a role in how the pixels of the projector's image plane map to points in the scene.
@@
 Augmented reality (AR) \cite{van2010survey,zhou2008trends} promises to enable use cases in industrial settings that include the embedding of assembly instructions directly into the scene \cite{schlund2018moglichkeiten,masood2019augmented,gattullo2019towards}, potentially reducing or altogether doing away with the need for workers to refer to instructions in paper form or on a screen.
-Typically, AR works by embedding the augmentation in an image of the scene acquired from the viewpoint of a single individual, with the resulting augmented image in turn displayed using some form of head mounted display.
-Reliance on head mounted displays has two adverse consequences: (i) a head mounted display must be worn by each individual wishing to view the augmentation, and (ii) such glasses---in some cases taking the form of a helmet in order to house multiple sensors in support of accurately tracking the viewpoint of the viewer relative to the scene---can be obtrusive.
-Spatial AR is a form of augmented reality carried out not by embedding the augmentation in an image of the scene as with head mounted displays, but by projection to the scene itself \cite{bimber2019spatial}, thus eliminating both aforementioned problems.
+Typically, AR works by embedding the augmentation in an image of the scene acquired from the viewpoint of a single individual, with the resulting augmented image in turn displayed using some form of head-mounted display.
+Reliance on head-mounted displays has two adverse consequences: (i) a head-mounted display must be worn by each individual wishing to view the augmentation, and (ii) such a head-mounted display---in some cases taking the form of a helmet in order to house multiple sensors in support of accurately tracking the viewpoint of the viewer relative to the scene---can be obtrusive.
+Spatial AR is a form of augmented reality carried out not by embedding the augmentation in an image of the scene as with head-mounted displays, but by projection to the scene itself \cite{bimber2019spatial}, thus eliminating both aforementioned problems.
 Yet considering for the moment a planar surface to be augmented, unless the projector faces the surface frontally, the bounds of a projected rectangular image will not appear rectangular; more generally, they will instead appear trapezoidal (i.e., the image will appear distorted).
@@
 The setup was mounted on the ceiling of the Pilotfabrik\footnote{\url{https://www.pilotfabrik.at/}} of TU Vienna, a collaborative space for research on Industry 4.0 topics situated in Vienna, Austria.
-The floorspace used for our experiments measured dimensions of X~m~$\times$~Y~m; the projector was mounted at approximately the center of this space, at a height of ca.\ XX m.
+The floorspace used for our experiments measured dimensions of ca.\ X~m~$\times$~Y~m; the projector was mounted at approximately the center of this space, at a height of ca.\ XX m.
 
@@
 
-Calibration of a projector (or camera) in the sense we employ the term here\footnote{We are referring to a geometric calibration; not, e.g., to a color calibration.} entails that one become able to project a scene point~$\mathbf{X}$ to its corresponding pixel~$\mathbf{x}$ in the projector's (or camera's) image plane, or compute the `back-projection' of $\mathbf{x}$, i.e., the ray from the projector's (or camera's) center of projection~$\mathbf{C}$ through $\mathbf{x}$ along which $\mathbf{X}$ must lie.
+Calibration of a projector (or camera) in the sense we employ the term here\footnote{We are referring to a geometric calibration; not, e.g., to a color calibration.} entails that one become able to project a scene point~$\mathbf{X} \in \mathbb{R}^3$ to its corresponding pixel~$\mathbf{x} \in \mathbb{R}^2$ in the projector's (or camera's) image plane, or compute the `back-projection' of $\mathbf{x}$ (cf.\ Figure~\ref{fig:3d}), i.e., the ray from the projector's (or camera's) center of projection~$\mathbf{C} \in \mathbb{R}^3$ through $\mathbf{x}$ along which $\mathbf{X}$ must lie.
 Such a calibration can be expressed in terms of (i) a $3~\times{}~3$ calibration matrix~$\mathtt{K}$ derived from the projector's (or camera's) focal length and principal point \cite{Hartley2004}, and (ii) the coefficients of a lens distortion model to correct for radial or tangential distortions caused by the lens system \cite{duane1971close}.
 
-In practice, calibrating a camera relies on (i) establishing 2D-3D correspondences between pixels in the camera's image plane and corresponding points in the scene, and on (ii) using those correspondences as input to an optimization procedure that relies on bundle adjustment \cite{triggs1999bundle} to output the calibration matrix $\mathtt{K}$, the associated lens distortion model coefficients, and, for each calibration image, the pose (i.e., position and orientation) of the camera relative to the 3D points \cite{Hartley2004,zhang2000flexible}.
+In practice, calibrating a camera relies on (i) establishing 2D-3D correspondences between pixels in the camera's image plane and corresponding points in the scene, and on (ii) using those correspondences as input to an optimization procedure that relies on bundle adjustment \cite{triggs1999bundle} to output the calibration matrix~$\mathtt{K}$, the associated lens distortion model coefficients, and, for each calibration image, the pose (i.e., position and orientation) of the camera relative to the 3D points \cite{Hartley2004,zhang2000flexible}.
 To calibrate a camera, a calibration surface such as a chessboard pattern is used to identify the correspondences.
@@
 
-\paragraph{Camera calibration} The recovery of 2D-3D correspondences in support of camera calibration is carried out by relying on a planar calibration surface of known scale to automatically identify correspondences between the 3D points on the calibration surface and their 2D correspondences in the image plane.
+\paragraph{Camera calibration} The recovery of 2D-3D correspondences in support of camera calibration is carried out by relying on a planar calibration surface to automatically identify correspondences between the 3D points on the calibration surface and their 2D correspondences in the image plane.
 The classical calibration surface is a chessboard pattern.
@@
 Note that the units of the chessboard's 3D points give the units of the camera calibration, and---owing to how our projector calibration relies on the camera calibration---of the projector calibration as well.}, requiring knowledge only of the dimensions of the chessboard pattern and of the length of a side of a chessboard square.
-The corresponding 2D points are obtained, in the same order, using a specialized algorithm.
+The corresponding 2D points are obtained, in the same order, using a specialized algorithm \cite{bradski2000opencv}.
 A set of calibration images is acquired, each with the calibration pattern visible in a different part of the image plane, and such that the center and all corners and edges of the image plane are covered.
@@
 We project the circles pattern image to each of the target locations, and use the camera to acquire a projector calibration image for each.
-Given a projector calibration image, we again detect the centers of the \textit{projected} circles pattern; given the scene plane (a point to which we shall return in the paragraph that follows), then given such a 2D circle center $\mathbf{x}$, its 3D correspondence is obtained by intersecting the back-projection of $\mathbf{x}$ with the scene plane.
+Given a projector calibration image, we again detect the centers of the \textit{projected} circles pattern; given the scene plane (the recovery of which we shall turn to in the paragraph that follows), then given such a 2D circle center $\mathbf{x}$, its 3D correspondence is obtained by intersecting the back-projection of $\mathbf{x}$ with the scene plane.
 Since the algorithm that yields 2D circle centers does so in a consistent ordering, we thus obtain the 2D-3D correspondences between the projector's image plane and the scene required for projector calibration, yielding the projector calibration matrix~$\mathtt{K}_\text{proj}$.

### 2021-06-22 paper/elsarticle/elsarticle-template.tex
paper

@@
 
-Calibration of a projector (or camera) in the sense we employ the term here\footnote{We are referring to a geometric calibration; not, e.g., to a color calibration.} entails that one become able to project a scene point~$\mathbf{X} \in \mathbb{R}^3$ to its corresponding pixel~$\mathbf{x} \in \mathbb{R}^2$ in the projector's (or camera's) image plane, or compute the `back-projection' of $\mathbf{x}$ (cf.\ Figure~\ref{fig:3d}), i.e., the ray from the projector's (or camera's) center of projection~$\mathbf{C} \in \mathbb{R}^3$ through $\mathbf{x}$ along which $\mathbf{X}$ must lie.
+Calibration of a projector (or camera) in the sense we employ the term here\footnote{We are referring to a geometric calibration; not, e.g., to a color calibration.} renders one able to project a scene point~$\mathbf{X} \in \mathbb{R}^3$ to its corresponding pixel~$\mathbf{x} \in \mathbb{R}^2$ in the projector's (or camera's) image plane, or compute the `back-projection' of $\mathbf{x}$ (cf.\ Figure~\ref{fig:3d}), i.e., the ray from the projector's (or camera's) center of projection through $\mathbf{x}$ along which $\mathbf{X}$ must lie.
 Such a calibration can be expressed in terms of (i) a $3~\times{}~3$ calibration matrix~$\mathtt{K}$ derived from the projector's (or camera's) focal length and principal point \cite{Hartley2004}, and (ii) the coefficients of a lens distortion model to correct for radial or tangential distortions caused by the lens system \cite{duane1971close}.
@@
 \paragraph{Projector calibration} As in the case of camera calibration, projector calibration relies on 2D-3D correspondences, yet we obtain them in this case by projecting a pattern of circles.
-We rely on an algorithm to detect the circle pattern center points in the circles pattern image in the image plane of the projector, giving the 2D positions of our 2D-3D correspondences.
+We rely on an algorithm to detect the circle pattern center points in the circles pattern image in the image plane of the projector, giving the 2D positions of our 2D-3D correspondences (cf.\ Figure~\ref{fig:2d}(a)).
 We project the circles pattern image to each of the target locations, and use the camera to acquire a projector calibration image for each.
-Given a projector calibration image, we again detect the centers of the \textit{projected} circles pattern; given the scene plane (the recovery of which we shall turn to in the paragraph that follows), then given such a 2D circle center $\mathbf{x}$, its 3D correspondence is obtained by intersecting the back-projection of $\mathbf{x}$ with the scene plane.
+Given a projector calibration image, we again detect the centers of the \textit{projected} circles pattern (cf.\ Figure~\ref{fig:2d}(b)); given the scene plane (the recovery of which we shall turn to in the paragraph that follows) and such a 2D circle center $\mathbf{x}$, its 3D correspondence is obtained by intersecting the back-projection of $\mathbf{x}$ with the scene plane (cf.\ Figure~\ref{fig:3d}).
 Since the algorithm that yields 2D circle centers does so in a consistent ordering, we thus obtain the 2D-3D correspondences between the projector's image plane and the scene required for projector calibration, yielding the projector calibration matrix~$\mathtt{K}_\text{proj}$.

### 2021-06-22 paper/elsarticle/elsarticle-template.tex
paper

@@
 %% Group authors per affiliation:
-\author{Michael Horn\'{a}\v{c}ek}
+\author{Michael Horn\'{a}\v{c}ek}\corref{mycorrespondingauthor}
+\cortext[mycorrespondingauthor]{Corresponding author}
+\ead{michael.hornacek@tuwien.ac.at}
 \author{Hans K\"{u}ffner-McCauley}
@@
 We propose a system that produces the effect of keystone correction analytically, and that intuitively places the desired augmentations in a manner aligned with the axes of an image of the scene acquired by a camera.
-Moreover, our system is able to handle a projector equipped with a steerable mirror, enabling wide-area factory floor augmentation exceeding the projector's own immediate field of view.
+Moreover, our system is able to handle a projector equipped with a steerable mirror, enabling wide-area factory floor augmentation exceeding the bounds of the projector's own immediate field of view.
 \end{abstract}
@@
 We recover the scene plane via spatial resection by applying a PnP algorithm \cite{terzakis2020consistently} to the with respect to 2D-3D correspondences of a chessboard pattern.
-Note that this step is separate from camera calibration, yet could well be carried out using the same pattern used in the camera calibration step.\footnote{The critical point is that the pattern should ideally be coplanar with the local scene plane, meaning its height above the scene plane should not exceed a few millimeters.} While a single image of such a chessboard pattern placed on the floor could be sufficient if the floor is even, we place a chessboard pattern in close proximity to the projected circles pattern in each projector calibration image in order to recover the scene plane locally to each target location, in order to account for the possibility of an uneven floor.
+Note that this step is separate from camera calibration, yet could well be carried out using the same calibration pattern used in the camera calibration step.\footnote{The critical point is that the pattern should ideally be coplanar with the local scene plane, meaning its height above the scene plane should not exceed a few millimeters.} While a single image of such a chessboard pattern placed on the floor could be sufficient if the floor is even, we place a chessboard pattern in close proximity to the projected circles pattern in each projector calibration image in order to recover the scene plane locally to each target location, in order to account for the possibility of an uneven floor.
 
-The pose (i.e., position and orientation) of the projector, for each projector calibration image, is provided alongside $\mathtt{K}_\text{proj}$ by the aforementioned optimization.
+The pose (i.e., position and orientation) of the projector, for each projector calibration image, is provided alongside $\mathtt{K}_\text{proj}$ by the aforementioned optimization procedure.
 Note that for a fixed projector with steerable mirror, given a projector calibration image, the recovered projector's pose is the pose the projector would have to have had to project to the given target location \textit{in the absence of the mirror}.
+As this is sufficient for our needs in Section~\ref{sec:approach:homography}, it is in this sense that our system is able to handle a projector equipped with a steerable mirror, without need for modeling the steerable mirror explicitly.
 
@@
 
+We presented a system that produces the effect of keystone correction analytically, and that intuitively places the desired augmentations in a manner aligned with the axes of an image of the scene acquired by a camera.
+Moreover, we showed our system to be able to handle a projector equipped with a steerable mirror, enabling wide-area factory floor augmentation exceeding the bounds of the projector's own immediate field of view.
+Our evaluation demonstrated ...
+
 \bibliography{mybibfile}

### 2021-06-22 paper/elsarticle/elsarticle-template.tex
paper

@@
 
+A natural extension of this work would be to extend it to handle planar and non-planar scenes alike.
+To handle non-planar scenes would call for a change in how scene geometry is recovered and how distortion of the image to be projected is carried out; projector calibration---presented here as a one-time step---could, however, remain unchanged.
+
 \bibliography{mybibfile}

### 2021-06-22 paper/elsarticle/elsarticle-template.tex
paper

@@
 
-We presented a system that produces the effect of keystone correction analytically, and that intuitively places the desired augmentations in a manner aligned with the axes of an image of the scene acquired by a camera.
+We presented a spatial AR system for planar scenes that produces the effect of keystone correction analytically, and that intuitively places the desired augmentations in a manner aligned with the axes of an image of the scene acquired by a camera.
 Moreover, we showed our system to be able to handle a projector equipped with a steerable mirror, enabling wide-area factory floor augmentation exceeding the bounds of the projector's own immediate field of view.

### 2021-06-22 paper/elsarticle/elsarticle-template.tex
Update elsarticle-template.tex

@@
 
-We recover the scene plane via spatial resection by applying a PnP algorithm \cite{terzakis2020consistently} to the with respect to 2D-3D correspondences of a chessboard pattern.
+We recover the scene plane via spatial resection by applying a PnP algorithm \cite{terzakis2020consistently} to the 2D-3D correspondences obtained using a chessboard pattern.
 Note that this step is separate from camera calibration, yet could well be carried out using the same calibration pattern used in the camera calibration step.\footnote{The critical point is that the pattern should ideally be coplanar with the local scene plane, meaning its height above the scene plane should not exceed a few millimeters.} While a single image of such a chessboard pattern placed on the floor could be sufficient if the floor is even, we place a chessboard pattern in close proximity to the projected circles pattern in each projector calibration image in order to recover the scene plane locally to each target location, in order to account for the possibility of an uneven floor.

### 2021-06-24 paper/elsarticle/elsarticle-template.tex
Update elsarticle-template.tex
paper

@@
 \author{Hans K\"{u}ffner-McCauley}
+\author{Majesa Trimmel}
 \author{Sebastian Schlund}
@@
 
-If the projector is calibrated and the scene plane is known in the coordinate frame of the projector, a `virtual' projector (with the same calibration $\mathtt{K}$ and lens distortion coefficients) can be placed elsewhere relative to the scene plane.
-The projection of an image from the virtual projector to the scene plane can in turn be projected to the image plane of the original projector;
+If the projector is calibrated and its pose relative to the scene plane is known, a `virtual' projector (with the same calibration $\mathtt{K}$ and lens distortion coefficients) can be placed elsewhere relative to the scene plane.
+If we for a moment imagine that the projector---at its recovered `original' pose---functions as a camera,\footnote{Recall that the calibration matrix~$\mathtt{K}$ enables computing the projection of a scene point to the image plane (the function of a camera), or computing the back-projection of a pixel in the image plane, giving a ray into the scene (along which a projector illuminates the scene with the given pixel).} then (i) projecting an image to the scene plane \textit{from the viewpoint of the virtual projector} and (ii) acquiring the resulting projected image from the viewpoint of the recovered {original projector} gives the desired distortion that is to be applied, if the image is to be projected from the viewpoint of the original projector.
+Projecting an image distorted in this manner to the scene plane \textit{from the viewpoint of the original projector} then has the same effect as projecting the undistorted image to the scene plane from the viewpoint of the virtual projector.
+Given the scene plane, the projector, and the virtual projector, this distortion can computed using a plane-induced homography.
 
@@
 
-A natural extension of this work would be to extend it to handle planar and non-planar scenes alike.
+A natural extension of this work would be to address non-planar scenes.
 To handle non-planar scenes would call for a change in how scene geometry is recovered and how distortion of the image to be projected is carried out; projector calibration---presented here as a one-time step---could, however, remain unchanged.

### 2021-06-24 paper/elsarticle/elsarticle-template.tex
paper

@@
     \centering
-		\subfloat[\centering Circles pattern image, in image plane of projector (detected 2D circles pattern center points overlain).]{{\includegraphics[height=3.25cm]{images/circles.png} }}
+		\subfloat[\centering Circles pattern image, in image plane of projector.]{{\includegraphics[height=3.25cm]{images/circles.png} }}
     \qquad
-    \subfloat[\centering Projector calibration image (one for each target location), in image plane of camera (detected 2D projected circles pattern center points and chessboard corners overlain).]{{\includegraphics[height=3.25cm]{images/detections.png}}}
+    \subfloat[\centering Projector calibration image (one for each target location), in image plane of camera.]{{\includegraphics[height=3.25cm]{images/detections.png}}}
 		\caption{Recovering 2D positions in support of projector calibration.
 (a) 2D positions of the 2D-3D correspondences to be used for calibrating the projector are obtained by detecting---in the image plane of the projector---the circle centers in the circles pattern image projected by the projector to each of the target locations in the scene plane.
-(b) For each such target location, an image is acquired from the viewpoint of the camera and the circle centers of the projected circles pattern are detected.
-A chessboard pattern to be used for recovering the local scene plane is placed near the projected pattern, whose corners are likewise detected.} %At this stage, the chessboard is used to recover the local scene plane for each target location.
+(b) For each such target location, an image is acquired from the viewpoint of the camera and the circle centers of the projected circles pattern are detected, in the image plane of the camera.
+A chessboard pattern to be used for recovering the local scene plane is placed near the projected pattern, whose corners are likewise detected.
+Detected 2D projected circles pattern center points and chessboard corners overlain for illustration.} %At this stage, the chessboard is used to recover the local scene plane for each target location.
 Since the projected circles pattern lies in this plane, we are in a subsequent step able to rely on that plane to derive the 3D positions corresponding to the 2D projected circles pattern center points detected in the image.
@@
 
-\paragraph{Projector calibration} As in the case of camera calibration, projector calibration relies on 2D-3D correspondences, yet we obtain them in this case by projecting a pattern of circles.
-We rely on an algorithm to detect the circle pattern center points in the circles pattern image in the image plane of the projector, giving the 2D positions of our 2D-3D correspondences (cf.\ Figure~\ref{fig:2d}(a)).
+\paragraph{Projector calibration} As with camera calibration, projector calibration relies on 2D-3D correspondences, yet we obtain them in this case by projecting a pattern of circles.
+We rely on an algorithm to detect the circle pattern center points in the circles pattern image in the image plane of the projector \cite{bradski2000opencv}, giving the 2D positions of our 2D-3D correspondences (cf.\ Figure~\ref{fig:2d}(a)).
 We project the circles pattern image to each of the target locations, and use the camera to acquire a projector calibration image for each.
-Given a projector calibration image, we again detect the centers of the \textit{projected} circles pattern (cf.\ Figure~\ref{fig:2d}(b)); given the scene plane (the recovery of which we shall turn to in the paragraph that follows) and such a 2D circle center $\mathbf{x}$, its 3D correspondence is obtained by intersecting the back-projection of $\mathbf{x}$ with the scene plane (cf.\ Figure~\ref{fig:3d}).
+Given a projector calibration image, we again detect the circle centers of the \textit{projected} circles pattern (cf.\ Figure~\ref{fig:2d}(b)); given the scene plane (the recovery of which we shall turn to in the paragraph that follows) and such a 2D circle center $\mathbf{x}$, its 3D correspondence is obtained by intersecting the back-projection of $\mathbf{x}$ with the scene plane (cf.\ Figure~\ref{fig:3d}).
 Since the algorithm that yields 2D circle centers does so in a consistent ordering, we thus obtain the 2D-3D correspondences between the projector's image plane and the scene required for projector calibration, yielding the projector calibration matrix~$\mathtt{K}_\text{proj}$.
@@
 If the projector is calibrated and its pose relative to the scene plane is known, a `virtual' projector (with the same calibration $\mathtt{K}$ and lens distortion coefficients) can be placed elsewhere relative to the scene plane.
-If we for a moment imagine that the projector---at its recovered `original' pose---functions as a camera,\footnote{Recall that the calibration matrix~$\mathtt{K}$ enables computing the projection of a scene point to the image plane (the function of a camera), or computing the back-projection of a pixel in the image plane, giving a ray into the scene (along which a projector illuminates the scene with the given pixel).} then (i) projecting an image to the scene plane \textit{from the viewpoint of the virtual projector} and (ii) acquiring the resulting projected image from the viewpoint of the recovered {original projector} gives the desired distortion that is to be applied, if the image is to be projected from the viewpoint of the original projector.
+If we for a moment imagine that the projector---at its recovered `original' pose---functions as a camera,\footnote{Recall that the calibration matrix~$\mathtt{K}$ enables computing both (i) the projection of a scene point to the image plane (the function of a camera), or (ii) the back-projection of a pixel in the image plane, giving a ray into the scene (along which a projector illuminates the scene with the given pixel).} then (i) projecting an image to the scene plane \textit{from the viewpoint of the virtual projector} and (ii) acquiring the resulting projected image from the viewpoint of the original projector gives the desired distortion.
 Projecting an image distorted in this manner to the scene plane \textit{from the viewpoint of the original projector} then has the same effect as projecting the undistorted image to the scene plane from the viewpoint of the virtual projector.
-Given the scene plane, the projector, and the virtual projector, this distortion can computed using a plane-induced homography.
+This distortion can computed using a plane-induced homography, computed analytically as a function of the scene plane, the projector, and the virtual projector.
 

### 2021-06-25 paper/elsarticle/elsarticle-template.tex
paper

@@
 \author{Majesa Trimmel}
+\author{Patrick Rupprecht}
 \author{Sebastian Schlund}
@@
 In turn, \textit{spatial} AR is a form of AR whereby the augmentation of the scene is carried out using a projector, with the advantage of rendering the augmentation visible to all onlookers simultaneously without calling for each to wear some form of head-mounted display.
-However, care must be taken to distort the images to be projected in a manner that they appear undistorted to the viewer, since the geometry of the scene as it relates to the geometry of the projector plays a role in how the pixels of the projector's image plane map to points in the scene.
+Care must be taken, however, to distort the images to be projected in a manner that they appear undistorted to the viewer, since the geometry of the scene as it relates to the geometry of the projector plays a role in how the pixels of the projector's image plane map to points in the scene.
 For planar scene geometry (such as a floor, wall, or table), this can be done in a cumbersome manual process called keystone correction, often using software bundled with the projector.
@@
 
-Correcting for projective distortions of the sort outlined in Section~\ref{sec:intro} calls for knowledge of the manner in which the respective rays through the pixels of the projector's image plane fan out into the scene (i.e., we must `calibrate' the projector) and the geometry of the scene itself (i.e., we must recover the scene plane relative to the projector) within at least the projector's field of view.
+Correcting for projective distortions of the sort outlined in Section~\ref{sec:intro} can be achieved by modeling the manner in which the respective rays through the pixels of the projector's image plane fan out into the scene (i.e., we `calibrate' the projector) and the geometry of the scene itself (i.e., we recover the scene plane relative to the projector) within at least the projector's field of view.
 This is because the scene point `illuminated' by a pixel in the projector's image plane is given by intersecting its corresponding ray with the geometry of the scene surface.
@@
 Calibration of a projector (or camera) in the sense we employ the term here\footnote{We are referring to a geometric calibration; not, e.g., to a color calibration.} renders one able to project a scene point~$\mathbf{X} \in \mathbb{R}^3$ to its corresponding pixel~$\mathbf{x} \in \mathbb{R}^2$ in the projector's (or camera's) image plane, or compute the `back-projection' of $\mathbf{x}$ (cf.\ Figure~\ref{fig:3d}), i.e., the ray from the projector's (or camera's) center of projection through $\mathbf{x}$ along which $\mathbf{X}$ must lie.
-Such a calibration can be expressed in terms of (i) a $3~\times{}~3$ calibration matrix~$\mathtt{K}$ derived from the projector's (or camera's) focal length and principal point \cite{Hartley2004}, and (ii) the coefficients of a lens distortion model to correct for radial or tangential distortions caused by the lens system \cite{duane1971close}.
+Such a calibration can be expressed in terms of (i) a $3~\times{}~3$ calibration matrix~$\mathtt{K}$ derived from the projector's (or camera's) focal length and principal point \cite{Hartley2004}, and (ii) the coefficients of a lens distortion model used to correct for radial or tangential distortions caused by the lens system \cite{duane1971close}.
 
@@
 To calibrate a camera, a calibration surface such as a chessboard pattern is used to identify the correspondences.
-Calibration of a projector can be carried out in precisely the same manner insofar as step (ii) is concerned; the major difference in projector calibration relative to camera calibration concerns the manner in which 2D-3D correspondences are identified, i.e., between pixels in the image plane of the projector and the corresponding points in the scene.
-A convenient consequences of the approach we take to identifying the 3D points of the 2D-3D correspondences needed for projector calibration is, for each target location, recovery of the pose (i.e., position and orientation) of (i) the projector and of (ii) the scene plane, both relative to the coordinate frame of the camera.
+Calibration of a projector can be carried out in precisely the same manner insofar as step (ii) is concerned; the major difference in projector calibration relative to the calibrating a camera concerns the manner in which 2D-3D correspondences are identified, i.e., between pixels in the image plane of the projector and the corresponding points in the scene.
+A convenient consequence of the approach we take to identifying the 3D points of the 2D-3D correspondences needed for projector calibration is, for each target location, recovery of the pose (i.e., position and orientation) of (i) the projector and of (ii) the scene plane, both relative to the coordinate frame of the camera.
 
@@
 We project the circles pattern image to each of the target locations, and use the camera to acquire a projector calibration image for each.
-Given a projector calibration image, we again detect the circle centers of the \textit{projected} circles pattern (cf.\ Figure~\ref{fig:2d}(b)); given the scene plane (the recovery of which we shall turn to in the paragraph that follows) and such a 2D circle center $\mathbf{x}$, its 3D correspondence is obtained by intersecting the back-projection of $\mathbf{x}$ with the scene plane (cf.\ Figure~\ref{fig:3d}).
+Given a projector calibration image, we detect the circle centers of the \textit{projected} circles pattern (cf.\ Figure~\ref{fig:2d}(b)); given the scene plane (the recovery of which we shall turn to in the paragraph that follows) and such a 2D circle center $\mathbf{x}$, its 3D correspondence is obtained by intersecting the back-projection of $\mathbf{x}$ with the scene plane (cf.\ Figure~\ref{fig:3d}).
 Since the algorithm that yields 2D circle centers does so in a consistent ordering, we thus obtain the 2D-3D correspondences between the projector's image plane and the scene required for projector calibration, yielding the projector calibration matrix~$\mathtt{K}_\text{proj}$.

### 2021-06-25 paper/elsarticle/elsarticle-template.tex
paper

@@
 
-We propose a system that produces the effect of keystone correction analytically, and that intuitively places the desired augmentations in a manner aligned with the axes of an image of the scene acquired by a camera.
+We propose a spatial AR system that produces the effect of keystone correction analytically, and that intuitively places the desired augmentations in a manner aligned with the axes of an image of the scene acquired by a camera.
 Moreover, our system is able to handle a projector equipped with a steerable mirror, enabling wide-area factory floor augmentation exceeding the bounds of the projector's own immediate field of view.
@@
 Typically, AR works by embedding the augmentation in an image of the scene acquired from the viewpoint of a single individual, with the resulting augmented image in turn displayed using some form of head-mounted display.
-Reliance on head-mounted displays has two adverse consequences: (i) a head-mounted display must be worn by each individual wishing to view the augmentation, and (ii) such a head-mounted display---in some cases taking the form of a helmet in order to house multiple sensors in support of accurately tracking the viewpoint of the viewer relative to the scene---can be obtrusive.
+Reliance on head-mounted displays, however, has two adverse consequences: (i) a head-mounted display must be worn by each individual wishing to view the augmentation, and (ii) such a head-mounted display---in some cases taking the form of a helmet in order to house multiple sensors in support of accurately tracking the viewpoint of the viewer relative to the scene---can be obtrusive.
 Spatial AR is a form of augmented reality carried out not by embedding the augmentation in an image of the scene as with head-mounted displays, but by projection to the scene itself \cite{bimber2019spatial}, thus eliminating both aforementioned problems.
-Yet considering for the moment a planar surface to be augmented, unless the projector faces the surface frontally, the bounds of a projected rectangular image will not appear rectangular; more generally, they will instead appear trapezoidal (i.e., the image will appear distorted).
+Yet considering for the moment a planar surface to be augmented, unless the projector faces the surface frontally, the bounds of a projected rectangular image will not appear rectangular; more generally, they will instead appear trapezoidal (i.e., the projected image will appear distorted).
 Such distortions can be eliminated by carrying out a cumbersome manual process called keystone correction, often using software bundled with the projector.
 
-Using the $X$- and $Y$-axes of an image of the scene as a proxy, our contribution is to propose a system that produces the effect of keystone correction analytically, and in a manner aligning the axes of the augmentation with those of the proxy image.
+Using the $X$- and $Y$-axes of an image of the scene as a proxy, our contribution is to propose a spatial AR system that produces the effect of keystone correction analytically, and in a manner aligning the axes of the augmentation with those of the proxy image.
 We achieve this by distorting the image to be projected using a plane-induced homography computed to produce the effect of projecting the image not from the actual projector viewpoint, but in accordance with the viewpoint of a \textit{virtual} projector (i) facing directly downwards to the scene plane and (ii) rotated to place the axes of the image plane of the virtual projector in line with those of the camera.
@@
 
-Correcting for projective distortions of the sort outlined in Section~\ref{sec:intro} can be achieved by modeling the manner in which the respective rays through the pixels of the projector's image plane fan out into the scene (i.e., we `calibrate' the projector) and the geometry of the scene itself (i.e., we recover the scene plane relative to the projector) within at least the projector's field of view.
+Correcting for projective distortions of the sort outlined in Section~\ref{sec:intro} can be achieved by modeling the manner in which the respective rays through the pixels of the projector's image plane fan out into the scene (i.e., by `calibrating' the projector) and the geometry of the scene itself (i.e., by recovering the scene plane relative to the projector) within at least the projector's field of view.
 This is because the scene point `illuminated' by a pixel in the projector's image plane is given by intersecting its corresponding ray with the geometry of the scene surface.
@@
 		\caption{Recovering 2D positions in support of projector calibration.
-(a) 2D positions of the 2D-3D correspondences to be used for calibrating the projector are obtained by detecting---in the image plane of the projector---the circle centers in the circles pattern image projected by the projector to each of the target locations in the scene plane.
+(a) 2D positions of the 2D-3D correspondences to be used for calibrating the projector are obtained by detecting---in the image plane of the projector---the circle centers in the circles pattern image, projected by the projector to each of the target locations in the scene plane.
 (b) For each such target location, an image is acquired from the viewpoint of the camera and the circle centers of the projected circles pattern are detected, in the image plane of the camera.
@@
 If the projector is calibrated and its pose relative to the scene plane is known, a `virtual' projector (with the same calibration $\mathtt{K}$ and lens distortion coefficients) can be placed elsewhere relative to the scene plane.
-If we for a moment imagine that the projector---at its recovered `original' pose---functions as a camera,\footnote{Recall that the calibration matrix~$\mathtt{K}$ enables computing both (i) the projection of a scene point to the image plane (the function of a camera), or (ii) the back-projection of a pixel in the image plane, giving a ray into the scene (along which a projector illuminates the scene with the given pixel).} then (i) projecting an image to the scene plane \textit{from the viewpoint of the virtual projector} and (ii) acquiring the resulting projected image from the viewpoint of the original projector gives the desired distortion.
-Projecting an image distorted in this manner to the scene plane \textit{from the viewpoint of the original projector} then has the same effect as projecting the undistorted image to the scene plane from the viewpoint of the virtual projector.
+If we for a moment imagine that the projector---at its recovered pose---functions as a camera,\footnote{Recall that the calibration matrix~$\mathtt{K}$ enables computing both (i) the projection of a scene point to the image plane (the function of a camera), or (ii) the back-projection of a pixel in the image plane, giving a ray into the scene (along which a projector illuminates the scene with the given pixel).} then (i) projecting an image to the scene plane \textit{from the viewpoint of the virtual projector} and (ii) acquiring the resulting projected image from the viewpoint of the recovered projector gives the desired distortion.
+Projecting an image distorted in this manner to the scene plane \textit{from the viewpoint of the recovered projector} then has the same effect as projecting the undistorted image to the scene plane from the viewpoint of the virtual projector.
 This distortion can computed using a plane-induced homography, computed analytically as a function of the scene plane, the projector, and the virtual projector.
@@
 
-\paragraph{Plane-induced homography} Let $\mathtt{K}_\text{proj}$ express the $3~\times~3$ calibration matrix of the projector and $(\mathtt{R}', \mathbf{t}') \in SE(3)$ the rigid body transformation that transforms points from the coordinate frame of the projector to that of the virtual projector.
-
+\paragraph{Plane-induced homography} Let $\mathtt{K}_\text{proj}$ express the calibration matrix of the recovered projector and $(\mathtt{R}, \mathbf{t}) \in SE(3)$ the rigid body transformation that transforms points from the coordinate frame of the recovered projector to that of the virtual projector, for a given target location.
+Moreover, let $(\mathbf{n}^\top, -d)^\top$ give the scene plane, expressed in the coordinate frame of the recovered projector, where $\mathbf{n} \in \mathbb{R}^3$ is the scene plane's normal vector and $d = \mathbf{n}^\top\mathbf{X}$, for any point $\mathbf{X}$ in the plane.
+The transformation that distorts the image to be projected to the scene plane by the recovered projector such that it appear as if were projected to the scene plane by the virtual projector is given the by the $3~\times~3$ matrix
 \begin{equation}
-\mathtt{H} = \mathtt{K}_\text{proj}\left(\mathtt{R}' - \frac{\mathbf{t}'\mathbf{n}^\top}{d}\right)\mathtt{K}_\text{proj}^{-1},
+\mathtt{H} = \mathtt{K}_\text{proj}\left(\mathtt{R} - \frac{\mathbf{t}\mathbf{n}^\top}{d}\right)\mathtt{K}_\text{proj}^{-1},
 \label{homgen}
 \end{equation}
-where $(\mathbf{n}^\top, -d)^\top$ gives the scene plane and $d = \mathbf{n}^\top\mathbf{X}$, for any point $\mathbf{X}$ in the plane.
+a form of `plane-induced' homography \cite{Hartley2004}.
 

### 2021-06-26 paper/elsarticle/elsarticle-template.tex
paper

@@
 
-\paragraph{Virtual projector}
+\paragraph{Virtual projector} The placement of the virtual projector determines from which pose the image to be projected is to \textit{appear} to have been projected.
+This placement is produced according to a small handful of steps.
+First, we (i) intersect the scene plane with the optical axis (i.e., the ray from the projector's center of projection through the center of the image plane); next, we (ii) rotate the projector's placement about that point of intersection, aligning it with the scene plane's normal vector and giving an initial virtual projector.
+Separately, we (iii) carry out steps (i) and (ii) for the camera, giving a virtual camera likewise facing directly downwards to the scene plane.
+Finally, we (iv) rotate the $X$- and $Y$-axes of the initial virtual projector with those of the virtual camera, which gives the virtual projector.
+The virtual projector is thus rendered fronto-parallel with the scene plane, enabling projection to the scene plane absent of projective distortions.
+Moreover, the virtual projector's axes are aligned with the axes of an image of the scene acquired by a camera, enabling intuitive placement of augmentations in the scene.
 
@@
 A natural extension of this work would be to address non-planar scenes.
-To handle non-planar scenes would call for a change in how scene geometry is recovered and how distortion of the image to be projected is carried out; projector calibration---presented here as a one-time step---could, however, remain unchanged.
+To handle non-planar scenes would call for a change in how scene geometry is recovered and how distortion of the image to be projected is carried out; projector calibration could, however, be left unchanged.
 

### 2021-06-26 paper/elsarticle/elsarticle-template.tex
paper

@@
 This placement is produced according to a small handful of steps.
-First, we (i) intersect the scene plane with the optical axis (i.e., the ray from the projector's center of projection through the center of the image plane); next, we (ii) rotate the projector's placement about that point of intersection, aligning it with the scene plane's normal vector and giving an initial virtual projector.
+First, we (i) intersect the scene plane with the optical axis (i.e., the ray from the projector's center of projection through the center of the image plane); next, we (ii) rotate the projector's placement about that point of intersection, aligning the optical axis with the scene plane's normal vector and giving an initial virtual projector.
 Separately, we (iii) carry out steps (i) and (ii) for the camera, giving a virtual camera likewise facing directly downwards to the scene plane.
@@
 The virtual projector is thus rendered fronto-parallel with the scene plane, enabling projection to the scene plane absent of projective distortions.
-Moreover, the virtual projector's axes are aligned with the axes of an image of the scene acquired by a camera, enabling intuitive placement of augmentations in the scene.
+Moreover, the virtual projector's axes are aligned with the axes of an image of the scene acquired by a camera, enabling intuitive placement of augmentations in the scene, in accordance with the placement of the camera.
 
@@
 a form of `plane-induced' homography \cite{Hartley2004}.
+For convenience, we enable optional scaling or rotation of the image to be projected \textit{before} applying the homography; that rotation, parameterized in degrees, is thus in effect carried out intuitively relative to the placement of the camera.
 

### 2021-06-28 paper/elsarticle/elsarticle-template.tex
Update elsarticle-template.tex

@@
 The setup was mounted on the ceiling of the Pilotfabrik\footnote{\url{https://www.pilotfabrik.at/}} of TU Vienna, a collaborative space for research on Industry 4.0 topics situated in Vienna, Austria.
-The floorspace used for our experiments measured dimensions of ca.\ X~m~$\times$~Y~m; the projector was mounted at approximately the center of this space, at a height of ca.\ XX m.
+The floorspace used for our experiments measured dimensions of ca.\ X~m~$\times$~Y~m; the projector was mounted at approximately the center of this space, at a height of ca.\ 3.5 m.
 

### 2021-07-02 paper/elsarticle/elsarticle-template.tex
Update elsarticle-template.tex

@@
 
-We propose a spatial AR system that produces the effect of keystone correction analytically, and that intuitively places the desired augmentations in a manner aligned with the axes of an image of the scene acquired by a camera.
+We propose a spatial AR system that produces the effect of keystone correction analytically, and that facilitates placement by intuitively placing the desired augmentations in a manner aligned with the axes of an image of the scene acquired by a camera.
 Moreover, our system is able to handle a projector equipped with a steerable mirror, enabling wide-area factory floor augmentation exceeding the bounds of the projector's own immediate field of view.

### 2021-07-02 paper/elsarticle/elsarticle-template.tex
Update elsarticle-template.tex

@@
 We achieve this by distorting the image to be projected using a plane-induced homography computed to produce the effect of projecting the image not from the actual projector viewpoint, but in accordance with the viewpoint of a \textit{virtual} projector (i) facing directly downwards to the scene plane and (ii) rotated to place the axes of the image plane of the virtual projector in line with those of the camera.
-This enables intuitive placement of augmentations in the scene, and eliminates the need for manual keystone correction.
-Moreover, a consequence of our approach is that our system is able to handle a projector equipped with a steerable mirror (without need for explicitly modeling the action of the steerable mirror on the projector), thereby enabling wide-area applications.
+This facilitates placement of augmentations by intuitively placing them in accordance with the axes of an image of the scene, and eliminates the need for manual keystone correction.
+Moreover, a consequence of our approach is that our system is able to handle a projector equipped with a steerable mirror (without need for explicitly modeling the action of the steerable mirror on the projector), thereby enabling wide-area factory floor applications.
 

### 2021-07-02 paper/elsarticle/elsarticle-template.tex
Update elsarticle-template.tex

@@
 Yet considering for the moment a planar surface to be augmented, unless the projector faces the surface frontally, the bounds of a projected rectangular image will not appear rectangular; more generally, they will instead appear trapezoidal (i.e., the projected image will appear distorted).
-Such distortions can be eliminated by carrying out a cumbersome manual process called keystone correction, often using software bundled with the projector.
+Such distortions can be eliminated by carrying out a cumbersome manual process called keystone correction to appropriately warp the image to be projected, often using software bundled with the projector.
 

### 2021-07-04 paper/elsarticle/elsarticle-template.tex
Update elsarticle-template.tex

@@
 Augmented reality (AR) promises to enable use cases in industrial settings that include the embedding of assembly instructions directly into the scene, potentially reducing or altogether obviating the need for workers to refer to instructions in paper form or on a screen.
-In turn, \textit{spatial} AR is a form of AR whereby the augmentation of the scene is carried out using a projector, with the advantage of rendering the augmentation visible to all onlookers simultaneously without calling for each to wear some form of head-mounted display.
+\textit{Spatial} AR, in turn, is a form of AR whereby the augmentation of the scene is carried out using a projector, with the advantage of rendering the augmentation visible to all onlookers simultaneously without calling for each to wear some form of head-mounted display.
 Care must be taken, however, to distort the images to be projected in a manner that they appear undistorted to the viewer, since the geometry of the scene as it relates to the geometry of the projector plays a role in how the pixels of the projector's image plane map to points in the scene.

### 2021-07-05 paper/elsarticle/elsarticle-template.tex
paper

--- 2021-07-02 paper/elsarticle/elsarticle-template.tex
@@
 
-todo - maybe Hans would like to help?
+We evaluate our approach by considering the relative merits of augmenting 15 locations across the floorspace at the Pilotfabrik using our approach and two variants of a baseline approach.
+At each of the 15 locations, we aim to place an image aligned with the principal axes of the floorspace and such that the image appear in each instance to be free of projective distortions and to have the same dimensions, such that the horizontal axis of the image measure 30 cm.
+
+ We carry out all steps of our approach within 15 minutes, which we set Our baseline in both instances calls for using the software bundled with the steerable mirror system mounted alongside our projector
 

### 2021-07-05 paper/elsarticle/elsarticle-template.tex
paper

@@
 
-The hardware setup employed in this work comprised a Panasonic PT-RZ660BE projector with a steerable mirror system manufactured by Dynamic Projection Institute.
+The hardware setup employed in this work comprised a Panasonic PT-RZ660BE projector with a steerable mirror system manufactured by Dynamic Projection Institute \cite{rupprecht2020information}.
 In addition, we used a Zed 2 stereo camera manufactured by Stereolabs, yet relied only on the left view.

### 2021-07-05 paper/elsarticle/elsarticle-template.tex
Update elsarticle-template.tex

@@
 We project the circles pattern image to each of the target locations, and use the camera to acquire a projector calibration image for each.
-Given a projector calibration image, we detect the circle centers of the \textit{projected} circles pattern (cf.\ Figure~\ref{fig:2d}(b)); given the scene plane (the recovery of which we shall turn to in the paragraph that follows) and such a 2D circle center $\mathbf{x}$, its 3D correspondence is obtained by intersecting the back-projection of $\mathbf{x}$ with the scene plane (cf.\ Figure~\ref{fig:3d}).
+Given a projector calibration image, we detect the circle centers of the \textit{projected} circles pattern (cf.\ Figure~\ref{fig:2d}(b)); given the scene plane (the recovery of which we shall return to in the paragraph that follows) and such a 2D circle center $\mathbf{x}$, its 3D correspondence is obtained by intersecting the back-projection of $\mathbf{x}$ with the scene plane (cf.\ Figure~\ref{fig:3d}).
 Since the algorithm that yields 2D circle centers does so in a consistent ordering, we thus obtain the 2D-3D correspondences between the projector's image plane and the scene required for projector calibration, yielding the projector calibration matrix~$\mathtt{K}_\text{proj}$.

### 2021-07-06 paper/elsarticle/elsarticle-template.tex
paper

--- 2021-07-05 paper/elsarticle/elsarticle-template.tex
@@
 
-We propose a spatial AR system that produces the effect of keystone correction analytically, and that facilitates placement by intuitively placing the desired augmentations in a manner aligned with the axes of an image of the scene acquired by a camera.
+We propose a spatial AR system that produces the effect of keystone correction analytically, and that facilitates placement by intuitively placing the desired augmentations in a manner aligned with the axes of an image of the scene acquired by a camera facing downward toward the scene plane.
 Moreover, our system is able to handle a projector equipped with a steerable mirror, enabling wide-area factory floor augmentation exceeding the bounds of the projector's own immediate field of view.
@@
 Reliance on head-mounted displays, however, has two adverse consequences: (i) a head-mounted display must be worn by each individual wishing to view the augmentation, and (ii) such a head-mounted display---in some cases taking the form of a helmet in order to house multiple sensors in support of accurately tracking the viewpoint of the viewer relative to the scene---can be obtrusive.
-Spatial AR is a form of augmented reality carried out not by embedding the augmentation in an image of the scene as with head-mounted displays, but by projection to the scene itself \cite{bimber2019spatial}, thus eliminating both aforementioned problems.
+\textit{Spatial} AR is a form of augmented reality carried out not by embedding the augmentation in an image of the scene as with head-mounted displays, but by projection to the scene itself \cite{bimber2019spatial}, thus eliminating both aforementioned problems.
 Yet considering for the moment a planar surface to be augmented, unless the projector faces the surface frontally, the bounds of a projected rectangular image will not appear rectangular; more generally, they will instead appear trapezoidal (i.e., the projected image will appear distorted).
@@
 We evaluate our approach by considering the relative merits of augmenting 15 locations across the floorspace at the Pilotfabrik using our approach and two variants of a baseline approach.
-At each of the 15 locations, we aim to place an image aligned with the principal axes of the floorspace and such that the image appear in each instance to be free of projective distortions and to have the same dimensions, such that the horizontal axis of the image measure 30 cm.
+At each of the 15 locations, we aim to place an image aligned with the principal axes of the floorspace and such that the image appear in each instance to be free of projective distortions and to have the same dimensions, with the horizontal axis of the image as projected to the floor measuring 30 cm.
 

### 2021-07-06 paper/elsarticle/elsarticle-template.tex
paper

@@
 The virtual projector is thus rendered fronto-parallel with the scene plane, enabling projection to the scene plane absent of projective distortions.
-Moreover, the virtual projector's axes are aligned with the axes of an image of the scene acquired by a camera, enabling intuitive placement of augmentations in the scene, in accordance with the placement of the camera.
+
+Owing to the manner in which we place the virtual projector, the virtual projector's axes and thus the augmentation are aligned with the axes of an image of the scene acquired by a camera, enabling intuitive placement of augmentations in the scene in accordance with the placement of the camera.
+Note further that a consequence of placing the virtual projector by rotating about the point of intersection of the projector's optical axis with the scene plane is that the center of the projector's image plane remains invariant to projective distortion, i.e., a steerable mirror can be aimed with respect to a point projected from the center of the projector's image plane, further facilitating placement.
 
@@
 
-We evaluate our approach by considering the relative merits of augmenting 15 locations across the floorspace at the Pilotfabrik using our approach and two variants of a baseline approach.
+We evaluate our approach by augmenting 15 locations across the floorspace at the Pilotfabrik using our approach and two variants of a baseline approach.
 At each of the 15 locations, we aim to place an image aligned with the principal axes of the floorspace and such that the image appear in each instance to be free of projective distortions and to have the same dimensions, with the horizontal axis of the image as projected to the floor measuring 30 cm.
-
- We carry out all steps of our approach within 15 minutes, which we set Our baseline in both instances calls for using the software bundled with the steerable mirror system mounted alongside our projector
+We managed to carry out all steps using our approach within a time budget of 15 min.
+The first variant of the baseline approach is to carry out a manual keystone correction using the software bundled with the steerable mirror system mounted alongside our projector within the same 15 min time budget; we then compare the quality of the augmentations produced with this baseline with those produced using out approach.
+The second variant of the baseline approach is to record how much time is necessary to produce augmentations within a tolerance of what is desired.
 

### 2021-07-06 paper/elsarticle/elsarticle-template.tex
paper

@@
 First, we (i) intersect the scene plane with the optical axis (i.e., the ray from the projector's center of projection through the center of the image plane); next, we (ii) rotate the projector's placement about that point of intersection, aligning the optical axis with the scene plane's normal vector and giving an initial virtual projector.
-Separately, we (iii) carry out steps (i) and (ii) for the camera, giving a virtual camera likewise facing directly downwards to the scene plane.
+Separately, we (iii) carry out steps (i) and (ii) for the downward-facing camera, giving a virtual camera likewise facing directly\footnote{A physical camera placed to face downwards is likely to not face downwards precisely; in contrast, the virtual camera's optical axis is aligned exactly with the scene plane's normal vector.} downwards to the scene plane.
 Finally, we (iv) rotate the $X$- and $Y$-axes of the initial virtual projector with those of the virtual camera, which gives the virtual projector.
@@
 
-Owing to the manner in which we place the virtual projector, the virtual projector's axes and thus the augmentation are aligned with the axes of an image of the scene acquired by a camera, enabling intuitive placement of augmentations in the scene in accordance with the placement of the camera.
-Note further that a consequence of placing the virtual projector by rotating about the point of intersection of the projector's optical axis with the scene plane is that the center of the projector's image plane remains invariant to projective distortion, i.e., a steerable mirror can be aimed with respect to a point projected from the center of the projector's image plane, further facilitating placement.
+Owing to the manner in which we place the virtual projector, the virtual projector's axes and thus the augmentation are aligned with the axes of the downward-facing camera; the placement of the camera thereby intuitively determines the principal axes according to which augmentations are placed.
+Note further that a consequence of placing the virtual projector by rotating about the point of intersection of the projector's optical axis with the scene plane is that the center of the projector's image plane remains invariant to the placement of the virtual projector, i.e., a steerable mirror can be aimed with respect to a point projected from the center of the projector's image plane, further facilitating placement.
 

### 2021-07-06 paper/elsarticle/elsarticle-template.tex
paper

@@
 
-\section{Hardware Setup}
+\section{Hardware Setup}\label{sec:hw}
 
@@
 
-We evaluate our approach by augmenting 15 locations across the floorspace at the Pilotfabrik using our approach and two variants of a baseline approach.
-At each of the 15 locations, we aim to place an image aligned with the principal axes of the floorspace and such that the image appear in each instance to be free of projective distortions and to have the same dimensions, with the horizontal axis of the image as projected to the floor measuring 30 cm.
-We managed to carry out all steps using our approach within a time budget of 15 min.
-The first variant of the baseline approach is to carry out a manual keystone correction using the software bundled with the steerable mirror system mounted alongside our projector within the same 15 min time budget; we then compare the quality of the augmentations produced with this baseline with those produced using out approach.
-The second variant of the baseline approach is to record how much time is necessary to produce augmentations within a tolerance of what is desired.
+We evaluate our approach by augmenting 15 locations across the floorspace at the Pilotfabrik using our approach (`Our Approach') and two variants of a baseline approach (`Baseline A' and `Baseline B').
+At each of the 15 locations, we aim to place an image aligned with the principal axes of the floorspace and such that the image appear in each instance to be free of projective distortions and to have the same dimensions (30 cm $\times{}$ 20 cm).
+Using our approach, we carried out all steps---which we further detail below---within a time budget of 15 min.
+Accordingly, we set 15 min to be the time budget of the first variant of the baseline approach (`Baseline A'), in which we carry out a \texit{manual} keystone correction using the software bundled with the steerable mirror system mounted alongside our projector (cf.\ Section~\ref{sec:hw}).
+We then compare the quality of the augmentations produced with this baseline with those produced using our approach.
+The second variant of the baseline approach (`Baseline B') is to record how much time is necessary to produce augmentations within a mean tolerance of 1 cm.
+All experiments were carried out by the same experienced technician and assistant.
 

### 2021-07-07 paper/elsarticle/elsarticle-template.tex
paper

@@
 
-\section{Hardware Setup}\label{sec:hw}
-
-The hardware setup employed in this work comprised a Panasonic PT-RZ660BE projector with a steerable mirror system manufactured by Dynamic Projection Institute \cite{rupprecht2020information}.
-In addition, we used a Zed 2 stereo camera manufactured by Stereolabs, yet relied only on the left view.
-The setup was mounted on the ceiling of the Pilotfabrik\footnote{\url{https://www.pilotfabrik.at/}} of TU Vienna, a collaborative space for research on Industry 4.0 topics situated in Vienna, Austria.
-The floorspace used for our experiments measured dimensions of ca.\ X~m~$\times$~Y~m; the projector was mounted at approximately the center of this space, at a height of ca.\ 3.5 m.
-
 \section{Approach}
@@
 This is because the scene point `illuminated' by a pixel in the projector's image plane is given by intersecting its corresponding ray with the geometry of the scene surface.
-To model this interaction, we (i) carry out a one-time projector calibration, which in our approach calls for additionally calibrating a camera overlooking the scene and includes recovery of the scene plane as a convenient side effect.
+To model this interaction, we (i) carry out a one-time projector calibration, which in our approach calls for additionally calibrating a camera facing downward to the scene plane and includes recovery of the scene plane as a convenient side effect.
 Next, we use the relative camera-projector-scene plane geometry to (ii) compute a plane-induced homography that distorts the image to be projected in a manner that it appear undistorted to the viewer, and placed in alignment with the axes of a proxy image of the scene.
@@
 First, we (i) intersect the scene plane with the optical axis (i.e., the ray from the projector's center of projection through the center of the image plane); next, we (ii) rotate the projector's placement about that point of intersection, aligning the optical axis with the scene plane's normal vector and giving an initial virtual projector.
-Separately, we (iii) carry out steps (i) and (ii) for the downward-facing camera, giving a virtual camera likewise facing directly\footnote{A physical camera placed to face downwards is likely to not face downwards precisely; in contrast, the virtual camera's optical axis is aligned exactly with the scene plane's normal vector.} downwards to the scene plane.
+Separately, we (iii) carry out steps (i) and (ii) for the downward-facing camera, giving a virtual camera likewise facing directly\footnote{A physical camera placed to face downwards is likely to not face downwards precisely; in contrast, the virtual camera's optical axis is aligned exactly with the scene plane's normal vector, rendering it fronto-parallel with respect to the scene plane.} downwards to the scene plane.
 Finally, we (iv) rotate the $X$- and $Y$-axes of the initial virtual projector with those of the virtual camera, which gives the virtual projector.
@@
 
-We evaluate our approach by augmenting 15 locations across the floorspace at the Pilotfabrik using our approach (`Our Approach') and two variants of a baseline approach (`Baseline A' and `Baseline B').
-At each of the 15 locations, we aim to place an image aligned with the principal axes of the floorspace and such that the image appear in each instance to be free of projective distortions and to have the same dimensions (30 cm $\times{}$ 20 cm).
-Using our approach, we carried out all steps---which we further detail below---within a time budget of 15 min.
-Accordingly, we set 15 min to be the time budget of the first variant of the baseline approach (`Baseline A'), in which we carry out a \texit{manual} keystone correction using the software bundled with the steerable mirror system mounted alongside our projector (cf.\ Section~\ref{sec:hw}).
+We evaluate our approach by augmenting 15 locations across the floorspace at the Pilotfabrik using our approach and two variants of a baseline approach (`Baseline A' and `Baseline B').
+At each of the 15 locations, we aim to augment the scene by placing an image aligned with the principal axes of the floorspace and such that the image appear in each instance to be free of projective distortions and to have the same dimensions (30 cm $\times{}$ 20 cm).
+Using our approach to carry out keystone correction automatically, we carried out all steps---which we further detail in Figure X---within a time budget of 15 min.
+Accordingly, we set 15 min to be the time budget of the first variant of the baseline approach (`Baseline A'), in which we carry out a \texit{manual} keystone correction using the software bundled with the steerable mirror system mounted alongside our projector.
 We then compare the quality of the augmentations produced with this baseline with those produced using our approach.
-The second variant of the baseline approach (`Baseline B') is to record how much time is necessary to produce augmentations within a mean tolerance of 1 cm.
-All experiments were carried out by the same experienced technician and assistant.
+The second variant of the baseline approach (`Baseline B') is to again use the bundled software to carry out keystone correction manually, but to record how much time is necessary to produce augmentations within a mean tolerance of 1 cm.
+The steps for the baseline approaches are detailed in Figure Y.
+All experiments were carried out jointly by the same experienced technician and assistant.
+
+The hardware setup employed in the evaluation comprised a Panasonic PT-RZ660BE projector with a steerable mirror system manufactured by Dynamic Projection Institute \cite{rupprecht2020information}.
+In addition, we used a Zed 2 stereo camera manufactured by Stereolabs, yet relied only on the left view.
+The setup was mounted on the ceiling of the Pilotfabrik\footnote{\url{https://www.pilotfabrik.at/}} of TU Vienna, a collaborative space for research on Industry 4.0 topics situated in Vienna, Austria.
+The floorspace used for our experiments measured dimensions of ca.\ X~m~$\times$~Y~m; the projector was mounted at approximately the center of this space, at a height of ca.\ 3.5 m.
+
+\paragraph{Our approach} ...
+
+\paragraph{Baseline A} ...
+
+\paragraph{Baseline B} ...
 

### 2021-07-08 paper/elsarticle/elsarticle-template.tex
paper

@@
 
-We propose a spatial AR system that produces the effect of keystone correction analytically, and that facilitates placement by intuitively placing the desired augmentations in a manner aligned with the axes of an image of the scene acquired by a camera facing downward toward the scene plane.
+We propose a spatial AR system that produces the effect of keystone correction analytically, and that facilitates placement by enabling intuitively placing the desired augmentations in a manner aligned with the axes of an image of the scene acquired by a camera facing downward toward the scene plane.
 Moreover, our system is able to handle a projector equipped with a steerable mirror, enabling wide-area factory floor augmentation exceeding the bounds of the projector's own immediate field of view.
@@
     \caption{Scene plane (gray) obtained via spatial resection with respect to 2D-3D correspondences obtained using a chessboard pattern (blue); 3D circles pattern points obtained by intersection with the scene plane of back-projections (likewise gray) of circles pattern center points detected in the image plane (red).
-Note that the points in the figure are the points recovered for the projector calibration image in Figure~\ref{fig:2d}(b), acquired by the camera (frustum in black, with up vector in red).}
+Note that the points in the figure are the points recovered for the projector calibration image in Figure~\ref{fig:2d}(b), acquired by the camera (frustum of the camera in black, with up vector in red).}
     \label{fig:3d}
@@
 First, we (i) intersect the scene plane with the optical axis (i.e., the ray from the projector's center of projection through the center of the image plane); next, we (ii) rotate the projector's placement about that point of intersection, aligning the optical axis with the scene plane's normal vector and giving an initial virtual projector.
-Separately, we (iii) carry out steps (i) and (ii) for the downward-facing camera, giving a virtual camera likewise facing directly\footnote{A physical camera placed to face downwards is likely to not face downwards precisely; in contrast, the virtual camera's optical axis is aligned exactly with the scene plane's normal vector, rendering it fronto-parallel with respect to the scene plane.} downwards to the scene plane.
+Separately, we (iii) rotate the camera about its center of projection to align its optical axis with the normal vector of the scene plane, giving a virtual camera likewise facing directly\footnote{A physical camera placed to face downwards is likely to not face downwards precisely; in contrast, the virtual camera's optical axis is aligned exactly with the scene plane's normal vector, rendering it fronto-parallel with respect to the scene plane.} downwards to the scene plane.
 Finally, we (iv) rotate the $X$- and $Y$-axes of the initial virtual projector with those of the virtual camera, which gives the virtual projector.
 The virtual projector is thus rendered fronto-parallel with the scene plane, enabling projection to the scene plane absent of projective distortions.
+Optionally, we additionally adjust the height above the scene plane of the virtual projector, in order to satisfy desired projected image dimensions provided in metric units.
 
@@
 \paragraph{Plane-induced homography} Let $\mathtt{K}_\text{proj}$ express the calibration matrix of the recovered projector and $(\mathtt{R}, \mathbf{t}) \in SE(3)$ the rigid body transformation that transforms points from the coordinate frame of the recovered projector to that of the virtual projector, for a given target location.
-Moreover, let $(\mathbf{n}^\top, -d)^\top$ give the scene plane, expressed in the coordinate frame of the recovered projector, where $\mathbf{n} \in \mathbb{R}^3$ is the scene plane's normal vector and $d = \mathbf{n}^\top\mathbf{X}$, for any point $\mathbf{X}$ in the plane.
+Moreover, let $(\mathbf{n}^\top, -d)^\top$ give the scene plane, expressed in the coordinate frame of the recovered projector, where $\mathbf{n} \in \mathbb{R}^3$ is the scene plane's normal vector and $d = \mathbf{n}^\top\mathbf{X}$ for any point~$\mathbf{X}$ in the plane, so that $(\mathbf{n}^\top, -d) (\mathbf{X}^\top, 1)^\top = 0$.
 The transformation that distorts the image to be projected to the scene plane by the recovered projector such that it appear as if were projected to the scene plane by the virtual projector is given the by the $3~\times~3$ matrix
@@
 a form of `plane-induced' homography \cite{Hartley2004}.
-For convenience, we enable optional scaling or rotation of the image to be projected \textit{before} applying the homography; that rotation, parameterized in degrees, is thus in effect carried out intuitively relative to the placement of the camera.
+For convenience, we enable optional rotation of the image to be projected \textit{before} applying the homography; that rotation, parameterized in degrees, is thus in effect likewise carried out intuitively relative to the placement of the camera.
 
@@
 
-We evaluate our approach by augmenting 15 locations across the floorspace at the Pilotfabrik using our approach and two variants of a baseline approach (`Baseline A' and `Baseline B').
-At each of the 15 locations, we aim to augment the scene by placing an image aligned with the principal axes of the floorspace and such that the image appear in each instance to be free of projective distortions and to have the same dimensions (30 cm $\times{}$ 20 cm).
-Using our approach to carry out keystone correction automatically, we carried out all steps---which we further detail in Figure X---within a time budget of 15 min.
+We evaluate our approach by augmenting 15 locations across the floorspace at the Pilotfabrik\footnote{\url{https://www.pilotfabrik.at/}} of TU Vienna, a collaborative space for research on Industry 4.0 topics situated in Vienna, Austria.
+We contrast our approach with two variants of a baseline approach involving manual keystone correction (`Baseline A' and `Baseline B'), by aiming to place an image for each location aligned with the principal axes of the floorspace, absent projective distortions, and with the same dimensions (30 cm $\times{}$ 20 cm).
+Using our approach to carry out keystone correction analytically, we carried out all steps within a time budget of 15 min.
 Accordingly, we set 15 min to be the time budget of the first variant of the baseline approach (`Baseline A'), in which we carry out a \texit{manual} keystone correction using the software bundled with the steerable mirror system mounted alongside our projector.
-We then compare the quality of the augmentations produced with this baseline with those produced using our approach.
-The second variant of the baseline approach (`Baseline B') is to again use the bundled software to carry out keystone correction manually, but to record how much time is necessary to produce augmentations within a mean tolerance of 1 cm.
-The steps for the baseline approaches are detailed in Figure Y.
+The second variant of the baseline approach (`Baseline B') is to again use the bundled software to carry out keystone correction manually, but to record how much time is necessary to produce augmentations within a mean tolerance of XX.
 All experiments were carried out jointly by the same experienced technician and assistant.
 
-The hardware setup employed in the evaluation comprised a Panasonic PT-RZ660BE projector with a steerable mirror system manufactured by Dynamic Projection Institute \cite{rupprecht2020information}.
-In addition, we used a Zed 2 stereo camera manufactured by Stereolabs, yet relied only on the left view.
-The setup was mounted on the ceiling of the Pilotfabrik\footnote{\url{https://www.pilotfabrik.at/}} of TU Vienna, a collaborative space for research on Industry 4.0 topics situated in Vienna, Austria.
+The hardware setup employed in the evaluation comprised a Panasonic PT-RZ660BE projector with a steerable mirror system---used to point the projection to each of the 15 locations---manufactured by Dynamic Projection Institute \cite{rupprecht2020information}, bundled with the MDC-X software for steering the mirror, loading imagery, and optionally carrying out manual keystone correction, such that each position and (warped) image can be registered as a preset.
+In addition, we used a downward-facing Zed 2 stereo camera manufactured by Stereolabs, yet relied only on the left view.
+The setup was mounted on the ceiling of the Pilotfabrik.
 The floorspace used for our experiments measured dimensions of ca.\ X~m~$\times$~Y~m; the projector was mounted at approximately the center of this space, at a height of ca.\ 3.5 m.
 
-\paragraph{Our approach} ...
+\paragraph{Our approach} We begin by carrying out a one-time\footnote{Note that once the camera is calibrated, that calibration can be reused if the camera's pose relative to the projector-steerable mirror system remains fixed.} calibration of the camera, acquiring 10 camera calibration images (cf.\ Section~\ref{sec:approach:geometry}) and feeding the images as input to our camera calibration module.
+Separately, for each of the 15 target locations, we produce a projector calibration image (cf.\ again Section~\ref{sec:approach:geometry}) by projecting the circles pattern image to the location in question using the steerable mirror, placing a checkerboard pattern beside the projected pattern, and acquire the image using the downward-facing camera.
+We then fed these images alongside the output of the camera calibration module to our projector calibration module.
+For each of the target locations, the steerable mirror is made to point to that location, the circles pattern image is projected to the scene plane, an image is acquired using the camera, and the location is registered in the MDC-X software as a preset.
+The output of the projector calibration module is a homography per input projector calibration image (cf.\ Section~\ref{sec:approach:homography}).
+Next, we warp the images to be projected to the respective locations using their corresponding homography, using a third dedicated module.
+These warped images are finally imported into the MDC-X software and associated with their respective location presets.
+
+The total amount of time to carry out all the above steps amounted to ca.\ 15 min, with ca.\ 2 min going to acquisition of the camera calibration images, and ca.\ 4 min going to that of projector calibration images.
+The remainder of the time was spent running our modules or working with the MDC-X software.
 

### 2021-07-08 src/apps/calibrateProj/calibrateProj.cpp

@@ -204,11 +204,11 @@ void display()
                         glVertex3f(CCDWidth_half_mm * points[ptIdx].x / points[ptIdx].z, CCDWidth_half_mm * points[ptIdx].y / points[ptIdx].z, CCDWidth_half_mm);
                     glEnd();
 
-                    glBegin(GL_LINES);
-                        glColor3f(0.9, 0.9, 0.9);
-                        glVertex3f(0, 0, 0);
-                        glVertex3f(points[ptIdx].x, points[ptIdx].y, points[ptIdx].z);
-                    glEnd();
+                    //glBegin(GL_LINES);
+                    //    glColor3f(0.9, 0.9, 0.9);
+                    //    glVertex3f(0, 0, 0);
+                    //    glVertex3f(points[ptIdx].x, points[ptIdx].y, points[ptIdx].z);
+                    //glEnd();
 
                     glBegin(GL_POINTS);
                         glPointSize(0.5 * pointSize);
@@ -233,8 +233,8 @@ void display()
         float offset = 0.02;
         for (int i = 0; i < cams.size(); i++)
         {
-            //if (i != 0)
-            //    continue;
+            if (i == 1 || i == 4)
+                continue;
 
             float r = 0.5;
             float g = 0.5;
@@ -251,17 +251,17 @@ void display()
 
             char* str;
             if (i == 0)
-                str = "cam0";
+                str = "camera"; //"cam0";
             else if (i == 1)
                 str = "cam1";
             else if (i == 2)
                 str = "projector";
             else if (i == 3)
-                str = "projector virtual (downward)";
+                str = "initial virtual projector"; //"projector virtual (downward)";
             else if (i == 4)
                 str = "projector virtual (final)";
             else if (i == 5)
-                str = "cam0 virtual (downward)";
+                str = "virtual camera"; //"cam0 virtual (downward)";
 
             cv::Vec3d C = cams[i].getC();
 
@@ -277,14 +277,14 @@ void display()
             if (i == 5)
                 additionalOffset = 0.03;
 
-            displayText(C[0] + offset, C[1] + offset, C[2] - offset + additionalOffset, 0, 0, 0, str);
+            //displayText(C[0] + offset, C[1] + offset, C[2] - offset + additionalOffset, 0, 0, 0, str);
         }
 
-        displayText(
-            centroidChessboardObjectPts[0] + offset,
[…]

### 2021-07-08 paper/elsarticle/elsarticle-template.tex
paper

@@
     \centering
-    \subfloat[\centering Circles pattern image, in image plane of projector.]{{\includegraphics[height=3.25cm]{images/circles.png} }}
+    \subfloat[\centering Circles pattern image, in image plane of projector (detections overlain).]{{\includegraphics[height=3.25cm]{images/circles.png} }}
     \qquad
-    \subfloat[\centering Projector calibration image (one for each target location), in image plane of camera.]{{\includegraphics[height=3.25cm]{images/detections.png}}}
+    \subfloat[\centering Projector calibration image (one for each target location), in image plane of camera (detections overlain).]{{\includegraphics[height=3.25cm]{images/detections.png}}}
     \caption{Recovering 2D positions in support of projector calibration.

### 2021-07-08 paper/elsarticle/elsarticle-template.tex
paper

@@
 
+\begin{figure}
+    \centering
+    %\subfloat[\centering Initial virtual projector.]{{\includegraphics[height=3.25cm]{images/1.png} }}
+    %\qquad
+    \subfloat[\centering Virtual camera.]{{\includegraphics[height=4.5cm]{images/2.png}}}
+    \qquad
+    \subfloat[\centering Virtual projector.]{{\includegraphics[height=4.5cm]{images/3.png}}}
+    \caption{Obtaining the virtual projector.
+(a) The virtual camera is obtained by rotating the camera about its center of projection such that its optical axis be made parallel with the normal vector of the scene plane.
+(b) The virtual projector is obtained by (i) rotating the projector about the point of intersection of its optical axis with the scene plane such that its optical axis be made parallel with the scene plane's normal, (ii) rotating the $X$- and $Y$-axes to render them parallel with those of the virtual camera, and (iii) translating to achieve the desired projected image dimensions.}
+    \label{fig:virtualproj}
+\end{figure}
+
 \paragraph{Virtual projector} The placement of the virtual projector determines from which pose the image to be projected is to \textit{appear} to have been projected.

### 2021-07-08 src/apps/calibrateProj/calibrateProj.cpp

@@ -46,6 +46,7 @@ GLdouble cameraDist, cameraYaw, cameraRoll;
 GLdouble * cameraPos;
 
 int camIdx;
+int viewIdx;
 vector<Camera> cams;
 int imWidth, imHeight;
 
@@ -68,6 +69,7 @@ std::vector<Plane> planes;
 
 cv::Vec3d projCamIntersection;
 cv::Vec3d projPlaneIntersection;
+cv::Vec3d camPlaneIntersection;
 cv::Vec3d virtualCamPlaneIntersection;
 
 /* Previous x and y cursor coordinates (used to determine change in camera position) */
@@ -194,7 +196,12 @@ void display()
                 std::vector<cv::Point3f> points2 = pointCloud->getPoints();
 
                 float f = cams[0].getf();
-                float CCDWidth_half_mm = 20.0 * 0.01 * 0.5;
+                float CCDWidth_half_mm = 20.0 * 0.02 * 0.5;
+
+                float fProj = cams[2].getf();
+                cv::Vec3d projC = cams[2].getC();
+                cv::Mat Rt44 = cams[2].getRt44();
+                cv::Mat Rt44Inv = cams[2].getRt44Inv();
 
                 for (int ptIdx = 0; ptIdx < points.size(); ptIdx++)
                 {
@@ -210,11 +217,23 @@ void display()
                     //    glVertex3f(points[ptIdx].x, points[ptIdx].y, points[ptIdx].z);
                     //glEnd();
 
+                    cv::Vec3d pt = Ancillary::Mat44dTimesVec3dHomog(Rt44, cv::Vec3d(points[ptIdx].x, points[ptIdx].y, points[ptIdx].z));
+
+                    cv::Vec3d projPt(CCDWidth_half_mm * pt[0] / pt[2], CCDWidth_half_mm * pt[1] / pt[2], CCDWidth_half_mm);
+                    cv::Vec3d projPtLocal = Ancillary::Mat44dTimesVec3dHomog(Rt44Inv, projPt);
+
+
                     glBegin(GL_POINTS);
                         glPointSize(0.5 * pointSize);
                         glColor3f(1.0, 0.0, 0.0);
-                        glVertex3f(CCDWidth_half_mm * points[ptIdx].x / points[ptIdx].z, CCDWidth_half_mm * points[ptIdx].y / points[ptIdx].z, CCDWidth_half_mm);
+                        glVertex3f(projPtLocal[0], projPtLocal[1], projPtLocal[2]);
[…]

### 2021-07-08 paper/elsarticle/elsarticle-template.tex
paper

@@
 (a) The virtual camera is obtained by rotating the camera about its center of projection such that its optical axis be made parallel with the normal vector of the scene plane.
-(b) The virtual projector is obtained by (i) rotating the projector about the point of intersection of its optical axis with the scene plane such that its optical axis be made parallel with the scene plane's normal, (ii) rotating the $X$- and $Y$-axes to render them parallel with those of the virtual camera, and (iii) translating to achieve the desired projected image dimensions.}
+(b) The virtual projector is obtained by (i) rotating the projector about the point of intersection of its optical axis with the scene plane such that its optical axis be made parallel with the scene plane's normal, (ii) rotating the $X$- and $Y$-axes to render them parallel with those of the virtual camera, and (iii) optionally translating to achieve the desired projected image dimensions.
+The virtual projector is thus made fronto-parallel with the ground plane and axially aligned with the virtual camera.}
     \label{fig:virtualproj}

### 2021-07-08 paper/elsarticle/elsarticle-template.tex
paper

--- 2021-07-08 paper/elsarticle/elsarticle-template.tex
@@
 (a) The virtual camera is obtained by rotating the camera about its center of projection such that its optical axis be made parallel with the normal vector of the scene plane.
-(b) The virtual projector is obtained by (i) rotating the projector about the point of intersection of its optical axis with the scene plane such that its optical axis be made parallel with the scene plane's normal, (ii) rotating the $X$- and $Y$-axes to render them parallel with those of the virtual camera, and (iii) translating to achieve the desired projected image dimensions.}
+(b) The virtual projector is obtained by (i) rotating the projector about the point of intersection of its optical axis with the scene plane such that its optical axis be made parallel with the scene plane's normal vector, (ii) rotating the $X$- and $Y$-axes to render them parallel with those of the virtual camera, and (iii) optionally translating along the normal direction to achieve the desired projected image dimensions.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
     \label{fig:virtualproj}

### 2021-07-09 paper/elsarticle/elsarticle-template.tex
paper

--- 2021-07-08 paper/elsarticle/elsarticle-template.tex
@@
     \centering
-    %\subfloat[\centering Initial virtual projector.]{{\includegraphics[height=3.25cm]{images/1.png} }}
-    %\qquad
-    \subfloat[\centering Virtual camera.]{{\includegraphics[height=4.5cm]{images/2.png}}}
+    \subfloat[\centering Oblique view.]{{\includegraphics[height=4.25cm]{images/3.png}}}
     \qquad
-    \subfloat[\centering Virtual projector.]{{\includegraphics[height=4.5cm]{images/3.png}}}
-    \caption{Obtaining the virtual projector.
-(a) The virtual camera is obtained by rotating the camera about its center of projection such that its optical axis be made parallel with the normal vector of the scene plane.
-(b) The virtual projector is obtained by (i) rotating the projector about the point of intersection of its optical axis with the scene plane such that its optical axis be made parallel with the scene plane's normal, (ii) rotating the $X$- and $Y$-axes to render them parallel with those of the virtual camera, and (iii) translating to achieve the desired projected image dimensions.}
+    \subfloat[\centering Downwards view.]{{\includegraphics[height=4.25cm]{images/3_above3.png}}}
+    \caption{The virtual camera is obtained by rotating the camera (top left, black) about its center of projection such that its optical axis be made parallel with the normal vector of the scene plane.
+The virtual projector is obtained by (i) rotating the projector (bottom left, black) about the point of intersection of its optical axis with the scene plane such that its optical axis be made parallel with the scene plane's normal vector, (ii) rotating the $X$- and $Y$-axes to render them parallel with those of the virtual camera, and (iii) optionally translating along the normal direction to achieve the desired projected image dimensions.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
     \label{fig:virtualproj}
@@
 This placement is produced according to a small handful of steps.
-First, we (i) intersect the scene plane with the optical axis (i.e., the ray from the projector's center of projection through the center of the image plane); next, we (ii) rotate the projector's placement about that point of intersection, aligning the optical axis with the scene plane's normal vector and giving an initial virtual projector.
-Separately, we (iii) rotate the camera about its center of projection to align its optical axis with the normal vector of the scene plane, giving a virtual camera likewise facing directly\footnote{A physical camera placed to face downwards is likely to not face downwards precisely; in contrast, the virtual camera's optical axis is aligned exactly with the scene plane's normal vector, rendering it fronto-parallel with respect to the scene plane.} downwards to the scene plane.
-Finally, we (iv) rotate the $X$- and $Y$-axes of the initial virtual projector with those of the virtual camera, which gives the virtual projector.
+First, we (i) rotate the camera about its center of projection to align its optical axis with the normal vector of the scene plane, giving a virtual camera likewise facing directly\footnote{A physical camera placed to face downwards is likely to not face downwards precisely; in contrast, the virtual camera's optical axis is aligned exactly with the scene plane's normal vector, rendering it fronto-parallel with respect to the scene plane.} downwards to the scene plane (cf.\ Figure~\ref{fig:virtualproj}(a)).
+Next, we (ii) intersect the scene plane with the optical axis (i.e., the ray from the projector's center of projection through the center of the image plane); then, we (iii) rotate the projector's placement about that point of intersection, aligning the optical axis with the scene plane's normal vector and giving an initial virtual projector.
+Finally, we (iv) rotate the $X$- and $Y$-axes of the initial virtual projector with those of the virtual camera, which gives the virtual projector (cf.\ Figure~\ref{fig:virtualproj}(b)).
 The virtual projector is thus rendered fronto-parallel with the scene plane, enabling projection to the scene plane absent of projective distortions.

### 2021-07-09 paper/elsarticle/elsarticle-template.tex
paper

@@
     \centering
-    \subfloat[\centering Oblique view.]{{\includegraphics[height=4.25cm]{images/3.png}}}
+    \subfloat[\centering Oblique view.]{{\includegraphics[height=4.2cm]{images/3.png}}}
     \qquad
-    \subfloat[\centering Downwards view.]{{\includegraphics[height=4.25cm]{images/3_above3.png}}}
+    \subfloat[\centering Downwards view.]{{\includegraphics[height=4.2cm]{images/3_above3.png}}}
     \caption{The virtual camera is obtained by rotating the camera (top left, black) about its center of projection such that its optical axis be made parallel with the normal vector of the scene plane.

### 2021-07-09 paper/elsarticle/elsarticle-template.tex
paper

@@
     \caption{The virtual camera is obtained by rotating the camera (top left, black) about its center of projection such that its optical axis be made parallel with the normal vector of the scene plane.
-The virtual projector is obtained by (i) rotating the projector (bottom left, black) about the point of intersection of its optical axis with the scene plane such that its optical axis be made parallel with the scene plane's normal vector, (ii) rotating the $X$- and $Y$-axes to render them parallel with those of the virtual camera, and (iii) optionally translating along the normal direction to achieve the desired projected image dimensions.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
+The virtual projector is obtained by (i) rotating the projector (bottom left, black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (ii) rotating the $X$- and $Y$-axes to render them parallel with those of the virtual camera, and (iii) optionally translating along the normal direction to achieve the desired projected image dimensions.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
     \label{fig:virtualproj}
@@
 This placement is produced according to a small handful of steps.
-First, we (i) rotate the camera about its center of projection to align its optical axis with the normal vector of the scene plane, giving a virtual camera likewise facing directly\footnote{A physical camera placed to face downwards is likely to not face downwards precisely; in contrast, the virtual camera's optical axis is aligned exactly with the scene plane's normal vector, rendering it fronto-parallel with respect to the scene plane.} downwards to the scene plane (cf.\ Figure~\ref{fig:virtualproj}(a)).
+First, we (i) rotate the camera about its center of projection to align its optical axis with the normal vector of the scene plane, giving a virtual camera likewise facing directly\footnote{A physical camera placed to face downwards is likely to not face downwards precisely; in contrast, the virtual camera's optical axis is aligned exactly with the scene plane's normal vector, rendering it fronto-parallel with respect to the scene plane.} downwards to the scene plane (cf.\ Figure~\ref{fig:virtualproj}).
 Next, we (ii) intersect the scene plane with the optical axis (i.e., the ray from the projector's center of projection through the center of the image plane); then, we (iii) rotate the projector's placement about that point of intersection, aligning the optical axis with the scene plane's normal vector and giving an initial virtual projector.
-Finally, we (iv) rotate the $X$- and $Y$-axes of the initial virtual projector with those of the virtual camera, which gives the virtual projector (cf.\ Figure~\ref{fig:virtualproj}(b)).
+Finally, we (iv) rotate the $X$- and $Y$-axes of the initial virtual projector with those of the virtual camera, which gives the virtual projector (cf.\ again Figure~\ref{fig:virtualproj}).
 The virtual projector is thus rendered fronto-parallel with the scene plane, enabling projection to the scene plane absent of projective distortions.
@@
 a form of `plane-induced' homography \cite{Hartley2004}.
-For convenience, we enable optional rotation of the image to be projected \textit{before} applying the homography; that rotation, parameterized in degrees, is thus in effect likewise carried out intuitively relative to the placement of the camera.
+For convenience, we enable optional rotation of the image to be projected \textit{before} applying the homography, about the image center; that rotation, parameterized in degrees, is thus in effect likewise carried out intuitively relative to the placement of the camera.
 

### 2021-07-09 src/cpslib/libHomography/Homography.cpp
bug fix

@@ -109,7 +109,7 @@ void Homography::computeBirdsEyeViewVirtualCam(Plane& plane, Camera& cam, float 
     Eigen::Matrix3d invR = q.toRotationMatrix();
 
     // compute camera center C of virtual camera
-    Eigen::Vector3d C = centerPt + (distToPlane + verticalOffset) * n;
+    Eigen::Vector3d C = centerPt + (verticalOffset /*- distToPlane*/) * n;
 
     // compute (R, t) of virtual camera, relative to camera coordinate frame of input camera
     Eigen::Matrix3d R = invR.inverse();

### 2021-07-09 paper/elsarticle/elsarticle-template.tex
paper

@@
 
-\title{A Spatial AR System for Wide-area Axis-aligned Augmentation of Planar Scenes in Industrial Settings} %\title{Elsevier \LaTeX\ template\tnoteref{mytitlenote}}
+\title{A Spatial AR System for Axis-aligned Augmentation of Planar Scenes in Industrial Settings} %\title{Elsevier \LaTeX\ template\tnoteref{mytitlenote}}
 %\tnotetext[mytitlenote]{Fully documented templates are available in the elsarticle package on \href{http://www.ctan.org/tex-archive/macros/latex/contrib/elsarticle}{CTAN}.}
@@
 
-Using the $X$- and $Y$-axes of an image of the scene as a proxy, our contribution is to propose a spatial AR system that produces the effect of keystone correction analytically, and in a manner aligning the axes of the augmentation with those of the proxy image.
+Our contribution is to propose a spatial AR system that produces the effect of keystone correction analytically, and---using the $X$- and $Y$-axes of an image of the scene acquired by a downwards-facing camera as a proxy---in a manner aligning the axes of the augmentation with those of the proxy image.
 We achieve this by distorting the image to be projected using a plane-induced homography computed to produce the effect of projecting the image not from the actual projector viewpoint, but in accordance with the viewpoint of a \textit{virtual} projector (i) facing directly downwards to the scene plane and (ii) rotated to place the axes of the image plane of the virtual projector in line with those of the camera.
@@
 Correcting for projective distortions of the sort outlined in Section~\ref{sec:intro} can be achieved by modeling the manner in which the respective rays through the pixels of the projector's image plane fan out into the scene (i.e., by `calibrating' the projector) and the geometry of the scene itself (i.e., by recovering the scene plane relative to the projector) within at least the projector's field of view.
-This is because the scene point `illuminated' by a pixel in the projector's image plane is given by intersecting its corresponding ray with the geometry of the scene surface.
+This is because the scene point `illuminated' by a pixel in the projector's image plane can be recovered by intersecting its corresponding ray with the geometry of the scene surface.
 To model this interaction, we (i) carry out a one-time projector calibration, which in our approach calls for additionally calibrating a camera facing downward to the scene plane and includes recovery of the scene plane as a convenient side effect.
@@
     \centerline{\includegraphics[scale=.35]{images/2d3d.png}}
-    \caption{Scene plane (gray) obtained via spatial resection with respect to 2D-3D correspondences obtained using a chessboard pattern (blue); 3D circles pattern points obtained by intersection with the scene plane of back-projections (likewise gray) of circles pattern center points detected in the image plane (red).
-Note that the points in the figure are the points recovered for the projector calibration image in Figure~\ref{fig:2d}(b), acquired by the camera (frustum of the camera in black, with up vector in red).}
+    \caption{Scene plane (gray) recovered via spatial resection with respect to 2D-3D correspondences obtained using a chessboard pattern (blue); 3D circles pattern points obtained by intersection with the scene plane of back-projections (likewise gray) of circles pattern center points detected in the image plane (red).
+Note that the points in the figure are the points recovered for the projector calibration image in Figure~\ref{fig:2d}(b), acquired by the downwards-facing camera (frustum of the camera in black, with up vector in red).}
     \label{fig:3d}
@@
 The corresponding 2D points are obtained, in the same order, using a specialized algorithm \cite{bradski2000opencv}.
-A set of calibration images is acquired, each with the calibration pattern visible in a different part of the image plane, and such that the center and all corners and edges of the image plane are covered.
+A set of calibration images is acquired, each with the calibration pattern visible in a different part of the image plane, and such that the center and all corners and edges of the image plane are covered, the camera's autofocus setting is off, and the camera's zoom factor remain fixed.
 2D-3D correspondences are then recovered for each calibration image, and the resulting list is passed on as input to an optimization procedure that relies on bundle adjustment to yield the camera calibration matrix~$\mathtt{K}_\text{cam}$ and the associated lens distortion model coefficients.
@@
 If the projector is calibrated and its pose relative to the scene plane is known, a `virtual' projector (with the same calibration $\mathtt{K}$ and lens distortion coefficients) can be placed elsewhere relative to the scene plane.
-If we for a moment imagine that the projector---at its recovered pose---functions as a camera,\footnote{Recall that the calibration matrix~$\mathtt{K}$ enables computing both (i) the projection of a scene point to the image plane (the function of a camera), or (ii) the back-projection of a pixel in the image plane, giving a ray into the scene (along which a projector illuminates the scene with the given pixel).} then (i) projecting an image to the scene plane \textit{from the viewpoint of the virtual projector} and (ii) acquiring the resulting projected image from the viewpoint of the recovered projector gives the desired distortion.
-Projecting an image distorted in this manner to the scene plane \textit{from the viewpoint of the recovered projector} then has the same effect as projecting the undistorted image to the scene plane from the viewpoint of the virtual projector.
-This distortion can computed using a plane-induced homography, computed analytically as a function of the scene plane, the projector, and the virtual projector.
+If we for a moment imagine that the projector---at its recovered pose---functions as a camera,\footnote{Recall that the calibration matrix~$\mathtt{K}$ enables computing both (i) the projection of a scene point to the image plane (the function of a camera), or (ii) the back-projection of a pixel in the image plane, giving a ray into the scene (along which a projector illuminates the scene with the given pixel).} then (i) projecting an image to the scene plane \textit{from the viewpoint of the virtual projector} and (ii) acquiring the resulting projected image from the viewpoint of the recovered projector gives the desired corrective warp.
+Projecting an image warped in this manner to the scene plane \textit{from the viewpoint of the recovered projector} then has the same effect as projecting the unwarped image to the scene plane from the viewpoint of the virtual projector.
+This warp can computed using a plane-induced homography, computed analytically as a function of the scene plane, the projector, and the virtual projector.
 
@@
     \caption{The virtual camera is obtained by rotating the camera (top left, black) about its center of projection such that its optical axis be made parallel with the normal vector of the scene plane.
-The virtual projector is obtained by (i) rotating the projector (bottom left, black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (ii) rotating the $X$- and $Y$-axes to render them parallel with those of the virtual camera, and (iii) optionally translating along the normal direction to achieve the desired projected image dimensions.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
+The virtual projector is obtained by (i) rotating the projector (bottom left, black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (ii) rotating the $X$- and $Y$-axes to align them with those of the virtual camera, and (iii) optionally translating along the normal direction to achieve the desired projected image dimensions.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
     \label{fig:virtualproj}
@@
 Next, we (ii) intersect the scene plane with the optical axis (i.e., the ray from the projector's center of projection through the center of the image plane); then, we (iii) rotate the projector's placement about that point of intersection, aligning the optical axis with the scene plane's normal vector and giving an initial virtual projector.
-Finally, we (iv) rotate the $X$- and $Y$-axes of the initial virtual projector with those of the virtual camera, which gives the virtual projector (cf.\ again Figure~\ref{fig:virtualproj}).
+Finally, we (iv) align the $X$- and $Y$-axes of the initial virtual projector with those of the virtual camera, which gives the virtual projector (cf.\ again Figure~\ref{fig:virtualproj}).
 The virtual projector is thus rendered fronto-parallel with the scene plane, enabling projection to the scene plane absent of projective distortions.
@@
 We evaluate our approach by augmenting 15 locations across the floorspace at the Pilotfabrik\footnote{\url{https://www.pilotfabrik.at/}} of TU Vienna, a collaborative space for research on Industry 4.0 topics situated in Vienna, Austria.
-We contrast our approach with two variants of a baseline approach involving manual keystone correction (`Baseline A' and `Baseline B'), by aiming to place an image for each location aligned with the principal axes of the floorspace, absent projective distortions, and with the same dimensions (30 cm $\times{}$ 20 cm).
+We contrast our approach with two variants of a baseline approach involving manual keystone correction (`Baseline A' and `Baseline B'), by aiming to place an image for each location aligned with the principal axes of the floorspace, absent projective distortions, and with the same dimensions (50~cm~$\times$~31.25~cm).
 Using our approach to carry out keystone correction analytically, we carried out all steps within a time budget of 15 min.
@@
 
-The hardware setup employed in the evaluation comprised a Panasonic PT-RZ660BE projector with a steerable mirror system---used to point the projection to each of the 15 locations---manufactured by Dynamic Projection Institute \cite{rupprecht2020information}, bundled with the MDC-X software for steering the mirror, loading imagery, and optionally carrying out manual keystone correction, such that each position and (warped) image can be registered as a preset.
+The hardware setup employed in the evaluation comprised a Panasonic PT-RZ660BE projector with a steerable mirror system---used to point the projection to each of the 15 locations---manufactured by Dynamic Projection Institute \cite{rupprecht2020information}.
+The steerable mirror system was bundled with the MDC-X software for steering the mirror, loading imagery, and optionally carrying out manual keystone correction, such that each position and (warped) image can be registered as a preset.
 In addition, we used a downward-facing Zed 2 stereo camera manufactured by Stereolabs, yet relied only on the left view.
-The setup was mounted on the ceiling of the Pilotfabrik.
-The floorspace used for our experiments measured dimensions of ca.\ X~m~$\times$~Y~m; the projector was mounted at approximately the center of this space, at a height of ca.\ 3.5 m.
+The floorspace used for our experiments measured dimensions of ca.\ 9~m~$\times$~4~m; the projector was mounted at approximately the center of this space, at a height of ca.\ 3.5 m.
 
-\paragraph{Our approach} We begin by carrying out a one-time\footnote{Note that once the camera is calibrated, that calibration can be reused if the camera's pose relative to the projector-steerable mirror system remains fixed.} calibration of the camera, acquiring 10 camera calibration images (cf.\ Section~\ref{sec:approach:geometry}) and feeding the images as input to our camera calibration module.
-Separately, for each of the 15 target locations, we produce a projector calibration image (cf.\ again Section~\ref{sec:approach:geometry}) by projecting the circles pattern image to the location in question using the steerable mirror, placing a checkerboard pattern beside the projected pattern, and acquire the image using the downward-facing camera.
+\paragraph{Our approach} We began by carrying out a one-time\footnote{Note that once the camera is calibrated, that calibration can be reused if the camera's intrinsics remain fixed, i.e., no change is made to the zoom factor.} calibration of the camera, acquiring 10 camera calibration images (cf.\ Section~\ref{sec:approach:geometry}) and feeding the images as input to our camera calibration module.
+Separately, for each of the 15 target locations, we produced a projector calibration image (cf.\ again Section~\ref{sec:approach:geometry}) by projecting the circles pattern image to the location in question using the steerable mirror, placing a checkerboard pattern beside the projected pattern, and acquire the image using the downward-facing camera.
 We then fed these images alongside the output of the camera calibration module to our projector calibration module.
-For each of the target locations, the steerable mirror is made to point to that location, the circles pattern image is projected to the scene plane, an image is acquired using the camera, and the location is registered in the MDC-X software as a preset.
+For each of the target locations, the steerable mirror was made to point to that location, the circles pattern image was projected to the scene plane, an image was using the camera, and the location was registered in the MDC-X software as a preset.
 The output of the projector calibration module is a homography per input projector calibration image (cf.\ Section~\ref{sec:approach:homography}).
-Next, we warp the images to be projected to the respective locations using their corresponding homography, using a third dedicated module.
-These warped images are finally imported into the MDC-X software and associated with their respective location presets.
+Next, we warped the images to be projected to the respective locations using their corresponding homography, using a third dedicated custom module.
+These warped images were finally imported into the MDC-X software and associated with their respective location presets.
 
-The total amount of time to carry out all the above steps amounted to ca.\ 15 min, with ca.\ 2 min going to acquisition of the camera calibration images, and ca.\ 4 min going to that of projector calibration images.
+The total amount of time to carry out all the above steps amounted to ca.\ 15 min, with ca.\ 2 min going to acquisition of the camera calibration images, and ca.\ 5 min going to that of projector calibration images.
 The remainder of the time was spent running our modules or working with the MDC-X software.

### 2021-07-09 src/apps/applyHomography/applyHomography.cpp

@@ -4,12 +4,12 @@
  *   IMW-CPS TU Vienna, Austria
  *
  *   applyHomography <homographyPath> <inPath> <outPath> <scalingFactor> [<angleDeg> <width> <height>]
- * 
+ *
  *   Note that providing angleDeg (even if 0) will first scale the image so that rotating by an
  *   angle of 45 degrees with respect to the target dimensions of width, height will not cut out
  *   any of the input image contents, and then apply the input homography; not providing angleDeg,
  *   width, and height will simply apply the given homography
- * 
+ *
  *   Note additionally that if a video is provided, it must have extension .avi or .mp4 (case sensitive!)
  *
  *   Example invocations:
@@ -65,7 +65,9 @@ int main(int argc, char** argv)
     if (ext == ".avi" || ext == ".mp4")
         isVideo = true;
 
-    cv::VideoCapture cap(inPath);
+    cv::VideoCapture cap;
+    if (isVideo)
+        cap = cv::VideoCapture(inPath);
 
     bool hasVideoWriter = false;
     cv::VideoWriter videoWriter;
@@ -155,7 +157,7 @@ int main(int argc, char** argv)
 
             videoWriter.write(outIm);
         }
-        
+
         cv::imshow("warped", outIm);
         cv::waitKey(100);
 

### 2021-07-09 src/apps/calibrateProj/calibrateProj.cpp

@@ -3,10 +3,10 @@
  *   michael.hornacek@gmail.com
  *   IMW-CPS TU Vienna, Austria
  *
- *   calibrateProj <boardSqSize> <boardDimsX> <boardDimsY> <circlesDimsX> <circlesDimsY> <outDir> <numIms> <cam0ImDir> <cam1ImDir> <cam0Path> <cam1Path> <circlesImPath> <verticalNegOffset> [<visImIdx>] [<visIm>]
+ *   calibrateProj <boardSqSize> <boardDimsX> <boardDimsY> <circlesDimsX> <circlesDimsY> <outDir> <numIms> <cam0ImDir> <cam1ImDir> <cam0Path> <cam1Path> <circlesImPath> <targetWidth> [<visImIdx>] [<visIm>]
  *
  *   Example invocation:
- *   calibrateProj 0.0565 4 6 4 11 C:\Users\micha\Desktop\spatial-ar\in_out\calibrateProj\out 14 C:\Users\micha\Desktop\spatial-ar\in_out\splitZed\projCalib\outLeft C:\Users\micha\Desktop\spatial-ar\in_out\splitZed\projCalib\outRight C:\Users\micha\Desktop\spatial-ar\in_out\calibrateCam\out\cam_0.yml C:\Users\micha\Desktop\spatial-ar\in_out\calibrateCam\out\cam_1.yml C:\Users\micha\Desktop\spatial-ar\in_out\calibrateProj\acircles_pattern_960x600.png 1.75 3 C:\Users\micha\Desktop\spatial-ar\in_out\applyHomography\holodeck.png
+ *   calibrateProj 0.0565 4 6 4 11 C:\Users\micha\Desktop\spatial-ar\in_out\calibrateProj\out 11 C:\Users\micha\Desktop\spatial-ar\in_out\splitZed\projCalib\outLeft C:\Users\micha\Desktop\spatial-ar\in_out\splitZed\projCalib\outRight C:\Users\micha\Desktop\spatial-ar\in_out\calibrateCam\out\cam_0.yml  C:\Users\micha\Desktop\spatial-ar\in_out\calibrateCam\out\cam_1.yml C:\Users\micha\Desktop\spatial-ar\in_out\calibrateProj\acircles_pattern_960x600.png 1.0 3 C:\Users\micha\Desktop\spatial-ar\in_out\applyHomography\holodeck.png
  */
 
 
[…]

### 2021-07-09 paper/elsarticle/elsarticle-template.tex
paper

@@
 We propose a spatial AR system that produces the effect of keystone correction analytically, and that facilitates placement by enabling intuitively placing the desired augmentations in a manner aligned with the axes of an image of the scene acquired by a camera facing downward toward the scene plane.
-Moreover, our system is able to handle a projector equipped with a steerable mirror, enabling wide-area factory floor augmentation exceeding the bounds of the projector's own immediate field of view.
+Moreover, our system is able to handle a projector equipped with a steerable mirror, enabling factory floor augmentation exceeding the bounds of the projector's own immediate field of view.
 \end{abstract}
@@
 This facilitates placement of augmentations by intuitively placing them in accordance with the axes of an image of the scene, and eliminates the need for manual keystone correction.
-Moreover, a consequence of our approach is that our system is able to handle a projector equipped with a steerable mirror (without need for explicitly modeling the action of the steerable mirror on the projector), thereby enabling wide-area factory floor applications.
+Moreover, a consequence of our approach is that our system is able to handle a projector equipped with a steerable mirror (without need for explicitly modeling the action of the steerable mirror on the projector), thereby enabling factory floor applications exceeding the immediate field of view of the projector.
 
@@
 
-Calibration of a projector (or camera) in the sense we employ the term here\footnote{We are referring to a geometric calibration; not, e.g., to a color calibration.} renders one able to project a scene point~$\mathbf{X} \in \mathbb{R}^3$ to its corresponding pixel~$\mathbf{x} \in \mathbb{R}^2$ in the projector's (or camera's) image plane, or compute the `back-projection' of $\mathbf{x}$ (cf.\ Figure~\ref{fig:3d}), i.e., the ray from the projector's (or camera's) center of projection through $\mathbf{x}$ along which $\mathbf{X}$ must lie.
+Calibration of a projector (or camera) in the sense we employ the term here\footnote{We are referring to a geometric calibration; not, e.g., to a color calibration.} renders one able to project a scene point~$\mathbf{X} \in \mathbb{R}^3$ to its corresponding pixel~$\mathbf{x} \in \mathbb{R}^2$ in the projector's (or camera's) image plane, or to compute the `back-projection' of $\mathbf{x}$ (cf.\ Figure~\ref{fig:3d}), i.e., the ray from the projector's (or camera's) center of projection through $\mathbf{x}$ along which $\mathbf{X}$ must lie.
 Such a calibration can be expressed in terms of (i) a $3~\times{}~3$ calibration matrix~$\mathtt{K}$ derived from the projector's (or camera's) focal length and principal point \cite{Hartley2004}, and (ii) the coefficients of a lens distortion model used to correct for radial or tangential distortions caused by the lens system \cite{duane1971close}.
@@
     \centering
-    \subfloat[\centering Oblique view.]{{\includegraphics[height=4.2cm]{images/3.png}}}
+    \subfloat[\centering Oblique view.]{{\includegraphics[height=4.2cm]{images/1.pdf}}}
     \qquad
     \subfloat[\centering Downwards view.]{{\includegraphics[height=4.2cm]{images/3_above3.png}}}
-    \caption{The virtual camera is obtained by rotating the camera (top left, black) about its center of projection such that its optical axis be made parallel with the normal vector of the scene plane.
-The virtual projector is obtained by (i) rotating the projector (bottom left, black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (ii) rotating the $X$- and $Y$-axes to align them with those of the virtual camera, and (iii) optionally translating along the normal direction to achieve the desired projected image dimensions.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
+    \caption{The virtual camera is obtained by (i) rotating the camera (top left, black) about its center of projection such that its optical axis be made parallel with the normal vector of the scene plane.
+The virtual projector is obtained by (ii) rotating the projector (bottom left, black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (iii) rotating the $X$- and $Y$-axes to align them with those of the virtual camera, and (iv) optionally translating along the normal direction to achieve the desired projected image dimensions.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
     \label{fig:virtualproj}
@@
 We presented a spatial AR system for planar scenes that produces the effect of keystone correction analytically, and that intuitively places the desired augmentations in a manner aligned with the axes of an image of the scene acquired by a camera.
-Moreover, we showed our system to be able to handle a projector equipped with a steerable mirror, enabling wide-area factory floor augmentation exceeding the bounds of the projector's own immediate field of view.
+Moreover, we showed our system to be able to handle a projector equipped with a steerable mirror, enabling factory floor augmentation exceeding the bounds of the projector's own immediate field of view.
 Our evaluation demonstrated ...

### 2021-07-10 paper/elsarticle/elsarticle-template.tex
paper

@@
 
-We propose a spatial AR system that produces the effect of keystone correction analytically, and that facilitates placement by enabling intuitively placing the desired augmentations in a manner aligned with the axes of an image of the scene acquired by a camera facing downward toward the scene plane.
+We propose a spatial AR system that produces the effect of keystone correction analytically, and that facilitates placement by enabling intuitively placing the desired augmentations in a manner aligned with the axes of an image of the scene acquired by a camera facing downwards towards the scene plane.
 Moreover, our system is able to handle a projector equipped with a steerable mirror, enabling factory floor augmentation exceeding the bounds of the projector's own immediate field of view.
@@
     \caption{The virtual camera is obtained by (i) rotating the camera (top left, black) about its center of projection such that its optical axis be made parallel with the normal vector of the scene plane.
-The virtual projector is obtained by (ii) rotating the projector (bottom left, black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (iii) rotating the $X$- and $Y$-axes to align them with those of the virtual camera, and (iv) optionally translating along the normal direction to achieve the desired projected image dimensions.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
+The virtual projector is obtained by (ii) rotating the projector (bottom left, black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (iii) rotating the $X$- and $Y$-axes to align them with those of the virtual camera, and optionally (iv) translating along the normal direction to achieve the desired projected image dimensions.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
     \label{fig:virtualproj}
@@
 First, we (i) rotate the camera about its center of projection to align its optical axis with the normal vector of the scene plane, giving a virtual camera likewise facing directly\footnote{A physical camera placed to face downwards is likely to not face downwards precisely; in contrast, the virtual camera's optical axis is aligned exactly with the scene plane's normal vector, rendering it fronto-parallel with respect to the scene plane.} downwards to the scene plane (cf.\ Figure~\ref{fig:virtualproj}).
-Next, we (ii) intersect the scene plane with the optical axis (i.e., the ray from the projector's center of projection through the center of the image plane); then, we (iii) rotate the projector's placement about that point of intersection, aligning the optical axis with the scene plane's normal vector and giving an initial virtual projector.
-Finally, we (iv) align the $X$- and $Y$-axes of the initial virtual projector with those of the virtual camera, which gives the virtual projector (cf.\ again Figure~\ref{fig:virtualproj}).
+Next, we (ii) intersect the scene plane with the optical axis (i.e., the ray from the projector's center of projection through the center of the image plane) and rotate the projector's placement about that point of intersection, aligning the optical axis with the scene plane's normal vector and giving an initial virtual projector.
+Finally, we (iii) align the $X$- and $Y$-axes of the initial virtual projector with those of the virtual camera, which gives the virtual projector (cf.\ again Figure~\ref{fig:virtualproj}).
 The virtual projector is thus rendered fronto-parallel with the scene plane, enabling projection to the scene plane absent of projective distortions.
-Optionally, we additionally adjust the height above the scene plane of the virtual projector, in order to satisfy desired projected image dimensions provided in metric units.
+Optionally, we additionally (iv) adjust the height above the scene plane of the virtual projector, in order to satisfy desired projected image dimensions provided in metric units.
 
@@
 We evaluate our approach by augmenting 15 locations across the floorspace at the Pilotfabrik\footnote{\url{https://www.pilotfabrik.at/}} of TU Vienna, a collaborative space for research on Industry 4.0 topics situated in Vienna, Austria.
-We contrast our approach with two variants of a baseline approach involving manual keystone correction (`Baseline A' and `Baseline B'), by aiming to place an image for each location aligned with the principal axes of the floorspace, absent projective distortions, and with the same dimensions (50~cm~$\times$~31.25~cm).
-Using our approach to carry out keystone correction analytically, we carried out all steps within a time budget of 15 min.
-Accordingly, we set 15 min to be the time budget of the first variant of the baseline approach (`Baseline A'), in which we carry out a \texit{manual} keystone correction using the software bundled with the steerable mirror system mounted alongside our projector.
-The second variant of the baseline approach (`Baseline B') is to again use the bundled software to carry out keystone correction manually, but to record how much time is necessary to produce augmentations within a mean tolerance of XX.
-All experiments were carried out jointly by the same experienced technician and assistant.
+We contrast our approach with a baseline approach involving manual keystone correction, by aiming with both approaches to place an image for each location aligned with the principal axes of the floorspace, absent of projective distortions, and with the same dimensions (50~cm~$\times$~31.25~cm).
+All experiments were carried out by the same experienced technician.
 
@@
 In addition, we used a downward-facing Zed 2 stereo camera manufactured by Stereolabs, yet relied only on the left view.
-The floorspace used for our experiments measured dimensions of ca.\ 9~m~$\times$~4~m; the projector was mounted at approximately the center of this space, at a height of ca.\ 3.5 m.
+The floorspace used for our experiments measured dimensions of ca.\ 6~m~$\times$~4~m; the projector was mounted at approximately the center of this space, at a height of ca.\ 3.5 m.
 
@@
 
-The total amount of time to carry out all the above steps amounted to ca.\ 15 min, with ca.\ 2 min going to acquisition of the camera calibration images, and ca.\ 5 min going to that of projector calibration images.
+The total amount of time to carry out all the above steps amounted to ca.\ 20 min, with ca.\ 2 min going to acquisition of the camera calibration images, and ca.\ 5 min going to that of projector calibration images.
 The remainder of the time was spent running our modules or working with the MDC-X software.
 
-\paragraph{Baseline A} ...
-
-\paragraph{Baseline B} ...
+\paragraph{Baseline} ...
 

### 2021-07-10 paper/elsarticle/elsarticle-template.tex
paper

@@
 
+\begin{figure}
+    \centering
+    \subfloat[\centering Projecting original image.]{{\includegraphics[height=4.2cm]{images/proj_projim.png}}}
+    \qquad
+    \subfloat[\centering Projecting warped image.]{{\includegraphics[height=4.2cm]{images/proj_projim_waped.png}}}
+    \caption{Projection from recovered projector viewpoint (bottom left, black) to scene plane of original image and of warped image.
+(a) Projecting the original 960~pixel~$\times$~600~pixel image ...
+(b) After warping the original image accoring to the plane-induced homography for the given target location, the image is projected in a manner that appears ...}
+    \label{fig:warp}
+\end{figure}
+
 \paragraph{Plane-induced homography} Let $\mathtt{K}_\text{proj}$ express the calibration matrix of the recovered projector and $(\mathtt{R}, \mathbf{t}) \in SE(3)$ the rigid body transformation that transforms points from the coordinate frame of the recovered projector to that of the virtual projector, for a given target location.
 Moreover, let $(\mathbf{n}^\top, -d)^\top$ give the scene plane, expressed in the coordinate frame of the recovered projector, where $\mathbf{n} \in \mathbb{R}^3$ is the scene plane's normal vector and $d = \mathbf{n}^\top\mathbf{X}$ for any point~$\mathbf{X}$ in the plane, so that $(\mathbf{n}^\top, -d) (\mathbf{X}^\top, 1)^\top = 0$.
-The transformation that distorts the image to be projected to the scene plane by the recovered projector such that it appear as if were projected to the scene plane by the virtual projector is given the by the $3~\times~3$ matrix
+The transformation that distorts the image to be projected to the scene plane by the recovered projector such that it appear as if were projected to the scene plane by the virtual projector (cf.\ Figure~\ref{fig:warp}(b)) is given the by the $3~\times~3$ matrix
 \begin{equation}

### 2021-07-10 paper/elsarticle/elsarticle-template.tex
paper

@@
     \subfloat[\centering Projecting warped image.]{{\includegraphics[height=4.2cm]{images/proj_projim_waped.png}}}
-    \caption{Projection from recovered projector viewpoint (bottom left, black) to scene plane of original image and of warped image.
-(a) Projecting the original 960~pixel~$\times$~600~pixel image ...
-(b) After warping the original image accoring to the plane-induced homography for the given target location, the image is projected in a manner that appears ...}
+    \caption{Projection from the recovered projector viewpoint (bottom left, black) to the scene plane of the original image and of the warped image.
+(a) Projecting the original image to the scene plane.
+(b) After warping the original image according to our plane-induced homography for the given target location, the image is projected in a manner that appears free of perspective distortions, aligned with the axes of the virtual camera, and to have the desired dimensions in the scene plane, expressed in metric units.
+Note that background in the projected image is shown set to black.}
     \label{fig:warp}

### 2021-07-10 src/apps/calibrateProj/calibrateProj.cpp

@@ -1145,9 +1145,9 @@ int main(int argc, char** argv)
         cv::waitKey(5);
 
         vector<Point3f> imVisPts, imVisColors;
-        for (int y = 0; y < projSize.height; y += 3)
+        for (int y = 0; y < projSize.height; y += 1)
         {
-            for (int x = 0; x < projSize.width; x += 3)
+            for (int x = 0; x < projSize.width; x += 1)
             {
                 cv::Vec3d intersectionLocal = planeProjLocal.intersect(cams[2].backprojectLocal(cv::Point2f(x, y)));
                 cv::Vec3d intersectionGlobal = Ancillary::Mat44dTimesVec3dHomog(cams[2].getRt44Inv(), intersectionLocal);
@@ -1172,9 +1172,9 @@ int main(int argc, char** argv)
         pointCloud2Circles = new PointCloud(circlesObjectPtsVirtual_);
 
         vector<Point3f> imVis2Pts, imVis2Colors;
-        for (int y = 0; y < projSize.height; y += 3)
+        for (int y = 0; y < projSize.height; y += 1)
         {
-            for (int x = 0; x < projSize.width; x += 3)
+            for (int x = 0; x < projSize.width; x += 1)
             {
                 cv::Vec3d intersectionLocal = planeProjLocal.intersect(cams[2].backprojectLocal(cv::Point2f(x, y)));
                 cv::Vec3d intersectionGlobal = Ancillary::Mat44dTimesVec3dHomog(cams[2].getRt44Inv(), intersectionLocal);

### 2021-07-10 paper/elsarticle/elsarticle-template.tex
paper

@@
     \qquad
-    \subfloat[\centering Projecting warped image.]{{\includegraphics[height=4.2cm]{images/proj_projim_waped.png}}}
-    \caption{Projection from the recovered projector viewpoint (bottom left, black) to the scene plane of the original image and of the warped image.
+    \subfloat[\centering Projecting our warped image.]{{\includegraphics[height=4.2cm]{images/proj_projim_waped.png}}}
+    \caption{Projection from the recovered projector viewpoint (bottom left, black; virtual projector in gray) to the scene plane of the original image and of the warped image.
 (a) Projecting the original image to the scene plane.

### 2021-07-10 src/apps/calibrateProj/calibrateProj.cpp

@@ -261,6 +261,9 @@ void display()
             if (viewIdx == 1 && (i == 4))
                 continue;
 
+            if (viewIdx > 1 && (i == 0 || i == 1 || i == 3 || i == 5))
+                continue;
+
             float r = 0.5;
             float g = 0.5;
             float b = 0.5;

### 2021-07-10 paper/elsarticle/elsarticle-template.tex
paper

@@
     \centerline{\includegraphics[scale=.35]{images/2d3d.png}}
-    \caption{Scene plane (gray) recovered via spatial resection with respect to 2D-3D correspondences obtained using a chessboard pattern (blue); 3D circles pattern points obtained by intersection with the scene plane of back-projections (likewise gray) of circles pattern center points detected in the image plane (red).
+    \caption{Scene plane (gray) recovered via spatial resection with respect to 2D-3D correspondences obtained using a chessboard pattern (blue); 3D circles pattern points---i.e., the 3D positions of the
+2D-3D correspondences to be used for calibrating the projector---obtained by intersection with the scene plane of back-projections (likewise gray) of circles pattern center points detected in the image plane (red).
 Note that the points in the figure are the points recovered for the projector calibration image in Figure~\ref{fig:2d}(b), acquired by the downwards-facing camera (frustum of the camera in black, with up vector in red).}

### 2021-07-10 paper/elsarticle/elsarticle-template.tex
paper

--- 2021-07-10 paper/elsarticle/elsarticle-template.tex
@@
     \centerline{\includegraphics[scale=.35]{images/2d3d.png}}
-    \caption{Scene plane (gray) recovered via spatial resection with respect to 2D-3D correspondences obtained using a chessboard pattern (blue); 3D circles pattern points obtained by intersection with the scene plane of back-projections (likewise gray) of circles pattern center points detected in the image plane (red).
+    \caption{Scene plane (gray) recovered via spatial resection with respect to 2D-3D correspondences obtained using a chessboard pattern (blue); 3D circles pattern points---i.e., the 3D positions of the
+2D-3D correspondences to be used for calibrating the projector---obtained by intersection with the scene plane of back-projections (likewise gray) of the 2D circles pattern center points detected in the image plane (red).
 Note that the points in the figure are the points recovered for the projector calibration image in Figure~\ref{fig:2d}(b), acquired by the downwards-facing camera (frustum of the camera in black, with up vector in red).}

### 2021-07-11 paper/elsarticle/elsarticle-template.tex
paper

@@
 
-We propose a spatial AR system that produces the effect of keystone correction analytically, and that facilitates placement by enabling intuitively placing the desired augmentations in a manner aligned with the axes of an image of the scene acquired by a camera facing downwards towards the scene plane.
+We propose a spatial AR system for planar scenes that produces the effect of keystone correction analytically, and that facilitates placement by enabling intuitively placing the desired augmentations in a manner aligned with the axes of an image of the scene acquired by a camera facing downwards towards the scene plane.
 Moreover, our system is able to handle a projector equipped with a steerable mirror, enabling factory floor augmentation exceeding the bounds of the projector's own immediate field of view.
@@
 
-\paragraph{Camera calibration} The recovery of 2D-3D correspondences in support of camera calibration is carried out by relying on a planar calibration surface to automatically identify correspondences between the 3D points on the calibration surface and their 2D correspondences in the image plane.
+\paragraph{Camera calibration} We recover 2D-3D correspondences in support of calibrating the downwards-facing camera is carried out by relying on a planar calibration surface to automatically identify correspondences between the 3D points on the calibration surface and their 2D correspondences in the image plane.
 The classical calibration surface is a chessboard pattern.
@@
 The corresponding 2D points are obtained, in the same order, using a specialized algorithm \cite{bradski2000opencv}.
-A set of calibration images is acquired, each with the calibration pattern visible in a different part of the image plane, and such that the center and all corners and edges of the image plane are covered, the camera's autofocus setting is off, and the camera's zoom factor remain fixed.
+A set of calibration images is acquired, each with the calibration pattern visible in a different part of the image plane, and such that the center and all corners and edges of the image plane are covered, the camera's autofocus setting be off, and the camera's zoom factor remain fixed.
 2D-3D correspondences are then recovered for each calibration image, and the resulting list is passed on as input to an optimization procedure that relies on bundle adjustment to yield the camera calibration matrix~$\mathtt{K}_\text{cam}$ and the associated lens distortion model coefficients.
@@
 \paragraph{Projector calibration} As with camera calibration, projector calibration relies on 2D-3D correspondences, yet we obtain them in this case by projecting a pattern of circles.
-We rely on an algorithm to detect the circle pattern center points in the circles pattern image in the image plane of the projector \cite{bradski2000opencv}, giving the 2D positions of our 2D-3D correspondences (cf.\ Figure~\ref{fig:2d}(a)).
+We rely on an algorithm to detect the circle pattern center points in the circles pattern image in the image plane of the projector \cite{bradski2000opencv}, giving the 2D positions of our 2D-3D correspondences for calibrating the projector (cf.\ Figure~\ref{fig:2d}(a)).
 We project the circles pattern image to each of the target locations, and use the camera to acquire a projector calibration image for each.
-Given a projector calibration image, we detect the circle centers of the \textit{projected} circles pattern (cf.\ Figure~\ref{fig:2d}(b)); given the scene plane (the recovery of which we shall return to in the paragraph that follows) and such a 2D circle center $\mathbf{x}$, its 3D correspondence is obtained by intersecting the back-projection of $\mathbf{x}$ with the scene plane (cf.\ Figure~\ref{fig:3d}).
+Given a projector calibration image acquired using the downwards-facing camera, we detect the circle centers of the \textit{projected} circles pattern (cf.\ Figure~\ref{fig:2d}(b)); given the scene plane (the recovery of which we shall return to in the paragraph that follows) and such a 2D circle center $\mathbf{x}$, its 3D correspondence is obtained by intersecting the back-projection of $\mathbf{x}$ with the scene plane (cf.\ Figure~\ref{fig:3d}).
 Since the algorithm that yields 2D circle centers does so in a consistent ordering, we thus obtain the 2D-3D correspondences between the projector's image plane and the scene required for projector calibration, yielding the projector calibration matrix~$\mathtt{K}_\text{proj}$.
@@
     \subfloat[\centering Projecting our warped image.]{{\includegraphics[height=4.2cm]{images/proj_projim_waped.png}}}
-    \caption{Projection from the recovered projector viewpoint (bottom left, black; virtual projector in gray) to the scene plane of the original image and of the warped image.
+    \caption{Projection from the recovered projector viewpoint (bottom left, black; virtual projector in center, gray) to the scene plane of the original image and of the warped image.
 (a) Projecting the original image to the scene plane.
-(b) After warping the original image according to our plane-induced homography for the given target location, the image is projected in a manner that appears free of perspective distortions, aligned with the axes of the virtual camera, and to have the desired dimensions in the scene plane, expressed in metric units.
+(b) After warping the original image according to our plane-induced homography for the given target location, the image is projected in a manner that appears free of perspective distortions, aligned with the axes of the virtual camera (via the virtual projector), and to have the desired dimensions in the scene plane, expressed in metric units.
 Note that background in the projected image is shown set to black.}
@@
 We evaluate our approach by augmenting 15 locations across the floorspace at the Pilotfabrik\footnote{\url{https://www.pilotfabrik.at/}} of TU Vienna, a collaborative space for research on Industry 4.0 topics situated in Vienna, Austria.
-We contrast our approach with a baseline approach involving manual keystone correction, by aiming with both approaches to place an image for each location aligned with the principal axes of the floorspace, absent of projective distortions, and with the same dimensions (50~cm~$\times$~31.25~cm).
+We contrast our approach with a baseline approach involving manual keystone correction, by aiming with both approaches to place an image for each location aligned with the principal axes of the floorspace, absent of projective distortions, and with the same dimensions (50~cm~$\times$~31.25~cm\footnote{}).
 All experiments were carried out by the same experienced technician.
@@
 
-\paragraph{Our approach} We began by carrying out a one-time\footnote{Note that once the camera is calibrated, that calibration can be reused if the camera's intrinsics remain fixed, i.e., no change is made to the zoom factor.} calibration of the camera, acquiring 10 camera calibration images (cf.\ Section~\ref{sec:approach:geometry}) and feeding the images as input to our camera calibration module.
+\paragraph{Our approach} We began by carrying out a one-time\footnote{Note that once the camera is calibrated, that calibration can be reused if the camera's intrinsics remain fixed, e.g., no change is made to the zoom factor.} calibration of the camera, acquiring 10 camera calibration images (cf.\ Section~\ref{sec:approach:geometry}) and feeding the images as input to our camera calibration module.
 Separately, for each of the 15 target locations, we produced a projector calibration image (cf.\ again Section~\ref{sec:approach:geometry}) by projecting the circles pattern image to the location in question using the steerable mirror, placing a checkerboard pattern beside the projected pattern, and acquire the image using the downward-facing camera.

### 2021-07-11 paper/elsarticle/elsarticle-template.tex
paper

@@
 We propose a spatial AR system for planar scenes that produces the effect of keystone correction analytically, and that facilitates placement by enabling intuitively placing the desired augmentations in a manner aligned with the axes of an image of the scene acquired by a camera facing downwards towards the scene plane.
-Moreover, our system is able to handle a projector equipped with a steerable mirror, enabling factory floor augmentation exceeding the bounds of the projector's own immediate field of view.
+Moreover, our system is able to handle a projector equipped with a steerable mirror, enabling factory floor augmentation exceeding the bounds of the projector's own immediate field of view, using only a single projector.
 \end{abstract}
@@
 This facilitates placement of augmentations by intuitively placing them in accordance with the axes of an image of the scene, and eliminates the need for manual keystone correction.
-Moreover, a consequence of our approach is that our system is able to handle a projector equipped with a steerable mirror (without need for explicitly modeling the action of the steerable mirror on the projector), thereby enabling factory floor applications exceeding the immediate field of view of the projector.
+Moreover, a consequence of our approach is that our system is able to handle a projector equipped with a steerable mirror (without need for explicitly modeling the action of the steerable mirror on the projector), thereby enabling factory floor applications exceeding the immediate field of view of the projector without needing to rely on multiple projectors.
 
@@
 
-todo - maybe Majesa would like to help? Sebastian's inputs would be valuable for more general background story.
+An early spatial AR system for planar scenes using a projector with a steerable mirror is the IBM Everywhere Displays prototype of Pinhanez \cite{pinhanez2001everywhere}, yet compute the warps applied to imagery to be projected manually.
+Rather than rely on a steerable mirror to support spatial AR to multiple locations using a single projector, some works mount the projector and a camera on a rigid rig and subject the rig to motion \cite{ehnes2004projected,borkowski2004spatial}.
+Steering only a mirror, however, places humbler requirements on the system from a hardware standpoint than if an entire camera-projector rig is to be subject to steering.
+...
+
+%todo: It consists of a projector and a small mirror with 2 axes of rotation.
+In contrast with the present work, Everywhere Displays is limited to projecting on a discrete set of planar surfaces which are calibrated beforehand, each modeled by a 2D homography.
+The Everywhere Displays project explored many applications of steerable displays, particularly in the retail domain.
+With the addition of a pan and tilt video camera, a number of gesture-based interactions are demonstrated, including touch-based manipulation of projected graphical widgets [12], as well as person tracking to enable graphics to follow the user.
+A small number of later works contribute steerable displays designs which mount the projector and camera directly onto motorized platform with pan and tilt axes of motion [8, 6, 11].
+Butz et al.
+[7] contribute a number of interactions in the tradition of ubiquitous computing, such as the ability to
 
@@
 
-\begin{figure}
-    \centering
-    \subfloat[\centering Circles pattern image, in image plane of projector (detections overlain).]{{\includegraphics[height=3.25cm]{images/circles.png} }}
-    \qquad
-    \subfloat[\centering Projector calibration image (one for each target location), in image plane of camera (detections overlain).]{{\includegraphics[height=3.25cm]{images/detections.png}}}
-    \caption{Recovering 2D positions in support of projector calibration.
-(a) 2D positions of the 2D-3D correspondences to be used for calibrating the projector are obtained by detecting---in the image plane of the projector---the circle centers in the circles pattern image, projected by the projector to each of the target locations in the scene plane.
-(b) For each such target location, an image is acquired from the viewpoint of the camera and the circle centers of the projected circles pattern are detected, in the image plane of the camera.
-A chessboard pattern to be used for recovering the local scene plane is placed near the projected pattern, whose corners are likewise detected.
-Detected 2D projected circles pattern center points and chessboard corners overlain for illustration.} %At this stage, the chessboard is used to recover the local scene plane for each target location.
-Since the projected circles pattern lies in this plane, we are in a subsequent step able to rely on that plane to derive the 3D positions corresponding to the 2D projected circles pattern center points detected in the image.
-    \label{fig:2d}
-\end{figure}
-
-\begin{figure}
-    \centerline{\includegraphics[scale=.35]{images/2d3d.png}}
-    \caption{Scene plane (gray) recovered via spatial resection with respect to 2D-3D correspondences obtained using a chessboard pattern (blue); 3D circles pattern points---i.e., the 3D positions of the
-2D-3D correspondences to be used for calibrating the projector---obtained by intersection with the scene plane of back-projections (likewise gray) of the 2D circles pattern center points detected in the image plane (red).
-Note that the points in the figure are the points recovered for the projector calibration image in Figure~\ref{fig:2d}(b), acquired by the downwards-facing camera (frustum of the camera in black, with up vector in red).}
-    \label{fig:3d}
-\end{figure}
-
 Calibration of a projector (or camera) in the sense we employ the term here\footnote{We are referring to a geometric calibration; not, e.g., to a color calibration.} renders one able to project a scene point~$\mathbf{X} \in \mathbb{R}^3$ to its corresponding pixel~$\mathbf{x} \in \mathbb{R}^2$ in the projector's (or camera's) image plane, or to compute the `back-projection' of $\mathbf{x}$ (cf.\ Figure~\ref{fig:3d}), i.e., the ray from the projector's (or camera's) center of projection through $\mathbf{x}$ along which $\mathbf{X}$ must lie.
@@
 The classical calibration surface is a chessboard pattern.
-The 3D corner points of the chessboard are obtained \textit{a priori} in a coordinate system defined in the plane of the chessboard\footnote{E.g., $(0,0,0), (1.5,0,0), (3,0,0), \dots, (9,7.5,0)$ for a chessboard with $7~\times{}~6$ corners, with each square of length and width of 1.5 unit, respectively.
+The 3D corner points of the chessboard are obtained \textit{a priori} in a coordinate system defined in the plane of the chessboard\footnote{E.g., $(0,0,0), (1.5,0,0), (3,0,0), \dots, (9,7.5,0)$ for a chessboard with $7~\times{}~6$ corners ($8~\times{}~7$ squares), with each square of length and width of 1.5 unit, respectively.
 Note that the units of the chessboard's 3D points give the units of the camera calibration, and---owing to how our projector calibration relies on the camera calibration---of the projector calibration as well.}, requiring knowledge only of the dimensions of the chessboard pattern and of the length of a side of a chessboard square.
@@
 2D-3D correspondences are then recovered for each calibration image, and the resulting list is passed on as input to an optimization procedure that relies on bundle adjustment to yield the camera calibration matrix~$\mathtt{K}_\text{cam}$ and the associated lens distortion model coefficients.
+
+\begin{figure}
+    \centering
+    \subfloat[\centering Circles pattern image, in image plane of projector (detections overlain).]{{\includegraphics[height=3.25cm]{images/circles.png} }}
+    \qquad
+    \subfloat[\centering Projector calibration image (one for each target location), in image plane of camera (detections overlain).]{{\includegraphics[height=3.25cm]{images/detections.png}}}
+    \caption{Recovering 2D positions in support of projector calibration.
+(a) 2D positions of the 2D-3D correspondences to be used for calibrating the projector are obtained by detecting---in the image plane of the projector---the circle centers in the circles pattern image, projected by the projector to each of the target locations in the scene plane.
+(b) For each such target location, an image is acquired from the viewpoint of the camera and the circle centers of the projected circles pattern are detected, in the image plane of the camera.
+A chessboard pattern to be used for recovering the local scene plane is placed near the projected pattern, whose corners are likewise detected.
+Detected 2D projected circles pattern center points and chessboard corners overlain for illustration.}
+    \label{fig:2d}
+\end{figure}
 
@@
 As this is sufficient for our needs in Section~\ref{sec:approach:homography}, it is in this sense that our system is able to handle a projector equipped with a steerable mirror, without need for modeling the steerable mirror explicitly.
+
+\begin{figure}
+    \centerline{\includegraphics[scale=.35]{images/2d3d.png}}
+    \caption{Scene plane (gray) recovered via spatial resection with respect to 2D-3D correspondences obtained using a chessboard pattern (blue); 3D circles pattern points---i.e., the 3D positions of the
+2D-3D correspondences to be used for calibrating the projector---obtained by intersection with the scene plane of back-projections (likewise gray) of the 2D circles pattern center points detected in the image plane (red).
+Note that the points in the figure are the points recovered for the projector calibration image in Figure~\ref{fig:2d}(b), acquired by the downwards-facing camera (frustum of the camera in black, with up vector in red).}
+    \label{fig:3d}
+\end{figure}
 
@@
 We evaluate our approach by augmenting 15 locations across the floorspace at the Pilotfabrik\footnote{\url{https://www.pilotfabrik.at/}} of TU Vienna, a collaborative space for research on Industry 4.0 topics situated in Vienna, Austria.
-We contrast our approach with a baseline approach involving manual keystone correction, by aiming with both approaches to place an image for each location aligned with the principal axes of the floorspace, absent of projective distortions, and with the same dimensions (50~cm~$\times$~31.25~cm\footnote{}).
-All experiments were carried out by the same experienced technician.
+We contrast our approach with a baseline approach involving manual keystone correction, by aiming with both approaches to place the same image for each location aligned with the principal axes of the floorspace, absent of projective distortions, and with the same metric dimensions (50~cm~$\times$~31.25~cm\footnote{The dimensions in pixels of the image we project are 960~pixels~$\times$~600~pixels; we chose for our experiments to set the projected metric length of the horizontal axis of the image to 50~cm, which implies 31.25~cm for the vertical axis.}).
+All experiments were carried out by the same technician, experienced in both approaches.
 
-The hardware setup employed in the evaluation comprised a Panasonic PT-RZ660BE projector with a steerable mirror system---used to point the projection to each of the 15 locations---manufactured by Dynamic Projection Institute \cite{rupprecht2020information}.
+The hardware setup employed in the evaluation comprised a Panasonic PT-RZ660BE projector with a steerable mirror system---used in our experiments to point the projection to each of the 15 locations---manufactured by Dynamic Projection Institute \cite{rupprecht2020information}.
 The steerable mirror system was bundled with the MDC-X software for steering the mirror, loading imagery, and optionally carrying out manual keystone correction, such that each position and (warped) image can be registered as a preset.
@@
 
-\paragraph{Our approach} We began by carrying out a one-time\footnote{Note that once the camera is calibrated, that calibration can be reused if the camera's intrinsics remain fixed, e.g., no change is made to the zoom factor.} calibration of the camera, acquiring 10 camera calibration images (cf.\ Section~\ref{sec:approach:geometry}) and feeding the images as input to our camera calibration module.
-Separately, for each of the 15 target locations, we produced a projector calibration image (cf.\ again Section~\ref{sec:approach:geometry}) by projecting the circles pattern image to the location in question using the steerable mirror, placing a checkerboard pattern beside the projected pattern, and acquire the image using the downward-facing camera.
+\paragraph{Our approach} We began by carrying out a calibration of the camera, acquiring 10 camera calibration images (cf.\ Section~\ref{sec:approach:geometry}) of a chessboard calibration pattern with $6~\times~4$ corners ($7~\times~5$ squares) and feeding the images as input to our camera calibration module.
+Separately, for each of the 15 target locations, we produced a projector calibration image (cf.\ again Section~\ref{sec:approach:geometry}) by projecting a $11~\times~4$ circles pattern image to the location in question using the steerable mirror, placing a checkerboard pattern beside the projected pattern, and acquiring the image using the downward-facing camera.
 We then fed these images alongside the output of the camera calibration module to our projector calibration module.
@@
 The remainder of the time was spent running our modules or working with the MDC-X software.
+Note that once the camera is calibrated, that calibration can be reused if the camera's intrinsics remain fixed, in particular if no change is made to the zoom factor of the camera.
 
@@
 Moreover, we showed our system to be able to handle a projector equipped with a steerable mirror, enabling factory floor augmentation exceeding the bounds of the projector's own immediate field of view.
-Our evaluation demonstrated ...
+Our evaluation demonstrated our approach to produce compelling results at less time than the more cumbersome traditional manual approach to keystone correction.
 
-A natural extension of this work would be to address non-planar scenes.
+A subject for a followup paper would be consider in greater detail the causes of and propose a solution to the apparent growth in discrepancy between intended metric projected dimensions and actual metric projected dimensions as the mirror is pointed further away from directly downwards.
+A further natural extension of this work would be to address non-planar scenes.
 To handle non-planar scenes would call for a change in how scene geometry is recovered and how distortion of the image to be projected is carried out; projector calibration could, however, be left unchanged.

### 2021-07-11 src/apps/calibrateProj/calibrateProj.cpp

@@ -261,8 +261,8 @@ void display()
             if (viewIdx == 1 && (i == 4))
                 continue;
 
-            if (viewIdx > 1 && (i == 0 || i == 1 || i == 3 || i == 5))
-                continue;
+            //if (viewIdx > 1 && (i == 0 || i == 1 || i == 3 || i == 5))
+            //    continue;
 
             float r = 0.5;
             float g = 0.5;

### 2021-07-11 paper/elsarticle/elsarticle-template.tex
paper

@@
 
-An early spatial AR system for planar scenes using a projector with a steerable mirror is the IBM Everywhere Displays prototype of Pinhanez \cite{pinhanez2001everywhere}, yet compute the warps applied to imagery to be projected manually.
-Rather than rely on a steerable mirror to support spatial AR to multiple locations using a single projector, some works mount the projector and a camera on a rigid rig and subject the rig to motion \cite{ehnes2004projected,borkowski2004spatial}.
+An early spatial AR system for using a projector with a steerable mirror is the IBM Everywhere Displays prototype of Pinhanez \cite{pinhanez2001everywhere}, yet carry out keystone correction manually.
+Rather than rely on a steerable mirror to support spatial AR to multiple locations using a single projector, some works mount the projector and a camera on a rigid rig and subject the rig to motion \cite{ehnes2004projected,borkowski2004spatial,butz2006applying}.
 Steering only a mirror, however, places humbler requirements on the system from a hardware standpoint than if an entire camera-projector rig is to be subject to steering.
@@
 
+\begin{figure}
+    \centering
+    \subfloat[\centering Circles pattern image, in image plane of projector (detections overlain).]{{\includegraphics[height=3.25cm]{images/circles.png} }}
+    \qquad
+    \subfloat[\centering Projector calibration image (one for each target location), in image plane of camera (detections overlain).]{{\includegraphics[height=3.25cm]{images/detections.png}}}
+    \caption{Recovering 2D positions in support of projector calibration.
+(a) 2D positions of the 2D-3D correspondences to be used for calibrating the projector are obtained by detecting---in the image plane of the projector---the circle centers in the circles pattern image, projected by the projector to each of the target locations in the scene plane.
+(b) For each such target location, an image is acquired from the viewpoint of the camera and the circle centers of the projected circles pattern are detected, in the image plane of the camera.
+A chessboard pattern to be used for recovering the local scene plane is placed near the projected pattern, whose corners are likewise detected.
+Detected 2D projected circles pattern center points and chessboard corners overlain for illustration.}
+    \label{fig:2d}
+\end{figure}
+
 Calibration of a projector (or camera) in the sense we employ the term here\footnote{We are referring to a geometric calibration; not, e.g., to a color calibration.} renders one able to project a scene point~$\mathbf{X} \in \mathbb{R}^3$ to its corresponding pixel~$\mathbf{x} \in \mathbb{R}^2$ in the projector's (or camera's) image plane, or to compute the `back-projection' of $\mathbf{x}$ (cf.\ Figure~\ref{fig:3d}), i.e., the ray from the projector's (or camera's) center of projection through $\mathbf{x}$ along which $\mathbf{X}$ must lie.
@@
 \begin{figure}
-    \centering
-    \subfloat[\centering Circles pattern image, in image plane of projector (detections overlain).]{{\includegraphics[height=3.25cm]{images/circles.png} }}
-    \qquad
-    \subfloat[\centering Projector calibration image (one for each target location), in image plane of camera (detections overlain).]{{\includegraphics[height=3.25cm]{images/detections.png}}}
-    \caption{Recovering 2D positions in support of projector calibration.
-(a) 2D positions of the 2D-3D correspondences to be used for calibrating the projector are obtained by detecting---in the image plane of the projector---the circle centers in the circles pattern image, projected by the projector to each of the target locations in the scene plane.
-(b) For each such target location, an image is acquired from the viewpoint of the camera and the circle centers of the projected circles pattern are detected, in the image plane of the camera.
-A chessboard pattern to be used for recovering the local scene plane is placed near the projected pattern, whose corners are likewise detected.
-Detected 2D projected circles pattern center points and chessboard corners overlain for illustration.}
-    \label{fig:2d}
+    \centerline{\includegraphics[scale=.35]{images/2d3d.png}}
+    \caption{Scene plane (gray) recovered via spatial resection with respect to 2D-3D correspondences obtained using a chessboard pattern (blue); 3D circles pattern points---i.e., the 3D positions of the
+2D-3D correspondences to be used for calibrating the projector---obtained by intersection with the scene plane of back-projections (likewise gray) of the 2D circles pattern center points detected in the image plane (red).
+Note that the points in the figure are the points recovered for the projector calibration image in Figure~\ref{fig:2d}(b), acquired by the downwards-facing camera (frustum of the camera in black, with up vector in red).}
+    \label{fig:3d}
 \end{figure}
@@
 As this is sufficient for our needs in Section~\ref{sec:approach:homography}, it is in this sense that our system is able to handle a projector equipped with a steerable mirror, without need for modeling the steerable mirror explicitly.
-
-\begin{figure}
-    \centerline{\includegraphics[scale=.35]{images/2d3d.png}}
-    \caption{Scene plane (gray) recovered via spatial resection with respect to 2D-3D correspondences obtained using a chessboard pattern (blue); 3D circles pattern points---i.e., the 3D positions of the
-2D-3D correspondences to be used for calibrating the projector---obtained by intersection with the scene plane of back-projections (likewise gray) of the 2D circles pattern center points detected in the image plane (red).
-Note that the points in the figure are the points recovered for the projector calibration image in Figure~\ref{fig:2d}(b), acquired by the downwards-facing camera (frustum of the camera in black, with up vector in red).}
-    \label{fig:3d}
-\end{figure}
 

### 2021-07-11 paper/elsarticle/elsarticle-template.tex
paper

@@
 
-We propose a spatial AR system for planar scenes that produces the effect of keystone correction analytically, and that facilitates placement by enabling intuitively placing the desired augmentations in a manner aligned with the axes of an image of the scene acquired by a camera facing downwards towards the scene plane.
+We propose a spatial AR system for planar scenes that produces the effect of keystone correction analytically.
+Our system does so in a manner that facilitates placement by enabling intuitively placing the desired augmentations in a manner aligned with the axes of an image of the scene acquired by a camera facing downwards towards the scene plane, and reasoning about the dimensions of the projected augmentations in metric terms.
 Moreover, our system is able to handle a projector equipped with a steerable mirror, enabling factory floor augmentation exceeding the bounds of the projector's own immediate field of view, using only a single projector.
@@
 
-An early spatial AR system for using a projector with a steerable mirror is the IBM Everywhere Displays prototype of Pinhanez \cite{pinhanez2001everywhere}, yet carry out keystone correction manually.
+Keystone correction for planar scenes can be carried out using a 2D homography \cite{Hartley2004}---an invertible transformation that preserves colinearity---and is ultimately the approach we take as well.
+One approach to calibrating a projector-camera system using a homography reduces to automatically establishing keypoint matches between the image to be projected in the projector's image plane and the projected image in the camera's image plane, and then to use those matches to automatically compute the homography that warps the one image to the other \cite{sukthankar2001smarter}.
+While such an approach could be extended to enable placement to multiple locations in accordance with a steerable mirror, using keypoint matches alone does not enable reasoning about the target dimensions of augmentations in metric terms.
+
+An early system explicitly using a projector with a steerable mirror is the IBM Everywhere Displays prototype of Pinhanez \cite{pinhanez2001everywhere}; the system, however, has keystone correction carried out manually.
 Rather than rely on a steerable mirror to support spatial AR to multiple locations using a single projector, some works mount the projector and a camera on a rigid rig and subject the rig to motion \cite{ehnes2004projected,borkowski2004spatial,butz2006applying}.
-Steering only a mirror, however, places humbler requirements on the system from a hardware standpoint than if an entire camera-projector rig is to be subject to steering.
-...
+Steering only a mirror, however, places humbler requirements on the system from a hardware standpoint than if an entire camera-projector rig is to be subject to motion.
 
@@
 [7] contribute a number of interactions in the tradition of ubiquitous computing, such as the ability to
-
-\section{Approach}
-
-Correcting for projective distortions of the sort outlined in Section~\ref{sec:intro} can be achieved by modeling the manner in which the respective rays through the pixels of the projector's image plane fan out into the scene (i.e., by `calibrating' the projector) and the geometry of the scene itself (i.e., by recovering the scene plane relative to the projector) within at least the projector's field of view.
-This is because the scene point `illuminated' by a pixel in the projector's image plane can be recovered by intersecting its corresponding ray with the geometry of the scene surface.
-To model this interaction, we (i) carry out a one-time projector calibration, which in our approach calls for additionally calibrating a camera facing downward to the scene plane and includes recovery of the scene plane as a convenient side effect.
-Next, we use the relative camera-projector-scene plane geometry to (ii) compute a plane-induced homography that distorts the image to be projected in a manner that it appear undistorted to the viewer, and placed in alignment with the axes of a proxy image of the scene.
-These two points are treated in Sections~\ref{sec:approach:geometry} and \ref{sec:approach:homography}, respectively.
-
-\subsection{Recovering Geometry}\label{sec:approach:geometry}
 
@@
 \end{figure}
+
+\section{Approach}
+
+Correcting for projective distortions of the sort outlined in Section~\ref{sec:intro} can be achieved by modeling the manner in which the respective rays through the pixels of the projector's image plane fan out into the scene (i.e., by `calibrating' the projector) and the geometry of the scene itself (i.e., by recovering the scene plane relative to the projector) within at least the projector's field of view.
+This is because the scene point `illuminated' by a pixel in the projector's image plane can be recovered by intersecting its corresponding ray with the geometry of the scene surface.
+To model this interaction, we (i) carry out a one-time projector calibration, which in our approach calls for additionally calibrating a camera facing downward to the scene plane and includes recovery of the scene plane as a convenient side effect.
+Next, we use the relative camera-projector-scene plane geometry to (ii) compute a plane-induced homography that distorts the image to be projected in a manner that it appear undistorted to the viewer, and placed in alignment with the axes of a proxy image of the scene.
+These two points are treated in Sections~\ref{sec:approach:geometry} and \ref{sec:approach:homography}, respectively.
+
+\subsection{Recovering Geometry}\label{sec:approach:geometry}
 

### 2021-07-11 paper/elsarticle/elsarticle-template.tex
Update elsarticle-template.tex

@@
 Keystone correction for planar scenes can be carried out using a 2D homography \cite{Hartley2004}---an invertible transformation that preserves colinearity---and is ultimately the approach we take as well.
-One approach to calibrating a projector-camera system using a homography reduces to automatically establishing keypoint matches between the image to be projected in the projector's image plane and the projected image in the camera's image plane, and then to use those matches to automatically compute the homography that warps the one image to the other \cite{sukthankar2001smarter}.
+One approach to calibrating a projector-camera system using a homography reduces to automatically establishing keypoint matches between the image to be projected in the projector's image plane and the projected image in the camera's image plane, and then using those matches to automatically compute the homography that warps the one image to the other \cite{sukthankar2001smarter}.
 While such an approach could be extended to enable placement to multiple locations in accordance with a steerable mirror, using keypoint matches alone does not enable reasoning about the target dimensions of augmentations in metric terms.

### 2021-07-12 paper/elsarticle/elsarticle-template.tex
paper

--- 2021-07-11 paper/elsarticle/elsarticle-template.tex
@@
 
-An early spatial AR system for using a projector with a steerable mirror is the IBM Everywhere Displays prototype of Pinhanez \cite{pinhanez2001everywhere}, yet carry out keystone correction manually.
+Keystone correction for planar scenes can be carried out using a 2D homography \cite{Hartley2004}---an invertible transformation that preserves colinearity, hence alternatively termed a `colineation'---and is ultimately the approach we take as well.
+One approach to calibrating a projector-camera system in a fully automatic fashion using a homography reduces to automatically establishing keypoint matches between the image to be projected in the projector's image plane and the projected image in the camera's image plane, and then to use those matches to automatically compute the homography that warps the one image to the other \cite{sukthankar2001smarter}.
+While such an approach could be extended to enable placement to multiple locations in accordance with a steerable mirror \cite{okatani2005autocalibration}, relying on keypoint matches alone does not enable reasoning about the target dimensions of augmentations in metric terms, or offer the convenience of placing annotations in terms of the principal axes of camera facing the scene plane.
+
+An early spatial AR system explicitly using a projector with a steerable mirror is the IBM Everywhere Displays prototype of Pinhanez \cite{pinhanez2001everywhere}; the system, however, has keystone correction carried out manually.
 Rather than rely on a steerable mirror to support spatial AR to multiple locations using a single projector, some works mount the projector and a camera on a rigid rig and subject the rig to motion \cite{ehnes2004projected,borkowski2004spatial,butz2006applying}.
-Steering only a mirror, however, places humbler requirements on the system from a hardware standpoint than if an entire camera-projector rig is to be subject to steering.
-...
+Steering only a mirror, however, places humbler requirements on the system from a hardware standpoint than if an entire camera-projector rig is to be subject to rigid motion.
 
@@
 
+Calibration of a projector (or camera) in the sense we employ the term here\footnote{We are referring to a geometric calibration; not, e.g., to a color calibration.} renders one able to project a scene point~$\mathbf{X} \in \mathbb{R}^3$ to its corresponding pixel~$\mathbf{x} \in \mathbb{R}^2$ in the projector's (or camera's) image plane, or to compute the `back-projection' of $\mathbf{x}$, i.e., the ray from the projector's (or camera's) center of projection through $\mathbf{x}$ along which $\mathbf{X}$ must lie.
+Such a calibration can be expressed in terms of (i) a $3~\times{}~3$ calibration matrix~$\mathtt{K}$ derived from the projector's (or camera's) focal length and principal point \cite{Hartley2004}, and (ii) the coefficients of a lens distortion model used to correct for radial or tangential distortions caused by the lens system \cite{duane1971close}.
+
 \begin{figure}
@@
 \end{figure}
-
-Calibration of a projector (or camera) in the sense we employ the term here\footnote{We are referring to a geometric calibration; not, e.g., to a color calibration.} renders one able to project a scene point~$\mathbf{X} \in \mathbb{R}^3$ to its corresponding pixel~$\mathbf{x} \in \mathbb{R}^2$ in the projector's (or camera's) image plane, or to compute the `back-projection' of $\mathbf{x}$ (cf.\ Figure~\ref{fig:3d}), i.e., the ray from the projector's (or camera's) center of projection through $\mathbf{x}$ along which $\mathbf{X}$ must lie.
-Such a calibration can be expressed in terms of (i) a $3~\times{}~3$ calibration matrix~$\mathtt{K}$ derived from the projector's (or camera's) focal length and principal point \cite{Hartley2004}, and (ii) the coefficients of a lens distortion model used to correct for radial or tangential distortions caused by the lens system \cite{duane1971close}.
 
@@
 
+\paragraph{Projector calibration} As with camera calibration, projector calibration relies on 2D-3D correspondences, yet we obtain them in this case by projecting a pattern of circles.
+We rely on an algorithm to detect the circle pattern center points in the circles pattern image in the image plane of the projector \cite{bradski2000opencv}, giving the 2D positions of our 2D-3D correspondences for calibrating the projector (cf.\ Figure~\ref{fig:2d}(a)).
+We project the circles pattern image to each of the target locations, and use the camera to acquire a projector calibration image for each.
+Given a projector calibration image acquired using the downwards-facing camera, we detect the circle centers of the \textit{projected} circles pattern (cf.\ Figure~\ref{fig:2d}(b)); given the scene plane (the recovery of which we shall return to in the paragraph that follows) and such a 2D circle center $\mathbf{x}$, its 3D correspondence is obtained by intersecting the back-projection of $\mathbf{x}$ with the scene plane (cf.\ Figure~\ref{fig:3d}).
+Since the algorithm that yields 2D circle centers does so in a consistent ordering, we thus obtain the 2D-3D correspondences between the projector's image plane and the scene required for projector calibration, yielding the projector calibration matrix~$\mathtt{K}_\text{proj}$.
+
 \begin{figure}
@@
 
-\paragraph{Projector calibration} As with camera calibration, projector calibration relies on 2D-3D correspondences, yet we obtain them in this case by projecting a pattern of circles.
-We rely on an algorithm to detect the circle pattern center points in the circles pattern image in the image plane of the projector \cite{bradski2000opencv}, giving the 2D positions of our 2D-3D correspondences for calibrating the projector (cf.\ Figure~\ref{fig:2d}(a)).
-We project the circles pattern image to each of the target locations, and use the camera to acquire a projector calibration image for each.
-Given a projector calibration image acquired using the downwards-facing camera, we detect the circle centers of the \textit{projected} circles pattern (cf.\ Figure~\ref{fig:2d}(b)); given the scene plane (the recovery of which we shall return to in the paragraph that follows) and such a 2D circle center $\mathbf{x}$, its 3D correspondence is obtained by intersecting the back-projection of $\mathbf{x}$ with the scene plane (cf.\ Figure~\ref{fig:3d}).
-Since the algorithm that yields 2D circle centers does so in a consistent ordering, we thus obtain the 2D-3D correspondences between the projector's image plane and the scene required for projector calibration, yielding the projector calibration matrix~$\mathtt{K}_\text{proj}$.
-
 We recover the scene plane via spatial resection by applying a PnP algorithm \cite{terzakis2020consistently} to the 2D-3D correspondences obtained using a chessboard pattern.
 Note that this step is separate from camera calibration, yet could well be carried out using the same calibration pattern used in the camera calibration step.\footnote{The critical point is that the pattern should ideally be coplanar with the local scene plane, meaning its height above the scene plane should not exceed a few millimeters.} While a single image of such a chessboard pattern placed on the floor could be sufficient if the floor is even, we place a chessboard pattern in close proximity to the projected circles pattern in each projector calibration image in order to recover the scene plane locally to each target location, in order to account for the possibility of an uneven floor.
+Note that in principle, we could project a chessboard pattern instead of a circles pattern to obtain the 2D-3D correspondences needed for projector calibration; it is, however, because we rely on detecting the corners of a chessboard pattern to recover the scene plane that we opt instead for a alternative pattern.
 
@@
 Projecting an image warped in this manner to the scene plane \textit{from the viewpoint of the recovered projector} then has the same effect as projecting the unwarped image to the scene plane from the viewpoint of the virtual projector.
-This warp can computed using a plane-induced homography, computed analytically as a function of the scene plane, the projector, and the virtual projector.
+This warp can effected using a plane-induced homography, computed analytically as a function of the scene plane, the projector, and the virtual projector.
 

### 2021-07-14 paper/elsarticle/elsarticle-template.tex
paper

@@
 Keystone correction for planar scenes can be carried out using a 2D homography \cite{Hartley2004}---an invertible transformation that preserves colinearity, hence alternatively termed a `colineation'---and is ultimately the approach we take as well.
-One approach to calibrating a projector-camera system in a fully automatic fashion using a homography reduces to automatically establishing keypoint matches between the image to be projected in the projector's image plane and the projected image in the camera's image plane, and then to use those matches to automatically compute the homography that warps the one image to the other \cite{sukthankar2001smarter}.
-While such an approach could be extended to enable placement to multiple locations in accordance with a steerable mirror \cite{okatani2005autocalibration}, relying on keypoint matches alone does not enable reasoning about the target dimensions of augmentations in metric terms, or offer the convenience of placing annotations in terms of the principal axes of camera facing the scene plane.
+The intuition for why a transformation that maps lines to lines can serve to model the necessary image warp can be drawn from considering a planar chessboard pattern: looking at an image of a chessboard acquired from an oblique angle, one observes that parallel lines appear to meet in respective vanishing points; looking at an image of the same chessboard acquired frontally with respect to the plane of the chessboard, lines parallel in the chessboard appear parallel in the image (i.e., they are said to meet `at infinity').
+To warp the former image (where lines parallel in the scene meet in respective vanishing points) such that the lines of the chessboard appear as in the latter image (where lines parallel in the scene meet at infinity), a transformation that maps lines to lines that would give the desired effect.
+
+One approach to calibrating a projector-camera system in a fully automatic fashion using a homography reduces to (i) automatically establishing keypoint matches between the image to be projected in the projector's image plane and the projected image in the camera's image plane, and then to (ii) using those matches to automatically compute the homography that warps the one image to the other \cite{sukthankar2001smarter}.
+Relying on keypoint matches alone, however, does not enable reasoning about the target dimensions of augmentations in metric terms, or offer the convenience of placing annotations in terms of the principal axes of camera facing the scene plane.
+For this, calibration of the camera and of the projector is necessary: \cite{raskar2001self}...
 
@@
 Steering only a mirror, however, places humbler requirements on the system from a hardware standpoint than if an entire camera-projector rig is to be subject to rigid motion.
+
+Mention circles pattern: \cite{chiu2011novel}
+
+...
 
@@
 
-Calibration of a projector (or camera) in the sense we employ the term here\footnote{We are referring to a geometric calibration; not, e.g., to a color calibration.} renders one able to project a scene point~$\mathbf{X} \in \mathbb{R}^3$ to its corresponding pixel~$\mathbf{x} \in \mathbb{R}^2$ in the projector's (or camera's) image plane, or to compute the `back-projection' of $\mathbf{x}$, i.e., the ray from the projector's (or camera's) center of projection through $\mathbf{x}$ along which $\mathbf{X}$ must lie.
-Such a calibration can be expressed in terms of (i) a $3~\times{}~3$ calibration matrix~$\mathtt{K}$ derived from the projector's (or camera's) focal length and principal point \cite{Hartley2004}, and (ii) the coefficients of a lens distortion model used to correct for radial or tangential distortions caused by the lens system \cite{duane1971close}.
-
 \begin{figure}
@@
 \end{figure}
+
+Calibration of a projector (or camera) in the sense we employ the term here\footnote{We are referring to a geometric calibration; not, e.g., to a color calibration.} renders one able to project a scene point~$\mathbf{X} \in \mathbb{R}^3$ to its corresponding pixel~$\mathbf{x} \in \mathbb{R}^2$ in the projector's (or camera's) image plane, or to compute the `back-projection' of $\mathbf{x}$, i.e., the ray from the projector's (or camera's) center of projection through $\mathbf{x}$ along which $\mathbf{X}$ must lie.
+Such a calibration can be expressed in terms of (i) a $3~\times{}~3$ calibration matrix~$\mathtt{K}$ derived from the projector's (or camera's) focal length and principal point \cite{Hartley2004}, and (ii) the coefficients of a lens distortion model used to correct for radial or tangential distortions caused by the lens system \cite{duane1971close}.
 

### 2021-07-14 paper/elsarticle/elsarticle-template.tex
Paper

@@
 Keystone correction for planar scenes can be carried out using a 2D homography \cite{Hartley2004}---an invertible transformation that preserves colinearity, hence alternatively termed a `colineation'---and is ultimately the approach we take as well.
-The intuition for why a transformation that maps lines to lines can serve to model the necessary image warp can be drawn from considering a planar chessboard pattern: looking at an image of a chessboard acquired from an oblique angle, one observes that parallel lines appear to meet in respective vanishing points; looking at an image of the same chessboard acquired frontally with respect to the plane of the chessboard, lines parallel in the chessboard appear parallel in the image (i.e., they are said to meet `at infinity').
+The intuition for why a transformation that maps lines to lines can serve to model the appropriate image warp can be drawn from considering a planar chessboard pattern: looking at an image of a chessboard acquired from an oblique angle, one observes that lines parallel in the chessboard appear to meet in respective vanishing points; looking at an image of the same chessboard acquired frontally with respect to the plane of the chessboard, lines parallel in the chessboard appear parallel in the image (i.e., they are said to meet `at infinity').
 To warp the former image (where lines parallel in the scene meet in respective vanishing points) such that the lines of the chessboard appear as in the latter image (where lines parallel in the scene meet at infinity), a transformation that maps lines to lines that would give the desired effect.

### 2021-07-14 paper/elsarticle/elsarticle-template.tex
paper

@@
 The intuition for why a transformation that maps lines to lines can serve to model the appropriate image warp can be drawn from considering a planar chessboard pattern: looking at an image of a chessboard acquired from an oblique angle, one observes that lines parallel in the chessboard appear to meet in respective vanishing points; looking at an image of the same chessboard acquired frontally with respect to the plane of the chessboard, lines parallel in the chessboard appear parallel in the image (i.e., they are said to meet `at infinity').
-To warp the former image (where lines parallel in the scene meet in respective vanishing points) such that the lines of the chessboard appear as in the latter image (where lines parallel in the scene meet at infinity), a transformation that maps lines to lines that would give the desired effect.
+To warp the former image (where lines parallel in the scene meet in respective vanishing points) such that the lines of the chessboard appear as in the latter image (where lines parallel in the scene meet at infinity), a transformation maps lines to lines that would give the desired effect.
 

### 2021-07-14 paper/elsarticle/elsarticle-template.tex
paper

@@
 This placement is produced according to a small handful of steps.
-First, we (i) rotate the camera about its center of projection to align its optical axis with the normal vector of the scene plane, giving a virtual camera likewise facing directly\footnote{A physical camera placed to face downwards is likely to not face downwards precisely; in contrast, the virtual camera's optical axis is aligned exactly with the scene plane's normal vector, rendering it fronto-parallel with respect to the scene plane.} downwards to the scene plane (cf.\ Figure~\ref{fig:virtualproj}).
+First, we (i) rotate the camera about its center of projection to align its optical axis with the normal vector of the scene plane, giving a virtual camera likewise facing directly\footnote{A physical camera placed to face downwards is almost certain to not face downwards precisely; in contrast, the virtual camera's optical axis is aligned exactly with the scene plane's normal vector, rendering it fronto-parallel with respect to the scene plane.} downwards to the scene plane (cf.\ Figure~\ref{fig:virtualproj}).
 Next, we (ii) intersect the scene plane with the optical axis (i.e., the ray from the projector's center of projection through the center of the image plane) and rotate the projector's placement about that point of intersection, aligning the optical axis with the scene plane's normal vector and giving an initial virtual projector.

### 2021-07-14 paper/elsarticle/elsarticle-template.tex
paper

@@
 
-Augmented reality (AR) \cite{van2010survey,zhou2008trends} promises to enable use cases in industrial settings that include the embedding of assembly instructions directly into the scene \cite{schlund2018moglichkeiten,masood2019augmented,gattullo2019towards}, potentially reducing or altogether doing away with the need for workers to refer to instructions in paper form or on a screen.
+Augmented reality (AR) \cite{van2010survey,zhou2008trends} promises to enable use cases in industrial settings that include the embedding of assembly instructions directly into the scene \cite{schlund2018moglichkeiten,masood2019augmented,gattullo2019towards,rupprecht2020information,Rupprecht2021}, potentially reducing or altogether doing away with the need for workers to refer to instructions in paper form or on a screen.
 Typically, AR works by embedding the augmentation in an image of the scene acquired from the viewpoint of a single individual, with the resulting augmented image in turn displayed using some form of head-mounted display.
@@
 
-\subsection{Correcting for Projective Distortion}\label{sec:approach:homography}
-
-If the projector is calibrated and its pose relative to the scene plane is known, a `virtual' projector (with the same calibration $\mathtt{K}$ and lens distortion coefficients) can be placed elsewhere relative to the scene plane.
-If we for a moment imagine that the projector---at its recovered pose---functions as a camera,\footnote{Recall that the calibration matrix~$\mathtt{K}$ enables computing both (i) the projection of a scene point to the image plane (the function of a camera), or (ii) the back-projection of a pixel in the image plane, giving a ray into the scene (along which a projector illuminates the scene with the given pixel).} then (i) projecting an image to the scene plane \textit{from the viewpoint of the virtual projector} and (ii) acquiring the resulting projected image from the viewpoint of the recovered projector gives the desired corrective warp.
-Projecting an image warped in this manner to the scene plane \textit{from the viewpoint of the recovered projector} then has the same effect as projecting the unwarped image to the scene plane from the viewpoint of the virtual projector.
-This warp can effected using a plane-induced homography, computed analytically as a function of the scene plane, the projector, and the virtual projector.
-
 \begin{figure}
@@
 \end{figure}
+
+\subsection{Correcting for Projective Distortion}\label{sec:approach:homography}
+
+If the projector is calibrated and its pose relative to the scene plane is known, a `virtual' projector (with the same calibration $\mathtt{K}$ and lens distortion coefficients) can be placed elsewhere relative to the scene plane.
+If we for a moment imagine that the projector---at its recovered pose---functions as a camera,\footnote{Recall that the calibration matrix~$\mathtt{K}$ enables computing both (i) the projection of a scene point to the image plane (the function of a camera), or (ii) the back-projection of a pixel in the image plane, giving a ray into the scene (along which a projector illuminates the scene with the given pixel).} then (i) projecting an image to the scene plane \textit{from the viewpoint of the virtual projector} and (ii) acquiring the resulting projected image from the viewpoint of the recovered projector gives the desired corrective warp.
+Projecting an image warped in this manner to the scene plane \textit{from the viewpoint of the recovered projector} then has the same effect as projecting the unwarped image to the scene plane from the viewpoint of the virtual projector.
+This warp can effected using a plane-induced homography, computed analytically as a function of the scene plane, the projector, and the virtual projector.
 
@@
 
-The hardware setup employed in the evaluation comprised a Panasonic PT-RZ660BE projector with a steerable mirror system---used in our experiments to point the projection to each of the 15 locations---manufactured by Dynamic Projection Institute \cite{rupprecht2020information}.
+The hardware setup employed in the evaluation comprised a Panasonic PT-RZ660BE projector with a steerable mirror system---used in our experiments to point the projection to each of the 15 locations---manufactured by Dynamic Projection Institute \cite{rupprecht2020information,Rupprecht2021}.
 The steerable mirror system was bundled with the MDC-X software for steering the mirror, loading imagery, and optionally carrying out manual keystone correction, such that each position and (warped) image can be registered as a preset.

### 2021-07-14 paper/elsarticle/elsarticle-template.tex
paper

@@
 \author{Sebastian Schlund}
-\address{Human Centered Cyber Physical Production and Assembly Systems, Institute for Management Sciences, TU Vienna, Austria}
+\address{Human Centered Cyber Physical Production and Assembly Systems, Institute for Management Sciences, TU Wien, Vienna, Austria}
 
@@
 We propose a spatial AR system for planar scenes that produces the effect of keystone correction analytically.
-Our system does so in a manner that facilitates placement by enabling intuitively placing the desired augmentations in a manner aligned with the axes of an image of the scene acquired by a camera facing downwards towards the scene plane, and reasoning about the dimensions of the projected augmentations in metric terms.
+Our system does so in a manner that facilitates placement of augmentations in the scene by enabling intuitively placing the desired augmentations in a manner aligned with the axes of an image of the scene acquired by a camera facing downwards towards the scene plane, and reasoning about the dimensions of the projected augmentations in metric terms.
 Moreover, our system is able to handle a projector equipped with a steerable mirror, enabling factory floor augmentation exceeding the bounds of the projector's own immediate field of view, using only a single projector.
@@
 Reliance on head-mounted displays, however, has two adverse consequences: (i) a head-mounted display must be worn by each individual wishing to view the augmentation, and (ii) such a head-mounted display---in some cases taking the form of a helmet in order to house multiple sensors in support of accurately tracking the viewpoint of the viewer relative to the scene---can be obtrusive.
-\textit{Spatial} AR is a form of augmented reality carried out not by embedding the augmentation in an image of the scene as with head-mounted displays, but by projection to the scene itself \cite{bimber2019spatial}, thus eliminating both aforementioned problems.
+\textit{Spatial} AR\cite{bimber2019spatial} is a form of augmented reality carried out not by embedding the augmentation in an image of the scene as with head-mounted displays, but by projection to the scene itself , thus eliminating both aforementioned problems.
 Yet considering for the moment a planar surface to be augmented, unless the projector faces the surface frontally, the bounds of a projected rectangular image will not appear rectangular; more generally, they will instead appear trapezoidal (i.e., the projected image will appear distorted).
@@
 
-We evaluate our approach by augmenting 15 locations across the floorspace at the Pilotfabrik\footnote{\url{https://www.pilotfabrik.at/}} of TU Vienna, a collaborative space for research on Industry 4.0 topics situated in Vienna, Austria.
+We evaluate our approach by augmenting 15 locations across the floorspace at the Pilotfabrik\footnote{\url{https://www.pilotfabrik.at/}} of TU Wien, a collaborative space for research on Industry 4.0 topics situated in Vienna, Austria.
 We contrast our approach with a baseline approach involving manual keystone correction, by aiming with both approaches to place the same image for each location aligned with the principal axes of the floorspace, absent of projective distortions, and with the same metric dimensions (50~cm~$\times$~31.25~cm\footnote{The dimensions in pixels of the image we project are 960~pixels~$\times$~600~pixels; we chose for our experiments to set the projected metric length of the horizontal axis of the image to 50~cm, which implies 31.25~cm for the vertical axis.}).

### 2021-07-14 paper/elsarticle/elsarticle-template.tex
paper

@@
 Reliance on head-mounted displays, however, has two adverse consequences: (i) a head-mounted display must be worn by each individual wishing to view the augmentation, and (ii) such a head-mounted display---in some cases taking the form of a helmet in order to house multiple sensors in support of accurately tracking the viewpoint of the viewer relative to the scene---can be obtrusive.
-\textit{Spatial} AR\cite{bimber2019spatial} is a form of augmented reality carried out not by embedding the augmentation in an image of the scene as with head-mounted displays, but by projection to the scene itself , thus eliminating both aforementioned problems.
+In turn, \textit{spatial} AR\cite{bimber2019spatial} is a form of augmented reality carried out not by embedding the augmentation in an image of the scene as with head-mounted displays, but by projection to the scene itself, thus eliminating both aforementioned problems.
 Yet considering for the moment a planar surface to be augmented, unless the projector faces the surface frontally, the bounds of a projected rectangular image will not appear rectangular; more generally, they will instead appear trapezoidal (i.e., the projected image will appear distorted).

### 2021-07-15 paper/elsarticle/elsarticle-template.tex
paper

@@
 
-\title{A Spatial AR System for Axis-aligned Augmentation of Planar Scenes in Industrial Settings} %\title{Elsevier \LaTeX\ template\tnoteref{mytitlenote}}
+\title{A Spatial AR System for Axis-aligned Metric Augmentation of Planar Scenes} %\title{Elsevier \LaTeX\ template\tnoteref{mytitlenote}}
 %\tnotetext[mytitlenote]{Fully documented templates are available in the elsarticle package on \href{http://www.ctan.org/tex-archive/macros/latex/contrib/elsarticle}{CTAN}.}
@@
 \textit{Spatial} AR, in turn, is a form of AR whereby the augmentation of the scene is carried out using a projector, with the advantage of rendering the augmentation visible to all onlookers simultaneously without calling for each to wear some form of head-mounted display.
-Care must be taken, however, to distort the images to be projected in a manner that they appear undistorted to the viewer, since the geometry of the scene as it relates to the geometry of the projector plays a role in how the pixels of the projector's image plane map to points in the scene.
+Care must be taken in spatial AR, however, to distort the images to be projected in a manner that they appear undistorted to the viewer, since the geometry of the scene as it relates to the geometry of the projector plays a role in how the pixels of the projector's image plane map to points in the scene.
 For planar scene geometry (such as a floor, wall, or table), this can be done in a cumbersome manual process called keystone correction, often using software bundled with the projector.
@@
 The intuition for why a transformation that maps lines to lines can serve to model the appropriate image warp can be drawn from considering a planar chessboard pattern: looking at an image of a chessboard acquired from an oblique angle, one observes that lines parallel in the chessboard appear to meet in respective vanishing points; looking at an image of the same chessboard acquired frontally with respect to the plane of the chessboard, lines parallel in the chessboard appear parallel in the image (i.e., they are said to meet `at infinity').
-To warp the former image (where lines parallel in the scene meet in respective vanishing points) such that the lines of the chessboard appear as in the latter image (where lines parallel in the scene meet at infinity), a transformation maps lines to lines that would give the desired effect.
+To warp the former image (where lines parallel in the scene meet in respective vanishing points) such that the lines of the chessboard appear as in the latter image (where lines parallel in the scene meet at infinity), a transformation that maps lines to lines would give the desired effect.
 
-One approach to calibrating a projector-camera system in a fully automatic fashion using a homography reduces to (i) automatically establishing keypoint matches between the image to be projected in the projector's image plane and the projected image in the camera's image plane, and then to (ii) using those matches to automatically compute the homography that warps the one image to the other \cite{sukthankar2001smarter}.
-Relying on keypoint matches alone, however, does not enable reasoning about the target dimensions of augmentations in metric terms, or offer the convenience of placing annotations in terms of the principal axes of camera facing the scene plane.
-For this, calibration of the camera and of the projector is necessary: \cite{raskar2001self}...
+One way to compute a homography is by identifying at least four correspondences between pixel positions in two images of a planar surface.
+The keystone correction approach of Sukthankar \etal \cite{sukthankar2001smarter} reduces to using a segmentation approach to identify the four corners of projection screen in an image acquired by a camera, and computing a homography that maps the resulting four corners of the projection screen to the four corners of the projector's image plane.
+Calibrated approach: \cite{raskar2001self}...
 
@@
 We evaluate our approach by augmenting 15 locations across the floorspace at the Pilotfabrik\footnote{\url{https://www.pilotfabrik.at/}} of TU Wien, a collaborative space for research on Industry 4.0 topics situated in Vienna, Austria.
-We contrast our approach with a baseline approach involving manual keystone correction, by aiming with both approaches to place the same image for each location aligned with the principal axes of the floorspace, absent of projective distortions, and with the same metric dimensions (50~cm~$\times$~31.25~cm\footnote{The dimensions in pixels of the image we project are 960~pixels~$\times$~600~pixels; we chose for our experiments to set the projected metric length of the horizontal axis of the image to 50~cm, which implies 31.25~cm for the vertical axis.}).
+We contrast our approach with a baseline approach involving manual keystone correction, by aiming with both approaches to place the same image for each location aligned with the principal axes of the floorspace, absent of projective distortions, and with the same metric dimensions (50~cm~$\times$~31.25~cm\footnote{The dimensions in pixels of the image we project are 960~$\times$~600; we chose for our experiments to set the projected metric length of the horizontal axis of the image to 50~cm, which implies 31.25~cm for the vertical axis if aspect ratio is to be preserved.}).
 All experiments were carried out by the same technician, experienced in both approaches.
@@
 
-We presented a spatial AR system for planar scenes that produces the effect of keystone correction analytically, and that intuitively places the desired augmentations in a manner aligned with the axes of an image of the scene acquired by a camera.
+We presented a spatial AR system for planar scenes that produces the effect of keystone correction analytically, in a manner that enables intuitive placement the augmentations in accordance with the axes of an image of the scene acquired by a downwards-facing camera and such that the desired dimensions of augmentations can be specified in metric terms.
 Moreover, we showed our system to be able to handle a projector equipped with a steerable mirror, enabling factory floor augmentation exceeding the bounds of the projector's own immediate field of view.

### 2021-07-15 paper/elsarticle/elsarticle-template.tex
paper

@@
 Typically, AR works by embedding the augmentation in an image of the scene acquired from the viewpoint of a single individual, with the resulting augmented image in turn displayed using some form of head-mounted display.
-Reliance on head-mounted displays, however, has two adverse consequences: (i) a head-mounted display must be worn by each individual wishing to view the augmentation, and (ii) such a head-mounted display---in some cases taking the form of a helmet in order to house multiple sensors in support of accurately tracking the viewpoint of the viewer relative to the scene---can be obtrusive.
-In turn, \textit{spatial} AR\cite{bimber2019spatial} is a form of augmented reality carried out not by embedding the augmentation in an image of the scene as with head-mounted displays, but by projection to the scene itself, thus eliminating both aforementioned problems.
-Yet considering for the moment a planar surface to be augmented, unless the projector faces the surface frontally, the bounds of a projected rectangular image will not appear rectangular; more generally, they will instead appear trapezoidal (i.e., the projected image will appear distorted).
+Reliance on head-mounted displays, however, has two adverse consequences: (i) a head-mounted display must be worn by each individual wishing to partake in the augmentation, and (ii) such a head-mounted display---in some cases taking the form of a helmet in order to house multiple sensors in support of accurately tracking the viewpoint of the viewer relative to the scene---can be obtrusive.
+In turn, \textit{spatial} AR\cite{bimber2019spatial} is a form of augmented reality carried out not by embedding the augmentation in an image of the scene as with a head-mounted display, but by projection to the scene itself, thus eliminating both aforementioned problems.
+Yet considering a planar surface to be augmented (e.g., a floor, wall, or table), unless the projector faces the surface frontally, the bounds of a projected rectangular image will not appear rectangular; more generally, they will instead appear trapezoidal (i.e., the projected image will appear distorted).
 Such distortions can be eliminated by carrying out a cumbersome manual process called keystone correction to appropriately warp the image to be projected, often using software bundled with the projector.
 
-Our contribution is to propose a spatial AR system that produces the effect of keystone correction analytically, and---using the $X$- and $Y$-axes of an image of the scene acquired by a downwards-facing camera as a proxy---in a manner aligning the axes of the augmentation with those of the proxy image.
+Our contribution is to propose a spatial AR system for planar scenes that produces the effect of keystone correction analytically, and---using the $X$- and $Y$-axes of an image of the scene acquired by a downwards-facing camera as a proxy---in a manner aligning the axes of the augmentation with those of the proxy image.
 We achieve this by distorting the image to be projected using a plane-induced homography computed to produce the effect of projecting the image not from the actual projector viewpoint, but in accordance with the viewpoint of a \textit{virtual} projector (i) facing directly downwards to the scene plane and (ii) rotated to place the axes of the image plane of the virtual projector in line with those of the camera.
@@
     \subfloat[\centering Projecting our warped image.]{{\includegraphics[height=4.2cm]{images/proj_projim_waped.png}}}
-    \caption{Projection from the recovered projector viewpoint (bottom left, black; virtual projector in center, gray) to the scene plane of the original image and of the warped image.
+    \caption{Projection to the scene plane from the recovered projector viewpoint (bottom left, black; virtual projector in center, gray) of the original image and of the warped image.
 (a) Projecting the original image to the scene plane.

### 2021-07-17 paper/elsarticle/elsarticle-template.tex
paper

@@
 \author{Patrick Rupprecht}
-\author{Sebastian Schlund}
+\author{Sebastian~Schlund}
 \address{Human Centered Cyber Physical Production and Assembly Systems, Institute for Management Sciences, TU Wien, Vienna, Austria}
@@
 In turn, \textit{spatial} AR\cite{bimber2019spatial} is a form of augmented reality carried out not by embedding the augmentation in an image of the scene as with a head-mounted display, but by projection to the scene itself, thus eliminating both aforementioned problems.
-Yet considering a planar surface to be augmented (e.g., a floor, wall, or table), unless the projector faces the surface frontally, the bounds of a projected rectangular image will not appear rectangular; more generally, they will instead appear trapezoidal (i.e., the projected image will appear distorted).
+Yet considering a planar surface to be augmented (e.g., a floor, wall, or table), unless the projector faces the surface frontally, the bounds of a projected rectangular image will not appear rectangular.
 Such distortions can be eliminated by carrying out a cumbersome manual process called keystone correction to appropriately warp the image to be projected, often using software bundled with the projector.
@@
 The keystone correction approach of Sukthankar \etal \cite{sukthankar2001smarter} reduces to using a segmentation approach to identify the four corners of projection screen in an image acquired by a camera, and computing a homography that maps the resulting four corners of the projection screen to the four corners of the projector's image plane.
-Calibrated approach: \cite{raskar2001self}...
+Tilt sensors can be used to recover the projector's gravity vector, and be integrated in the computation of a homography as in Raskar and Beardsley \cite{raskar2001self}.
 
@@
 Mention circles pattern: \cite{chiu2011novel}
-
-...
 
@@
 
-We recover the scene plane via spatial resection by applying a PnP algorithm \cite{terzakis2020consistently} to the 2D-3D correspondences obtained using a chessboard pattern.
+We recover the scene plane via spatial resection by applying a PnP algorithm \cite{collins2014infinitesimal} to the 2D-3D correspondences obtained using a chessboard pattern.
 Note that this step is separate from camera calibration, yet could well be carried out using the same calibration pattern used in the camera calibration step.\footnote{The critical point is that the pattern should ideally be coplanar with the local scene plane, meaning its height above the scene plane should not exceed a few millimeters.} While a single image of such a chessboard pattern placed on the floor could be sufficient if the floor is even, we place a chessboard pattern in close proximity to the projected circles pattern in each projector calibration image in order to recover the scene plane locally to each target location, in order to account for the possibility of an uneven floor.
@@
 
-We presented a spatial AR system for planar scenes that produces the effect of keystone correction analytically, in a manner that enables intuitive placement the augmentations in accordance with the axes of an image of the scene acquired by a downwards-facing camera and such that the desired dimensions of augmentations can be specified in metric terms.
+We presented a spatial AR system for planar scenes that produces the effect of keystone correction analytically, in a manner that enables intuitive placement the augmentations in accordance with the axes of an image of the scene acquired by a downwards-facing camera, and such that the desired dimensions of augmentations can be specified in metric terms.
 Moreover, we showed our system to be able to handle a projector equipped with a steerable mirror, enabling factory floor augmentation exceeding the bounds of the projector's own immediate field of view.

### 2021-07-17 src/apps/calibrateProj/calibrateProj.cpp

@@ -472,8 +472,8 @@ void reshape(int width, int height)
     // todo: compare with http://cvrr.ucsd.edu/publications/2008/MurphyChutorian_Trivedi_CVGPU08.pdf
     double matTransp[] = { 2*f/w,     0,         0,                                          0,
                            0,         2*f/h,     0,                                          0,
-                          -2*cx/w+1,  2*cy/h-1, -(farPlane+nearPlane)/(farPlane-nearPlane), -1,
-                           0,         0,        -2*farPlane*nearPlane/(farPlane-nearPlane),  0};
+                          1-2*cx/w,  -1 + (2*cy + 2)/h, (farPlane+nearPlane)/(nearPlane - farPlane), -1,
+                           0,         0,        2*farPlane*nearPlane/(nearPlane-farPlane),  0};
 	
     glLoadMatrixd(matTransp);
 	
@@ -921,43 +921,47 @@ int main(int argc, char** argv)
 
     // get circles and chessboard image points determine corresponding plane parameters
     std::vector<std::vector<cv::Point2f>> cam0ChessboardImPts, cam0CirclesImPts;
+    cv::Mat cam0ImVis;
     CalibPattern::findChessboardAndCirclesImPts(cam0ImDir, numIms,
         chessboardPatternSize, circlesPatternSize,
         cam0ChessboardImPts, cam0CirclesImPts, cv::Size(),
-        true, cam0K, cam0DistCoeffs);
+        true, cam0K, cam0DistCoeffs, cam0ImVis, visImIdx);
 
     std::vector<std::vector<cv::Point2f>> cam1ChessboardImPts, cam1CirclesImPts;
+    cv::Mat cam1ImVis;
     CalibPattern::findChessboardAndCirclesImPts(cam1ImDir, numIms,
         chessboardPatternSize, circlesPatternSize,
         cam1ChessboardImPts, cam1CirclesImPts, cv::Size(),
-        true, cam1K, cam1DistCoeffs);
-
-    std::vector<cv::Mat> planeRs, planeTs;
-
-    // use not to calibrate camera (we provide camera parameters and keep them fixed), but to obtain plane for each image in cam coordinate frame
[…]

### 2021-07-17 src/cpslib/libCalib/CalibPattern.cpp

@@ -8,7 +8,7 @@ void CalibPattern::computeChessboardObjPts(int numIms, float chessboardSquareSiz
     for (int i = 0; i < chessboardPatternSize.height; i++)
         for (int j = 0; j < chessboardPatternSize.width; j++)
             chessboardObjectPts_.push_back(
-                cv::Point3f(float(j * chessboardSquareSize), float(i * chessboardSquareSize), 0));
+                cv::Point3f(j * chessboardSquareSize, i * chessboardSquareSize, 0.0));
 
     // store copy of chessboard object points once per input image
     for (int numIm = 0; numIm < numIms; numIm++)
@@ -57,10 +57,8 @@ void CalibPattern::findChessboardImPts(std::string inDir, int numIms, cv::Size c
         ss1 << inDir << "\\" << numIm << ".png";
         std::cout << ss1.str() << std::endl;
 
-        cv::Mat im = cv::imread(ss1.str());
-
         cv::Mat imVis;
-        findChessboardImPts(im, chessboardPatternSize,
+        findChessboardImPts(cv::imread(ss1.str()), chessboardPatternSize,
             outChessboardImPts, outImSize, imVis,
             applyIntrinsics, K, distCoeffs);
 
@@ -108,7 +106,7 @@ void CalibPattern::findCirclesImPts(cv::Mat& im, cv::Size circlesPatternSize,
 
 void CalibPattern::findChessboardAndCirclesImPts(std::string inDir, int numIms, cv::Size chessboardPatternSize, cv::Size circlesPatternSize,
     std::vector<std::vector<cv::Point2f>>& outChessboardImPts, std::vector<std::vector<cv::Point2f>>& outCirclesImPts, cv::Size& outImSize,
-    bool applyIntrinsics, cv::Mat& K, cv::Mat& distCoeffs)
+    bool applyIntrinsics, cv::Mat& K, cv::Mat& distCoeffs, cv::Mat& imVis, int visImIdx)
 {
     for (int numIm = 0; numIm < numIms; numIm++)
     {
@@ -116,23 +114,24 @@ void CalibPattern::findChessboardAndCirclesImPts(std::string inDir, int numIms, 
         ss1 << inDir << "\\" << numIm << ".png";
         std::cout << ss1.str() << std::endl;
 
-        cv::Mat im = cv::imread(ss1.str());
-
-        cv::Mat imVis;
-        findChessboardImPts(im, chessboardPatternSize,
[…]

### 2021-07-17 src/cpslib/libCalib/CalibPattern.h

@@ -24,7 +24,7 @@ public:
         bool applyIntrinsics = true, cv::Mat& K = cv::Mat(), cv::Mat& distCoeffs = cv::Mat());
     static void findChessboardAndCirclesImPts(std::string inDir, int numIms, cv::Size chessboardPatternSize, cv::Size circlesPatternSize,
         std::vector<std::vector<cv::Point2f>>& outChessboardImPts, std::vector<std::vector<cv::Point2f>>& outCirclesImPts, cv::Size& outImSize,
-        bool applyIntrinsics = false, cv::Mat& K = cv::Mat(), cv::Mat& distCoeffs = cv::Mat());
+        bool applyIntrinsics = false, cv::Mat& K = cv::Mat(), cv::Mat& distCoeffs = cv::Mat(), cv::Mat& imVis = cv::Mat(), int visImIdx = 0);
 
 private:
 

### 2021-07-17 src/cpslib/libGeom/Plane.cpp

@@ -14,7 +14,14 @@ Plane::Plane(cv::Mat & rigid)
 	distance_ = 0;
 
 	rigid_ = cv::Mat::eye(cv::Size(4, 4), CV_64F);
-	rigidTransform(rigid);
+	cv::Vec3d pt = -normal_ * distance_;
+
+	pt = Ancillary::Mat44dTimesVec3dHomog(rigid, pt);
+
+	normal_ = Ancillary::Mat33dTimesVec3d(rigid, normal_);
+	distance_ = -normal_.dot(pt);
+
+	rigid.copyTo(rigid_);
 }
 
 Plane::Plane(cv::Vec3d normal, float distance)

### 2021-07-18 paper/elsarticle/elsarticle-template.tex
paper

@@
 
-\title{A Spatial AR System for Axis-aligned Metric Augmentation of Planar Scenes} %\title{Elsevier \LaTeX\ template\tnoteref{mytitlenote}}
+\title{A Spatial AR System for Wide-area Axis-aligned Metric Augmentation of Planar Scenes} %\title{Elsevier \LaTeX\ template\tnoteref{mytitlenote}}
 %\tnotetext[mytitlenote]{Fully documented templates are available in the elsarticle package on \href{http://www.ctan.org/tex-archive/macros/latex/contrib/elsarticle}{CTAN}.}
@@
 
-We propose a spatial AR system for planar scenes that produces the effect of keystone correction analytically.
-Our system does so in a manner that facilitates placement of augmentations in the scene by enabling intuitively placing the desired augmentations in a manner aligned with the axes of an image of the scene acquired by a camera facing downwards towards the scene plane, and reasoning about the dimensions of the projected augmentations in metric terms.
-Moreover, our system is able to handle a projector equipped with a steerable mirror, enabling factory floor augmentation exceeding the bounds of the projector's own immediate field of view, using only a single projector.
+We propose a spatial AR system for augmenting planar scene surfaces that produces the effect of keystone correction analytically.
+Our system does so in a manner that facilitates placement of augmentations in the scene by enabling intuitively (i) placing the desired augmentations in accordance with the axes of an image of the scene acquired by a camera placed to face downwards towards the scene plane, and (ii) reasoning about the desired dimensions of the projected augmentations in metric terms.
+Moreover, our system is able to handle a projector equipped with a steerable mirror, enabling wide-area factory floor augmentation exceeding the bounds of the projector's immediate field of view.
 \end{abstract}
@@
 In turn, \textit{spatial} AR\cite{bimber2019spatial} is a form of augmented reality carried out not by embedding the augmentation in an image of the scene as with a head-mounted display, but by projection to the scene itself, thus eliminating both aforementioned problems.
-Yet considering a planar surface to be augmented (e.g., a floor, wall, or table), unless the projector faces the surface frontally, the bounds of a projected rectangular image will not appear rectangular.
+Yet considering a planar surface to be augmented (e.g., a floor, wall, or table), unless the projector faces the surface frontally, the bounds of a projected rectangular image will not appear rectangular, but will instead appear distorted.
 Such distortions can be eliminated by carrying out a cumbersome manual process called keystone correction to appropriately warp the image to be projected, often using software bundled with the projector.
@@
 This facilitates placement of augmentations by intuitively placing them in accordance with the axes of an image of the scene, and eliminates the need for manual keystone correction.
-Moreover, a consequence of our approach is that our system is able to handle a projector equipped with a steerable mirror (without need for explicitly modeling the action of the steerable mirror on the projector), thereby enabling factory floor applications exceeding the immediate field of view of the projector without needing to rely on multiple projectors.
+Moreover, a consequence of our approach is that our system is able to handle a projector equipped with a steerable mirror (without need for explicitly modeling the action of the steerable mirror on the projector), thereby enabling wide-area factory floor applications exceeding the immediate field of view of the projector without needing to rely on multiple projectors.
 
@@
 We rely on an algorithm to detect the circle pattern center points in the circles pattern image in the image plane of the projector \cite{bradski2000opencv}, giving the 2D positions of our 2D-3D correspondences for calibrating the projector (cf.\ Figure~\ref{fig:2d}(a)).
-We project the circles pattern image to each of the target locations, and use the camera to acquire a projector calibration image for each.
-Given a projector calibration image acquired using the downwards-facing camera, we detect the circle centers of the \textit{projected} circles pattern (cf.\ Figure~\ref{fig:2d}(b)); given the scene plane (the recovery of which we shall return to in the paragraph that follows) and such a 2D circle center $\mathbf{x}$, its 3D correspondence is obtained by intersecting the back-projection of $\mathbf{x}$ with the scene plane (cf.\ Figure~\ref{fig:3d}).
+We project the circles pattern image to each of the target locations, and use the the downwards-facing camera to acquire a projector calibration image for each.
+Given a projector calibration image acquired using camera, we detect the circle centers of the \textit{projected} circles pattern (cf.\ Figure~\ref{fig:2d}(b)); given the scene plane (the recovery of which we shall return to in the paragraph that follows) and such a 2D circle center $\mathbf{x}$, its 3D correspondence is obtained by intersecting the back-projection of $\mathbf{x}$ with the scene plane (cf.\ Figure~\ref{fig:3d}).
 Since the algorithm that yields 2D circle centers does so in a consistent ordering, we thus obtain the 2D-3D correspondences between the projector's image plane and the scene required for projector calibration, yielding the projector calibration matrix~$\mathtt{K}_\text{proj}$.
@@
 2D-3D correspondences to be used for calibrating the projector---obtained by intersection with the scene plane of back-projections (likewise gray) of the 2D circles pattern center points detected in the image plane (red).
-Note that the points in the figure are the points recovered for the projector calibration image in Figure~\ref{fig:2d}(b), acquired by the downwards-facing camera (frustum of the camera in black, with up vector in red).}
+Note that as in the figures that follow, the rendering in the figure corresponds to the projection calibration image in Figure~\ref{fig:2d}(b), acquired by the downwards-facing camera (frustum of the camera in black, with up vector in red).}
     \label{fig:3d}
@@
     \caption{The virtual camera is obtained by (i) rotating the camera (top left, black) about its center of projection such that its optical axis be made parallel with the normal vector of the scene plane.
-The virtual projector is obtained by (ii) rotating the projector (bottom left, black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (iii) rotating the $X$- and $Y$-axes to align them with those of the virtual camera, and optionally (iv) translating along the normal direction to achieve the desired projected image dimensions.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
+The virtual projector is obtained by (ii) rotating the projector (bottom left, black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (iii) rotating the $X$- and $Y$-axes to align them with those of the virtual camera, and (iv) translating along the normal direction to achieve the desired projected image dimensions in metric units.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
     \label{fig:virtualproj}
@@
 The virtual projector is thus rendered fronto-parallel with the scene plane, enabling projection to the scene plane absent of projective distortions.
-Optionally, we additionally (iv) adjust the height above the scene plane of the virtual projector, in order to satisfy desired projected image dimensions provided in metric units.
+We additionally (iv) adjust the height above the scene plane of the virtual projector, in order to satisfy desired projected image dimensions provided in metric units.
 
 Owing to the manner in which we place the virtual projector, the virtual projector's axes and thus the augmentation are aligned with the axes of the downward-facing camera; the placement of the camera thereby intuitively determines the principal axes according to which augmentations are placed.
-Note further that a consequence of placing the virtual projector by rotating about the point of intersection of the projector's optical axis with the scene plane is that the center of the projector's image plane remains invariant to the placement of the virtual projector, i.e., a steerable mirror can be aimed with respect to a point projected from the center of the projector's image plane, further facilitating placement.
+Note further that a consequence of placing the virtual projector by rotating about the point of intersection of the projector's optical axis with the scene plane is that the center of the projector's image plane remains invariant to the placement of the virtual projector, i.e., a steerable mirror can be aimed with respect to a point projected from the center of the projector's image plane, further facilitating placement of augmentations.
 
@@
 
-A subject for a followup paper would be consider in greater detail the causes of and propose a solution to the apparent growth in discrepancy between intended metric projected dimensions and actual metric projected dimensions as the mirror is pointed further away from directly downwards.
-A further natural extension of this work would be to address non-planar scenes.
-To handle non-planar scenes would call for a change in how scene geometry is recovered and how distortion of the image to be projected is carried out; projector calibration could, however, be left unchanged.
+A natural extension of this work would be to address non-planar scenes.
+To handle non-planar scenes would call for a change in how scene geometry is recovered and how distortion of the image to be projected is carried out; the methodology we proposed for projector calibration could, however, be left unchanged.
+
+\section{Acknowledgments}
+
+This work was supported by the Austrian Research Promotion Agency (FFG) through its endowed professorship in Human Centered Cyber Physical Production and Assembly Systems at TU Wien (FFG-852789).
 
@@
 \end{document}
+
+% idea for next work: LM-based approach to globally recover scene plane from projected patterns (try to make a patent of this!)...
+compute intersection of optical axes of all recovered projectors - it is about this point that the projector can be steered in terms of angles; once scene plane is availabe, intrinsics of projector are available, and the 'steering point' is available, we can have a GUI with which you click on an undistorted image of the scene plane to control the steerable mirror

### 2021-07-19 paper/elsarticle/elsarticle-template.tex
paper

@@
 \usepackage{colortbl}
+\usepackage{amsmath}
 \modulolinenumbers[5]
@@
 We propose a spatial AR system for augmenting planar scene surfaces that produces the effect of keystone correction analytically.
-Our system does so in a manner that facilitates placement of augmentations in the scene by enabling intuitively (i) placing the desired augmentations in accordance with the axes of an image of the scene acquired by a camera placed to face downwards towards the scene plane, and (ii) reasoning about the desired dimensions of the projected augmentations in metric terms.
-Moreover, our system is able to handle a projector equipped with a steerable mirror, enabling wide-area factory floor augmentation exceeding the bounds of the projector's immediate field of view.
+Our system renders the placement of augmentations in the scene intuitive in two ways.
+First, (i) placement of the desired augmentations is carried out in accordance with the axes of an image of the scene acquired by a camera placed to face downwards towards the scene plane; second, (ii) the desired dimensions of the projected augmentations is given in metric terms.
+Moreover, our system is able to handle projector equipped with a steerable mirror, enabling wide-area factory floor augmentation exceeding the bounds of a single projector's immediate field~of~view.
 \end{abstract}
@@
 Keystone correction for planar scenes can be carried out using a 2D homography \cite{Hartley2004}---an invertible transformation that preserves colinearity, hence alternatively termed a `colineation'---and is ultimately the approach we take as well.
-The intuition for why a transformation that maps lines to lines can serve to model the appropriate image warp can be drawn from considering a planar chessboard pattern: looking at an image of a chessboard acquired from an oblique angle, one observes that lines parallel in the chessboard appear to meet in respective vanishing points; looking at an image of the same chessboard acquired frontally with respect to the plane of the chessboard, lines parallel in the chessboard appear parallel in the image (i.e., they are said to meet `at infinity').
-To warp the former image (where lines parallel in the scene meet in respective vanishing points) such that the lines of the chessboard appear as in the latter image (where lines parallel in the scene meet at infinity), a transformation that maps lines to lines would give the desired effect.
+The intuition for why it is that a transformation that maps lines to lines can serve to model the appropriate image warp can be drawn from considering a planar chessboard pattern: looking at an image of a chessboard acquired from an oblique angle, one observes that lines parallel in the chessboard appear to meet in respective vanishing points; looking at an image of the same chessboard acquired frontally with respect to the plane of the chessboard, lines parallel in the chessboard appear parallel in the image (i.e., they are said to meet `at infinity').
+To warp the former image (where lines parallel in the scene meet in respective vanishing points) such that the lines of the chessboard appear as in the latter image (where lines parallel in the scene meet at infinity), a transformation that maps lines to lines would be sufficient.
 
@@
 In practice, calibrating a camera relies on (i) establishing 2D-3D correspondences between pixels in the camera's image plane and corresponding points in the scene, and on (ii) using those correspondences as input to an optimization procedure that relies on bundle adjustment \cite{triggs1999bundle} to output the calibration matrix~$\mathtt{K}$, the associated lens distortion model coefficients, and, for each calibration image, the pose (i.e., position and orientation) of the camera relative to the 3D points \cite{Hartley2004,zhang2000flexible}.
-To calibrate a camera, a calibration surface such as a chessboard pattern is used to identify the correspondences.
+To calibrate a camera, images of a calibration surface such as a chessboard pattern are used to identify the 2D-3D correspondences $\{\mathbf{x}_{i,j} \leftrightarrow \mathbf{X}_i\}$, $i \in \{ 1, \dots, n_\text{pt} \}, j \in \{ 1, \dots, n_\text{im} \}$, where $n_\text{pt}$ gives the number of correspondences obtained from one calibration image of the pattern, and $n_\text{im}$ the number of such images.
+Bundle adjustment is used to obtain the calibration matrix~$\mathtt{K}$ and rigid body transformations $g = (\mathtt{R}_j, \mathbf{t}_j) \in SE(3)$ that minimize the reprojection error
+\begin{equation}
+\sum_{i,j} d\big(\mathbf{x}_{i,j}, \mathtt{K}(\mathtt{R}_j\mathbf{X}_i + \mathbf{t}_j)\big),
+\label{re}
+\end{equation}
+where $\mathbf{X}_{i,j} = \mathtt{R}_j\mathbf{X}_i + \mathbf{t}_j$ expresses $\mathbf{X}_i$ in the camera coordinate frame corresponding to the $j^{\text{th}}$ calibration image, $\mathtt{K}\mathbf{X}_{i,j} \in \mathbb{R}^2$ gives the projection of $\mathbf{X}_i$ to that image, and $d$ is a distance function.
+The inverse rigid body transformation~$g^{-1}$ gives the pose of the $j^{\text{th}}$ camera relative to the points $\{\mathbf{X}_i\}$.
+
 Calibration of a projector can be carried out in precisely the same manner insofar as step (ii) is concerned; the major difference in projector calibration relative to the calibrating a camera concerns the manner in which 2D-3D correspondences are identified, i.e., between pixels in the image plane of the projector and the corresponding points in the scene.
@@
 
-\paragraph{Camera calibration} We recover 2D-3D correspondences in support of calibrating the downwards-facing camera is carried out by relying on a planar calibration surface to automatically identify correspondences between the 3D points on the calibration surface and their 2D correspondences in the image plane.
+\paragraph{Camera calibration} We recover 2D-3D correspondences in support of calibrating the downwards-facing camera by relying on a planar calibration surface to automatically identify correspondences between the 3D points on the calibration surface and their 2D correspondences in the image plane.
 The classical calibration surface is a chessboard pattern.
@@
     \caption{The virtual camera is obtained by (i) rotating the camera (top left, black) about its center of projection such that its optical axis be made parallel with the normal vector of the scene plane.
-The virtual projector is obtained by (ii) rotating the projector (bottom left, black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (iii) rotating the $X$- and $Y$-axes to align them with those of the virtual camera, and (iv) translating along the normal direction to achieve the desired projected image dimensions in metric units.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
+The virtual projector is obtained by (ii) rotating the projector (bottom left, black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (iii) rotating the $X$- and $Y$-axes to align them with those of the virtual camera, and (iv) translating along the normal direction to achieve the desired metric projected image dimensions.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
     \label{fig:virtualproj}

### 2021-07-19 paper/elsarticle/elsarticle-template.tex
paper

@@
 In practice, calibrating a camera relies on (i) establishing 2D-3D correspondences between pixels in the camera's image plane and corresponding points in the scene, and on (ii) using those correspondences as input to an optimization procedure that relies on bundle adjustment \cite{triggs1999bundle} to output the calibration matrix~$\mathtt{K}$, the associated lens distortion model coefficients, and, for each calibration image, the pose (i.e., position and orientation) of the camera relative to the 3D points \cite{Hartley2004,zhang2000flexible}.
-To calibrate a camera, images of a calibration surface such as a chessboard pattern are used to identify the 2D-3D correspondences $\{\mathbf{x}_{i,j} \leftrightarrow \mathbf{X}_i\}$, $i \in \{ 1, \dots, n_\text{pt} \}, j \in \{ 1, \dots, n_\text{im} \}$, where $n_\text{pt}$ gives the number of correspondences obtained from one calibration image of the pattern, and $n_\text{im}$ the number of such images.
+To calibrate a camera, images of a calibration surface such as a chessboard pattern are used to identify the 2D-3D correspondences $\{\mathbf{x}_{i,j} \leftrightarrow \mathbf{X}_{i,j}\}$, $i \in \{ 1, \dots, n_\text{pt} \}, j \in \{ 1, \dots, n_\text{im} \}$, where $n_\text{pt}$ gives the number of correspondences obtained from one calibration image of the pattern, $n_\text{im}$ the number of such images, and $\{\mathbf{X}_{i,j}\} = \{\mathbf{X}_{i,j'}\}$, for $j, j' \in \{ 1, \dots, n_\text{im} \}$.
 Bundle adjustment is used to obtain the calibration matrix~$\mathtt{K}$ and rigid body transformations $g = (\mathtt{R}_j, \mathbf{t}_j) \in SE(3)$ that minimize the reprojection error
 \begin{equation}
-\sum_{i,j} d\big(\mathbf{x}_{i,j}, \mathtt{K}(\mathtt{R}_j\mathbf{X}_i + \mathbf{t}_j)\big),
+\sum_{i,j} d\big(\mathbf{x}_{i,j}, \mathtt{K}(\mathtt{R}_j\mathbf{X}_{i,j} + \mathbf{t}_j)\big),
 \label{re}
 \end{equation}
-where $\mathbf{X}_{i,j} = \mathtt{R}_j\mathbf{X}_i + \mathbf{t}_j$ expresses $\mathbf{X}_i$ in the camera coordinate frame corresponding to the $j^{\text{th}}$ calibration image, $\mathtt{K}\mathbf{X}_{i,j} \in \mathbb{R}^2$ gives the projection of $\mathbf{X}_i$ to that image, and $d$ is a distance function.
-The inverse rigid body transformation~$g^{-1}$ gives the pose of the $j^{\text{th}}$ camera relative to the points $\{\mathbf{X}_i\}$.
+where $\mathbf{X}_{i,j}' = \mathtt{R}_j\mathbf{X}_{i,j} + \mathbf{t}_j$ expresses $\mathbf{X}_i$ in the camera coordinate frame corresponding to the $j^{\text{th}}$ calibration image, $\mathtt{K}\mathbf{X}_{i,j}' \in \mathbb{R}^2$ gives the projection of $\mathbf{X}_{i,j}$ to that image, and $d$ is a distance function.
+The inverse rigid body transformation~$g^{-1}$ gives the pose of the $j^{\text{th}}$ camera relative to the static 3D points of the calibration pattern.
 
 Calibration of a projector can be carried out in precisely the same manner insofar as step (ii) is concerned; the major difference in projector calibration relative to the calibrating a camera concerns the manner in which 2D-3D correspondences are identified, i.e., between pixels in the image plane of the projector and the corresponding points in the scene.
-A convenient consequence of the approach we take to identifying the 3D points of the 2D-3D correspondences needed for projector calibration is, for each target location, recovery of the pose (i.e., position and orientation) of (i) the projector and of (ii) the scene plane, both relative to the coordinate frame of the camera.
+What remains of this section is concerned primarily with the recovery of these correspondences.
+A consequence of the approach we take to identifying the 3D points of the 2D-3D correspondences we use for projector calibration is, for each target location, recovery of the pose of the scene plane relative to the coordinate frame of the camera.
 

### 2021-07-19 paper/elsarticle/elsarticle-template.tex
paper

@@
 
-Calibration of a projector can be carried out in precisely the same manner insofar as step (ii) is concerned; the major difference in projector calibration relative to the calibrating a camera concerns the manner in which 2D-3D correspondences are identified, i.e., between pixels in the image plane of the projector and the corresponding points in the scene.
+Calibration of a projector can be carried out in precisely the same manner as calibrating a camera insofar as step (ii) is concerned; the major difference in projector calibration relative to the calibrating a camera concerns the manner in which 2D-3D correspondences are identified, i.e., between pixels in the image plane of the projector and the corresponding points in the scene.
 What remains of this section is concerned primarily with the recovery of these correspondences.

### 2021-07-19 paper/elsarticle/elsarticle-template.tex
paper

@@
 \end{equation}
-where $\mathbf{X}_{i,j}' = \mathtt{R}_j\mathbf{X}_{i,j} + \mathbf{t}_j$ expresses $\mathbf{X}_i$ in the camera coordinate frame corresponding to the $j^{\text{th}}$ calibration image, $\mathtt{K}\mathbf{X}_{i,j}' \in \mathbb{R}^2$ gives the projection of $\mathbf{X}_{i,j}$ to that image, and $d$ is a distance function.
+where $\mathbf{X}_{i,j}' = \mathtt{R}_j\mathbf{X}_{i,j} + \mathbf{t}_j$ expresses $\mathbf{X}_{i,j}$ in the camera coordinate frame corresponding to the $j^{\text{th}}$ calibration image, $\mathtt{K}\mathbf{X}_{i,j}' \in \mathbb{R}^2$ gives the projection of $\mathbf{X}_{i,j}$ to that image, and $d$ is a distance function.
 The inverse rigid body transformation~$g^{-1}$ gives the pose of the $j^{\text{th}}$ camera relative to the static 3D points of the calibration pattern.

### 2021-07-19 paper/elsarticle/elsarticle-template.tex
paper

@@
 To calibrate a camera, images of a calibration surface such as a chessboard pattern are used to identify the 2D-3D correspondences $\{\mathbf{x}_{i,j} \leftrightarrow \mathbf{X}_{i,j}\}$, $i \in \{ 1, \dots, n_\text{pt} \}, j \in \{ 1, \dots, n_\text{im} \}$, where $n_\text{pt}$ gives the number of correspondences obtained from one calibration image of the pattern, $n_\text{im}$ the number of such images, and $\{\mathbf{X}_{i,j}\} = \{\mathbf{X}_{i,j'}\}$, for $j, j' \in \{ 1, \dots, n_\text{im} \}$.
-Bundle adjustment is used to obtain the calibration matrix~$\mathtt{K}$ and rigid body transformations $g = (\mathtt{R}_j, \mathbf{t}_j) \in SE(3)$ that minimize the reprojection error
+Bundle adjustment is used to obtain the lens distortion coefficients, calibration matrix~$\mathtt{K}$, and rigid body transformations $g = (\mathtt{R}_j, \mathbf{t}_j) \in SE(3)$ that minimize the reprojection error
 \begin{equation}

### 2021-07-20 paper/elsarticle/elsarticle-template.tex
paper

--- 2021-07-19 paper/elsarticle/elsarticle-template.tex
@@
 \textit{Spatial} AR, in turn, is a form of AR whereby the augmentation of the scene is carried out using a projector, with the advantage of rendering the augmentation visible to all onlookers simultaneously without calling for each to wear some form of head-mounted display.
-Care must be taken in spatial AR, however, to distort the images to be projected in a manner that they appear undistorted to the viewer, since the geometry of the scene as it relates to the geometry of the projector plays a role in how the pixels of the projector's image plane map to points in the scene.
+Care must be taken in carrying out spatial AR, however, to warp the images to be projected in a manner that they appear free of distortions to the viewer, since the geometry of the scene as it relates to the geometry of the projector plays a role in how the pixels of the projector's image plane map to points in the scene.
 For planar scene geometry (such as a floor, wall, or table), this can be done in a cumbersome manual process called keystone correction, often using software bundled with the projector.
 
-We propose a spatial AR system for augmenting planar scene surfaces that produces the effect of keystone correction analytically.
-Our system renders the placement of augmentations in the scene intuitive in two ways.
-First, (i) placement of the desired augmentations is carried out in accordance with the axes of an image of the scene acquired by a camera placed to face downwards towards the scene plane; second, (ii) the desired dimensions of the projected augmentations is given in metric terms.
-Moreover, our system is able to handle projector equipped with a steerable mirror, enabling wide-area factory floor augmentation exceeding the bounds of a single projector's immediate field~of~view.
+We propose a spatial AR system for wide-area metric augmentation of planar scene surfaces that produces the effect of keystone correction analytically, by using a projector equipped with a steerable mirror and a downwards-facing camera.
+Our system renders the placement of augmentations in the scene more intuitive than manual keystone correction in two ways.
+First, (i) placement of the desired augmentations is carried out in accordance with the axes of an image of the scene acquired by a camera placed to face downwards towards the scene plane; second, (ii) the desired dimensions of the projected augmentations is specified in metric terms.
+Our system produces such augmentations accurately, with compelling time savings relative to the manual approach.
 \end{abstract}
@@
 
-Our contribution is to propose a spatial AR system for planar scenes that produces the effect of keystone correction analytically, and---using the $X$- and $Y$-axes of an image of the scene acquired by a downwards-facing camera as a proxy---in a manner aligning the axes of the augmentation with those of the proxy image.
-We achieve this by distorting the image to be projected using a plane-induced homography computed to produce the effect of projecting the image not from the actual projector viewpoint, but in accordance with the viewpoint of a \textit{virtual} projector (i) facing directly downwards to the scene plane and (ii) rotated to place the axes of the image plane of the virtual projector in line with those of the camera.
+Our contribution is to propose a wide-area spatial AR system for planar scenes that produces the effect of keystone correction analytically, and---using the $X$- and $Y$-axes of an image of the scene acquired by a downwards-facing camera as a proxy---in a manner aligning the axes of the augmentation with those of the proxy image.
+Moreover, our system enables specifying the dimensions of augmentations in metric terms.
+We achieve this by warping the image to be projected using a plane-induced homography computed to produce the effect of projecting the image not from the actual projector viewpoint, but in accordance with the viewpoint of a \textit{virtual} projector (i) facing directly downwards to the scene plane and (ii) rotated to place the axes of the image plane of the virtual projector in line with those of the camera.
 This facilitates placement of augmentations by intuitively placing them in accordance with the axes of an image of the scene, and eliminates the need for manual keystone correction.
-Moreover, a consequence of our approach is that our system is able to handle a projector equipped with a steerable mirror (without need for explicitly modeling the action of the steerable mirror on the projector), thereby enabling wide-area factory floor applications exceeding the immediate field of view of the projector without needing to rely on multiple projectors.
+Finally, our system is able to handle a projector equipped with a steerable mirror (without need for explicitly modeling the action of the steerable mirror on the projector), thereby enabling wide-area factory floor applications exceeding the immediate field of view of the projector without needing to rely on multiple projectors.
 
@@
 In practice, calibrating a camera relies on (i) establishing 2D-3D correspondences between pixels in the camera's image plane and corresponding points in the scene, and on (ii) using those correspondences as input to an optimization procedure that relies on bundle adjustment \cite{triggs1999bundle} to output the calibration matrix~$\mathtt{K}$, the associated lens distortion model coefficients, and, for each calibration image, the pose (i.e., position and orientation) of the camera relative to the 3D points \cite{Hartley2004,zhang2000flexible}.
-To calibrate a camera, images of a calibration surface such as a chessboard pattern are used to identify the 2D-3D correspondences $\{\mathbf{x}_{i,j} \leftrightarrow \mathbf{X}_{i,j}\}$, $i \in \{ 1, \dots, n_\text{pt} \}, j \in \{ 1, \dots, n_\text{im} \}$, where $n_\text{pt}$ gives the number of correspondences obtained from one calibration image of the pattern, $n_\text{im}$ the number of such images, and $\{\mathbf{X}_{i,j}\} = \{\mathbf{X}_{i,j'}\}$, for $j, j' \in \{ 1, \dots, n_\text{im} \}$.
-Bundle adjustment is used to obtain the calibration matrix~$\mathtt{K}$ and rigid body transformations $g = (\mathtt{R}_j, \mathbf{t}_j) \in SE(3)$ that minimize the reprojection error
+To calibrate a camera, images of a calibration surface such as a chessboard pattern are used to identify the 2D-3D correspondences $\{\mathbf{x}_{i,j} \leftrightarrow \mathbf{X}_{i,j}\}$, $i \in \{ 1, \dots, n_\text{pt} \}, j \in \{ 1, \dots, n_\text{im} \}$, where $n_\text{pt}$ gives the number of correspondences obtained from one calibration image of the pattern, $n_\text{im}$ the number of such images, and $\mathbf{X}_{i,j} = \mathbf{X}_{i,j'}$, for $j, j' \in \{ 1, \dots, n_\text{im} \}$.
+Bundle adjustment is used to obtain the lens distortion model coefficients, calibration matrix~$\mathtt{K}$, and rigid body transformations $g = (\mathtt{R}_j, \mathbf{t}_j) \in SE(3)$ that minimize the reprojection error
 \begin{equation}
@@
 \end{equation}
-where $\mathbf{X}_{i,j}' = \mathtt{R}_j\mathbf{X}_{i,j} + \mathbf{t}_j$ expresses $\mathbf{X}_i$ in the camera coordinate frame corresponding to the $j^{\text{th}}$ calibration image, $\mathtt{K}\mathbf{X}_{i,j}' \in \mathbb{R}^2$ gives the projection of $\mathbf{X}_{i,j}$ to that image, and $d$ is a distance function.
+where $\mathbf{X}_{i,j}' = \mathtt{R}_j\mathbf{X}_{i,j} + \mathbf{t}_j$ expresses $\mathbf{X}_{i,j}$ in the camera coordinate frame corresponding to the $j^{\text{th}}$ calibration image, $\mathtt{K}\mathbf{X}_{i,j}' \in \mathbb{R}^2$ gives the projection of $\mathbf{X}_{i,j}$ to that image, and $d$ is a distance function that corrects for lens distortion using the recovered lens distortion model coefficients and computes a distance with respect to the two undistorted pixel positions.
 The inverse rigid body transformation~$g^{-1}$ gives the pose of the $j^{\text{th}}$ camera relative to the static 3D points of the calibration pattern.
@@
 Calibration of a projector can be carried out in precisely the same manner as calibrating a camera insofar as step (ii) is concerned; the major difference in projector calibration relative to the calibrating a camera concerns the manner in which 2D-3D correspondences are identified, i.e., between pixels in the image plane of the projector and the corresponding points in the scene.
-What remains of this section is concerned primarily with the recovery of these correspondences.
+What remains of this section is concerned primarily with the recovery of 2D-3D correspondences in support of calibrating cameras and projectors.
 A consequence of the approach we take to identifying the 3D points of the 2D-3D correspondences we use for projector calibration is, for each target location, recovery of the pose of the scene plane relative to the coordinate frame of the camera.
@@
 A set of calibration images is acquired, each with the calibration pattern visible in a different part of the image plane, and such that the center and all corners and edges of the image plane are covered, the camera's autofocus setting be off, and the camera's zoom factor remain fixed.
-2D-3D correspondences are then recovered for each calibration image, and the resulting list is passed on as input to an optimization procedure that relies on bundle adjustment to yield the camera calibration matrix~$\mathtt{K}_\text{cam}$ and the associated lens distortion model coefficients.
+2D-3D correspondences are then recovered for each calibration image, and the resulting list is passed on as input to an optimization procedure that relies on bundle adjustment to yield the camera calibration matrix~$\mathtt{K}_\text{cam}$.
 
-\paragraph{Projector calibration} As with camera calibration, projector calibration relies on 2D-3D correspondences, yet we obtain them in this case by projecting a pattern of circles.
-We rely on an algorithm to detect the circle pattern center points in the circles pattern image in the image plane of the projector \cite{bradski2000opencv}, giving the 2D positions of our 2D-3D correspondences for calibrating the projector (cf.\ Figure~\ref{fig:2d}(a)).
-We project the circles pattern image to each of the target locations, and use the the downwards-facing camera to acquire a projector calibration image for each.
-Given a projector calibration image acquired using camera, we detect the circle centers of the \textit{projected} circles pattern (cf.\ Figure~\ref{fig:2d}(b)); given the scene plane (the recovery of which we shall return to in the paragraph that follows) and such a 2D circle center $\mathbf{x}$, its 3D correspondence is obtained by intersecting the back-projection of $\mathbf{x}$ with the scene plane (cf.\ Figure~\ref{fig:3d}).
+\paragraph{Projector calibration} As with camera calibration, projector calibration relies on 2D-3D correspondences, yet we obtain them in this case by \textit{projecting} a calibration pattern.
+The pattern we project is one of circles (cf.\ Figure~\ref{fig:2d}(a)), and we rely on an algorithm to detect the circle pattern center points in the circles pattern image in the image plane of the projector \cite{bradski2000opencv}, giving the 2D positions of our 2D-3D correspondences for calibrating the projector.
+We project the circles pattern image to each of the target locations, and use the the calibrated downwards-facing camera to acquire a projector calibration image for each.
+Given a projector calibration image acquired using camera, we detect the circle centers of the \textit{projected} circles pattern (cf.\ Figure~\ref{fig:2d}(b)); given the scene plane (the recovery of which we shall return to in the paragraph that follows) and such a 2D circle center $\mathbf{x}$, we obtain its 3D correspondence by intersecting the back-projection~$\mathtt{K}_\text{cam}^{-1}(\mathbf{x}^\top, 1)^\top \in \mathbb{P}^2$ of $\mathbf{x}$ with the recovered scene plane (cf.\ Figure~\ref{fig:3d}).
 Since the algorithm that yields 2D circle centers does so in a consistent ordering, we thus obtain the 2D-3D correspondences between the projector's image plane and the scene required for projector calibration, yielding the projector calibration matrix~$\mathtt{K}_\text{proj}$.
@@
 
-The pose (i.e., position and orientation) of the projector, for each projector calibration image, is provided alongside $\mathtt{K}_\text{proj}$ by the aforementioned optimization procedure.
+The pose of the projector, for each projector calibration image, is provided relative to the 3D points of the pattern---and thus in the coordinate frame of the camera---alongside $\mathtt{K}_\text{proj}$ by the aforementioned optimization procedure.
 Note that for a fixed projector with steerable mirror, given a projector calibration image, the recovered projector's pose is the pose the projector would have to have had to project to the given target location \textit{in the absence of the mirror}.
-As this is sufficient for our needs in Section~\ref{sec:approach:homography}, it is in this sense that our system is able to handle a projector equipped with a steerable mirror, without need for modeling the steerable mirror explicitly.
+This is sufficient for our needs in Section~\ref{sec:approach:homography}, and it is in this sense that our system is able to handle a projector equipped with a steerable mirror, without need for modeling the steerable mirror explicitly.
 
@@
 If we for a moment imagine that the projector---at its recovered pose---functions as a camera,\footnote{Recall that the calibration matrix~$\mathtt{K}$ enables computing both (i) the projection of a scene point to the image plane (the function of a camera), or (ii) the back-projection of a pixel in the image plane, giving a ray into the scene (along which a projector illuminates the scene with the given pixel).} then (i) projecting an image to the scene plane \textit{from the viewpoint of the virtual projector} and (ii) acquiring the resulting projected image from the viewpoint of the recovered projector gives the desired corrective warp.
-Projecting an image warped in this manner to the scene plane \textit{from the viewpoint of the recovered projector} then has the same effect as projecting the unwarped image to the scene plane from the viewpoint of the virtual projector.
+Projecting an image warped in this manner to the scene plane \textit{from the viewpoint of the recovered projector} then has the same effect as projecting the original image to the scene plane from the viewpoint of the virtual projector.
 This warp can effected using a plane-induced homography, computed analytically as a function of the scene plane, the projector, and the virtual projector.
@@
 a form of `plane-induced' homography \cite{Hartley2004}.
-For convenience, we enable optional rotation of the image to be projected \textit{before} applying the homography, about the image center; that rotation, parameterized in degrees, is thus in effect likewise carried out intuitively relative to the placement of the camera.
+For convenience, we enable optional rotation of the image to be projected \textit{before} applying $\mathtt{H}$, about the image center; that rotation, parameterized in degrees, is thus in effect likewise carried out intuitively relative to the placement of the camera.
 
@@
 
+\paragraph{Baseline} ...
+
 \paragraph{Our approach} We began by carrying out a calibration of the camera, acquiring 10 camera calibration images (cf.\ Section~\ref{sec:approach:geometry}) of a chessboard calibration pattern with $6~\times~4$ corners ($7~\times~5$ squares) and feeding the images as input to our camera calibration module.
@@
 Note that once the camera is calibrated, that calibration can be reused if the camera's intrinsics remain fixed, in particular if no change is made to the zoom factor of the camera.
-
-\paragraph{Baseline} ...
 

### 2021-07-20 paper/elsarticle/elsarticle-template.tex
paper

@@
 We recover the scene plane via spatial resection by applying a PnP algorithm \cite{collins2014infinitesimal} to the 2D-3D correspondences obtained using a chessboard pattern.
-Note that this step is separate from camera calibration, yet could well be carried out using the same calibration pattern used in the camera calibration step.\footnote{The critical point is that the pattern should ideally be coplanar with the local scene plane, meaning its height above the scene plane should not exceed a few millimeters.} While a single image of such a chessboard pattern placed on the floor could be sufficient if the floor is even, we place a chessboard pattern in close proximity to the projected circles pattern in each projector calibration image in order to recover the scene plane locally to each target location, in order to account for the possibility of an uneven floor.
+Note that this step is separate from camera calibration, yet could well be carried out using the same calibration pattern used in the camera calibration step.\footnote{The critical point is that the pattern should ideally be coplanar with the local scene plane, meaning its height above the scene plane should not exceed a few millimeters.} While a single image of such a chessboard pattern placed on the floor could be sufficient if the floor is even, we place a chessboard pattern in close proximity to the projected circles pattern in each projector calibration image in order to recover the scene plane locally to each target location, in order to account for the possibility of an uneven floor (cf.\ Figure~\ref{fig:2d}(b)).
 Note that in principle, we could project a chessboard pattern instead of a circles pattern to obtain the 2D-3D correspondences needed for projector calibration; it is, however, because we rely on detecting the corners of a chessboard pattern to recover the scene plane that we opt instead for a alternative pattern.

### 2021-07-22 paper/elsarticle/elsarticle-template.tex
paper

@@
 \begin{keyword}
-Spatial augmented reality (SAR) \sep Industry 4.0 \sep Pilotfabrik
+Spatial augmented reality (SAR) \sep computer vision \sep projector calibration \sep Industry 4.0 \sep Pilotfabrik
 \end{keyword}
@@
 To model this interaction, we (i) carry out a one-time projector calibration, which in our approach calls for additionally calibrating a camera facing downward to the scene plane and includes recovery of the scene plane as a convenient side effect.
-Next, we use the relative camera-projector-scene plane geometry to (ii) compute a plane-induced homography that distorts the image to be projected in a manner that it appear undistorted to the viewer, and placed in alignment with the axes of a proxy image of the scene.
+Next, we use the relative camera-projector-scene plane geometry to (ii) compute a plane-induced homography that warps the image to be projected in a manner that it appear undistorted to the viewer, and placed in alignment with the axes of a proxy image of the scene.
 These two points are treated in Sections~\ref{sec:approach:geometry} and \ref{sec:approach:homography}, respectively.
@@
 To calibrate a camera, images of a calibration surface such as a chessboard pattern are used to identify the 2D-3D correspondences $\{\mathbf{x}_{i,j} \leftrightarrow \mathbf{X}_{i,j}\}$, $i \in \{ 1, \dots, n_\text{pt} \}, j \in \{ 1, \dots, n_\text{im} \}$, where $n_\text{pt}$ gives the number of correspondences obtained from one calibration image of the pattern, $n_\text{im}$ the number of such images, and $\mathbf{X}_{i,j} = \mathbf{X}_{i,j'}$, for $j, j' \in \{ 1, \dots, n_\text{im} \}$.
-Bundle adjustment is used to obtain the lens distortion model coefficients, calibration matrix~$\mathtt{K}$, and rigid body transformations $g = (\mathtt{R}_j, \mathbf{t}_j) \in SE(3)$ that minimize the reprojection error
+Bundle adjustment is used to obtain the lens distortion model coefficients, calibration matrix~$\mathtt{K}$, and rigid body transformations $(\mathtt{R}_j, \mathbf{t}_j) \in SE(3)$ that minimize the reprojection error
 \begin{equation}
@@
 where $\mathbf{X}_{i,j}' = \mathtt{R}_j\mathbf{X}_{i,j} + \mathbf{t}_j$ expresses $\mathbf{X}_{i,j}$ in the camera coordinate frame corresponding to the $j^{\text{th}}$ calibration image, $\mathtt{K}\mathbf{X}_{i,j}' \in \mathbb{R}^2$ gives the projection of $\mathbf{X}_{i,j}$ to that image, and $d$ is a distance function that corrects for lens distortion using the recovered lens distortion model coefficients and computes a distance with respect to the two undistorted pixel positions.
-The inverse rigid body transformation~$g^{-1}$ gives the pose of the $j^{\text{th}}$ camera relative to the static 3D points of the calibration pattern.
+The inverse rigid body transformation~$(\mathtt{R}_j^{-1}, -\mathtt{R}_j^{-1}\mathbf{t}_j^{})$ gives the pose of the $j^{\text{th}}$ camera relative to the static 3D points of the calibration pattern.
 
@@
 
-If the projector is calibrated and its pose relative to the scene plane is known, a `virtual' projector (with the same calibration $\mathtt{K}$ and lens distortion coefficients) can be placed elsewhere relative to the scene plane.
+If the projector is calibrated and its pose relative to the scene plane is known, a `virtual' projector (with the same calibration $\mathtt{K}$ and lens distortion model coefficients) can be placed elsewhere relative to the scene plane.
 If we for a moment imagine that the projector---at its recovered pose---functions as a camera,\footnote{Recall that the calibration matrix~$\mathtt{K}$ enables computing both (i) the projection of a scene point to the image plane (the function of a camera), or (ii) the back-projection of a pixel in the image plane, giving a ray into the scene (along which a projector illuminates the scene with the given pixel).} then (i) projecting an image to the scene plane \textit{from the viewpoint of the virtual projector} and (ii) acquiring the resulting projected image from the viewpoint of the recovered projector gives the desired corrective warp.
@@
 (a) Projecting the original image to the scene plane.
-(b) After warping the original image according to our plane-induced homography for the given target location, the image is projected in a manner that appears free of perspective distortions, aligned with the axes of the virtual camera (via the virtual projector), and to have the desired dimensions in the scene plane, expressed in metric units.
+(b) After warping the original image according to our plane-induced homography for the given target location, the image is projected in a manner that appears free of projective distortions, aligned with the axes of the virtual camera (via the virtual projector), and to have the desired dimensions in the scene plane, expressed in metric units.
 Note that background in the projected image is shown set to black.}
@@
 Moreover, let $(\mathbf{n}^\top, -d)^\top$ give the scene plane, expressed in the coordinate frame of the recovered projector, where $\mathbf{n} \in \mathbb{R}^3$ is the scene plane's normal vector and $d = \mathbf{n}^\top\mathbf{X}$ for any point~$\mathbf{X}$ in the plane, so that $(\mathbf{n}^\top, -d) (\mathbf{X}^\top, 1)^\top = 0$.
-The transformation that distorts the image to be projected to the scene plane by the recovered projector such that it appear as if were projected to the scene plane by the virtual projector (cf.\ Figure~\ref{fig:warp}(b)) is given the by the $3~\times~3$ matrix
+The transformation that warps the image to be projected to the scene plane by the recovered projector such that it appear as if were projected to the scene plane by the virtual projector (cf.\ Figure~\ref{fig:warp}(b)) is given the by the $3~\times~3$ matrix
 \begin{equation}
@@
 
-\paragraph{Baseline} ...
+\paragraph{Manual approach} ...
 
@@
 A natural extension of this work would be to address non-planar scenes.
-To handle non-planar scenes would call for a change in how scene geometry is recovered and how distortion of the image to be projected is carried out; the methodology we proposed for projector calibration could, however, be left unchanged.
+To handle non-planar scenes would call for a change in how scene geometry is recovered and how warping of the image to be projected is carried out; the methodology we proposed for projector calibration could, however, be left unchanged.
 

### 2021-07-22 src/cpslib/libCalib/CalibPattern.cpp

@@ -34,6 +34,10 @@ void CalibPattern::findChessboardImPts(cv::Mat& im, cv::Size chessboardPatternSi
     cv::cvtColor(imUndistort, imGrayUndistort, cv::COLOR_BGR2GRAY);
     // cv::threshold(candidateIm, candidateImThresh, 100, 255, THRESH_BINARY);
 
+    //cv::Ptr<cv::CLAHE> clahe = cv::createCLAHE();
+    //clahe->setClipLimit(4);
+    //clahe->apply(imGrayUndistort, imGrayUndistort);
+
     outImSize = cv::Size(im.cols, im.rows);
 
     std::vector<cv::Point2f> candidateChessboardImPts;
@@ -78,6 +82,8 @@ void CalibPattern::findCirclesImPts(cv::Mat& im, cv::Size circlesPatternSize,
 {
     cv::Mat imUndistort, imGrayUndistort, imGrayUndistortInvert;
 
+    std::cout << 1 << std::endl;
+
     if (applyIntrinsics)
     {
         cv::undistort(im, imUndistort, K, distCoeffs);
@@ -85,13 +91,33 @@ void CalibPattern::findCirclesImPts(cv::Mat& im, cv::Size circlesPatternSize,
     else
         im.copyTo(imUndistort);
 
+    std::cout << 2 << std::endl;
+
     if (outImVis.size != imUndistort.size)
         imUndistort.copyTo(outImVis);
 
+    std::cout << 3 << std::endl;
+
     cv::cvtColor(imUndistort, imGrayUndistort, cv::COLOR_BGR2GRAY);
+
+    std::cout << 4 << std::endl;
+
+    //cv::Ptr<cv::CLAHE> clahe = cv::createCLAHE();
+    //clahe->setClipLimit(4);
+    //clahe->apply(imGrayUndistort, imGrayUndistort);
+
     cv::bitwise_not(imGrayUndistort, imGrayUndistortInvert);
 
+    std::cout << 5 << std::endl;
+
     outImSize = cv::Size(im.cols, im.rows);
+
+    std::cout << 6 << std::endl;
+
+    cv::Mat imGrayUndistortInvertVis;
+    cv::resize(imGrayUndistortInvert, imGrayUndistortInvertVis, cv::Size(imGrayUndistortInvert.cols * 0.5, imGrayUndistortInvert.rows * 0.5));
+    cv::imshow("imGrayUndistortInvert", imGrayUndistortInvertVis);
+    cv::waitKey(10);
 
[…]

### 2021-07-23 paper/elsarticle/elsarticle-template.tex
Update elsarticle-template.tex

@@
 where $\mathbf{X}_{i,j}' = \mathtt{R}_j\mathbf{X}_{i,j} + \mathbf{t}_j$ expresses $\mathbf{X}_{i,j}$ in the camera coordinate frame corresponding to the $j^{\text{th}}$ calibration image, $\mathtt{K}\mathbf{X}_{i,j}' \in \mathbb{R}^2$ gives the projection of $\mathbf{X}_{i,j}$ to that image, and $d$ is a distance function that corrects for lens distortion using the recovered lens distortion model coefficients and computes a distance with respect to the two undistorted pixel positions.
-The inverse rigid body transformation~$(\mathtt{R}_j^{-1}, -\mathtt{R}_j^{-1}\mathbf{t}_j^{})$ gives the pose of the $j^{\text{th}}$ camera relative to the static 3D points of the calibration pattern.
+The inverse rigid body transformation~$(\mathtt{R}_j^{-1}, -\mathtt{R}_j^{-1}\mathbf{t}_j^{})$ gives the pose of the $j^{\text{th}}$ camera relative to the 3D points of the calibration pattern.
 

### 2021-07-24 paper/elsarticle/elsarticle-template.tex
paper

@@
 \author{Patrick Rupprecht}
-\author{Sebastian~Schlund}
+\author{Sebastian Schlund}
 \address{Human Centered Cyber Physical Production and Assembly Systems, Institute for Management Sciences, TU Wien, Vienna, Austria}
@@
 One way to compute a homography is by identifying at least four correspondences between pixel positions in two images of a planar surface.
-The keystone correction approach of Sukthankar \etal \cite{sukthankar2001smarter} reduces to using a segmentation approach to identify the four corners of projection screen in an image acquired by a camera, and computing a homography that maps the resulting four corners of the projection screen to the four corners of the projector's image plane.
+The keystone correction approach of Sukthankar \textit{et al.} \cite{sukthankar2001smarter} reduces to using a segmentation approach to identify the four corners of projection screen in an image acquired by a camera, and computing a homography that maps the resulting four corners of the projection screen to the four corners of the projector's image plane.
 Tilt sensors can be used to recover the projector's gravity vector, and be integrated in the computation of a homography as in Raskar and Beardsley \cite{raskar2001self}.
@@
 Calibration of a projector (or camera) in the sense we employ the term here\footnote{We are referring to a geometric calibration; not, e.g., to a color calibration.} renders one able to project a scene point~$\mathbf{X} \in \mathbb{R}^3$ to its corresponding pixel~$\mathbf{x} \in \mathbb{R}^2$ in the projector's (or camera's) image plane, or to compute the `back-projection' of $\mathbf{x}$, i.e., the ray from the projector's (or camera's) center of projection through $\mathbf{x}$ along which $\mathbf{X}$ must lie.
-Such a calibration can be expressed in terms of (i) a $3~\times{}~3$ calibration matrix~$\mathtt{K}$ derived from the projector's (or camera's) focal length and principal point \cite{Hartley2004}, and (ii) the coefficients of a lens distortion model used to correct for radial or tangential distortions caused by the lens system \cite{duane1971close}.
+Such a calibration can be expressed in terms of (i) a $3~\times{}~3$ calibration matrix~$\mathtt{K}$ derived from the projector's (or camera's) focal length~$f$ and principal point~$\mathbf{p}_0 = (x_0, y_0)^\top \in \mathbb{R}^2$ \cite{Hartley2004}, and (ii) the coefficients of a lens distortion model used to correct for radial or tangential distortions caused by the lens system \cite{duane1971close}; the calibration matrix~$\mathtt{K}$ is then
+\begin{equation}
+\mathtt{K} = \begin{bmatrix}
+f & 0 & x_0 \\
+0 & f & y_0 \\
+0 & 0 & 1
+\end{bmatrix},
+\end{equation}
+where $f$ and $(p_x, p_y)^\top$ are both expressed in units of pixels.
 
 In practice, calibrating a camera relies on (i) establishing 2D-3D correspondences between pixels in the camera's image plane and corresponding points in the scene, and on (ii) using those correspondences as input to an optimization procedure that relies on bundle adjustment \cite{triggs1999bundle} to output the calibration matrix~$\mathtt{K}$, the associated lens distortion model coefficients, and, for each calibration image, the pose (i.e., position and orientation) of the camera relative to the 3D points \cite{Hartley2004,zhang2000flexible}.
-To calibrate a camera, images of a calibration surface such as a chessboard pattern are used to identify the 2D-3D correspondences $\{\mathbf{x}_{i,j} \leftrightarrow \mathbf{X}_{i,j}\}$, $i \in \{ 1, \dots, n_\text{pt} \}, j \in \{ 1, \dots, n_\text{im} \}$, where $n_\text{pt}$ gives the number of correspondences obtained from one calibration image of the pattern, $n_\text{im}$ the number of such images, and $\mathbf{X}_{i,j} = \mathbf{X}_{i,j'}$, for $j, j' \in \{ 1, \dots, n_\text{im} \}$.
+To calibrate a camera, images of a calibration surface such as a chessboard pattern are used to identify the 2D-3D correspondences~$\{\mathbf{x}_{i,j} \leftrightarrow \mathbf{X}_{i,j}\}$, $i \in \{ 1, \dots, n_\text{pt} \}, j \in \{ 1, \dots, n_\text{im} \}$, where $n_\text{pt}$ gives the number of correspondences obtained from one calibration image of the pattern, $n_\text{im}$ the number of such images, and $\mathbf{X}_{i,j} = \mathbf{X}_{i,j'}$, for $j, j' \in \{ 1, \dots, n_\text{im} \}$.
 Bundle adjustment is used to obtain the lens distortion model coefficients, calibration matrix~$\mathtt{K}$, and rigid body transformations $(\mathtt{R}_j, \mathbf{t}_j) \in SE(3)$ that minimize the reprojection error
@@
 \end{equation}
-where $\mathbf{X}_{i,j}' = \mathtt{R}_j\mathbf{X}_{i,j} + \mathbf{t}_j$ expresses $\mathbf{X}_{i,j}$ in the camera coordinate frame corresponding to the $j^{\text{th}}$ calibration image, $\mathtt{K}\mathbf{X}_{i,j}' \in \mathbb{R}^2$ gives the projection of $\mathbf{X}_{i,j}$ to that image, and $d$ is a distance function that corrects for lens distortion using the recovered lens distortion model coefficients and computes a distance with respect to the two undistorted pixel positions.
-The inverse rigid body transformation~$(\mathtt{R}_j^{-1}, -\mathtt{R}_j^{-1}\mathbf{t}_j^{})$ gives the pose of the $j^{\text{th}}$ camera relative to the 3D points of the calibration pattern.
+where $\mathbf{X}_{i,j}' = \mathtt{R}_j\mathbf{X}_{i,j} + \mathbf{t}_j \in \mathbb{R}^3$ expresses $\mathbf{X}_{i,j}$ in the camera coordinate frame of the camera corresponding to the $j^{\text{th}}$ calibration image, $(\mathbf{x}_{i,j}', 1)^\top \sim \mathtt{K}\mathbf{X}_{i,j}' \in \mathbb{P}^2$ gives the projection~$\mathbf{x}'_{i,j} \in \mathbb{R}^2$ of $\mathbf{X}_{i,j}'$ to that image, and $d$ is a distance function that~(i) corrects for lens distortion using the recovered lens distortion model coefficients~\cite{duane1971close} and (ii) returns a distance with respect to the two undistorted pixel positions.
+The inverse rigid body transformation~$(\mathtt{R}_j^{-1}, -\mathtt{R}_j^{-1}\mathbf{t}_j^{})$ gives the pose of the $j^{\text{th}}$~camera relative to the 3D points of the calibration pattern.
 

### 2021-07-24 paper/elsarticle/elsarticle-template.tex
paper

@@
 
-Calibration of a projector (or camera) in the sense we employ the term here\footnote{We are referring to a geometric calibration; not, e.g., to a color calibration.} renders one able to project a scene point~$\mathbf{X} \in \mathbb{R}^3$ to its corresponding pixel~$\mathbf{x} \in \mathbb{R}^2$ in the projector's (or camera's) image plane, or to compute the `back-projection' of $\mathbf{x}$, i.e., the ray from the projector's (or camera's) center of projection through $\mathbf{x}$ along which $\mathbf{X}$ must lie.
-Such a calibration can be expressed in terms of (i) a $3~\times{}~3$ calibration matrix~$\mathtt{K}$ derived from the projector's (or camera's) focal length~$f$ and principal point~$\mathbf{p}_0 = (x_0, y_0)^\top \in \mathbb{R}^2$ \cite{Hartley2004}, and (ii) the coefficients of a lens distortion model used to correct for radial or tangential distortions caused by the lens system \cite{duane1971close}; the calibration matrix~$\mathtt{K}$ is then
+Calibration of a projector (or camera) in the sense we employ the term here\footnote{We are referring to a geometric calibration; not, e.g., to a color calibration.} renders one able to project a scene point~$\mathbf{X} \in \mathbb{R}^3$ to its corresponding pixel~$\mathbf{x} \in \mathbb{R}^2$ in the projector's (or camera's) image plane, or to compute the `back-projection' of $\mathbf{x}$, i.e., the ray from the projector's (or camera's) center of projection through $\mathbf{x}$ along which the projecting point~$\mathbf{X}$ must lie.
+Such a calibration can be expressed in terms of (i) a $3~\times{}~3$ calibration matrix~$\mathtt{K}$ derived from the projector's (or camera's) focal length~$f$ and principal point~$\mathbf{p}_0 = (x_0, y_0)^\top \in \mathbb{R}^2$ \cite{Hartley2004}, and (ii)~the coefficients of a lens distortion model used to correct for radial or tangential distortions caused by the lens system \cite{duane1971close,weng1992camera}; the calibration matrix~$\mathtt{K}$ is then
 \begin{equation}
@@
 \end{equation}
-where $\mathbf{X}_{i,j}' = \mathtt{R}_j\mathbf{X}_{i,j} + \mathbf{t}_j \in \mathbb{R}^3$ expresses $\mathbf{X}_{i,j}$ in the camera coordinate frame of the camera corresponding to the $j^{\text{th}}$ calibration image, $(\mathbf{x}_{i,j}', 1)^\top \sim \mathtt{K}\mathbf{X}_{i,j}' \in \mathbb{P}^2$ gives the projection~$\mathbf{x}'_{i,j} \in \mathbb{R}^2$ of $\mathbf{X}_{i,j}'$ to that image, and $d$ is a distance function that~(i) corrects for lens distortion using the recovered lens distortion model coefficients~\cite{duane1971close} and (ii) returns a distance with respect to the two undistorted pixel positions.
+where $\mathbf{X}_{i,j}' = \mathtt{R}_j\mathbf{X}_{i,j} + \mathbf{t}_j \in \mathbb{R}^3$ expresses $\mathbf{X}_{i,j}$ in the camera coordinate frame of the camera corresponding to the $j^{\text{th}}$ calibration image, $(\mathbf{x}_{i,j}', 1)^\top \sim \mathtt{K}\mathbf{X}_{i,j}' \in \mathbb{P}^2$ gives the projection~$\mathbf{x}'_{i,j} \in \mathbb{R}^2$ of $\mathbf{X}_{i,j}'$ to that image, and $d$ is a distance function that~(i) corrects for lens distortion using the lens distortion model coefficients~\cite{duane1971close,weng1992camera} and (ii) returns a distance with respect to the two undistorted pixel positions.
 The inverse rigid body transformation~$(\mathtt{R}_j^{-1}, -\mathtt{R}_j^{-1}\mathbf{t}_j^{})$ gives the pose of the $j^{\text{th}}$~camera relative to the 3D points of the calibration pattern.
@@
 The 3D corner points of the chessboard are obtained \textit{a priori} in a coordinate system defined in the plane of the chessboard\footnote{E.g., $(0,0,0), (1.5,0,0), (3,0,0), \dots, (9,7.5,0)$ for a chessboard with $7~\times{}~6$ corners ($8~\times{}~7$ squares), with each square of length and width of 1.5 unit, respectively.
-Note that the units of the chessboard's 3D points give the units of the camera calibration, and---owing to how our projector calibration relies on the camera calibration---of the projector calibration as well.}, requiring knowledge only of the dimensions of the chessboard pattern and of the length of a side of a chessboard square.
+Note that the units of the chessboard's 3D points give the units in terms of which the camera calibration is carried out, and---owing to how our projector calibration relies on the camera calibration---the units of the projector calibration as well.}, requiring knowledge only of the dimensions of the chessboard pattern and of the length of a side of a chessboard square.
 The corresponding 2D points are obtained, in the same order, using a specialized algorithm \cite{bradski2000opencv}.

### 2021-07-24 paper/elsarticle/elsarticle-template.tex
paper

@@
 
-Mention circles pattern: \cite{chiu2011novel}
+Mention asymmetrical circle pattern: \cite{chiu2011novel}
 
@@
     \centering
-    \subfloat[\centering Circles pattern image, in image plane of projector (detections overlain).]{{\includegraphics[height=3.25cm]{images/circles.png} }}
+    \subfloat[\centering Asymetrical circle pattern image, in image plane of projector (detections overlain).]{{\includegraphics[height=3.25cm]{images/circles.png} }}
     \qquad
@@
     \caption{Recovering 2D positions in support of projector calibration.
-(a) 2D positions of the 2D-3D correspondences to be used for calibrating the projector are obtained by detecting---in the image plane of the projector---the circle centers in the circles pattern image, projected by the projector to each of the target locations in the scene plane.
-(b) For each such target location, an image is acquired from the viewpoint of the camera and the circle centers of the projected circles pattern are detected, in the image plane of the camera.
+(a) 2D positions of the 2D-3D correspondences to be used for calibrating the projector are obtained by detecting---in the image plane of the projector---the circle centers in the asymmetrical circle pattern image, to be projected to each of the target locations in the scene plane.
+(b) For each such target location, an image is acquired from the viewpoint of the camera and the circle centers of the projected asymmetrical circle pattern are detected, in the image plane of the camera.
 A chessboard pattern to be used for recovering the local scene plane is placed near the projected pattern, whose corners are likewise detected.
-Detected 2D projected circles pattern center points and chessboard corners overlain for illustration.}
+Detected 2D projected asymmetrical circle pattern circle center points and chessboard corners overlain for illustration.}
     \label{fig:2d}
@@
 \paragraph{Projector calibration} As with camera calibration, projector calibration relies on 2D-3D correspondences, yet we obtain them in this case by \textit{projecting} a calibration pattern.
-The pattern we project is one of circles (cf.\ Figure~\ref{fig:2d}(a)), and we rely on an algorithm to detect the circle pattern center points in the circles pattern image in the image plane of the projector \cite{bradski2000opencv}, giving the 2D positions of our 2D-3D correspondences for calibrating the projector.
-We project the circles pattern image to each of the target locations, and use the the calibrated downwards-facing camera to acquire a projector calibration image for each.
-Given a projector calibration image acquired using camera, we detect the circle centers of the \textit{projected} circles pattern (cf.\ Figure~\ref{fig:2d}(b)); given the scene plane (the recovery of which we shall return to in the paragraph that follows) and such a 2D circle center $\mathbf{x}$, we obtain its 3D correspondence by intersecting the back-projection~$\mathtt{K}_\text{cam}^{-1}(\mathbf{x}^\top, 1)^\top \in \mathbb{P}^2$ of $\mathbf{x}$ with the recovered scene plane (cf.\ Figure~\ref{fig:3d}).
+The pattern we project is one of circles (cf.\ Figure~\ref{fig:2d}(a)), and we rely on an algorithm to detect the circle center points in the asymmetrical circle pattern in the image plane of the projector \cite{bradski2000opencv}, giving the 2D positions of our 2D-3D correspondences for calibrating the projector.
+We project the asymmetrical circle pattern to each of the target locations, and use the the calibrated downwards-facing camera to acquire a projector calibration image for each.
+Given a projector calibration image acquired using camera, we detect the circle centers of the \textit{projected} asymmetrical circle pattern (cf.\ Figure~\ref{fig:2d}(b)); given the scene plane (the recovery of which we shall return to in the paragraph that follows) and such a 2D circle center $\mathbf{x}$, we obtain its 3D correspondence by intersecting the back-projection~$\mathtt{K}_\text{cam}^{-1}(\mathbf{x}^\top, 1)^\top \in \mathbb{P}^2$ of $\mathbf{x}$ with the recovered scene plane (cf.\ Figure~\ref{fig:3d}).
 Since the algorithm that yields 2D circle centers does so in a consistent ordering, we thus obtain the 2D-3D correspondences between the projector's image plane and the scene required for projector calibration, yielding the projector calibration matrix~$\mathtt{K}_\text{proj}$.
@@
     \centerline{\includegraphics[scale=.35]{images/2d3d.png}}
-    \caption{Scene plane (gray) recovered via spatial resection with respect to 2D-3D correspondences obtained using a chessboard pattern (blue); 3D circles pattern points---i.e., the 3D positions of the
-2D-3D correspondences to be used for calibrating the projector---obtained by intersection with the scene plane of back-projections (likewise gray) of the 2D circles pattern center points detected in the image plane (red).
+    \caption{Scene plane (gray) recovered via spatial resection with respect to 2D-3D correspondences obtained using a chessboard pattern (blue); 3D circle center points of the asymmetrical circle pattern---i.e., the 3D positions of the 2D-3D correspondences to be used for calibrating the projector---obtained by intersection with the scene plane of back-projections (likewise gray) of the 2D circle center points of the asymmetrical circle pattern detected in the image plane (red).
 Note that as in the figures that follow, the rendering in the figure corresponds to the projection calibration image in Figure~\ref{fig:2d}(b), acquired by the downwards-facing camera (frustum of the camera in black, with up vector in red).}
@@
 We recover the scene plane via spatial resection by applying a PnP algorithm \cite{collins2014infinitesimal} to the 2D-3D correspondences obtained using a chessboard pattern.
-Note that this step is separate from camera calibration, yet could well be carried out using the same calibration pattern used in the camera calibration step.\footnote{The critical point is that the pattern should ideally be coplanar with the local scene plane, meaning its height above the scene plane should not exceed a few millimeters.} While a single image of such a chessboard pattern placed on the floor could be sufficient if the floor is even, we place a chessboard pattern in close proximity to the projected circles pattern in each projector calibration image in order to recover the scene plane locally to each target location, in order to account for the possibility of an uneven floor (cf.\ Figure~\ref{fig:2d}(b)).
-Note that in principle, we could project a chessboard pattern instead of a circles pattern to obtain the 2D-3D correspondences needed for projector calibration; it is, however, because we rely on detecting the corners of a chessboard pattern to recover the scene plane that we opt instead for a alternative pattern.
+Note that this step is separate from camera calibration, yet could well be carried out using the same calibration pattern used in the camera calibration step.\footnote{The critical point is that the pattern should ideally be coplanar with the local scene plane, meaning its height above the scene plane should not exceed a few millimeters.} While a single image of such a chessboard pattern placed on the floor could be sufficient if the floor is even, we place a chessboard pattern in close proximity to the projected asymmetrical circle pattern in each projector calibration image in order to recover the scene plane locally to each target location, in order to account for the possibility of an uneven floor (cf.\ Figure~\ref{fig:2d}(b)).
+Note that in principle, we could project a chessboard pattern instead of an asymmetrical circle pattern to obtain the 2D-3D correspondences needed for projector calibration; it is, however, because we rely on detecting the corners of a chessboard pattern to recover the scene plane that we opt instead for a alternative pattern.
 
@@
 \paragraph{Our approach} We began by carrying out a calibration of the camera, acquiring 10 camera calibration images (cf.\ Section~\ref{sec:approach:geometry}) of a chessboard calibration pattern with $6~\times~4$ corners ($7~\times~5$ squares) and feeding the images as input to our camera calibration module.
-Separately, for each of the 15 target locations, we produced a projector calibration image (cf.\ again Section~\ref{sec:approach:geometry}) by projecting a $11~\times~4$ circles pattern image to the location in question using the steerable mirror, placing a checkerboard pattern beside the projected pattern, and acquiring the image using the downward-facing camera.
+Separately, for each of the 15 target locations, we produced a projector calibration image (cf.\ again Section~\ref{sec:approach:geometry}) by projecting a $11~\times~4$ asymmetrical circle pattern image to the location in question using the steerable mirror, placing a checkerboard pattern beside the projected pattern, and acquiring the image using the downward-facing camera.
 We then fed these images alongside the output of the camera calibration module to our projector calibration module.
-For each of the target locations, the steerable mirror was made to point to that location, the circles pattern image was projected to the scene plane, an image was using the camera, and the location was registered in the MDC-X software as a preset.
+For each of the target locations, the steerable mirror was made to point to that location, the asymmetrical circle pattern image was projected to the scene plane, an image was using the camera, and the location was registered in the MDC-X software as a preset.
 The output of the projector calibration module is a homography per input projector calibration image (cf.\ Section~\ref{sec:approach:homography}).

### 2021-07-24 paper/elsarticle/elsarticle-template.tex
paper

@@
 
-In practice, calibrating a camera relies on (i) establishing 2D-3D correspondences between pixels in the camera's image plane and corresponding points in the scene, and on (ii) using those correspondences as input to an optimization procedure that relies on bundle adjustment \cite{triggs1999bundle} to output the calibration matrix~$\mathtt{K}$, the associated lens distortion model coefficients, and, for each calibration image, the pose (i.e., position and orientation) of the camera relative to the 3D points \cite{Hartley2004,zhang2000flexible}.
+Calibrating a camera can be carried out by relying on (i) establishing 2D-3D correspondences between pixels in the camera's image plane and corresponding points in the scene, and on (ii) using those correspondences as input to an optimization procedure that relies on bundle adjustment \cite{triggs1999bundle} to output the calibration matrix~$\mathtt{K}$, the associated lens distortion model coefficients, and, for each calibration image, the pose (i.e., position and orientation) of the camera relative to the 3D points \cite{Hartley2004,zhang2000flexible}.
 To calibrate a camera, images of a calibration surface such as a chessboard pattern are used to identify the 2D-3D correspondences~$\{\mathbf{x}_{i,j} \leftrightarrow \mathbf{X}_{i,j}\}$, $i \in \{ 1, \dots, n_\text{pt} \}, j \in \{ 1, \dots, n_\text{im} \}$, where $n_\text{pt}$ gives the number of correspondences obtained from one calibration image of the pattern, $n_\text{im}$ the number of such images, and $\mathbf{X}_{i,j} = \mathbf{X}_{i,j'}$, for $j, j' \in \{ 1, \dots, n_\text{im} \}$.
-Bundle adjustment is used to obtain the lens distortion model coefficients, calibration matrix~$\mathtt{K}$, and rigid body transformations $(\mathtt{R}_j, \mathbf{t}_j) \in SE(3)$ that minimize the reprojection error
+Bundle adjustment is used to obtain the maximum likelihood estimate of the lens distortion model coefficients~$\omega$, calibration matrix~$\mathtt{K}$, and rigid body transformations $(\mathtt{R}_j, \mathbf{t}_j) \in SE(3)$ that minimize the cumulative reprojection error
 \begin{equation}
-\sum_{i,j} d\big(\mathbf{x}_{i,j}, \mathtt{K}(\mathtt{R}_j\mathbf{X}_{i,j} + \mathbf{t}_j)\big),
+\sum_{i=0}^{n_\text{pt}} \sum_{j=0}^{n_\text{im}} d_\omega\big(\mathbf{x}_{i,j}, \mathtt{K}(\mathtt{R}_j\mathbf{X}_{i,j} + \mathbf{t}_j)\big),
 \label{re}
 \end{equation}
-where $\mathbf{X}_{i,j}' = \mathtt{R}_j\mathbf{X}_{i,j} + \mathbf{t}_j \in \mathbb{R}^3$ expresses $\mathbf{X}_{i,j}$ in the camera coordinate frame of the camera corresponding to the $j^{\text{th}}$ calibration image, $(\mathbf{x}_{i,j}', 1)^\top \sim \mathtt{K}\mathbf{X}_{i,j}' \in \mathbb{P}^2$ gives the projection~$\mathbf{x}'_{i,j} \in \mathbb{R}^2$ of $\mathbf{X}_{i,j}'$ to that image, and $d$ is a distance function that~(i) corrects for lens distortion using the lens distortion model coefficients~\cite{duane1971close,weng1992camera} and (ii) returns a distance with respect to the two undistorted pixel positions.
+where $\mathbf{X}_{i,j}' = \mathtt{R}_j\mathbf{X}_{i,j} + \mathbf{t}_j \in \mathbb{R}^3$ expresses $\mathbf{X}_{i,j}$ in the camera coordinate frame of the camera corresponding to the $j^{\text{th}}$ calibration image, $(\mathbf{x}_{i,j}', 1)^\top \sim \mathtt{K}\mathbf{X}_{i,j}' \in \mathbb{P}^2$ gives the projection~$\mathbf{x}'_{i,j} \in \mathbb{R}^2$ of $\mathbf{X}_{i,j}'$ to that image, and $d_\omega$ is a distance function that~(i) corrects for lens distortion using the lens distortion model coefficients~$\omega$ \cite{duane1971close,weng1992camera} and (ii) returns a distance with respect to the two undistorted pixel positions.
 The inverse rigid body transformation~$(\mathtt{R}_j^{-1}, -\mathtt{R}_j^{-1}\mathbf{t}_j^{})$ gives the pose of the $j^{\text{th}}$~camera relative to the 3D points of the calibration pattern.

### 2021-07-24 paper/elsarticle/elsarticle-template.tex
paper

@@
 First, (i) placement of the desired augmentations is carried out in accordance with the axes of an image of the scene acquired by a camera placed to face downwards towards the scene plane; second, (ii) the desired dimensions of the projected augmentations is specified in metric terms.
-Our system produces such augmentations accurately, with compelling time savings relative to the manual approach.
+Our system produces such augmentations accurately, with compelling time savings relative manual keystone correction.
 \end{abstract}
@@
 In turn, \textit{spatial} AR\cite{bimber2019spatial} is a form of augmented reality carried out not by embedding the augmentation in an image of the scene as with a head-mounted display, but by projection to the scene itself, thus eliminating both aforementioned problems.
-Yet considering a planar surface to be augmented (e.g., a floor, wall, or table), unless the projector faces the surface frontally, the bounds of a projected rectangular image will not appear rectangular, but will instead appear distorted.
+Yet considering a planar surface to be augmented (e.g., a floor, wall, or table), unless the projector faces the surface frontally, the bounds of a projected rectangular image will not appear rectangular, but will instead be subject to projective distortions.
 Such distortions can be eliminated by carrying out a cumbersome manual process called keystone correction to appropriately warp the image to be projected, often using software bundled with the projector.
 
-Our contribution is to propose a wide-area spatial AR system for planar scenes that produces the effect of keystone correction analytically, and---using the $X$- and $Y$-axes of an image of the scene acquired by a downwards-facing camera as a proxy---in a manner aligning the axes of the augmentation with those of the proxy image.
-Moreover, our system enables specifying the dimensions of augmentations in metric terms.
-We achieve this by warping the image to be projected using a plane-induced homography computed to produce the effect of projecting the image not from the actual projector viewpoint, but in accordance with the viewpoint of a \textit{virtual} projector (i) facing directly downwards to the scene plane and (ii) rotated to place the axes of the image plane of the virtual projector in line with those of the camera.
-This facilitates placement of augmentations by intuitively placing them in accordance with the axes of an image of the scene, and eliminates the need for manual keystone correction.
-Finally, our system is able to handle a projector equipped with a steerable mirror (without need for explicitly modeling the action of the steerable mirror on the projector), thereby enabling wide-area factory floor applications exceeding the immediate field of view of the projector without needing to rely on multiple projectors.
+Our contribution is to propose a wide-area spatial AR system for planar scenes that produces the effect of keystone correction analytically.
+Most importantly, we do this in a manner placing the axes of the augmentation in accordance with the axes of a camera placed to face downwards towards the scene plane.
+This facilitates placement of augmentations, since establishing the principal axes of the augmentation thus reduces to placing the downwards-facing camera in a manner such that the $X$- and $Y$-axes of the image align with the intended $X$- and $Y$-axes of the augmentation.
+We achieve this by warping the image to be projected using a plane-induced homography computed to produce the effect of projecting the image not from the actual projector viewpoint, but in accordance with the viewpoint of a \textit{virtual} projector (i) facing directly downwards to the target location in the scene plane and (ii) rotated to place the axes of the virtual projector in accordance with those of the camera.
+Moreover, our system enables specifying the dimensions of augmentations in metric terms, which we achieve by placing the virtual projector at the appropriate height above the scene plane.
+Finally, our system set up to handle a projector equipped with a steerable mirror (without need for explicitly modeling the action of the steerable mirror on the projector), thereby enabling wide-area factory floor applications exceeding the immediate field of view of the projector without needing to rely on multiple projectors.
 
@@
     \centering
-    \subfloat[\centering Asymetrical circle pattern image, in image plane of projector (detections overlain).]{{\includegraphics[height=3.25cm]{images/circles.png} }}
+    \subfloat[\centering Asymmetrical circle pattern image, in image plane of projector (detections overlain).]{{\includegraphics[height=3.25cm]{images/circles.png} }}
     \qquad
@@
 Projecting an image warped in this manner to the scene plane \textit{from the viewpoint of the recovered projector} then has the same effect as projecting the original image to the scene plane from the viewpoint of the virtual projector.
-This warp can effected using a plane-induced homography, computed analytically as a function of the scene plane, the projector, and the virtual projector.
+This warp can be effected using a plane-induced homography, computed analytically as a function of the scene plane, the projector, and the virtual projector.
 
-\paragraph{Virtual projector} The placement of the virtual projector determines from which pose the image to be projected is to \textit{appear} to have been projected.
-This placement is produced according to a small handful of steps.
-First, we (i) rotate the camera about its center of projection to align its optical axis with the normal vector of the scene plane, giving a virtual camera likewise facing directly\footnote{A physical camera placed to face downwards is almost certain to not face downwards precisely; in contrast, the virtual camera's optical axis is aligned exactly with the scene plane's normal vector, rendering it fronto-parallel with respect to the scene plane.} downwards to the scene plane (cf.\ Figure~\ref{fig:virtualproj}).
+\paragraph{Virtual projector} The placement of the virtual projector is to determine from which pose the image to be projected is to \textit{appear} to have been projected.
+We carry out this placement according to a small handful of steps.
+First, we (i) rotate the camera about its center of projection to align its optical axis\footnote{Strictly speaking, the optical axis is the back-projection of the principal point; in our usage, we understand it to refer to the back-projection of the center of the image to be projected.} with the normal vector of the scene plane, giving a virtual camera likewise facing directly\footnote{A physical camera placed to face downwards is almost certain to not face downwards precisely; in contrast, the virtual camera's optical axis is aligned exactly with the scene plane's normal vector, rendering it fronto-parallel with respect to the scene plane.} downwards to the scene plane (cf.\ Figure~\ref{fig:virtualproj}).
 Next, we (ii) intersect the scene plane with the optical axis (i.e., the ray from the projector's center of projection through the center of the image plane) and rotate the projector's placement about that point of intersection, aligning the optical axis with the scene plane's normal vector and giving an initial virtual projector.
@@
 
-Owing to the manner in which we place the virtual projector, the virtual projector's axes and thus the augmentation are aligned with the axes of the downward-facing camera; the placement of the camera thereby intuitively determines the principal axes according to which augmentations are placed.
-Note further that a consequence of placing the virtual projector by rotating about the point of intersection of the projector's optical axis with the scene plane is that the center of the projector's image plane remains invariant to the placement of the virtual projector, i.e., a steerable mirror can be aimed with respect to a point projected from the center of the projector's image plane, further facilitating placement of augmentations.
+Owing to the manner in which we place the virtual projector, the virtual projector's axes and thus the augmentation are aligned with the axes of the downward-facing camera; the placement of the camera thereby intuitively determines the principal axes according to which augmentations are to be placed.
+Note further that a consequence of placing the virtual projector by rotating about the point of intersection of the projector's optical axis with the scene plane is that \textit{the center of the projector's image plane remains invariant} to the placement of the virtual projector, i.e., a steerable mirror can be aimed with respect to a point projected from the center of the projector's image plane, further facilitating placement of augmentations.
 
@@
 (b) After warping the original image according to our plane-induced homography for the given target location, the image is projected in a manner that appears free of projective distortions, aligned with the axes of the virtual camera (via the virtual projector), and to have the desired dimensions in the scene plane, expressed in metric units.
-Note that background in the projected image is shown set to black.}
+Note that unprojected background is shown set to black.}
     \label{fig:warp}

### 2021-07-26 paper/elsarticle/elsarticle-template.tex
paper

@@
 
-Calibrating a camera can be carried out by relying on (i) establishing 2D-3D correspondences between pixels in the camera's image plane and corresponding points in the scene, and on (ii) using those correspondences as input to an optimization procedure that relies on bundle adjustment \cite{triggs1999bundle} to output the calibration matrix~$\mathtt{K}$, the associated lens distortion model coefficients, and, for each calibration image, the pose (i.e., position and orientation) of the camera relative to the 3D points \cite{Hartley2004,zhang2000flexible}.
-To calibrate a camera, images of a calibration surface such as a chessboard pattern are used to identify the 2D-3D correspondences~$\{\mathbf{x}_{i,j} \leftrightarrow \mathbf{X}_{i,j}\}$, $i \in \{ 1, \dots, n_\text{pt} \}, j \in \{ 1, \dots, n_\text{im} \}$, where $n_\text{pt}$ gives the number of correspondences obtained from one calibration image of the pattern, $n_\text{im}$ the number of such images, and $\mathbf{X}_{i,j} = \mathbf{X}_{i,j'}$, for $j, j' \in \{ 1, \dots, n_\text{im} \}$.
-Bundle adjustment is used to obtain the maximum likelihood estimate of the lens distortion model coefficients~$\omega$, calibration matrix~$\mathtt{K}$, and rigid body transformations $(\mathtt{R}_j, \mathbf{t}_j) \in SE(3)$ that minimize the cumulative reprojection error
+Calibrating a camera can be carried out by relying on (i) establishing 2D-3D correspondences between pixels in the camera's image plane and corresponding points in the scene, and on (ii) using those correspondences as input to an optimization procedure that relies on bundle adjustment \cite{triggs1999bundle} to output maximum likelihood estimates of the camera calibration matrix, the associated lens distortion model coefficients, and, for each calibration image, the pose (i.e., position and orientation) of the camera relative to the 3D points \cite{Hartley2004,zhang2000flexible}.
+Images of a calibration surface such as a chessboard pattern are used to identify the 2D-3D correspondences~$\{\mathbf{x}_{i,j} \leftrightarrow \mathbf{X}_{i,j}\}$, $i \in \{ 1, \dots, n_\text{pt} \}, j \in \{ 1, \dots, n_\text{im} \}$, where $n_\text{pt}$ gives the number of correspondences obtained from one calibration image of the pattern, $n_\text{im}$ the number of such images, and $\mathbf{X}_{i,j} = \mathbf{X}_{i,j'}$, for $j, j' \in \{ 1, \dots, n_\text{im} \}$.
+Bundle adjustment is then used to obtain the maximum likelihood estimates~$(\hat{\omega}, \hat{\mathtt{K}}, \{(\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j)\})$ of the lens distortion model coefficients~$\omega$, calibration matrix~$\mathtt{K}$, and rigid body transformations $(\mathtt{R}_j, \mathbf{t}_j) \in SE(3)$ that minimize the cumulative reprojection error
 \begin{equation}
-\sum_{i=0}^{n_\text{pt}} \sum_{j=0}^{n_\text{im}} d_\omega\big(\mathbf{x}_{i,j}, \mathtt{K}(\mathtt{R}_j\mathbf{X}_{i,j} + \mathbf{t}_j)\big),
+%\sum_{i=0}^{n_\text{pt}} \sum_{j=0}^{n_\text{im}} d_\omega\big(\mathbf{x}_{i,j}, \pi_{\mathtt{K},\mathtt{R}_j,\mathbf{t}_j}(\mathbf{X}_{i,j})\big),
+\sum_{i=0}^{n_\text{pt}} \sum_{j=0}^{n_\text{im}} d_{\hat{\omega}}(\mathbf{x}_{i,j}, \hat{\mathbf{x}}_{i,j}'),
 \label{re}
 \end{equation}
-where $\mathbf{X}_{i,j}' = \mathtt{R}_j\mathbf{X}_{i,j} + \mathbf{t}_j \in \mathbb{R}^3$ expresses $\mathbf{X}_{i,j}$ in the camera coordinate frame of the camera corresponding to the $j^{\text{th}}$ calibration image, $(\mathbf{x}_{i,j}', 1)^\top \sim \mathtt{K}\mathbf{X}_{i,j}' \in \mathbb{P}^2$ gives the projection~$\mathbf{x}'_{i,j} \in \mathbb{R}^2$ of $\mathbf{X}_{i,j}'$ to that image, and $d_\omega$ is a distance function that~(i) corrects for lens distortion using the lens distortion model coefficients~$\omega$ \cite{duane1971close,weng1992camera} and (ii) returns a distance with respect to the two undistorted pixel positions.
-The inverse rigid body transformation~$(\mathtt{R}_j^{-1}, -\mathtt{R}_j^{-1}\mathbf{t}_j^{})$ gives the pose of the $j^{\text{th}}$~camera relative to the 3D points of the calibration pattern.
+where $\hat{\mathtt{R}}_j\mathbf{X}_{i,j} + \hat{\mathbf{t}}_j \in \mathbb{R}^3$ expresses $\mathbf{X}_{i,j}$ in the camera coordinate frame of the camera corresponding to the $j^{\text{th}}$~calibration image and $(\hat{\mathbf{x}}_{i,j}'^\top, 1)^\top \sim \hat{\mathtt{K}}(\hat{\mathtt{R}}_j\mathbf{X}_{i,j} + \hat{\mathbf{t}}_j) \in \mathbb{P}^2$ gives the projection~$\hat{\mathbf{x}}_{i,j}' \in \mathbb{R}^2$ of the point~$\mathbf{X}_{i,j}$ to the that image.
+The function~$d_\omega$ is a distance function that computes a distance with respect to the two pixel positions corrected for lens distortions, in accordance with the lens distortion model coefficients~$\hat{\omega}$ \cite{duane1971close,weng1992camera}.
+The inverse rigid body transformation~$(\hat{\mathtt{R}}_j^{-1}, -\hat{\mathtt{R}}_j^{-1}\hat{\mathbf{t}}_j^{})$ gives the pose of the $j^{\text{th}}$~camera relative to the 3D points of the calibration pattern.
 

### 2021-07-26 src/cpslib/libCalib/CalibPattern.cpp

@@ -82,8 +82,6 @@ void CalibPattern::findCirclesImPts(cv::Mat& im, cv::Size circlesPatternSize,
 {
     cv::Mat imUndistort, imGrayUndistort, imGrayUndistortInvert;
 
-    std::cout << 1 << std::endl;
-
     if (applyIntrinsics)
     {
         cv::undistort(im, imUndistort, K, distCoeffs);
@@ -91,28 +89,14 @@ void CalibPattern::findCirclesImPts(cv::Mat& im, cv::Size circlesPatternSize,
     else
         im.copyTo(imUndistort);
 
-    std::cout << 2 << std::endl;
-
     if (outImVis.size != imUndistort.size)
         imUndistort.copyTo(outImVis);
 
-    std::cout << 3 << std::endl;
-
     cv::cvtColor(imUndistort, imGrayUndistort, cv::COLOR_BGR2GRAY);
-
-    std::cout << 4 << std::endl;
-
-    //cv::Ptr<cv::CLAHE> clahe = cv::createCLAHE();
-    //clahe->setClipLimit(4);
-    //clahe->apply(imGrayUndistort, imGrayUndistort);
 
     cv::bitwise_not(imGrayUndistort, imGrayUndistortInvert);
 
-    std::cout << 5 << std::endl;
-
     outImSize = cv::Size(im.cols, im.rows);
-
-    std::cout << 6 << std::endl;
 
     cv::Mat imGrayUndistortInvertVis;
     cv::resize(imGrayUndistortInvert, imGrayUndistortInvertVis, cv::Size(imGrayUndistortInvert.cols * 0.5, imGrayUndistortInvert.rows * 0.5));
@@ -125,12 +109,8 @@ void CalibPattern::findCirclesImPts(cv::Mat& im, cv::Size circlesPatternSize,
         cv::cornerSubPix(imGrayUndistort, candidateCirclesImPts, cv::Size(5, 5), cv::Size(-1, -1),
             cv::TermCriteria(cv::TermCriteria::EPS + cv::TermCriteria::COUNT, 40, 0.001));
 
-        std::cout << 7 << std::endl;
-
         drawChessboardCorners(outImVis, circlesPatternSize, candidateCirclesImPts, true);
         outCirclesImPts.push_back(candidateCirclesImPts);
-
-        std::cout << 8 << std::endl;
     }
 }
 

### 2021-07-26 paper/elsarticle/elsarticle-template.tex
paper

@@
 %\sum_{i=0}^{n_\text{pt}} \sum_{j=0}^{n_\text{im}} d_\omega\big(\mathbf{x}_{i,j}, \pi_{\mathtt{K},\mathtt{R}_j,\mathbf{t}_j}(\mathbf{X}_{i,j})\big),
-\sum_{i=0}^{n_\text{pt}} \sum_{j=0}^{n_\text{im}} d_{\hat{\omega}}(\mathbf{x}_{i,j}, \hat{\mathbf{x}}_{i,j}'),
+\sum_{i=1}^{n_\text{pt}} \sum_{j=1}^{n_\text{im}} d_{\hat{\omega}}(\mathbf{x}_{i,j}, \hat{\mathbf{x}}_{i,j}'),
 \label{re}
@@
 where $\hat{\mathtt{R}}_j\mathbf{X}_{i,j} + \hat{\mathbf{t}}_j \in \mathbb{R}^3$ expresses $\mathbf{X}_{i,j}$ in the camera coordinate frame of the camera corresponding to the $j^{\text{th}}$~calibration image and $(\hat{\mathbf{x}}_{i,j}'^\top, 1)^\top \sim \hat{\mathtt{K}}(\hat{\mathtt{R}}_j\mathbf{X}_{i,j} + \hat{\mathbf{t}}_j) \in \mathbb{P}^2$ gives the projection~$\hat{\mathbf{x}}_{i,j}' \in \mathbb{R}^2$ of the point~$\mathbf{X}_{i,j}$ to the that image.
-The function~$d_\omega$ is a distance function that computes a distance with respect to the two pixel positions corrected for lens distortions, in accordance with the lens distortion model coefficients~$\hat{\omega}$ \cite{duane1971close,weng1992camera}.
+The function~$d_{\hat{\omega}}$ is a distance function that computes a distance with respect to two pixel positions that it first corrects for lens distortions, in accordance with the lens distortion model coefficients~$\hat{\omega}$ \cite{duane1971close,weng1992camera}.
 The inverse rigid body transformation~$(\hat{\mathtt{R}}_j^{-1}, -\hat{\mathtt{R}}_j^{-1}\hat{\mathbf{t}}_j^{})$ gives the pose of the $j^{\text{th}}$~camera relative to the 3D points of the calibration pattern.

### 2021-07-26 paper/elsarticle/elsarticle-template.tex
paper

@@
 
-Calibration of a projector (or camera) in the sense we employ the term here\footnote{We are referring to a geometric calibration; not, e.g., to a color calibration.} renders one able to project a scene point~$\mathbf{X} \in \mathbb{R}^3$ to its corresponding pixel~$\mathbf{x} \in \mathbb{R}^2$ in the projector's (or camera's) image plane, or to compute the `back-projection' of $\mathbf{x}$, i.e., the ray from the projector's (or camera's) center of projection through $\mathbf{x}$ along which the projecting point~$\mathbf{X}$ must lie.
+Calibration of a projector (or camera) in the sense we employ the term here\footnote{We are referring to a geometric calibration; not, e.g., to a color calibration.} renders one able to project a scene point~$\mathbf{X} \in \mathbb{R}^3$ to its corresponding pixel~$\mathbf{x} \in \mathbb{R}^2$ in the projector's (or camera's) image plane, or to compute the `back-projection' of $\mathbf{x}$, i.e., the ray from the projector's (or camera's) center of projection through $\mathbf{x}$ along which a projecting point~$\mathbf{X}$ must lie.
 Such a calibration can be expressed in terms of (i) a $3~\times{}~3$ calibration matrix~$\mathtt{K}$ derived from the projector's (or camera's) focal length~$f$ and principal point~$\mathbf{p}_0 = (x_0, y_0)^\top \in \mathbb{R}^2$ \cite{Hartley2004}, and (ii)~the coefficients of a lens distortion model used to correct for radial or tangential distortions caused by the lens system \cite{duane1971close,weng1992camera}; the calibration matrix~$\mathtt{K}$ is then
@@
 
-Calibrating a camera can be carried out by relying on (i) establishing 2D-3D correspondences between pixels in the camera's image plane and corresponding points in the scene, and on (ii) using those correspondences as input to an optimization procedure that relies on bundle adjustment \cite{triggs1999bundle} to output maximum likelihood estimates of the camera calibration matrix, the associated lens distortion model coefficients, and, for each calibration image, the pose (i.e., position and orientation) of the camera relative to the 3D points \cite{Hartley2004,zhang2000flexible}.
-Images of a calibration surface such as a chessboard pattern are used to identify the 2D-3D correspondences~$\{\mathbf{x}_{i,j} \leftrightarrow \mathbf{X}_{i,j}\}$, $i \in \{ 1, \dots, n_\text{pt} \}, j \in \{ 1, \dots, n_\text{im} \}$, where $n_\text{pt}$ gives the number of correspondences obtained from one calibration image of the pattern, $n_\text{im}$ the number of such images, and $\mathbf{X}_{i,j} = \mathbf{X}_{i,j'}$, for $j, j' \in \{ 1, \dots, n_\text{im} \}$.
+Calibrating a camera can be carried out by relying on (i) establishing 2D-3D correspondences between pixels in the camera's image plane and the corresponding points in the scene, and on (ii) using those correspondences as input to an optimization procedure that relies on bundle adjustment \cite{triggs1999bundle} to output maximum likelihood estimates of the calibration matrix, the associated lens distortion model coefficients, and, for each calibration image, the pose (i.e., position and orientation) of the camera relative to the 3D points \cite{Hartley2004,zhang2000flexible}.
+Calibration images of a calibration surface such as a chessboard pattern are acquired from varying viewpoints using the camera, to be used to establish the 2D-3D correspondences~$\{\mathbf{x}_{i,j} \leftrightarrow \mathbf{X}_{i,j}\}$, $i \in \{ 1, \dots, n_\text{pt} \}, j \in \{ 1, \dots, n_\text{im} \}$, where $n_\text{pt}$ gives the number of correspondences obtained from one calibration image of the pattern, $n_\text{im}$ the number of such images, and $\mathbf{X}_{i,j} = \mathbf{X}_{i,j'}$, for $j, j' \in \{ 1, \dots, n_\text{im} \}$.
 Bundle adjustment is then used to obtain the maximum likelihood estimates~$(\hat{\omega}, \hat{\mathtt{K}}, \{(\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j)\})$ of the lens distortion model coefficients~$\omega$, calibration matrix~$\mathtt{K}$, and rigid body transformations $(\mathtt{R}_j, \mathbf{t}_j) \in SE(3)$ that minimize the cumulative reprojection error
@@
 \end{equation}
-where $\hat{\mathtt{R}}_j\mathbf{X}_{i,j} + \hat{\mathbf{t}}_j \in \mathbb{R}^3$ expresses $\mathbf{X}_{i,j}$ in the camera coordinate frame of the camera corresponding to the $j^{\text{th}}$~calibration image and $(\hat{\mathbf{x}}_{i,j}'^\top, 1)^\top \sim \hat{\mathtt{K}}(\hat{\mathtt{R}}_j\mathbf{X}_{i,j} + \hat{\mathbf{t}}_j) \in \mathbb{P}^2$ gives the projection~$\hat{\mathbf{x}}_{i,j}' \in \mathbb{R}^2$ of the point~$\mathbf{X}_{i,j}$ to the that image.
+where $\hat{\mathtt{R}}_j\mathbf{X}_{i,j} + \hat{\mathbf{t}}_j \in \mathbb{R}^3$ expresses $\mathbf{X}_{i,j}$ in the camera coordinate frame of the camera corresponding to the $j^{\text{th}}$~calibration image and $(\hat{\mathbf{x}}_{i,j}'^\top, 1)^\top \sim \hat{\mathtt{K}}(\hat{\mathtt{R}}_j\mathbf{X}_{i,j} + \hat{\mathbf{t}}_j) \in \mathbb{P}^2$ gives the projection~$\hat{\mathbf{x}}_{i,j}' \in \mathbb{R}^2$ of the point~$\mathbf{X}_{i,j}$ to that image.
 The function~$d_{\hat{\omega}}$ is a distance function that computes a distance with respect to two pixel positions that it first corrects for lens distortions, in accordance with the lens distortion model coefficients~$\hat{\omega}$ \cite{duane1971close,weng1992camera}.

### 2021-07-26 paper/elsarticle/elsarticle-template.tex
Update elsarticle-template.tex

@@
 The function~$d_{\hat{\omega}}$ is a distance function that computes a distance with respect to two pixel positions that it first corrects for lens distortions, in accordance with the lens distortion model coefficients~$\hat{\omega}$ \cite{duane1971close,weng1992camera}.
-The inverse rigid body transformation~$(\hat{\mathtt{R}}_j^{-1}, -\hat{\mathtt{R}}_j^{-1}\hat{\mathbf{t}}_j^{})$ gives the pose of the $j^{\text{th}}$~camera relative to the 3D points of the calibration pattern.
+The inverse rigid body transformation~$(\hat{\mathtt{R}}_j^{-1}, -\hat{\mathtt{R}}_j^{-1}\hat{\mathbf{t}}_j^{})$ gives the pose of the $j^{\text{th}}$~camera relative to the 3D points~$\{ \mathbf{X}_{i,j} \}$ of the calibration pattern.
 

### 2021-07-26 paper/elsarticle/elsarticle-template.tex
Update elsarticle-template.tex

--- 2021-07-26 paper/elsarticle/elsarticle-template.tex
@@
 Calibration images of a calibration surface such as a chessboard pattern are acquired from varying viewpoints using the camera, to be used to establish the 2D-3D correspondences~$\{\mathbf{x}_{i,j} \leftrightarrow \mathbf{X}_{i,j}\}$, $i \in \{ 1, \dots, n_\text{pt} \}, j \in \{ 1, \dots, n_\text{im} \}$, where $n_\text{pt}$ gives the number of correspondences obtained from one calibration image of the pattern, $n_\text{im}$ the number of such images, and $\mathbf{X}_{i,j} = \mathbf{X}_{i,j'}$, for $j, j' \in \{ 1, \dots, n_\text{im} \}$.
-Bundle adjustment is then used to obtain the maximum likelihood estimates~$(\hat{\omega}, \hat{\mathtt{K}}, \{(\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j)\})$ of the lens distortion model coefficients~$\omega$, calibration matrix~$\mathtt{K}$, and rigid body transformations $(\mathtt{R}_j, \mathbf{t}_j) \in SE(3)$ that minimize the cumulative reprojection error
+Bundle adjustment is then used to obtain the maximum likelihood estimates~$(\hat{\omega}, \hat{\mathtt{K}}, \{(\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j)\}_{i,j})$ of the lens distortion model coefficients~$\omega$, calibration matrix~$\mathtt{K}$, and rigid body transformations $(\mathtt{R}_j, \mathbf{t}_j) \in SE(3)$ that minimize the cumulative reprojection error
 \begin{equation}
@@
 The function~$d_{\hat{\omega}}$ is a distance function that computes a distance with respect to two pixel positions that it first corrects for lens distortions, in accordance with the lens distortion model coefficients~$\hat{\omega}$ \cite{duane1971close,weng1992camera}.
-The inverse rigid body transformation~$(\hat{\mathtt{R}}_j^{-1}, -\hat{\mathtt{R}}_j^{-1}\hat{\mathbf{t}}_j^{})$ gives the pose of the $j^{\text{th}}$~camera relative to the 3D points of the calibration pattern.
+The inverse rigid body transformation~$(\hat{\mathtt{R}}_j^{-1}, -\hat{\mathtt{R}}_j^{-1}\hat{\mathbf{t}}_j^{})$ gives the pose of the $j^{\text{th}}$~camera relative to the 3D points~$\{ \mathbf{X}_{i,j} \}_{i=1}^{n_\text{pt}$ of the calibration pattern.
 

### 2021-07-29 paper/elsarticle/elsarticle-template.tex
Update elsarticle-template.tex

@@
 \paragraph{Plane-induced homography} Let $\mathtt{K}_\text{proj}$ express the calibration matrix of the recovered projector and $(\mathtt{R}, \mathbf{t}) \in SE(3)$ the rigid body transformation that transforms points from the coordinate frame of the recovered projector to that of the virtual projector, for a given target location.
-Moreover, let $(\mathbf{n}^\top, -d)^\top$ give the scene plane, expressed in the coordinate frame of the recovered projector, where $\mathbf{n} \in \mathbb{R}^3$ is the scene plane's normal vector and $d = \mathbf{n}^\top\mathbf{X}$ for any point~$\mathbf{X}$ in the plane, so that $(\mathbf{n}^\top, -d) (\mathbf{X}^\top, 1)^\top = 0$.
+Moreover, let $(\mathbf{n}^\top, -d)^\top \in \mathbb{P}^3$ give the scene plane, expressed in the coordinate frame of the recovered projector, where $\mathbf{n} \in \mathbb{R}^3$ is the scene plane's normal vector and $d = \mathbf{n}^\top\mathbf{X}$ for any point~$\mathbf{X}$ in the plane, so that $(\mathbf{n}^\top, -d) (\mathbf{X}^\top, 1)^\top = 0$.
 The transformation that warps the image to be projected to the scene plane by the recovered projector such that it appear as if were projected to the scene plane by the virtual projector (cf.\ Figure~\ref{fig:warp}(b)) is given the by the $3~\times~3$ matrix

### 2021-07-29 paper/elsarticle/elsarticle-template.tex
Update elsarticle-template.tex

@@
 Moreover, let $(\mathbf{n}^\top, -d)^\top \in \mathbb{P}^3$ give the scene plane, expressed in the coordinate frame of the recovered projector, where $\mathbf{n} \in \mathbb{R}^3$ is the scene plane's normal vector and $d = \mathbf{n}^\top\mathbf{X}$ for any point~$\mathbf{X}$ in the plane, so that $(\mathbf{n}^\top, -d) (\mathbf{X}^\top, 1)^\top = 0$.
-The transformation that warps the image to be projected to the scene plane by the recovered projector such that it appear as if were projected to the scene plane by the virtual projector (cf.\ Figure~\ref{fig:warp}(b)) is given the by the $3~\times~3$ matrix
+The transformation that warps the image to be projected to the scene plane from the viewpoint of the \textit{recovered} projector such that it appear as if were projected to the scene plane by the \textit{virtual} projector (cf.\ Figure~\ref{fig:warp}(b)) is given the by the $3~\times~3$ matrix
 \begin{equation}

### 2021-07-31 paper/elsarticle/elsarticle-template.tex
paper

--- 2021-07-26 paper/elsarticle/elsarticle-template.tex
@@
 \author{Majesa Trimmel}
-\author{Patrick Rupprecht}
+\author{Patrick~Rupprecht}
 \author{Sebastian Schlund}
@@
 \begin{keyword}
-Spatial augmented reality (SAR) \sep computer vision \sep projector calibration \sep Industry 4.0 \sep Pilotfabrik
+Spatial augmented reality (SAR) \sep computer vision \sep projector calibration \sep steerable mirror projection \sep Industry 4.0 \sep Pilotfabrik
 \end{keyword}
@@
 
-Our contribution is to propose a wide-area spatial AR system for planar scenes that produces the effect of keystone correction analytically.
+Our contribution is to propose a wide-area spatial AR system for planar scenes that produces the effect of keystone correction analytically, free of any manual interaction.
 Most importantly, we do this in a manner placing the axes of the augmentation in accordance with the axes of a camera placed to face downwards towards the scene plane.
@@
 
-Keystone correction for planar scenes can be carried out using a 2D homography \cite{Hartley2004}---an invertible transformation that preserves colinearity, hence alternatively termed a `colineation'---and is ultimately the approach we take as well.
+An early spatial AR system explicitly using a projector with a steerable mirror is the IBM Everywhere Displays prototype of Pinharez \cite{pinhanez2001everywhere}.
+For the purposes of the prototype, a projector mounted with a steerable mirror was set up with a camera to demonstrate a variety of use cases, including collaborative assembly encompassing the projection of assembly instructions and an interactive projected user interface \cite{kjeldsen2002interacting,pinhanez2003applications}.
+The authors, however, carry out keystone correction manually, by interactively adjusting for the position and orientation of a virtual scene plane and a scaling factor and computing a 2D homography accordingly.
+%Rather than rely on a steerable mirror to support spatial AR to multiple locations using a single projector, some works mount the projector and a camera on a rigid rig and subject the rig to motion \cite{ehnes2004projected,borkowski2004spatial,butz2006applying}.
+Steering only a mirror, however, places humbler requirements on the system from a hardware standpoint than if an entire camera-projector rig is to be subject to rigid motion.
+
+More generally, keystone correction for planar scenes reduces to computing a 2D homography \cite{Hartley2004}, an invertible transformation that preserves colinearity; this is the case whether computing the homography relies on manual interaction as in the case of Pinharez, or is obtained free of manual interaction as in ours.
 The intuition for why it is that a transformation that maps lines to lines can serve to model the appropriate image warp can be drawn from considering a planar chessboard pattern: looking at an image of a chessboard acquired from an oblique angle, one observes that lines parallel in the chessboard appear to meet in respective vanishing points; looking at an image of the same chessboard acquired frontally with respect to the plane of the chessboard, lines parallel in the chessboard appear parallel in the image (i.e., they are said to meet `at infinity').
@@
 Tilt sensors can be used to recover the projector's gravity vector, and be integrated in the computation of a homography as in Raskar and Beardsley \cite{raskar2001self}.
-
-An early spatial AR system explicitly using a projector with a steerable mirror is the IBM Everywhere Displays prototype of Pinhanez \cite{pinhanez2001everywhere}; the system, however, has keystone correction carried out manually.
-Rather than rely on a steerable mirror to support spatial AR to multiple locations using a single projector, some works mount the projector and a camera on a rigid rig and subject the rig to motion \cite{ehnes2004projected,borkowski2004spatial,butz2006applying}.
-Steering only a mirror, however, places humbler requirements on the system from a hardware standpoint than if an entire camera-projector rig is to be subject to rigid motion.
 
@@
 
-Calibrating a camera can be carried out by relying on (i) establishing 2D-3D correspondences between pixels in the camera's image plane and the corresponding points in the scene, and on (ii) using those correspondences as input to an optimization procedure that relies on bundle adjustment \cite{triggs1999bundle} to output maximum likelihood estimates of the calibration matrix, the associated lens distortion model coefficients, and, for each calibration image, the pose (i.e., position and orientation) of the camera relative to the 3D points \cite{Hartley2004,zhang2000flexible}.
-Calibration images of a calibration surface such as a chessboard pattern are acquired from varying viewpoints using the camera, to be used to establish the 2D-3D correspondences~$\{\mathbf{x}_{i,j} \leftrightarrow \mathbf{X}_{i,j}\}$, $i \in \{ 1, \dots, n_\text{pt} \}, j \in \{ 1, \dots, n_\text{im} \}$, where $n_\text{pt}$ gives the number of correspondences obtained from one calibration image of the pattern, $n_\text{im}$ the number of such images, and $\mathbf{X}_{i,j} = \mathbf{X}_{i,j'}$, for $j, j' \in \{ 1, \dots, n_\text{im} \}$.
-Bundle adjustment is then used to obtain the maximum likelihood estimates~$(\hat{\omega}, \hat{\mathtt{K}}, \{(\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j)\}_{i,j})$ of the lens distortion model coefficients~$\omega$, calibration matrix~$\mathtt{K}$, and rigid body transformations $(\mathtt{R}_j, \mathbf{t}_j) \in SE(3)$ that minimize the cumulative reprojection error
+Calibrating a camera can be carried out by relying on (i) establishing 2D-3D correspondences between pixels in the camera's image plane and the corresponding points in the scene, and on (ii) using those correspondences as input to an optimization procedure that relies on bundle adjustment \cite{triggs1999bundle} to output maximum likelihood estimates of the calibration matrix, the associated lens distortion model coefficients, and, for each calibration image, the pose (i.e., position and orientation) of the camera relative to the 3D scene points \cite{Hartley2004,zhang2000flexible}.
+Calibration images containing a calibration surface such as a chessboard pattern of known dimensions and scale are acquired from varying viewpoints using the camera, to be used to establish the 2D-3D correspondences~$\{\mathbf{x}_{i,j} \leftrightarrow \mathbf{X}_{i,j}\}$, $i \in \{ 1, \dots, n_\text{pt} \}, j \in \{ 1, \dots, n_\text{im} \}$, where $n_\text{pt}$ gives the number of correspondences obtained from one calibration image of the pattern, $n_\text{im}$ the number of such images, and $\mathbf{X}_{i,j} = \mathbf{X}_{i,j'}$, for $j, j' \in \{ 1, \dots, n_\text{im} \}$.
+Bundle adjustment is then used to obtain the maximum likelihood estimates~$(\hat{\omega}, \hat{\mathtt{K}}, \{(\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j)\}), j \in \{ 1, \dots, n_\text{im} \}$ of the lens distortion model coefficients~$\omega$, calibration matrix~$\mathtt{K}$, and rigid body transformations $(\mathtt{R}_j, \mathbf{t}_j) \in SE(3), j \in \{ 1, \dots, n_\text{im} \}$ that minimize the cumulative reprojection~error
 \begin{equation}
-%\sum_{i=0}^{n_\text{pt}} \sum_{j=0}^{n_\text{im}} d_\omega\big(\mathbf{x}_{i,j}, \pi_{\mathtt{K},\mathtt{R}_j,\mathbf{t}_j}(\mathbf{X}_{i,j})\big),
 \sum_{i=1}^{n_\text{pt}} \sum_{j=1}^{n_\text{im}} d_{\hat{\omega}}(\mathbf{x}_{i,j}, \hat{\mathbf{x}}_{i,j}'),
@@
 The function~$d_{\hat{\omega}}$ is a distance function that computes a distance with respect to two pixel positions that it first corrects for lens distortions, in accordance with the lens distortion model coefficients~$\hat{\omega}$ \cite{duane1971close,weng1992camera}.
-The inverse rigid body transformation~$(\hat{\mathtt{R}}_j^{-1}, -\hat{\mathtt{R}}_j^{-1}\hat{\mathbf{t}}_j^{})$ gives the pose of the $j^{\text{th}}$~camera relative to the 3D points~$\{ \mathbf{X}_{i,j} \}_{i=1}^{n_\text{pt}$ of the calibration pattern.
+The inverse rigid body transformation~$(\hat{\mathtt{R}}_j^{-1}, -\hat{\mathtt{R}}_j^{-1}\hat{\mathbf{t}}_j^{})$ gives the pose of the $j^{\text{th}}$~camera relative to the 3D points of the calibration pattern.
 
@@
 What remains of this section is concerned primarily with the recovery of 2D-3D correspondences in support of calibrating cameras and projectors.
-A consequence of the approach we take to identifying the 3D points of the 2D-3D correspondences we use for projector calibration is, for each target location, recovery of the pose of the scene plane relative to the coordinate frame of the camera.
+A consequence of the approach we take to identifying the 3D points of the 2D-3D correspondences we use for projector calibration is, for each target location, recovery of the scene plane.
 

### 2021-07-31 paper/elsarticle/elsarticle-template.tex
paper

@@
 Care must be taken in carrying out spatial AR, however, to warp the images to be projected in a manner that they appear free of distortions to the viewer, since the geometry of the scene as it relates to the geometry of the projector plays a role in how the pixels of the projector's image plane map to points in the scene.
-For planar scene geometry (such as a floor, wall, or table), this can be done in a cumbersome manual process called keystone correction, often using software bundled with the projector.
+For planar scene geometry (such as a floor, wall, or table), this can be done in a cumbersome manual process referred to as keystone correction, often using software bundled with the projector.
 
@@
 
-Augmented reality (AR) \cite{van2010survey,zhou2008trends} promises to enable use cases in industrial settings that include the embedding of assembly instructions directly into the scene \cite{schlund2018moglichkeiten,masood2019augmented,gattullo2019towards,rupprecht2020information,Rupprecht2021}, potentially reducing or altogether doing away with the need for workers to refer to instructions in paper form or on a screen.
+Augmented reality (AR) \cite{van2010survey,zhou2008trends} promises to enable use cases in industrial settings that include the embedding of assembly instructions directly into the scene \cite{schlund2018moglichkeiten,masood2019augmented,gattullo2019towards,uva2018evaluating,rupprecht2020information,Rupprecht2021}, potentially reducing or altogether doing away with the need for workers to refer to instructions in paper form or on a screen.
 Typically, AR works by embedding the augmentation in an image of the scene acquired from the viewpoint of a single individual, with the resulting augmented image in turn displayed using some form of head-mounted display.
@@
 
-An early spatial AR system explicitly using a projector with a steerable mirror is the IBM Everywhere Displays prototype of Pinharez \cite{pinhanez2001everywhere}.
+An early spatial AR system explicitly using a projector mounted with a steerable mirror is the IBM Everywhere Displays prototype of Pinharez \cite{pinhanez2001everywhere}.
 For the purposes of the prototype, a projector mounted with a steerable mirror was set up with a camera to demonstrate a variety of use cases, including collaborative assembly encompassing the projection of assembly instructions and an interactive projected user interface \cite{kjeldsen2002interacting,pinhanez2003applications}.
@@
 
-More generally, keystone correction for planar scenes reduces to computing a 2D homography \cite{Hartley2004}, an invertible transformation that preserves colinearity; this is the case whether computing the homography relies on manual interaction as in the case of Pinharez, or is obtained free of manual interaction as in ours.
+More generally, keystone correction for planar scenes can be said to reduce to computing a 2D homography \cite{Hartley2004}, an invertible transformation that preserves colinearity; this is the case whether computing the homography relies on manual interaction as in the case of Pinharez, or is obtained free of manual interaction as in ours.
 The intuition for why it is that a transformation that maps lines to lines can serve to model the appropriate image warp can be drawn from considering a planar chessboard pattern: looking at an image of a chessboard acquired from an oblique angle, one observes that lines parallel in the chessboard appear to meet in respective vanishing points; looking at an image of the same chessboard acquired frontally with respect to the plane of the chessboard, lines parallel in the chessboard appear parallel in the image (i.e., they are said to meet `at infinity').
-To warp the former image (where lines parallel in the scene meet in respective vanishing points) such that the lines of the chessboard appear as in the latter image (where lines parallel in the scene meet at infinity), a transformation that maps lines to lines would be sufficient.
+To warp the former image (where lines parallel in the scene meet in respective vanishing points) such that the lines of the chessboard appear as in the latter image (where lines parallel in the scene meet at infinity), a transformation that maps lines to lines is sufficient.
 
-One way to compute a homography is by identifying at least four correspondences between pixel positions in two images of a planar surface.
+One way to compute a homography is by identifying at least four correspondences between pixel positions in two images of a planar surface \cite{Hartley2004}.
 The keystone correction approach of Sukthankar \textit{et al.} \cite{sukthankar2001smarter} reduces to using a segmentation approach to identify the four corners of projection screen in an image acquired by a camera, and computing a homography that maps the resulting four corners of the projection screen to the four corners of the projector's image plane.
-Tilt sensors can be used to recover the projector's gravity vector, and be integrated in the computation of a homography as in Raskar and Beardsley \cite{raskar2001self}.
+Additional sensors can be used to inform the computation of a homography; Raskar and Beardsley \cite{raskar2001self} use a tilt sensor to recover the projector's gravity vector, which they use to carry out a rotational correction in projecting to a wall.
 
@@
 
-\begin{figure}
-    \centering
-    \subfloat[\centering Asymmetrical circle pattern image, in image plane of projector (detections overlain).]{{\includegraphics[height=3.25cm]{images/circles.png} }}
-    \qquad
-    \subfloat[\centering Projector calibration image (one for each target location), in image plane of camera (detections overlain).]{{\includegraphics[height=3.25cm]{images/detections.png}}}
-    \caption{Recovering 2D positions in support of projector calibration.
-(a) 2D positions of the 2D-3D correspondences to be used for calibrating the projector are obtained by detecting---in the image plane of the projector---the circle centers in the asymmetrical circle pattern image, to be projected to each of the target locations in the scene plane.
-(b) For each such target location, an image is acquired from the viewpoint of the camera and the circle centers of the projected asymmetrical circle pattern are detected, in the image plane of the camera.
-A chessboard pattern to be used for recovering the local scene plane is placed near the projected pattern, whose corners are likewise detected.
-Detected 2D projected asymmetrical circle pattern circle center points and chessboard corners overlain for illustration.}
-    \label{fig:2d}
-\end{figure}
-
 Calibration of a projector (or camera) in the sense we employ the term here\footnote{We are referring to a geometric calibration; not, e.g., to a color calibration.} renders one able to project a scene point~$\mathbf{X} \in \mathbb{R}^3$ to its corresponding pixel~$\mathbf{x} \in \mathbb{R}^2$ in the projector's (or camera's) image plane, or to compute the `back-projection' of $\mathbf{x}$, i.e., the ray from the projector's (or camera's) center of projection through $\mathbf{x}$ along which a projecting point~$\mathbf{X}$ must lie.
@@
 The function~$d_{\hat{\omega}}$ is a distance function that computes a distance with respect to two pixel positions that it first corrects for lens distortions, in accordance with the lens distortion model coefficients~$\hat{\omega}$ \cite{duane1971close,weng1992camera}.
-The inverse rigid body transformation~$(\hat{\mathtt{R}}_j^{-1}, -\hat{\mathtt{R}}_j^{-1}\hat{\mathbf{t}}_j^{})$ gives the pose of the $j^{\text{th}}$~camera relative to the 3D points of the calibration pattern.
+The inverse rigid body transformation~$(\hat{\mathtt{R}}_j^{-1}, -\hat{\mathtt{R}}_j^{-1}\hat{\mathbf{t}}_j^{}) \in SE(3)$ gives the pose of the $j^{\text{th}}$~camera relative to the 3D scene points of the calibration pattern.
 
@@
 What remains of this section is concerned primarily with the recovery of 2D-3D correspondences in support of calibrating cameras and projectors.
-A consequence of the approach we take to identifying the 3D points of the 2D-3D correspondences we use for projector calibration is, for each target location, recovery of the scene plane.
+A consequence of the approach we take to identifying the 3D points of the 2D-3D correspondences we use for projector calibration is, for each target location, recovery of the local scene plane.
 
@@
 2D-3D correspondences are then recovered for each calibration image, and the resulting list is passed on as input to an optimization procedure that relies on bundle adjustment to yield the camera calibration matrix~$\mathtt{K}_\text{cam}$.
+
+\begin{figure}
+    \centering
+    \subfloat[\centering Asymmetrical circle pattern image, in image plane of projector (detections overlain).]{{\includegraphics[height=3.25cm]{images/circles.png} }}
+    \qquad
+    \subfloat[\centering Projector calibration image (one for each target location), in image plane of camera (detections overlain).]{{\includegraphics[height=3.25cm]{images/detections.png}}}
+    \caption{Recovering 2D positions in support of projector calibration.
+(a) 2D positions of the 2D-3D correspondences to be used for calibrating the projector are obtained by detecting---in the image plane of the projector---the circle centers in the asymmetrical circle pattern image, to be projected to each of the target locations in the scene plane.
+(b) For each such target location, an image is acquired from the viewpoint of the camera and the circle centers of the projected asymmetrical circle pattern are detected, in the image plane of the camera.
+A chessboard pattern to be used for recovering the local scene plane is placed near the projected pattern, whose corners are likewise detected.
+Detected 2D projected asymmetrical circle pattern circle center points and chessboard corners overlain for illustration.}
+    \label{fig:2d}
+\end{figure}
 
@@
 \end{equation}
-a form of `plane-induced' homography \cite{Hartley2004}.
-For convenience, we enable optional rotation of the image to be projected \textit{before} applying $\mathtt{H}$, about the image center; that rotation, parameterized in degrees, is thus in effect likewise carried out intuitively relative to the placement of the camera.
+a form of `plane-induced' 2D homography \cite{Hartley2004}.
+For convenience, we enable optional rotation of the image to be projected \textit{before} applying $\mathtt{H}$, about the image center; that rotation, parameterized in degrees, is thus in effect likewise carried out intuitively in accordance with the placement of the camera.
 

### 2021-07-31 paper/elsarticle/elsarticle-template.tex
paper

@@
 \begin{keyword}
-Spatial augmented reality (SAR) \sep computer vision \sep projector calibration \sep steerable mirror projection \sep Industry 4.0 \sep Pilotfabrik
+Spatial augmented reality (SAR) \sep computer vision \sep projector calibration \sep steerable mirror projector \sep Industry 4.0 \sep Pilotfabrik
 \end{keyword}
@@
 More generally, keystone correction for planar scenes can be said to reduce to computing a 2D homography \cite{Hartley2004}, an invertible transformation that preserves colinearity; this is the case whether computing the homography relies on manual interaction as in the case of Pinharez, or is obtained free of manual interaction as in ours.
-The intuition for why it is that a transformation that maps lines to lines can serve to model the appropriate image warp can be drawn from considering a planar chessboard pattern: looking at an image of a chessboard acquired from an oblique angle, one observes that lines parallel in the chessboard appear to meet in respective vanishing points; looking at an image of the same chessboard acquired frontally with respect to the plane of the chessboard, lines parallel in the chessboard appear parallel in the image (i.e., they are said to meet `at infinity').
+The intuition for why it is that a transformation that preserves colinearity---i.e., maps lines to lines---can serve to model an appropriate corrective image warp can be drawn from considering a planar chessboard pattern: looking at an image of a chessboard acquired from an oblique angle, one observes that lines parallel in the chessboard appear to meet in respective vanishing points; looking at an image of the same chessboard acquired frontally with respect to the plane of the chessboard, lines parallel in the chessboard appear parallel in the image (i.e., they are said to meet `at infinity').
 To warp the former image (where lines parallel in the scene meet in respective vanishing points) such that the lines of the chessboard appear as in the latter image (where lines parallel in the scene meet at infinity), a transformation that maps lines to lines is sufficient.
@@
 
+\begin{figure}
+    \centering
+    \subfloat[\centering Hardware setup (projector, steerable mirror system, camera).]{{\includegraphics[height=3.3cm]{images/system2.png}}}
+    \quad
+    \subfloat[\centering Example projection of image warped for target location using our approach.]{{\includegraphics[height=3.3cm]{images/hans_undist.png}}}
+    \caption{Evaluation scenario.
+(a) Our hardware setup, comprised of a Panasonic PT-RZ660BE projector, a steerable mirror system manufactured by Dynamic Projection Institute, and a Stereolabs Zed 2 stereo camera, of which we used only the left view.
+(b) Example augmentation produced by projecting---to one of our 15 target locations---an image warped using our approach, as seen from the downwards-facing camera.
+Note that the warp places the axes of the augmentation in accordance with the axes of the camera, and that the dimensions of the augmentation are in line with the desired target dimensions (50~cm~$\times$~31.25~cm).
+Corners of the full projected image extent projected to the floor in light green.}
+    \label{fig:eval}
+\end{figure}
+
 We evaluate our approach by augmenting 15 locations across the floorspace at the Pilotfabrik\footnote{\url{https://www.pilotfabrik.at/}} of TU Wien, a collaborative space for research on Industry 4.0 topics situated in Vienna, Austria.

### 2021-08-01 paper/elsarticle/elsarticle-template.tex
paper

@@
 \textit{Spatial} AR, in turn, is a form of AR whereby the augmentation of the scene is carried out using a projector, with the advantage of rendering the augmentation visible to all onlookers simultaneously without calling for each to wear some form of head-mounted display.
-Care must be taken in carrying out spatial AR, however, to warp the images to be projected in a manner that they appear free of distortions to the viewer, since the geometry of the scene as it relates to the geometry of the projector plays a role in how the pixels of the projector's image plane map to points in the scene.
+In carrying out spatial AR, however, care must be taken to appropriately warp the images to be projected in a manner that they appear free of distortions to the viewer.
 For planar scene geometry (such as a floor, wall, or table), this can be done in a cumbersome manual process referred to as keystone correction, often using software bundled with the projector.
+%, since the geometry of the scene as it relates to the geometry of the projector plays a role in how the pixels of the projector's image plane map to points in the scene
 
-We propose a spatial AR system for wide-area metric augmentation of planar scene surfaces that produces the effect of keystone correction analytically, by using a projector equipped with a steerable mirror and a downwards-facing camera.
+We propose a spatial AR system for wide-area metric augmentation of planar scene surfaces that produces the effect of keystone correction analytically as a function of the recovered scene plane, by using a projector equipped with a steerable mirror and a camera facing the scene plane.
 Our system renders the placement of augmentations in the scene more intuitive than manual keystone correction in two ways.
-First, (i) placement of the desired augmentations is carried out in accordance with the axes of an image of the scene acquired by a camera placed to face downwards towards the scene plane; second, (ii) the desired dimensions of the projected augmentations is specified in metric terms.
-Our system produces such augmentations accurately, with compelling time savings relative manual keystone correction.
+First, (i) placement of the desired augmentations is carried out in accordance with the axes of an image of the scene acquired by the camera, thereby making setting those axes as simple as appropriately rotating the camera.
+Second, (ii) the desired dimensions of the projected augmentations are specified in metric terms, thereby allowing for consistent scaling across all target locations.
+%Our system produces such augmentations accurately, with compelling time savings relative manual keystone correction.
 \end{abstract}
@@
 \begin{keyword}
-Spatial augmented reality (SAR) \sep computer vision \sep projector calibration \sep steerable mirror projector \sep Industry 4.0 \sep Pilotfabrik
+Spatial augmented reality (SAR) \sep computer vision \sep projector-camera calibration \sep steerable mirror projector \sep Industry 4.0
 \end{keyword}
@@
 
-Mention asymmetrical circle pattern: \cite{chiu2011novel}
+%Mention asymmetrical circle pattern: \cite{chiu2011novel}
+
+%13. X. Zhang and L. Zhu, “Projector calibration from the camera
+%image point of view,” Opt. Eng. 48, 117208 (2009).
+%14. X. Chen, J. Xi, Y. Jin, and J. Sun, “Accurate calibration
+%for a camera-projector measurement system based on
+%structured light projection,” Opt. Laser Eng. 47, 310–319
+%(2009).
+%15. I. Din, H. Anwar, I. Syed, H. Zafar, and L. Hasan, “Projector
+%calibration for pattern projection systems,” J. Appl. Res.
+%Technol. 12, 80–86 (2014).
+%16. R. Orghidan, J. Salvi, M. Gordan, C. Florea, and J. Batlle,
+%“Structured light self-calibration with vanishing points,”
+%Mach. Vision Appl. 25, 489–500 (2014).
+%17. J. Huang, Z. Wang, Q. Xue, and J. Gao, “Calibration of a
+%camera-projector measurement system and error impact
+%analysis,” Meas. Sci. Technol. 23, 125402 (2012).
+%18. W. Gao, L. Wang, and Z. Hu, “Flexible calibration of a portable
+%structured light system through surface plane,” Acta Autom.
+%Sinica 34, 1358–1362 (2008).
+%19. G. Falcao, N. Hurtos, J. Massich, and D. Fofi, “Projectorcamera calibration toolbox,” Technology Report, 2009, available at http://code.google.com/p/procamcalib.
+%20. S. Zhang and P. Huang, “Novel method for structured light
+%system calibration,” Opt. Eng. 45, 083601 (2006).
+%21. F. Berry, O. Aider, and J. Mosnier, “A visual servoing-based
+%method for ProCam system calibration,” Sensors 13, 13318–
+%13333 (2013).
+%22. H. Anwar, I. Din, and K. Park, “Projector calibration for 3D
+%scanning using virtual target images,” Int. J. Precis. Eng.
+%Manuf. 13, 125–131 (2012).
 
@@
 
-Calibration of a projector can be carried out in precisely the same manner as calibrating a camera insofar as step (ii) is concerned; the major difference in projector calibration relative to the calibrating a camera concerns the manner in which 2D-3D correspondences are identified, i.e., between pixels in the image plane of the projector and the corresponding points in the scene.
+Calibration of a projector can be carried out in precisely the same manner as calibrating a camera insofar as step (ii) is concerned; the major difference in projector calibration relative to the calibrating a camera concerns the manner in which 2D-3D correspondences are identified, i.e., between pixels in the image plane of the projector and the corresponding points in the scene \cite{moreno2012simple,zhang2009projector,zhang2006novel}.
 What remains of this section is concerned primarily with the recovery of 2D-3D correspondences in support of calibrating cameras and projectors.
 A consequence of the approach we take to identifying the 3D points of the 2D-3D correspondences we use for projector calibration is, for each target location, recovery of the local scene plane.
-
-\paragraph{Camera calibration} We recover 2D-3D correspondences in support of calibrating the downwards-facing camera by relying on a planar calibration surface to automatically identify correspondences between the 3D points on the calibration surface and their 2D correspondences in the image plane.
-The classical calibration surface is a chessboard pattern.
-The 3D corner points of the chessboard are obtained \textit{a priori} in a coordinate system defined in the plane of the chessboard\footnote{E.g., $(0,0,0), (1.5,0,0), (3,0,0), \dots, (9,7.5,0)$ for a chessboard with $7~\times{}~6$ corners ($8~\times{}~7$ squares), with each square of length and width of 1.5 unit, respectively.
-Note that the units of the chessboard's 3D points give the units in terms of which the camera calibration is carried out, and---owing to how our projector calibration relies on the camera calibration---the units of the projector calibration as well.}, requiring knowledge only of the dimensions of the chessboard pattern and of the length of a side of a chessboard square.
-The corresponding 2D points are obtained, in the same order, using a specialized algorithm \cite{bradski2000opencv}.
-A set of calibration images is acquired, each with the calibration pattern visible in a different part of the image plane, and such that the center and all corners and edges of the image plane are covered, the camera's autofocus setting be off, and the camera's zoom factor remain fixed.
-2D-3D correspondences are then recovered for each calibration image, and the resulting list is passed on as input to an optimization procedure that relies on bundle adjustment to yield the camera calibration matrix~$\mathtt{K}_\text{cam}$.
 
@@
 
-\paragraph{Projector calibration} As with camera calibration, projector calibration relies on 2D-3D correspondences, yet we obtain them in this case by \textit{projecting} a calibration pattern.
+\paragraph{Camera calibration} We recover 2D-3D correspondences in support of calibrating the downwards-facing camera by relying on a planar calibration surface to automatically identify correspondences between the 3D points on the calibration surface and their 2D correspondences in the image plane.
+The classical calibration surface is a chessboard pattern.
+The 3D corner points of the chessboard are obtained \textit{a priori} in a coordinate system defined in the plane of the chessboard\footnote{E.g., $(0,0,0), (1.5,0,0), (3,0,0), \dots, (9,7.5,0)$ for a chessboard with $7~\times{}~6$ corners ($8~\times{}~7$ squares), with each square of length and width of 1.5 unit, respectively.
+Note that the units of the chessboard's 3D points give the units in terms of which the camera calibration is carried out, and---owing to how our projector calibration relies on the camera calibration---the units of the projector calibration as well.}, requiring knowledge only of the dimensions of the chessboard pattern and of the length of a side of a chessboard square.
+The corresponding 2D points are obtained, in the same order, using a specialized algorithm \cite{bradski2000opencv}.
+A set of calibration images is acquired, each with the calibration pattern visible in a different part of the image plane, and such that the center and all corners and edges of the image plane are covered, the camera's autofocus setting be off, and the camera's zoom factor remain fixed.
+2D-3D correspondences are then recovered for each calibration image, and the resulting list is passed on as input to an optimization procedure that relies on bundle adjustment to yield the camera calibration matrix~$\mathtt{K}_\text{cam}$.
+
+\paragraph{Projector calibration} As in camera calibration, we calibrate the projector by relying on 2D-3D correspondences, yet for projector calibration we obtain them by \textit{projecting} a calibration pattern that we detect using the calibrated downwards-facing camera.
 The pattern we project is one of circles (cf.\ Figure~\ref{fig:2d}(a)), and we rely on an algorithm to detect the circle center points in the asymmetrical circle pattern in the image plane of the projector \cite{bradski2000opencv}, giving the 2D positions of our 2D-3D correspondences for calibrating the projector.
@@
 We carry out this placement according to a small handful of steps.
-First, we (i) rotate the camera about its center of projection to align its optical axis\footnote{Strictly speaking, the optical axis is the back-projection of the principal point; in our usage, we understand it to refer to the back-projection of the center of the image to be projected.} with the normal vector of the scene plane, giving a virtual camera likewise facing directly\footnote{A physical camera placed to face downwards is almost certain to not face downwards precisely; in contrast, the virtual camera's optical axis is aligned exactly with the scene plane's normal vector, rendering it fronto-parallel with respect to the scene plane.} downwards to the scene plane (cf.\ Figure~\ref{fig:virtualproj}).
+First, we (i) rotate the camera about its center of projection to align its optical axis\footnote{Strictly speaking, the optical axis is the back-projection of the principal point; in our usage, we understand it to refer to the back-projection of the center of the image to be projected.} with the normal vector of the scene plane, giving a virtual camera likewise facing directly\footnote{A physical camera placed to face downwards is almost certain to not face downwards precisely; in contrast, the virtual camera's optical axis is aligned exactly with the scene plane's normal vector, rendering it genuinely fronto-parallel with respect to the scene plane.} downwards to the scene plane (cf.\ Figure~\ref{fig:virtualproj}).
 Next, we (ii) intersect the scene plane with the optical axis (i.e., the ray from the projector's center of projection through the center of the image plane) and rotate the projector's placement about that point of intersection, aligning the optical axis with the scene plane's normal vector and giving an initial virtual projector.
@@
 
-We evaluate our approach by augmenting 15 locations across the floorspace at the Pilotfabrik\footnote{\url{https://www.pilotfabrik.at/}} of TU Wien, a collaborative space for research on Industry 4.0 topics situated in Vienna, Austria.
-We contrast our approach with a baseline approach involving manual keystone correction, by aiming with both approaches to place the same image for each location aligned with the principal axes of the floorspace, absent of projective distortions, and with the same metric dimensions (50~cm~$\times$~31.25~cm\footnote{The dimensions in pixels of the image we project are 960~$\times$~600; we chose for our experiments to set the projected metric length of the horizontal axis of the image to 50~cm, which implies 31.25~cm for the vertical axis if aspect ratio is to be preserved.}).
+We evaluate our approach by using a 960 pixel~$\times$~600 pixel image of the TU Wien logo (cf.\ Figure \ref{fig:warp}) to augment 15 locations across the floorspace at the Pilotfabrik\footnote{\url{https://www.pilotfabrik.at/}} of TU Wien, a collaborative space for research on Industry 4.0 topics situated in Vienna, Austria.
+We contrast our approach with a baseline approach involving manual keystone correction, by aiming with both approaches to place the same image for each location aligned with the principal axes of the floorspace, absent of projective distortions, and with the same metric dimensions of 50~cm~$\times$~31.25~cm\footnote{The dimensions in pixels of the image we project are 960~$\times$~600; we chose for our experiments to set the projected metric length of the horizontal axis of the image to 50~cm, which implies 31.25~cm for the vertical axis if aspect ratio is to be preserved.}.
 All experiments were carried out by the same technician, experienced in both approaches.
 
-The hardware setup employed in the evaluation comprised a Panasonic PT-RZ660BE projector with a steerable mirror system---used in our experiments to point the projection to each of the 15 locations---manufactured by Dynamic Projection Institute \cite{rupprecht2020information,Rupprecht2021}.
+The hardware setup employed in the evaluation (cf.\ Figure \ref{fig:eval}(a)) comprised a Panasonic PT-RZ660BE projector with a steerable mirror system---used in our experiments to point the projection to each of the 15 locations---manufactured by Dynamic Projection Institute \cite{rupprecht2020information,Rupprecht2021}.
 The steerable mirror system was bundled with the MDC-X software for steering the mirror, loading imagery, and optionally carrying out manual keystone correction, such that each position and (warped) image can be registered as a preset.
@@
 
-\paragraph{Manual approach} ...
+\paragraph{Manual approach} Having pointed the steerable mirror to the desired location, manual keystone correction using the MDC-X software involves warping the image to be projected by manipulating the corners of the image until the desired effect is produced on the projection surface.
+In order to provide a template for manually carrying out keystone correction, we prepared a 50~cm~$\times$~31.25~cm piece of cardboard.
+For each target location, we placed this piece of cardboard in a manner that its axes were aligned with the principal axes of the scene (cf.\ the yellow and red lines in Figure \ref{fig:eval}(b)); we then manually warped the image for the given target location so that any corner of the augmentation deviated by at most 1 cm from the corresponding corner of the template.
+End-to-end, the process of setting the presets in the MDC-X software having carried out keystone correction manually took ca.\ XXX min for the 15 target locations.
 
 \paragraph{Our approach} We began by carrying out a calibration of the camera, acquiring 10 camera calibration images (cf.\ Section~\ref{sec:approach:geometry}) of a chessboard calibration pattern with $6~\times~4$ corners ($7~\times~5$ squares) and feeding the images as input to our camera calibration module.
-Separately, for each of the 15 target locations, we produced a projector calibration image (cf.\ again Section~\ref{sec:approach:geometry}) by projecting a $11~\times~4$ asymmetrical circle pattern image to the location in question using the steerable mirror, placing a checkerboard pattern beside the projected pattern, and acquiring the image using the downward-facing camera.
+Separately, for each of the 15 target locations, we produced a projector calibration image (cf.\ again Section~\ref{sec:approach:geometry}) by projecting a $11~\times~4$ asymmetrical circle pattern image to the location in question using the steerable mirror, placing a chessboard pattern beside the projected pattern, and acquiring the image using the downward-facing camera.
 We then fed these images alongside the output of the camera calibration module to our projector calibration module.
-For each of the target locations, the steerable mirror was made to point to that location, the asymmetrical circle pattern image was projected to the scene plane, an image was using the camera, and the location was registered in the MDC-X software as a preset.
+For each of the target locations, the steerable mirror was made to point to that location, the asymmetrical circle pattern image was projected to the scene plane, an image was acquired using the camera, and the location was registered in the MDC-X software as a preset.
 The output of the projector calibration module is a homography per input projector calibration image (cf.\ Section~\ref{sec:approach:homography}).
 Next, we warped the images to be projected to the respective locations using their corresponding homography, using a third dedicated custom module.
-These warped images were finally imported into the MDC-X software and associated with their respective location presets.
+These warped images were finally imported into the MDC-X software and associated with their respective location presets (cf.\ Figure \ref{fig:eval}(b)).
 

### 2021-08-01 paper/elsarticle/elsarticle-template.tex
paper

@@
 
-We propose a spatial AR system for wide-area metric augmentation of planar scene surfaces that produces the effect of keystone correction analytically as a function of the recovered scene plane, by using a projector equipped with a steerable mirror and a camera facing the scene plane.
+We propose a spatial AR system for wide-area metric augmentation of planar scene surfaces that produces the effect of keystone correction analytically as a function of the relative geometry of the projector and scene plane, by using a projector equipped with a steerable mirror and a camera facing the scene plane.
 Our system renders the placement of augmentations in the scene more intuitive than manual keystone correction in two ways.
@@
 \begin{keyword}
-Spatial augmented reality (SAR) \sep computer vision \sep projector-camera calibration \sep steerable mirror projector \sep Industry 4.0
+Spatial augmented reality (SAR) \sep computer vision \sep steerable mirror projector \sep projector-camera calibration \sep Industry 4.0
 \end{keyword}
@@
 Additional sensors can be used to inform the computation of a homography; Raskar and Beardsley \cite{raskar2001self} use a tilt sensor to recover the projector's gravity vector, which they use to carry out a rotational correction in projecting to a wall.
+Still another way to compute a homography is as a function of a plane and a pair of pinhole cameras; the effect produced by the homography is one of projecting an image to the plane from the viewpoint of the one camera, and acquiring the projected image from the viewpoint of the other.
+Such a homography is said to be `induced' by a plane \cite{Hartley2004}.
+We exploit the fact that both cameras and projectors can be modeled using a pinhole camera model \cite{bimber2019spatial}, proceed to recover the relative geometry relating the scene plane and projector, and compute a particular form of plane-induced homography to effect the intended corrective image warp.
 
@@
 where $f$ and $(p_x, p_y)^\top$ are both expressed in units of pixels.
+The calibration matrix~$\mathtt{K}$ models a so-called pinhole camera; such a model is applicable to an image acquired using a real-world camera assuming the image has been corrected for lens distortion effects, by having applied the lens distortion model coefficients.
 
-Calibrating a camera can be carried out by relying on (i) establishing 2D-3D correspondences between pixels in the camera's image plane and the corresponding points in the scene, and on (ii) using those correspondences as input to an optimization procedure that relies on bundle adjustment \cite{triggs1999bundle} to output maximum likelihood estimates of the calibration matrix, the associated lens distortion model coefficients, and, for each calibration image, the pose (i.e., position and orientation) of the camera relative to the 3D scene points \cite{Hartley2004,zhang2000flexible}.
+Calibrating a camera can be carried out by (i) establishing 2D-3D correspondences between pixels in the camera's image plane and the corresponding points in the scene, and on (ii) using those correspondences as input to an optimization procedure that relies on bundle adjustment \cite{triggs1999bundle} to output maximum likelihood estimates of the calibration matrix, the associated lens distortion model coefficients, and, for each calibration image, the pose (i.e., position and orientation) of the camera relative to the 3D scene points from among the 2D-3D correspondences \cite{Hartley2004,zhang2000flexible}.
 Calibration images containing a calibration surface such as a chessboard pattern of known dimensions and scale are acquired from varying viewpoints using the camera, to be used to establish the 2D-3D correspondences~$\{\mathbf{x}_{i,j} \leftrightarrow \mathbf{X}_{i,j}\}$, $i \in \{ 1, \dots, n_\text{pt} \}, j \in \{ 1, \dots, n_\text{im} \}$, where $n_\text{pt}$ gives the number of correspondences obtained from one calibration image of the pattern, $n_\text{im}$ the number of such images, and $\mathbf{X}_{i,j} = \mathbf{X}_{i,j'}$, for $j, j' \in \{ 1, \dots, n_\text{im} \}$.
-Bundle adjustment is then used to obtain the maximum likelihood estimates~$(\hat{\omega}, \hat{\mathtt{K}}, \{(\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j)\}), j \in \{ 1, \dots, n_\text{im} \}$ of the lens distortion model coefficients~$\omega$, calibration matrix~$\mathtt{K}$, and rigid body transformations $(\mathtt{R}_j, \mathbf{t}_j) \in SE(3), j \in \{ 1, \dots, n_\text{im} \}$ that minimize the cumulative reprojection~error
+Bundle adjustment is then used to obtain the maximum likelihood estimates~$(\hat{\omega}, \hat{\mathtt{K}}, \{(\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j)\}), j \in \{ 1, \dots, n_\text{im} \}$ of the lens distortion model coefficients~$\omega$, calibration matrix~$\mathtt{K}$, and rigid body transformations~$(\mathtt{R}_j, \mathbf{t}_j) \in SE(3), j \in \{ 1, \dots, n_\text{im} \}$ that minimize the cumulative reprojection~error
 \begin{equation}
@@
 A consequence of the approach we take to identifying the 3D points of the 2D-3D correspondences we use for projector calibration is, for each target location, recovery of the local scene plane.
+
+\paragraph{Camera calibration} We recover 2D-3D correspondences in support of calibrating the downwards-facing camera by relying on a planar calibration surface to automatically identify correspondences between the 3D points on the calibration surface and their 2D correspondences in the image plane.
+The classical calibration surface is a chessboard pattern.
+The 3D corner points of the chessboard are obtained \textit{a priori} in a coordinate system defined in the plane of the chessboard\footnote{E.g., $(0,0,0), (1.5,0,0), (3,0,0), \dots, (9,7.5,0)$ for a chessboard with $7~\times{}~6$ corners ($8~\times{}~7$ squares), with each square of length and width of 1.5 unit, respectively.
+Note that the units of the chessboard's 3D points give the units in terms of which the camera calibration is carried out, and---owing to how our projector calibration relies on the camera calibration---the units of the projector calibration as well.}, requiring knowledge only of the dimensions of the chessboard pattern and of the length of a side of a chessboard square.
+The corresponding 2D points are obtained, in the same order, using a specialized algorithm \cite{bradski2000opencv}.
+A set of calibration images is acquired, each with the calibration pattern visible in a different part of the image plane, and such that the center and all corners and edges of the image plane are covered, the camera's autofocus setting be off, and the camera's zoom factor remain fixed.
+2D-3D correspondences are then recovered for each calibration image, and the resulting list is passed on as input to an optimization procedure that relies on bundle adjustment to yield the camera calibration matrix~$\mathtt{K}_\text{cam}$.
 
@@
 \end{figure}
-
-\paragraph{Camera calibration} We recover 2D-3D correspondences in support of calibrating the downwards-facing camera by relying on a planar calibration surface to automatically identify correspondences between the 3D points on the calibration surface and their 2D correspondences in the image plane.
-The classical calibration surface is a chessboard pattern.
-The 3D corner points of the chessboard are obtained \textit{a priori} in a coordinate system defined in the plane of the chessboard\footnote{E.g., $(0,0,0), (1.5,0,0), (3,0,0), \dots, (9,7.5,0)$ for a chessboard with $7~\times{}~6$ corners ($8~\times{}~7$ squares), with each square of length and width of 1.5 unit, respectively.
-Note that the units of the chessboard's 3D points give the units in terms of which the camera calibration is carried out, and---owing to how our projector calibration relies on the camera calibration---the units of the projector calibration as well.}, requiring knowledge only of the dimensions of the chessboard pattern and of the length of a side of a chessboard square.
-The corresponding 2D points are obtained, in the same order, using a specialized algorithm \cite{bradski2000opencv}.
-A set of calibration images is acquired, each with the calibration pattern visible in a different part of the image plane, and such that the center and all corners and edges of the image plane are covered, the camera's autofocus setting be off, and the camera's zoom factor remain fixed.
-2D-3D correspondences are then recovered for each calibration image, and the resulting list is passed on as input to an optimization procedure that relies on bundle adjustment to yield the camera calibration matrix~$\mathtt{K}_\text{cam}$.
 

### 2021-08-01 paper/elsarticle/elsarticle-template.tex
paper

@@
 In carrying out spatial AR, however, care must be taken to appropriately warp the images to be projected in a manner that they appear free of distortions to the viewer.
-For planar scene geometry (such as a floor, wall, or table), this can be done in a cumbersome manual process referred to as keystone correction, often using software bundled with the projector.
+For planar scene geometry (such as a floor, wall, or table), this can be done in a cumbersome manual process called keystone correction, often using software bundled with the projector.
 %, since the geometry of the scene as it relates to the geometry of the projector plays a role in how the pixels of the projector's image plane map to points in the scene
@@
 Augmented reality (AR) \cite{van2010survey,zhou2008trends} promises to enable use cases in industrial settings that include the embedding of assembly instructions directly into the scene \cite{schlund2018moglichkeiten,masood2019augmented,gattullo2019towards,uva2018evaluating,rupprecht2020information,Rupprecht2021}, potentially reducing or altogether doing away with the need for workers to refer to instructions in paper form or on a screen.
-Typically, AR works by embedding the augmentation in an image of the scene acquired from the viewpoint of a single individual, with the resulting augmented image in turn displayed using some form of head-mounted display.
+Typically, AR works by embedding the augmentation in an image of the scene acquired from the viewpoint of a single individual, with the resulting augmented image in turn displayed typically using some form of head-mounted display.
 Reliance on head-mounted displays, however, has two adverse consequences: (i) a head-mounted display must be worn by each individual wishing to partake in the augmentation, and (ii) such a head-mounted display---in some cases taking the form of a helmet in order to house multiple sensors in support of accurately tracking the viewpoint of the viewer relative to the scene---can be obtrusive.
-In turn, \textit{spatial} AR\cite{bimber2019spatial} is a form of augmented reality carried out not by embedding the augmentation in an image of the scene as with a head-mounted display, but by projection to the scene itself, thus eliminating both aforementioned problems.
+In turn, a form of augmented reality referred to as \textit{spatial} AR\cite{bimber2019spatial} is carried out not by embedding the augmentation in an image of the scene as with a head-mounted display, but by projection to the scene itself, thus eliminating both aforementioned problems.
 Yet considering a planar surface to be augmented (e.g., a floor, wall, or table), unless the projector faces the surface frontally, the bounds of a projected rectangular image will not appear rectangular, but will instead be subject to projective distortions.
-Such distortions can be eliminated by carrying out a cumbersome manual process called keystone correction to appropriately warp the image to be projected, often using software bundled with the projector.
+Such distortions can be corrected for by carrying out a cumbersome manual process referred to as keystone correction to appropriately warp the image to be projected, often using software bundled with the projector.
 
@@
 
-More generally, keystone correction for planar scenes can be said to reduce to computing a 2D homography \cite{Hartley2004}, an invertible transformation that preserves colinearity; this is the case whether computing the homography relies on manual interaction as in the case of Pinharez, or is obtained free of manual interaction as in ours.
+More generally, keystone correction for planar scenes can be said to reduce to computing a 2D homography \cite{Hartley2004}, an invertible transformation that preserves colinearity and is thus fittingly alternatively referred to as a `colineation'; this is the case whether computing the homography relies on manual interaction as in the case of Pinharez, or is obtained free of manual interaction as in ours.
 The intuition for why it is that a transformation that preserves colinearity---i.e., maps lines to lines---can serve to model an appropriate corrective image warp can be drawn from considering a planar chessboard pattern: looking at an image of a chessboard acquired from an oblique angle, one observes that lines parallel in the chessboard appear to meet in respective vanishing points; looking at an image of the same chessboard acquired frontally with respect to the plane of the chessboard, lines parallel in the chessboard appear parallel in the image (i.e., they are said to meet `at infinity').
@@
 Such a homography is said to be `induced' by a plane \cite{Hartley2004}.
-We exploit the fact that both cameras and projectors can be modeled using a pinhole camera model \cite{bimber2019spatial}, proceed to recover the relative geometry relating the scene plane and projector, and compute a particular form of plane-induced homography to effect the intended corrective image warp.
+We exploit the fact that cameras and projectors alike can be modeled using a pinhole camera model \cite{bimber2019spatial}, proceed to recover the relative geometry relating the scene plane and projector, and compute a particular form of plane-induced homography to effect the intended corrective image warp.
 
@@
 where $f$ and $(p_x, p_y)^\top$ are both expressed in units of pixels.
-The calibration matrix~$\mathtt{K}$ models a so-called pinhole camera; such a model is applicable to an image acquired using a real-world camera assuming the image has been corrected for lens distortion effects, by having applied the lens distortion model coefficients.
+The calibration matrix~$\mathtt{K}$ models a so-called pinhole camera; such a model is applicable to an image acquired using a real-world camera assuming the image has been corrected for lens distortion effects, by having applied the appropriate lens distortion model coefficients.
 
@@
 \end{equation}
-where $\hat{\mathtt{R}}_j\mathbf{X}_{i,j} + \hat{\mathbf{t}}_j \in \mathbb{R}^3$ expresses $\mathbf{X}_{i,j}$ in the camera coordinate frame of the camera corresponding to the $j^{\text{th}}$~calibration image and $(\hat{\mathbf{x}}_{i,j}'^\top, 1)^\top \sim \hat{\mathtt{K}}(\hat{\mathtt{R}}_j\mathbf{X}_{i,j} + \hat{\mathbf{t}}_j) \in \mathbb{P}^2$ gives the projection~$\hat{\mathbf{x}}_{i,j}' \in \mathbb{R}^2$ of the point~$\mathbf{X}_{i,j}$ to that image.
+where $\hat{\mathtt{R}}_j\mathbf{X}_{i,j} + \hat{\mathbf{t}}_j \in \mathbb{R}^3$ expresses $\mathbf{X}_{i,j}$ in the camera coordinate frame of the camera corresponding to the $j^{\text{th}}$~calibration image and $(\hat{\mathbf{x}}_{i,j}'^\top, 1)^\top \sim \hat{\mathtt{K}}(\hat{\mathtt{R}}_j\mathbf{X}_{i,j} + \hat{\mathbf{t}}_j) \in \mathbb{P}^2$ gives the projection~$\hat{\mathbf{x}}_{i,j}' \in \mathbb{R}^2$ in pixels of the point~$\mathbf{X}_{i,j}$ to that image.
 The function~$d_{\hat{\omega}}$ is a distance function that computes a distance with respect to two pixel positions that it first corrects for lens distortions, in accordance with the lens distortion model coefficients~$\hat{\omega}$ \cite{duane1971close,weng1992camera}.
@@
 For each target location, we placed this piece of cardboard in a manner that its axes were aligned with the principal axes of the scene (cf.\ the yellow and red lines in Figure \ref{fig:eval}(b)); we then manually warped the image for the given target location so that any corner of the augmentation deviated by at most 1 cm from the corresponding corner of the template.
-End-to-end, the process of setting the presets in the MDC-X software having carried out keystone correction manually took ca.\ XXX min for the 15 target locations.
+End-to-end, the process of setting the presets in the MDC-X software having carried out keystone correction manually took ca.\ 32 min for the 15 target locations.
 

### 2021-08-02 paper/elsarticle/elsarticle-template.tex
paper

@@
 In carrying out spatial AR, however, care must be taken to appropriately warp the images to be projected in a manner that they appear free of distortions to the viewer.
-For planar scene geometry (such as a floor, wall, or table), this can be done in a cumbersome manual process called keystone correction, often using software bundled with the projector.
+For planar scene geometry (such as a floor, wall, or table), this can be done in a cumbersome manual process called keystone correction---often using software bundled with the projector---to compute an appropriate corrective image warp.
 %, since the geometry of the scene as it relates to the geometry of the projector plays a role in how the pixels of the projector's image plane map to points in the scene
@@
 This is because the scene point `illuminated' by a pixel in the projector's image plane can be recovered by intersecting its corresponding ray with the geometry of the scene surface.
-To model this interaction, we (i) carry out a one-time projector calibration, which in our approach calls for additionally calibrating a camera facing downward to the scene plane and includes recovery of the scene plane as a convenient side effect.
+To model this interaction, we (i) carry out a one-time projector calibration, which in our approach calls for additionally calibrating a camera facing downwards to the scene plane and includes recovery of the scene plane as a convenient side effect.
 Next, we use the relative camera-projector-scene plane geometry to (ii) compute a plane-induced homography that warps the image to be projected in a manner that it appear undistorted to the viewer, and placed in alignment with the axes of a proxy image of the scene.
@@
     \caption{Scene plane (gray) recovered via spatial resection with respect to 2D-3D correspondences obtained using a chessboard pattern (blue); 3D circle center points of the asymmetrical circle pattern---i.e., the 3D positions of the 2D-3D correspondences to be used for calibrating the projector---obtained by intersection with the scene plane of back-projections (likewise gray) of the 2D circle center points of the asymmetrical circle pattern detected in the image plane (red).
-Note that as in the figures that follow, the rendering in the figure corresponds to the projection calibration image in Figure~\ref{fig:2d}(b), acquired by the downwards-facing camera (frustum of the camera in black, with up vector in red).}
+Note that as in the figures of this section that follow, the rendering in the figure corresponds to the projection calibration image in Figure~\ref{fig:2d}(b), acquired by the downwards-facing camera (frustum of the camera in black, with up vector in red).}
     \label{fig:3d}
@@
 
-Owing to the manner in which we place the virtual projector, the virtual projector's axes and thus the augmentation are aligned with the axes of the downward-facing camera; the placement of the camera thereby intuitively determines the principal axes according to which augmentations are to be placed.
+Owing to the manner in which we place the virtual projector, the virtual projector's axes and thus the augmentation are aligned with the axes of the downwards-facing camera; the placement of the camera thereby intuitively determines the principal axes according to which augmentations are to be placed.
 Note further that a consequence of placing the virtual projector by rotating about the point of intersection of the projector's optical axis with the scene plane is that \textit{the center of the projector's image plane remains invariant} to the placement of the virtual projector, i.e., a steerable mirror can be aimed with respect to a point projected from the center of the projector's image plane, further facilitating placement of augmentations.
@@
 The steerable mirror system was bundled with the MDC-X software for steering the mirror, loading imagery, and optionally carrying out manual keystone correction, such that each position and (warped) image can be registered as a preset.
-In addition, we used a downward-facing Zed 2 stereo camera manufactured by Stereolabs, yet relied only on the left view.
+In addition, we used a downwards-facing Zed 2 stereo camera manufactured by Stereolabs, yet relied only on the left view.
 The floorspace used for our experiments measured dimensions of ca.\ 6~m~$\times$~4~m; the projector was mounted at approximately the center of this space, at a height of ca.\ 3.5 m.
@@
 \paragraph{Our approach} We began by carrying out a calibration of the camera, acquiring 10 camera calibration images (cf.\ Section~\ref{sec:approach:geometry}) of a chessboard calibration pattern with $6~\times~4$ corners ($7~\times~5$ squares) and feeding the images as input to our camera calibration module.
-Separately, for each of the 15 target locations, we produced a projector calibration image (cf.\ again Section~\ref{sec:approach:geometry}) by projecting a $11~\times~4$ asymmetrical circle pattern image to the location in question using the steerable mirror, placing a chessboard pattern beside the projected pattern, and acquiring the image using the downward-facing camera.
+Separately, for each of the 15 target locations, we produced a projector calibration image (cf.\ again Section~\ref{sec:approach:geometry}) by projecting a $11~\times~4$ asymmetrical circle pattern image to the location in question using the steerable mirror, placing a chessboard pattern beside the projected pattern, and acquiring the image using the downwards-facing camera.
 We then fed these images alongside the output of the camera calibration module to our projector calibration module.

### 2021-08-03 paper/elsarticle/elsarticle-template.tex
paper

@@
 \textit{Spatial} AR, in turn, is a form of AR whereby the augmentation of the scene is carried out using a projector, with the advantage of rendering the augmentation visible to all onlookers simultaneously without calling for each to wear some form of head-mounted display.
-In carrying out spatial AR, however, care must be taken to appropriately warp the images to be projected in a manner that they appear free of distortions to the viewer.
-For planar scene geometry (such as a floor, wall, or table), this can be done in a cumbersome manual process called keystone correction---often using software bundled with the projector---to compute an appropriate corrective image warp.
+In carrying out spatial AR, however, care must be taken to appropriately warp an image to be projected in a manner that it, when projected, appear free of distortions to the viewer.
+For planar scene geometry (such as a floor, wall, or table), a cumbersome manual process called keystone correction---often using software bundled with the projector---can be used to carry out an appropriate corrective image warp.
 %, since the geometry of the scene as it relates to the geometry of the projector plays a role in how the pixels of the projector's image plane map to points in the scene
 
-We propose a spatial AR system for wide-area metric augmentation of planar scene surfaces that produces the effect of keystone correction analytically as a function of the relative geometry of the projector and scene plane, by using a projector equipped with a steerable mirror and a camera facing the scene plane.
+We propose a spatial AR system for wide-area metric augmentation of planar scene surfaces that produces the effect of keystone correction analytically as a function of the relative geometry of the projector and scene plane, using a projector equipped with a steerable mirror and a camera facing the scene plane.
 Our system renders the placement of augmentations in the scene more intuitive than manual keystone correction in two ways.
@@
 Such a homography is said to be `induced' by a plane \cite{Hartley2004}.
-We exploit the fact that cameras and projectors alike can be modeled using a pinhole camera model \cite{bimber2019spatial}, proceed to recover the relative geometry relating the scene plane and projector, and compute a particular form of plane-induced homography to effect the intended corrective image warp.
+We exploit the fact that cameras and projectors alike can be modeled using a pinhole camera model \cite{bimber2019spatial}, proceed to recover the relative geometry relating the scene plane and projector, and compute a particular form of plane-induced homography to effect the intended corrective image warp, taking into account both desired in-plane orientation and metric scale.
 
@@
 
+\paragraph{Projector calibration} As in camera calibration, we calibrate the projector by relying on 2D-3D correspondences, yet for projector calibration we obtain them by \textit{projecting} a calibration pattern that we detect using the calibrated downwards-facing camera.
+The pattern we project is one of circles (cf.\ Figure~\ref{fig:2d}(a)), and we rely on an algorithm to detect the circle center points in the asymmetrical circle pattern in the image plane of the projector \cite{bradski2000opencv}, giving the 2D positions of our 2D-3D correspondences for calibrating the projector.
+We project the asymmetrical circle pattern to each of the target locations, and use the the calibrated downwards-facing camera to acquire a projector calibration image for each.
+Given a projector calibration image acquired using camera, we detect the circle centers of the \textit{projected} asymmetrical circle pattern (cf.\ Figure~\ref{fig:2d}(b)); given the scene plane (the recovery of which we shall return to in the paragraph that follows) and such a 2D circle center $\mathbf{x}$, we obtain its 3D correspondence by intersecting the back-projection~$\mathtt{K}_\text{cam}^{-1}(\mathbf{x}^\top, 1)^\top \in \mathbb{P}^2$ of $\mathbf{x}$ with the recovered scene plane (cf.\ Figure~\ref{fig:3d}).
+Since the algorithm that yields 2D circle centers does so in a consistent ordering, we thus obtain the 2D-3D correspondences between the projector's image plane and the scene required for projector calibration, yielding the projector calibration matrix~$\mathtt{K}_\text{proj}$.
+
 \begin{figure}
@@
 
-\paragraph{Projector calibration} As in camera calibration, we calibrate the projector by relying on 2D-3D correspondences, yet for projector calibration we obtain them by \textit{projecting} a calibration pattern that we detect using the calibrated downwards-facing camera.
-The pattern we project is one of circles (cf.\ Figure~\ref{fig:2d}(a)), and we rely on an algorithm to detect the circle center points in the asymmetrical circle pattern in the image plane of the projector \cite{bradski2000opencv}, giving the 2D positions of our 2D-3D correspondences for calibrating the projector.
-We project the asymmetrical circle pattern to each of the target locations, and use the the calibrated downwards-facing camera to acquire a projector calibration image for each.
-Given a projector calibration image acquired using camera, we detect the circle centers of the \textit{projected} asymmetrical circle pattern (cf.\ Figure~\ref{fig:2d}(b)); given the scene plane (the recovery of which we shall return to in the paragraph that follows) and such a 2D circle center $\mathbf{x}$, we obtain its 3D correspondence by intersecting the back-projection~$\mathtt{K}_\text{cam}^{-1}(\mathbf{x}^\top, 1)^\top \in \mathbb{P}^2$ of $\mathbf{x}$ with the recovered scene plane (cf.\ Figure~\ref{fig:3d}).
-Since the algorithm that yields 2D circle centers does so in a consistent ordering, we thus obtain the 2D-3D correspondences between the projector's image plane and the scene required for projector calibration, yielding the projector calibration matrix~$\mathtt{K}_\text{proj}$.
+We recover the scene plane via spatial resection by applying a PnP algorithm \cite{collins2014infinitesimal} to the 2D-3D correspondences obtained using a chessboard pattern.
+Note that this step is separate from camera calibration, yet could well be carried out using the same calibration pattern used in the camera calibration step.\footnote{The critical point is that the pattern should ideally be coplanar with the local scene plane, meaning its height above the scene plane should not exceed a few millimeters.} While a single image of such a chessboard pattern placed on the floor could be sufficient if the floor is even, we place a chessboard pattern in close proximity to the projected asymmetrical circle pattern in each projector calibration image in order to recover the scene plane locally to each target location, in order to account for the possibility of an uneven floor (cf.\ Figure~\ref{fig:2d}(b)).
+Note that in principle, we could project a chessboard pattern instead of an asymmetrical circle pattern to obtain the 2D-3D correspondences needed for projector calibration; it is, however, because we rely on detecting the corners of a chessboard pattern to recover the scene plane that we opt instead for a alternative pattern.
 
@@
 \end{figure}
-
-We recover the scene plane via spatial resection by applying a PnP algorithm \cite{collins2014infinitesimal} to the 2D-3D correspondences obtained using a chessboard pattern.
-Note that this step is separate from camera calibration, yet could well be carried out using the same calibration pattern used in the camera calibration step.\footnote{The critical point is that the pattern should ideally be coplanar with the local scene plane, meaning its height above the scene plane should not exceed a few millimeters.} While a single image of such a chessboard pattern placed on the floor could be sufficient if the floor is even, we place a chessboard pattern in close proximity to the projected asymmetrical circle pattern in each projector calibration image in order to recover the scene plane locally to each target location, in order to account for the possibility of an uneven floor (cf.\ Figure~\ref{fig:2d}(b)).
-Note that in principle, we could project a chessboard pattern instead of an asymmetrical circle pattern to obtain the 2D-3D correspondences needed for projector calibration; it is, however, because we rely on detecting the corners of a chessboard pattern to recover the scene plane that we opt instead for a alternative pattern.
 
@@
     \quad
-    \subfloat[\centering Example projection of image warped for target location using our approach.]{{\includegraphics[height=3.3cm]{images/hans_undist.png}}}
+    \subfloat[\centering Example projection of image warped for target location using our approach.]{{\includegraphics[height=3.3cm]{images/hans_undist_inset.png}}}
     \caption{Evaluation scenario.
@@
 \paragraph{Manual approach} Having pointed the steerable mirror to the desired location, manual keystone correction using the MDC-X software involves warping the image to be projected by manipulating the corners of the image until the desired effect is produced on the projection surface.
-In order to provide a template for manually carrying out keystone correction, we prepared a 50~cm~$\times$~31.25~cm piece of cardboard.
+In order to provide a template for manually carrying out keystone correction, we prepared a rectangular 50~cm~$\times$~31.25~cm cutout of cardboard.
 For each target location, we placed this piece of cardboard in a manner that its axes were aligned with the principal axes of the scene (cf.\ the yellow and red lines in Figure \ref{fig:eval}(b)); we then manually warped the image for the given target location so that any corner of the augmentation deviated by at most 1 cm from the corresponding corner of the template.
@@
 
-A natural extension of this work would be to address non-planar scenes.
+A natural extension of this work would be to address the augmentation of non-planar scenes.
 To handle non-planar scenes would call for a change in how scene geometry is recovered and how warping of the image to be projected is carried out; the methodology we proposed for projector calibration could, however, be left unchanged.

### 2021-08-04 paper/elsarticle/elsarticle-template.tex
paper

@@
 Our contribution is to propose a wide-area spatial AR system for planar scenes that produces the effect of keystone correction analytically, free of any manual interaction.
-Most importantly, we do this in a manner placing the axes of the augmentation in accordance with the axes of a camera placed to face downwards towards the scene plane.
+Most importantly, we do this in a manner placing the axes of the augmentation in accordance with the axes of a camera placed to face directly towards the scene plane (a camera we shall in what follows refer to as `downwards facing', for brevity).
 This facilitates placement of augmentations, since establishing the principal axes of the augmentation thus reduces to placing the downwards-facing camera in a manner such that the $X$- and $Y$-axes of the image align with the intended $X$- and $Y$-axes of the augmentation.

### 2021-08-05 paper/elsarticle/elsarticle-template.tex
paper
Update elsarticle-template.tex

@@
     \centering
-    \subfloat[\centering Asymmetrical circle pattern image, in image plane of projector (detections overlain).]{{\includegraphics[height=3.25cm]{images/circles.png} }}
+    \subfloat[\centering Asymmetrical circle pattern image, in image plane of projector (detections overlaid).]{{\includegraphics[height=3.25cm]{images/circles.png} }}
     \qquad
-    \subfloat[\centering Projector calibration image (one for each target location), in image plane of camera (detections overlain).]{{\includegraphics[height=3.25cm]{images/detections.png}}}
+    \subfloat[\centering Projector calibration image (one for each target location), in image plane of camera (detections overlaid).]{{\includegraphics[height=3.25cm]{images/detections.png}}}
     \caption{Recovering 2D positions in support of projector calibration.
@@
     \centering
-    \subfloat[\centering Oblique view.]{{\includegraphics[height=4.2cm]{images/1.pdf}}}
+    \subfloat[\centering Oblique view.]{{\includegraphics[height=4.2cm]{images/1_big.pdf}}}
     \qquad

### 2021-08-06 paper/elsarticle/elsarticle-template.tex
paper

--- 2021-08-04 paper/elsarticle/elsarticle-template.tex
@@
 Moreover, our system enables specifying the dimensions of augmentations in metric terms, which we achieve by placing the virtual projector at the appropriate height above the scene plane.
-Finally, our system set up to handle a projector equipped with a steerable mirror (without need for explicitly modeling the action of the steerable mirror on the projector), thereby enabling wide-area factory floor applications exceeding the immediate field of view of the projector without needing to rely on multiple projectors.
+Finally, our system set up to handle a projector equipped with a steerable mirror (without need for explicitly modeling the action of the steerable mirror on the projector), thereby enabling wide-area applications exceeding the immediate field of view of the projector without needing to rely on multiple projectors.
 
@@
 
-Calibrating a camera can be carried out by (i) establishing 2D-3D correspondences between pixels in the camera's image plane and the corresponding points in the scene, and on (ii) using those correspondences as input to an optimization procedure that relies on bundle adjustment \cite{triggs1999bundle} to output maximum likelihood estimates of the calibration matrix, the associated lens distortion model coefficients, and, for each calibration image, the pose (i.e., position and orientation) of the camera relative to the 3D scene points from among the 2D-3D correspondences \cite{Hartley2004,zhang2000flexible}.
+A camera can be calibrated by (i) establishing 2D-3D correspondences between pixels in the camera's image plane and the corresponding points in the scene, and on (ii) using those correspondences as input to an optimization procedure that relies on bundle adjustment \cite{triggs1999bundle} to output maximum likelihood estimates of the calibration matrix, the associated lens distortion model coefficients, and, for each calibration image, the pose (i.e., position and orientation) of the camera relative to the 3D scene points from among the 2D-3D correspondences \cite{Hartley2004,zhang2000flexible}.
 Calibration images containing a calibration surface such as a chessboard pattern of known dimensions and scale are acquired from varying viewpoints using the camera, to be used to establish the 2D-3D correspondences~$\{\mathbf{x}_{i,j} \leftrightarrow \mathbf{X}_{i,j}\}$, $i \in \{ 1, \dots, n_\text{pt} \}, j \in \{ 1, \dots, n_\text{im} \}$, where $n_\text{pt}$ gives the number of correspondences obtained from one calibration image of the pattern, $n_\text{im}$ the number of such images, and $\mathbf{X}_{i,j} = \mathbf{X}_{i,j'}$, for $j, j' \in \{ 1, \dots, n_\text{im} \}$.
@@
 \end{equation}
-where $\hat{\mathtt{R}}_j\mathbf{X}_{i,j} + \hat{\mathbf{t}}_j \in \mathbb{R}^3$ expresses $\mathbf{X}_{i,j}$ in the camera coordinate frame of the camera corresponding to the $j^{\text{th}}$~calibration image and $(\hat{\mathbf{x}}_{i,j}'^\top, 1)^\top \sim \hat{\mathtt{K}}(\hat{\mathtt{R}}_j\mathbf{X}_{i,j} + \hat{\mathbf{t}}_j) \in \mathbb{P}^2$ gives the projection~$\hat{\mathbf{x}}_{i,j}' \in \mathbb{R}^2$ in pixels of the point~$\mathbf{X}_{i,j}$ to that image.
+where $\hat{\mathtt{R}}_j\mathbf{X}_{i,j} + \hat{\mathbf{t}}_j \in \mathbb{R}^3$ expresses the point~$\mathbf{X}_{i,j}$ in the local coordinate frame of the camera corresponding to the $j^{\text{th}}$~calibration image and $(\hat{\mathbf{x}}_{i,j}'^\top, 1)^\top \sim \hat{\mathtt{K}}(\hat{\mathtt{R}}_j\mathbf{X}_{i,j} + \hat{\mathbf{t}}_j) \in \mathbb{P}^2$ gives the projection~$\hat{\mathbf{x}}_{i,j}' \in \mathbb{R}^2$ in pixels of $\mathbf{X}_{i,j}$ to that image.
 The function~$d_{\hat{\omega}}$ is a distance function that computes a distance with respect to two pixel positions that it first corrects for lens distortions, in accordance with the lens distortion model coefficients~$\hat{\omega}$ \cite{duane1971close,weng1992camera}.
-The inverse rigid body transformation~$(\hat{\mathtt{R}}_j^{-1}, -\hat{\mathtt{R}}_j^{-1}\hat{\mathbf{t}}_j^{}) \in SE(3)$ gives the pose of the $j^{\text{th}}$~camera relative to the 3D scene points of the calibration pattern.
+Inverting the rigid body transformation~$(\mathtt{R}_j, \mathbf{t}_j)$ gives the pose~$(\hat{\mathtt{R}}_j^{-1}, -\hat{\mathtt{R}}_j^{-1}\hat{\mathbf{t}}_j^{}) \in SE(3)$ of the $j^{\text{th}}$~camera relative to the 3D scene points of the calibration pattern, with $-\hat{\mathtt{R}}_j^{-1}\hat{\mathbf{t}}_j^{} \in \mathbb{R}^3$ the corresponding center of projection.
 
-Calibration of a projector can be carried out in precisely the same manner as calibrating a camera insofar as step (ii) is concerned; the major difference in projector calibration relative to the calibrating a camera concerns the manner in which 2D-3D correspondences are identified, i.e., between pixels in the image plane of the projector and the corresponding points in the scene \cite{moreno2012simple,zhang2009projector,zhang2006novel}.
-What remains of this section is concerned primarily with the recovery of 2D-3D correspondences in support of calibrating cameras and projectors.
+Calibrating a projector \cite{moreno2012simple,zhang2009projector,zhang2006novel} can be carried out in precisely the same manner as calibrating a camera insofar as step (ii) is concerned; the major difference in projector calibration relative to the calibrating a camera concerns the manner in which 2D-3D correspondences are identified, i.e., between pixels in the image plane of the projector and the corresponding points in the scene.
+Since our approach to projector calibration relies on a calibrated camera, what remains of this section is concerned primarily with the recovery of 2D-3D correspondences first in support of calibrating cameras, then of projectors.
 A consequence of the approach we take to identifying the 3D points of the 2D-3D correspondences we use for projector calibration is, for each target location, recovery of the local scene plane.
@@
 The floorspace used for our experiments measured dimensions of ca.\ 6~m~$\times$~4~m; the projector was mounted at approximately the center of this space, at a height of ca.\ 3.5 m.
+At this height, the pixel size (i.e., length or width of a pixel if projected to the scene plane) of the camera was ca.\ 3.5 mm.
 
@@
 
+... At height above ground of 3.5 m, the pixel size was approximately 3.5 mm. Standard deviation of all measured sides was in all instances below the pixel size.
+
 \section{Conclusion}
 
-We presented a spatial AR system for planar scenes that produces the effect of keystone correction analytically, in a manner that enables intuitive placement the augmentations in accordance with the axes of an image of the scene acquired by a downwards-facing camera, and such that the desired dimensions of augmentations can be specified in metric terms.
-Moreover, we showed our system to be able to handle a projector equipped with a steerable mirror, enabling factory floor augmentation exceeding the bounds of the projector's own immediate field of view.
-Our evaluation demonstrated our approach to produce compelling results at less time than the more cumbersome traditional manual approach to keystone correction.
+We presented a spatial AR system for planar scenes that produces the effect of keystone correction analytically as a function of the geometry relating projector and scene plane, exploiting a steerable mirror system to enable augmentation exceeding the bounds of the projector's own immediate field of view.
+Our method produces this effect in a manner that enables intuitive placement of the augmentations in accordance with the $X$- and $Y$-axes of a camera placed to face downwards towards the scene plane.
+Moreover, our method allows for specifying the desired dimensions of augmentations in metric terms, thus enabling consistent scaling.
+Our evaluation demonstrated our approach to produce compelling results---with mean and standard deviation of errors in measured lengths within the downwards-facing camera's pixel size---at less time than a more cumbersome manual approach to keystone correction.
 
-A natural extension of this work would be to address the augmentation of non-planar scenes.
+A natural extension of this work would be to consider the impact of varying calibration patterns, their dimensions, and their scale relative to the camera.
+Still another would be to address the augmentation of non-planar scenes.
 To handle non-planar scenes would call for a change in how scene geometry is recovered and how warping of the image to be projected is carried out; the methodology we proposed for projector calibration could, however, be left unchanged.

### 2021-08-08 paper/elsarticle/elsarticle-template.tex
paper

@@
 \usepackage{amsmath}
+\usepackage{array}
+\usepackage{booktabs}
 \modulolinenumbers[5]
@@
 The intuition for why it is that a transformation that preserves colinearity---i.e., maps lines to lines---can serve to model an appropriate corrective image warp can be drawn from considering a planar chessboard pattern: looking at an image of a chessboard acquired from an oblique angle, one observes that lines parallel in the chessboard appear to meet in respective vanishing points; looking at an image of the same chessboard acquired frontally with respect to the plane of the chessboard, lines parallel in the chessboard appear parallel in the image (i.e., they are said to meet `at infinity').
-To warp the former image (where lines parallel in the scene meet in respective vanishing points) such that the lines of the chessboard appear as in the latter image (where lines parallel in the scene meet at infinity), a transformation that maps lines to lines is sufficient.
+To warp the former image (where lines parallel in the scene meet in respective vanishing points) such that the lines of the chessboard appear as in the latter image (where lines parallel in the scene meet at infinity) calls for a transformation that maps lines to lines.
 
@@
 a form of `plane-induced' 2D homography \cite{Hartley2004}.
-For convenience, we enable optional rotation of the image to be projected \textit{before} applying $\mathtt{H}$, about the image center; that rotation, parameterized in degrees, is thus in effect likewise carried out intuitively in accordance with the placement of the camera.
+For convenience, we enable optional rotation of the image to be projected \textit{before} applying $\mathtt{H}$, about the image center; that rotation, parameterized in degrees, is thus in effect carried out---likewise intuitively---with respect to the ground plane.
 
@@
 The floorspace used for our experiments measured dimensions of ca.\ 6~m~$\times$~4~m; the projector was mounted at approximately the center of this space, at a height of ca.\ 3.5 m.
-At this height, the pixel size (i.e., length or width of a pixel if projected to the scene plane) of the camera was ca.\ 3.5 mm.
+At this height, the pixel size (i.e., length or width of a pixel if projected to the scene plane) of the camera was 3.5 mm.
 
@@
 End-to-end, the process of setting the presets in the MDC-X software having carried out keystone correction manually took ca.\ 32 min for the 15 target locations.
+
+\begin{table}[ht]
+\caption{Deviations in measured lengths from 50~cm~$\times$~31.25~cm across the 15 augmentations produced using our approach, provided in cm and in pixel units of the downwards-facing camera (pixel size 3.5 mm).
+Note that in no instance do mean and standard deviation exceed 1 pixel from the viewpoint of the camera.}
+\label{table:length}
+\centering
+{\small
+\begin{tabularx}{7.4cm}{r p{1cm} p{1cm} p{1cm} p{1cm}}
+\toprule
+  & Top & Bottom & Left & Right \\
+\midrule
+Mean {\tiny[cm]} & 0.03 & 0.07 & 0.09 & 0.06 \\
+Stdev {\tiny[cm]} & 0.33 & 0.17 & 0.17 & 0.17 \\
+\midrule
+Mean {\tiny[px]} & 0.1 & 0.19 & 0.26 & 0.18 \\
+Stdev {\tiny[px]} & 0.95 & 0.48 & 0.49 & 0.49 \\
+\bottomrule
+\end{tabularx}}
+\end{table}
+
+\begin{table}[ht]
+\caption{Deviations in measured angles from $90^\circ{}$ across the 15 augmentations produced using our approach.}
+\label{table:angle}
+\centering
+{\small
+\begin{tabularx}{6.4cm}{r p{1.75cm} p{1.75cm}}
+\toprule
+  & Top Right & Bottom Left\\
+\midrule
+Mean {\tiny[deg]} & -0.03 & 0.0 \\
+Stdev {\tiny[deg]} & 0.4 & 0.34 \\
+\bottomrule
+\end{tabularx}}
+\end{table}
 
@@
 Next, we warped the images to be projected to the respective locations using their corresponding homography, using a third dedicated custom module.
-These warped images were finally imported into the MDC-X software and associated with their respective location presets (cf.\ Figure \ref{fig:eval}(b)).
+These warped images were finally imported into the MDC-X software and associated with their respective location presets (cf.\ Figure~\ref{fig:eval}(b)).
 
-The total amount of time to carry out all the above steps amounted to ca.\ 20 min, with ca.\ 2 min going to acquisition of the camera calibration images, and ca.\ 5 min going to that of projector calibration images.
+Tables~\ref{table:length} and \ref{table:angle} give the deviations in measured lengths and angles, respectively, from the intended quantities.
+Lengths were measured using a measuring tape for the top, bottom, left, and right sides of the augmentations; angles were measured using an angle gauge at the augmentations' top-left and bottom-right corners (the measuring tape and angle gauge are both visible in Figure~\ref{fig:eval}(b)).
+Note that if deviations in measured lengths are expressed in units of pixels of the downwards-facing camera, mean and standard deviation in no instance exceed 1 pixel.
+The total amount of time expended for carrying out all the above steps amounted to ca.\ 20 min, with ca.\ 2 min going to acquisition of the camera calibration images, and ca.\ 5 min going to that of projector calibration images.
 The remainder of the time was spent running our modules or working with the MDC-X software.
-Note that once the camera is calibrated, that calibration can be reused if the camera's intrinsics remain fixed, in particular if no change is made to the zoom factor of the camera.
-
-... At height above ground of 3.5 m, the pixel size was approximately 3.5 mm. Standard deviation of all measured sides was in all instances below the pixel size.
+Note that once the camera is calibrated, that calibration can be reused if the camera's intrinsics remain fixed, in particular if no change is made to the zoom factor or focus settings of the camera.
 

### 2021-08-10 paper/elsarticle/elsarticle-template.tex
paper

@@
 Correcting for projective distortions of the sort outlined in Section~\ref{sec:intro} can be achieved by modeling the manner in which the respective rays through the pixels of the projector's image plane fan out into the scene (i.e., by `calibrating' the projector) and the geometry of the scene itself (i.e., by recovering the scene plane relative to the projector) within at least the projector's field of view.
-This is because the scene point `illuminated' by a pixel in the projector's image plane can be recovered by intersecting its corresponding ray with the geometry of the scene surface.
+This is because the scene point `illuminated' by a pixel in the projector's image plane can be recovered by intersecting its corresponding ray with the geometry of the scene surface; modeling the geometry that relates projector and scene plane thus enables one to reason about the problem of correcting for distortions in terms of geometry.
 To model this interaction, we (i) carry out a one-time projector calibration, which in our approach calls for additionally calibrating a camera facing downwards to the scene plane and includes recovery of the scene plane as a convenient side effect.
@@
 
-\paragraph{Virtual projector} The placement of the virtual projector is to determine from which pose the image to be projected is to \textit{appear} to have been projected.
-We carry out this placement according to a small handful of steps.
-First, we (i) rotate the camera about its center of projection to align its optical axis\footnote{Strictly speaking, the optical axis is the back-projection of the principal point; in our usage, we understand it to refer to the back-projection of the center of the image to be projected.} with the normal vector of the scene plane, giving a virtual camera likewise facing directly\footnote{A physical camera placed to face downwards is almost certain to not face downwards precisely; in contrast, the virtual camera's optical axis is aligned exactly with the scene plane's normal vector, rendering it genuinely fronto-parallel with respect to the scene plane.} downwards to the scene plane (cf.\ Figure~\ref{fig:virtualproj}).
-Next, we (ii) intersect the scene plane with the optical axis (i.e., the ray from the projector's center of projection through the center of the image plane) and rotate the projector's placement about that point of intersection, aligning the optical axis with the scene plane's normal vector and giving an initial virtual projector.
-Finally, we (iii) align the $X$- and $Y$-axes of the initial virtual projector with those of the virtual camera, which gives the virtual projector (cf.\ again Figure~\ref{fig:virtualproj}).
-The virtual projector is thus rendered fronto-parallel with the scene plane, enabling projection to the scene plane absent of projective distortions.
-We additionally (iv) adjust the height above the scene plane of the virtual projector, in order to satisfy desired projected image dimensions provided in metric units.
-
-Owing to the manner in which we place the virtual projector, the virtual projector's axes and thus the augmentation are aligned with the axes of the downwards-facing camera; the placement of the camera thereby intuitively determines the principal axes according to which augmentations are to be placed.
-Note further that a consequence of placing the virtual projector by rotating about the point of intersection of the projector's optical axis with the scene plane is that \textit{the center of the projector's image plane remains invariant} to the placement of the virtual projector, i.e., a steerable mirror can be aimed with respect to a point projected from the center of the projector's image plane, further facilitating placement of augmentations.
-
 \begin{figure}
@@
 
-\section{Evaluation}
+\paragraph{Virtual projector} The placement of the virtual projector is to determine from which pose the image to be projected is to \textit{appear} to have been projected.
+We carry out this placement according to a small handful of steps.
+First, we (i) rotate the camera about its center of projection to align its optical axis\footnote{Strictly speaking, the optical axis is the back-projection of the principal point; in our usage, we understand it to refer to the back-projection of the center of the image to be projected.} with the normal vector of the scene plane, giving a virtual camera likewise facing directly\footnote{A physical camera placed to face downwards is almost certain to not face downwards precisely; in contrast, the virtual camera's optical axis is aligned exactly with the scene plane's normal vector, rendering it genuinely fronto-parallel with respect to the scene plane.} downwards to the scene plane (cf.\ Figure~\ref{fig:virtualproj}).
+Next, we (ii) intersect the scene plane with the optical axis (i.e., the ray from the projector's center of projection through the center of the image plane) and rotate the projector's placement about that point of intersection, aligning the optical axis with the scene plane's normal vector and giving an initial virtual projector.
+Finally, we (iii) align the $X$- and $Y$-axes of the initial virtual projector with those of the virtual camera, which gives the virtual projector (cf.\ again Figure~\ref{fig:virtualproj}).
+The virtual projector is thus rendered fronto-parallel with the scene plane, enabling projection to the scene plane absent of projective distortions.
+We additionally (iv) adjust the height above the scene plane of the virtual projector, in order to satisfy desired projected image dimensions provided in metric units.
+
+Owing to the manner in which we place the virtual projector, the virtual projector's axes and thus the augmentation are aligned with the axes of the downwards-facing camera; the placement of the camera thereby intuitively determines the principal axes according to which augmentations are to be placed.
+Note further that a consequence of placing the virtual projector by rotating about the point of intersection of the projector's optical axis with the scene plane is that \textit{the center of the projector's image plane remains invariant} to the placement of the virtual projector, i.e., a steerable mirror can be aimed with respect to a point projected from the center of the projector's image plane, further facilitating placement of augmentations.
 
@@
 
+\section{Evaluation}
 We evaluate our approach by using a 960 pixel~$\times$~600 pixel image of the TU Wien logo (cf.\ Figure \ref{fig:warp}) to augment 15 locations across the floorspace at the Pilotfabrik\footnote{\url{https://www.pilotfabrik.at/}} of TU Wien, a collaborative space for research on Industry 4.0 topics situated in Vienna, Austria.
@@
 {\small
-\begin{tabularx}{6.4cm}{r p{1.75cm} p{1.75cm}}
+\begin{tabularx}{6.25cm}{r p{1.725cm} p{1.725cm}}
 \toprule
-  & Top Right & Bottom Left\\
+  & Top Left & Bottom Right \\
 \midrule

### 2021-08-12 paper/elsarticle/elsarticle-template.tex
paper

@@
 In carrying out spatial AR, however, care must be taken to appropriately warp an image to be projected in a manner that it, when projected, appear free of distortions to the viewer.
-For planar scene geometry (such as a floor, wall, or table), a cumbersome manual process called keystone correction---often using software bundled with the projector---can be used to carry out an appropriate corrective image warp.
+For planar scene geometry (such as a floor, wall, or table), a cumbersome manual process called keystone correction---involving warping an image by perturbing its corners until the desired effect is produced when projected to the scene---can be used to carry out an appropriate corrective image warp.
 %, since the geometry of the scene as it relates to the geometry of the projector plays a role in how the pixels of the projector's image plane map to points in the scene
@@
 
+%Automatic Calibration of a Multiple-Projector Spherical Fish Tank VR Display
+
 \section{Approach}
@@
 Calibration images containing a calibration surface such as a chessboard pattern of known dimensions and scale are acquired from varying viewpoints using the camera, to be used to establish the 2D-3D correspondences~$\{\mathbf{x}_{i,j} \leftrightarrow \mathbf{X}_{i,j}\}$, $i \in \{ 1, \dots, n_\text{pt} \}, j \in \{ 1, \dots, n_\text{im} \}$, where $n_\text{pt}$ gives the number of correspondences obtained from one calibration image of the pattern, $n_\text{im}$ the number of such images, and $\mathbf{X}_{i,j} = \mathbf{X}_{i,j'}$, for $j, j' \in \{ 1, \dots, n_\text{im} \}$.
-Bundle adjustment is then used to obtain the maximum likelihood estimates~$(\hat{\omega}, \hat{\mathtt{K}}, \{(\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j)\}), j \in \{ 1, \dots, n_\text{im} \}$ of the lens distortion model coefficients~$\omega$, calibration matrix~$\mathtt{K}$, and rigid body transformations~$(\mathtt{R}_j, \mathbf{t}_j) \in SE(3), j \in \{ 1, \dots, n_\text{im} \}$ that minimize the cumulative reprojection~error
+Bundle adjustment is then used to obtain the maximum likelihood estimates~$(\hat{\omega}, \hat{\mathtt{K}}, \{(\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j)\}), j \in \{ 1, \dots, n_\text{im} \}$ of the lens distortion model coefficients~$\omega$, calibration matrix~$\mathtt{K}$, and rigid body transformations~$(\mathtt{R}_j, \mathbf{t}_j) \in SE(3), j \in \{ 1, \dots, n_\text{im} \}$.
+These are obtained using bundle adjustment by minimizing the cumulative reprojection error
 \begin{equation}
@@
 The pose of the projector, for each projector calibration image, is provided relative to the 3D points of the pattern---and thus in the coordinate frame of the camera---alongside $\mathtt{K}_\text{proj}$ by the aforementioned optimization procedure.
-Note that for a fixed projector with steerable mirror, given a projector calibration image, the recovered projector's pose is the pose the projector would have to have had to project to the given target location \textit{in the absence of the mirror}.
+Note that for a fixed projector with steerable mirror, given a projector calibration image, the recovered projector's pose is the pose the projector would have to have had to project to the given target location \textit{in the absence of the mirror} (i.e., facing directly towards the target location).
 This is sufficient for our needs in Section~\ref{sec:approach:homography}, and it is in this sense that our system is able to handle a projector equipped with a steerable mirror, without need for modeling the steerable mirror explicitly.
@@
 The floorspace used for our experiments measured dimensions of ca.\ 6~m~$\times$~4~m; the projector was mounted at approximately the center of this space, at a height of ca.\ 3.5 m.
-At this height, the pixel size (i.e., length or width of a pixel if projected to the scene plane) of the camera was 3.5 mm.
+At this height, the pixel size (i.e., length or width of a pixel if projected to the scene plane) of the camera was ca.\ 3.5 mm.
 
@@
 {\small
-\begin{tabularx}{6.25cm}{r p{1.725cm} p{1.725cm}}
+\begin{tabularx}{6.75cm}{r p{1.95cm} p{1.95cm}}
 \toprule

### 2021-08-27 paper/elsarticle/elsarticle-template.tex
paper

@@
 This facilitates placement of augmentations, since establishing the principal axes of the augmentation thus reduces to placing the downwards-facing camera in a manner such that the $X$- and $Y$-axes of the image align with the intended $X$- and $Y$-axes of the augmentation.
-We achieve this by warping the image to be projected using a plane-induced homography computed to produce the effect of projecting the image not from the actual projector viewpoint, but in accordance with the viewpoint of a \textit{virtual} projector (i) facing directly downwards to the target location in the scene plane and (ii) rotated to place the axes of the virtual projector in accordance with those of the camera.
+We achieve this by warping the image to be projected using a plane-induced homography computed to produce the effect of projecting the image not from the actual projector viewpoint, but in accordance with the viewpoint of a \textit{virtual} projector.
+We place this virtual projector to (i) face directly downwards to the target location in the scene plane and (ii) rotate its axes in accordance with those of the camera.
 Moreover, our system enables specifying the dimensions of augmentations in metric terms, which we achieve by placing the virtual projector at the appropriate height above the scene plane.
@@
     \caption{Recovering 2D positions in support of projector calibration.
-(a) 2D positions of the 2D-3D correspondences to be used for calibrating the projector are obtained by detecting---in the image plane of the projector---the circle centers in the asymmetrical circle pattern image, to be projected to each of the target locations in the scene plane.
+(a) 2D positions of the 2D-3D correspondences to be used for calibrating the projector are obtained by detecting---in the image plane of the projector---the circle centers in the asymmetrical circle pattern image.
+It is this pattern that we project to each of the target locations in the scene plane.
 (b) For each such target location, an image is acquired from the viewpoint of the camera and the circle centers of the projected asymmetrical circle pattern are detected, in the image plane of the camera.
@@
 We recover the scene plane via spatial resection by applying a PnP algorithm \cite{collins2014infinitesimal} to the 2D-3D correspondences obtained using a chessboard pattern.
-Note that this step is separate from camera calibration, yet could well be carried out using the same calibration pattern used in the camera calibration step.\footnote{The critical point is that the pattern should ideally be coplanar with the local scene plane, meaning its height above the scene plane should not exceed a few millimeters.} While a single image of such a chessboard pattern placed on the floor could be sufficient if the floor is even, we place a chessboard pattern in close proximity to the projected asymmetrical circle pattern in each projector calibration image in order to recover the scene plane locally to each target location, in order to account for the possibility of an uneven floor (cf.\ Figure~\ref{fig:2d}(b)).
+Note that this step is separate from camera calibration, yet could well be carried out using the same calibration pattern used in the camera calibration step.\footnote{The critical point is that the pattern should ideally be coplanar with the local scene plane, meaning its height above the scene plane should not exceed a few millimeters.} While a single image of such a chessboard pattern placed on the floor could be sufficient if the floor is even, we account for the possibility of an uneven floor by placing a chessboard pattern in close proximity to the projected asymmetrical circle pattern in each projector calibration image in order to recover the scene plane locally to each target location (cf.\ Figure~\ref{fig:2d}(b)).
 Note that in principle, we could project a chessboard pattern instead of an asymmetrical circle pattern to obtain the 2D-3D correspondences needed for projector calibration; it is, however, because we rely on detecting the corners of a chessboard pattern to recover the scene plane that we opt instead for a alternative pattern.

### 2021-08-27 paper/elsarticle/elsarticle-template.tex
paper

@@
 Our system renders the placement of augmentations in the scene more intuitive than manual keystone correction in two ways.
-First, (i) placement of the desired augmentations is carried out in accordance with the axes of an image of the scene acquired by the camera, thereby making setting those axes as simple as appropriately rotating the camera.
+First, (i) placement of the desired augmentations is carried out in accordance with the axes of an image of the scene acquired by the camera, thereby making setting those axes as simple as appropriately rotating the camera relative to the scene.
 Second, (ii) the desired dimensions of the projected augmentations are specified in metric terms, thereby allowing for consistent scaling across all target locations.

### 2021-09-02 paper/elsarticle/elsarticle-template.tex
paper

@@
 \begin{abstract}
-Augmented reality (AR) promises to enable use cases in industrial settings that include the embedding of assembly instructions directly into the scene, potentially reducing or altogether obviating the need for workers to refer to instructions in paper form or on a screen.
-\textit{Spatial} AR, in turn, is a form of AR whereby the augmentation of the scene is carried out using a projector, with the advantage of rendering the augmentation visible to all onlookers simultaneously without calling for each to wear some form of head-mounted display.
+Augmented reality (AR) promises to enable use cases in industrial settings that include the embedding of assembly instructions directly into the scene, potentially reducing or altogether obviating the need for workers to refer to such instructions in paper form or on a statically situated screen.
+\textit{Spatial} AR, in turn, is a form of AR whereby the augmentation of the scene is carried out using a projector, with the advantage of rendering the augmentation visible to all onlookers simultaneously without calling for any to hold a handheld device such as a tablet or for each to wear some form of head-mounted display.
 In carrying out spatial AR, however, care must be taken to appropriately warp an image to be projected in a manner that it, when projected, appear free of distortions to the viewer.
-For planar scene geometry (such as a floor, wall, or table), a cumbersome manual process called keystone correction---involving warping an image by perturbing its corners until the desired effect is produced when projected to the scene---can be used to carry out an appropriate corrective image warp.
-%, since the geometry of the scene as it relates to the geometry of the projector plays a role in how the pixels of the projector's image plane map to points in the scene
+For planar scene geometry (such as a floor, wall, or table), a cumbersome manual process called keystone correction can be used to carry out an appropriate corrective image warp.
+Another drawback of conventional spatial AR relying only on a projector is that it is capable of augmenting only the portion of the scene within the projector's own static field of view, thereby hindering its applicability to use cases calling for augmentation of wide areas such as a factory floorspace.
 
-We propose a spatial AR system for wide-area metric augmentation of planar scene surfaces that produces the effect of keystone correction analytically as a function of the relative geometry of the projector and scene plane, using a projector equipped with a steerable mirror and a camera facing the scene plane.
+We propose a spatial AR system for wide-area metric augmentation of planar scene surfaces that produces the effect of keystone correction analytically as a function of the relative geometry of the projector and scene plane, using a projector equipped with a steerable mirror to direct the projection across varying target locations and a camera facing the scene plane in support of calibrating the system.
 Our system renders the placement of augmentations in the scene more intuitive than manual keystone correction in two ways.
@@
 
-Augmented reality (AR) \cite{van2010survey,zhou2008trends} promises to enable use cases in industrial settings that include the embedding of assembly instructions directly into the scene \cite{schlund2018moglichkeiten,masood2019augmented,gattullo2019towards,uva2018evaluating,rupprecht2020information,Rupprecht2021}, potentially reducing or altogether doing away with the need for workers to refer to instructions in paper form or on a screen.
-Typically, AR works by embedding the augmentation in an image of the scene acquired from the viewpoint of a single individual, with the resulting augmented image in turn displayed typically using some form of head-mounted display.
-Reliance on head-mounted displays, however, has two adverse consequences: (i) a head-mounted display must be worn by each individual wishing to partake in the augmentation, and (ii) such a head-mounted display---in some cases taking the form of a helmet in order to house multiple sensors in support of accurately tracking the viewpoint of the viewer relative to the scene---can be obtrusive.
-In turn, a form of augmented reality referred to as \textit{spatial} AR\cite{bimber2019spatial} is carried out not by embedding the augmentation in an image of the scene as with a head-mounted display, but by projection to the scene itself, thus eliminating both aforementioned problems.
-Yet considering a planar surface to be augmented (e.g., a floor, wall, or table), unless the projector faces the surface frontally, the bounds of a projected rectangular image will not appear rectangular, but will instead be subject to projective distortions.
-Such distortions can be corrected for by carrying out a cumbersome manual process referred to as keystone correction to appropriately warp the image to be projected, often using software bundled with the projector.
+Augmented reality (AR) \cite{van2010survey,zhou2008trends} promises to enable use cases in industrial settings that include the embedding of assembly instructions directly into the scene~\cite{schlund2018moglichkeiten,uva2018evaluating,masood2019augmented,gattullo2019towards,aschenbrenner2019comparing,mayrhofer2019one,rupprecht2020information,Rupprecht2021}, potentially reducing or altogether doing away with the need for workers to refer to such instructions in paper form or on a statically situated screen.
+Typically, AR works by embedding the augmentation in an image of the scene acquired from the viewpoint of a single individual, with the resulting augmented image in turn displayed typically using some form of handheld or head-mounted display.
+Handheld devices such as tablets do not enable the user to work hands-free while using the device, or force the user to wear additional gear.
+Reliance on head-mounted displays has two adverse consequences of its own: a head-mounted display must be worn by each individual wishing to partake in the augmentation, and such a head-mounted display---in some cases taking the form of a helmet in order to house multiple sensors in support of accurately tracking the viewpoint of the viewer relative to the scene---can be obtrusive.
+A Hololens 2, for instance, is such a head-mounted display; at a weight of 566~grams,\footnote{\url{https://www.microsoft.com/en-us/hololens/hardware}} it places an additional ergonomic load on the worker, in particular if worn over a longer duration or in bent positions.
+
+In turn, a form of augmented reality referred to as \textit{spatial} AR\cite{bimber2019spatial} is carried out not by embedding the augmentation in an image of the scene as with a head-mounted or handheld display, but by projection to the scene itself, thus eliminating the aforementioned problems.
+Spatial AR thus turns surfaces such as the product being manufactured, a desk, the floor, a wall, or even the ceiling (considering overhead work) into displays.
+Information provision---of the work plan or of quality-related information---can be can be carried out by projection to the given workspace, while directly referring to exact positions within that workspace.
+Moreover, the fact that this information can be viewed by multiple users simultaneously serves to promote collaboration \cite{aschenbrenner2019comparing}.
+
+While interest in spatial AR in industrial settings has been growing, its deployment in industrial settings remains well behind that of wearable AR systems such as handhelds and head-mounted displays.
+One of the main shortcomings of typical projector-based spatial AR relative to AR relying on wearables is the applicability of spatial AR only to augmenting scenes within the projector's own static field of view.
+Here, dynamic projection systems, using mirror heads to deflect the projection across wider areas has the potential to bridge the gap to wider industrial penetration.
+Yet the increase in the area to be covered comes with noteworthy challenges in terms of ensuring that the images to be projected appear free of distortions to the users.
+Considering a planar surface to be augmented (e.g., a floor, wall, or table), unless the projector faces the surface frontally, the bounds of a projected rectangular image will not appear rectangular, but will instead be subject to projective distortions.
+Such distortions can be corrected for by carrying out a cumbersome manual process referred to as keystone correction---involving warping an image by perturbing its corners until the desired effect is produced when projected to the scene---to appropriately warp the image to be projected, often using software bundled with the projector.
+Automating this manual, time-consuming, and error-prone process significantly reduces the time needed for the industrial engineering of a spatial AR application and therefore has potential to promote adoption by industry.
+Especially manufacturing processes of large products that remain at a static mounting position during the assembly process---i.e., industrial site assembly---offer a significant application potential for the use of spatial AR as part of worker assistance systems \cite{mayrhofer2019one}.
 
@@
 We exploit the fact that cameras and projectors alike can be modeled using a pinhole camera model \cite{bimber2019spatial}, proceed to recover the relative geometry relating the scene plane and projector, and compute a particular form of plane-induced homography to effect the intended corrective image warp, taking into account both desired in-plane orientation and metric scale.
-
-%Mention asymmetrical circle pattern: \cite{chiu2011novel}
-
-%13. X. Zhang and L. Zhu, “Projector calibration from the camera
-%image point of view,” Opt. Eng. 48, 117208 (2009).
-%14. X. Chen, J. Xi, Y. Jin, and J. Sun, “Accurate calibration
-%for a camera-projector measurement system based on
-%structured light projection,” Opt. Laser Eng. 47, 310–319
-%(2009).
-%15. I. Din, H. Anwar, I. Syed, H. Zafar, and L. Hasan, “Projector
-%calibration for pattern projection systems,” J. Appl. Res.
-%Technol. 12, 80–86 (2014).
-%16. R. Orghidan, J. Salvi, M. Gordan, C. Florea, and J. Batlle,
-%“Structured light self-calibration with vanishing points,”
-%Mach. Vision Appl. 25, 489–500 (2014).
-%17. J. Huang, Z. Wang, Q. Xue, and J. Gao, “Calibration of a
-%camera-projector measurement system and error impact
-%analysis,” Meas. Sci. Technol. 23, 125402 (2012).
-%18. W. Gao, L. Wang, and Z. Hu, “Flexible calibration of a portable
-%structured light system through surface plane,” Acta Autom.
-%Sinica 34, 1358–1362 (2008).
-%19. G. Falcao, N. Hurtos, J. Massich, and D. Fofi, “Projectorcamera calibration toolbox,” Technology Report, 2009, available at http://code.google.com/p/procamcalib.
-%20. S. Zhang and P. Huang, “Novel method for structured light
-%system calibration,” Opt. Eng. 45, 083601 (2006).
-%21. F. Berry, O. Aider, and J. Mosnier, “A visual servoing-based
-%method for ProCam system calibration,” Sensors 13, 13318–
-%13333 (2013).
-%22. H. Anwar, I. Din, and K. Park, “Projector calibration for 3D
-%scanning using virtual target images,” Int. J. Precis. Eng.
-%Manuf. 13, 125–131 (2012).
-
-%todo: It consists of a projector and a small mirror with 2 axes of rotation.
-In contrast with the present work, Everywhere Displays is limited to projecting on a discrete set of planar surfaces which are calibrated beforehand, each modeled by a 2D homography.
-The Everywhere Displays project explored many applications of steerable displays, particularly in the retail domain.
-With the addition of a pan and tilt video camera, a number of gesture-based interactions are demonstrated, including touch-based manipulation of projected graphical widgets [12], as well as person tracking to enable graphics to follow the user.
-A small number of later works contribute steerable displays designs which mount the projector and camera directly onto motorized platform with pan and tilt axes of motion [8, 6, 11].
-Butz et al.
-[7] contribute a number of interactions in the tradition of ubiquitous computing, such as the ability to
-
-%Automatic Calibration of a Multiple-Projector Spherical Fish Tank VR Display
 
@@
 
+\begin{figure}
+    \centering
+    \subfloat[\centering Asymmetrical circle pattern image, in image plane of projector (detections overlaid).]{{\includegraphics[height=3.25cm]{images/circles.png} }}
+    \qquad
+    \subfloat[\centering Projector calibration image (one for each target location), in image plane of camera (detections overlaid).]{{\includegraphics[height=3.25cm]{images/detections.png}}}
+    \caption{Recovering 2D positions in support of projector calibration.
+(a) 2D positions of the 2D-3D correspondences to be used for calibrating the projector are obtained by detecting---in the image plane of the projector---the circle centers in the asymmetrical circle pattern image.
+It is this pattern that we project to each of the target locations in the scene plane.
+(b) For each such target location, an image is acquired from the viewpoint of the camera and the circle centers of the projected asymmetrical circle pattern are detected, in the image plane of the camera.
+A chessboard pattern to be used for recovering the local scene plane is placed near the projected pattern, whose corners are likewise detected.
+Detected 2D projected asymmetrical circle pattern circle center points and chessboard corners overlain for illustration.}
+    \label{fig:2d}
+\end{figure}
+
 \paragraph{Camera calibration} We recover 2D-3D correspondences in support of calibrating the downwards-facing camera by relying on a planar calibration surface to automatically identify correspondences between the 3D points on the calibration surface and their 2D correspondences in the image plane.
@@
 \begin{figure}
-    \centering
-    \subfloat[\centering Asymmetrical circle pattern image, in image plane of projector (detections overlaid).]{{\includegraphics[height=3.25cm]{images/circles.png} }}
-    \qquad
-    \subfloat[\centering Projector calibration image (one for each target location), in image plane of camera (detections overlaid).]{{\includegraphics[height=3.25cm]{images/detections.png}}}
-    \caption{Recovering 2D positions in support of projector calibration.
-(a) 2D positions of the 2D-3D correspondences to be used for calibrating the projector are obtained by detecting---in the image plane of the projector---the circle centers in the asymmetrical circle pattern image.
-It is this pattern that we project to each of the target locations in the scene plane.
-(b) For each such target location, an image is acquired from the viewpoint of the camera and the circle centers of the projected asymmetrical circle pattern are detected, in the image plane of the camera.
-A chessboard pattern to be used for recovering the local scene plane is placed near the projected pattern, whose corners are likewise detected.
-Detected 2D projected asymmetrical circle pattern circle center points and chessboard corners overlain for illustration.}
-    \label{fig:2d}
+    \centerline{\includegraphics[scale=.35]{images/2d3d.png}}
+    \caption{Scene plane (gray) recovered via spatial resection with respect to 2D-3D correspondences obtained using a chessboard pattern (blue); 3D circle center points of the asymmetrical circle pattern---i.e., the 3D positions of the 2D-3D correspondences to be used for calibrating the projector---obtained by intersection with the scene plane of back-projections (likewise gray) of the 2D circle center points of the asymmetrical circle pattern detected in the image plane (red).
+Note that as in the figures of this section that follow, the rendering in the figure corresponds to the projection calibration image in Figure~\ref{fig:2d}(b), acquired by the downwards-facing camera (frustum of the camera in black, with up vector in red).}
+    \label{fig:3d}
 \end{figure}
@@
 Note that in principle, we could project a chessboard pattern instead of an asymmetrical circle pattern to obtain the 2D-3D correspondences needed for projector calibration; it is, however, because we rely on detecting the corners of a chessboard pattern to recover the scene plane that we opt instead for a alternative pattern.
-
-\begin{figure}
-    \centerline{\includegraphics[scale=.35]{images/2d3d.png}}
-    \caption{Scene plane (gray) recovered via spatial resection with respect to 2D-3D correspondences obtained using a chessboard pattern (blue); 3D circle center points of the asymmetrical circle pattern---i.e., the 3D positions of the 2D-3D correspondences to be used for calibrating the projector---obtained by intersection with the scene plane of back-projections (likewise gray) of the 2D circle center points of the asymmetrical circle pattern detected in the image plane (red).
-Note that as in the figures of this section that follow, the rendering in the figure corresponds to the projection calibration image in Figure~\ref{fig:2d}(b), acquired by the downwards-facing camera (frustum of the camera in black, with up vector in red).}
-    \label{fig:3d}
-\end{figure}
 

### 2021-09-06 paper/elsarticle/elsarticle-template.tex
paper

@@
 Reliance on head-mounted displays has two adverse consequences of its own: a head-mounted display must be worn by each individual wishing to partake in the augmentation, and such a head-mounted display---in some cases taking the form of a helmet in order to house multiple sensors in support of accurately tracking the viewpoint of the viewer relative to the scene---can be obtrusive.
-A Hololens 2, for instance, is such a head-mounted display; at a weight of 566~grams,\footnote{\url{https://www.microsoft.com/en-us/hololens/hardware}} it places an additional ergonomic load on the worker, in particular if worn over a longer duration or in bent positions.
+A HoloLens 2, for instance, is such a head-mounted display; at a weight of 566~grams \cite{hololens}, it places an additional ergonomic load on the worker, in particular if worn over a longer duration or in bent positions.
 
@@
 Spatial AR thus turns surfaces such as the product being manufactured, a desk, the floor, a wall, or even the ceiling (considering overhead work) into displays.
-Information provision---of the work plan or of quality-related information---can be can be carried out by projection to the given workspace, while directly referring to exact positions within that workspace.
+Information provision---of the work plan or of quality-related information---can be carried out by projection to the given workspace, while directly referring to exact positions within that workspace.
 Moreover, the fact that this information can be viewed by multiple users simultaneously serves to promote collaboration \cite{aschenbrenner2019comparing}.

### 2021-09-09 paper/elsarticle/elsarticle-template.tex
Update elsarticle-template.tex

@@
 In carrying out spatial AR, however, care must be taken to appropriately warp an image to be projected in a manner that it, when projected, appear free of distortions to the viewer.
-For planar scene geometry (such as a floor, wall, or table), a cumbersome manual process called keystone correction can be used to carry out an appropriate corrective image warp.
+For planar scene geometry (such as a floor, wall, or table), a manual process called keystone correction can be used to carry out an appropriate corrective image warp, a process that is cumbersome and time consuming.
 Another drawback of conventional spatial AR relying only on a projector is that it is capable of augmenting only the portion of the scene within the projector's own static field of view, thereby hindering its applicability to use cases calling for augmentation of wide areas such as a factory floorspace.

### 2021-09-09 paper/elsarticle/elsarticle-template.tex
Update elsarticle-template.tex

@@
 
-While interest in spatial AR in industrial settings has been growing, its deployment in industrial settings remains well behind that of wearable AR systems such as handhelds and head-mounted displays.
+While interest in spatial AR in industrial settings has been growing, its deployment in industrial settings remains well behind that of wearable AR systems such as handhelds or head-mounted displays.
 One of the main shortcomings of typical projector-based spatial AR relative to AR relying on wearables is the applicability of spatial AR only to augmenting scenes within the projector's own static field of view.
-Here, dynamic projection systems, using mirror heads to deflect the projection across wider areas has the potential to bridge the gap to wider industrial penetration.
+Here, dynamic projection systems---using mirror heads to deflect the projection across wider areas---has the potential to bridge the gap to wider industrial penetration.
 Yet the increase in the area to be covered comes with noteworthy challenges in terms of ensuring that the images to be projected appear free of distortions to the users.

### 2021-09-10 src/apps/calibrateProj/calibrateProj.cpp
plane via plane fit

@@ -61,11 +61,15 @@ cv::Vec3d centroidChessboardObjectPts;
 PointCloud * pointCloud;
 PointCloud * pointCloudCircles;
 PointCloud* pointCloud2Circles;
+PointCloud* triangulatedPointCloudCircles;
 PointCloud* pointCloudImVis;
 PointCloud* pointCloud2ImVis;
+
 Plane * planeVis;
+Plane* fittedPlaneVis;
 
 std::vector<Plane> planes;
+std::vector<Plane> fittedPlanes;
 
 cv::Vec3d projCamIntersection;
 cv::Vec3d projPlaneIntersection;
@@ -191,6 +195,7 @@ void display()
             else
             {
                 pointCloudCircles->display(pointSize, 1, 0, 0);
+                triangulatedPointCloudCircles->display(pointSize, 0, 1, 0);
 
                 std::vector<cv::Point3f> points = pointCloudCircles->getPoints();
                 std::vector<cv::Point3f> points2 = pointCloud->getPoints();
@@ -248,6 +253,7 @@ void display()
         }
 
         planeVis->display(2, pointSize);
+        fittedPlaneVis->display(2, pointSize);
 
         float offset = 0.02;
         for (int i = 0; i < cams.size(); i++)
@@ -293,7 +299,7 @@ void display()
 
             cv::Vec3d C = cams[i].getC();
 
-            if (camIdx == i)
+            //if (camIdx == i)
             {
                 std::stringstream ss;
                 ss << "Rendering according to " << string(str);
@@ -935,7 +941,7 @@ int main(int argc, char** argv)
         true, cam1K, cam1DistCoeffs, cam1ImVis, visImIdx);
 
     // compute circles object points by intersecting circle pixel back-projections with ground plane
-    std::vector<std::vector<cv::Point3f>> circlesObjectPts;
+    std::vector<std::vector<cv::Point3f>> circlesObjectPts, triangulatedCirclesObjectPts;
     for (int numIm = 0; numIm < numIms; numIm++)
     {
         // compute ground plane w.r.t. checkerboard pattern
@@ -970,9 +976,9 @@ int main(int argc, char** argv)
 
         fsPlane << "Rt" << planeRigid;
 
-        std::vector<cv::Point3f> circlesObjectPts_;
-        if (false)
-        {
+        std::vector<cv::Point3f> circlesObjectPts_, triangulatedCirclesObjectPts_;
+        //if (false)
+        //{
             cv::Mat intersectionsHomog;
[…]

### 2021-09-10 src/cpslib/libGeom/Plane.cpp

@@ -31,12 +31,71 @@ Plane::Plane(cv::Vec3d normal, float distance)
 
 	cv::Mat rot = Ancillary::getMinArclengthRotationMat33d(cv::Vec3d(0, 0, -1), normal_);
 
+	Eigen::Matrix3d rotEigen;
+	for (int y = 0; y < 3; y++)
+		for (int x = 0; x < 3; x++)
+			rotEigen(y, x) = rot.at<double>(y, x);
+
+	cv::Vec3d pt = -normal_ * distance_;
+
+	Eigen::Vector3d ptEigen(pt[0], pt[1], pt[2]);
+	Eigen::Vector3d tEigen = -rotEigen * ptEigen;
+
 	rigid_ = cv::Mat::eye(cv::Size(4, 4), CV_64F);
 	for (int y = 0; y < 3; y++)
 		for (int x = 0; x < 3; x++)
 			rigid_.at<double>(y, x) = rot.at<double>(y, x);
 
-	// todo: got R, but what about t?
+	for (int i = 0; i < 3; i++)
+		rigid_.at<double>(i, 3) = ptEigen[i];
+}
+
+// from https://stackoverflow.com/questions/40589802/eigen-best-fit-of-a-plane-to-n-points
+Plane::Plane(std::vector<cv::Point3f> points)
+{
+	// copy coordinates to  matrix in Eigen format
+	size_t numPts = points.size();
+	Eigen::MatrixXd pointsEigen(3, numPts);
+	for (size_t i = 0; i < numPts; ++i)
+		pointsEigen.col(i) = Eigen::Vector3d(points[i].x, points[i].y, points[i].z);
+
+	// calculate centroid
+	Eigen::Vector3d centroidEigen(pointsEigen.row(0).mean(), pointsEigen.row(1).mean(), pointsEigen.row(2).mean());
+
+	// subtract centroid
+	pointsEigen.row(0).array() -= centroidEigen(0);
+	pointsEigen.row(1).array() -= centroidEigen(1);
+	pointsEigen.row(2).array() -= centroidEigen(2);
+
+	// we only need the left-singular matrix here
+	//  http://math.stackexchange.com/questions/99299/best-fitting-plane-given-a-set-of-points
+
+	auto svd = pointsEigen.jacobiSvd(Eigen::ComputeThinU | Eigen::ComputeThinV);
[…]

### 2021-09-10 src/cpslib/libGeom/Plane.h

@@ -9,6 +9,7 @@ public:
 
 	Plane(cv::Mat & rigid);
 	Plane(cv::Vec3d normal, float distance);
+	Plane(std::vector<cv::Point3f> points);
 
 	void rigidTransform(cv::Mat & rigid44d);
 

### 2021-10-09 paper/elsarticle/elsarticle-template.tex
paper

@@
 For planar scene geometry (such as a floor, wall, or table), a manual process called keystone correction can be used to carry out an appropriate corrective image warp, a process that is cumbersome and time consuming.
-Another drawback of conventional spatial AR relying only on a projector is that it is capable of augmenting only the portion of the scene within the projector's own static field of view, thereby hindering its applicability to use cases calling for augmentation of wide areas such as a factory floorspace.
+Another drawback of conventional projector-based spatial AR is that it is capable of augmenting only the portion of the scene within the projector's own static field of view, thereby hindering its applicability to use cases calling for augmentation of wide areas such as a factory floorspace.
 
@@
 Our system renders the placement of augmentations in the scene more intuitive than manual keystone correction in two ways.
-First, (i) placement of the desired augmentations is carried out in accordance with the axes of an image of the scene acquired by the camera, thereby making setting those axes as simple as appropriately rotating the camera relative to the scene.
+First, (i) the horizontal and vertical axes of the desired augmentations is carried out in accordance with the horizontal and vertical image axes of the camera, thereby making setting those axes as simple as appropriately rotating the camera relative to the scene.
 Second, (ii) the desired dimensions of the projected augmentations are specified in metric terms, thereby allowing for consistent scaling across all target locations.
@@
 \end{equation}
-where $f$ and $(p_x, p_y)^\top$ are both expressed in units of pixels.
+where $f$ and $(x_0, y_0)^\top$ are both expressed in units of pixels.
 The calibration matrix~$\mathtt{K}$ models a so-called pinhole camera; such a model is applicable to an image acquired using a real-world camera assuming the image has been corrected for lens distortion effects, by having applied the appropriate lens distortion model coefficients.
@@
     \caption{Recovering 2D positions in support of projector calibration.
-(a) 2D positions of the 2D-3D correspondences to be used for calibrating the projector are obtained by detecting---in the image plane of the projector---the circle centers in the asymmetrical circle pattern image.
-It is this pattern that we project to each of the target locations in the scene plane.
-(b) For each such target location, an image is acquired from the viewpoint of the camera and the circle centers of the projected asymmetrical circle pattern are detected, in the image plane of the camera.
+(a) The asymmetrical circle pattern that we project to each of the target locations in the scene plane.
+2D positions of the 2D-3D correspondences to be used for calibrating the projector are detected automatically, as the respective circle centers in the asymmetrical circle pattern image and given in a fixed ordering.
+(b) For each such target location, a projector calibration image is acquired from the viewpoint of the camera and the circle centers of the projected asymmetrical circle pattern are detected using the same algorithm as in (a).
 A chessboard pattern to be used for recovering the local scene plane is placed near the projected pattern, whose corners are likewise detected.
@@
 
-\paragraph{Camera calibration} We recover 2D-3D correspondences in support of calibrating the downwards-facing camera by relying on a planar calibration surface to automatically identify correspondences between the 3D points on the calibration surface and their 2D correspondences in the image plane.
+\paragraph{2D-3D correspondences for camera calibration} We recover 2D-3D correspondences in support of calibrating the downwards-facing camera by relying on a planar calibration surface to automatically identify correspondences between the 3D points on the calibration surface and their 2D correspondences in the image plane.
 The classical calibration surface is a chessboard pattern.
@@
 
-\paragraph{Projector calibration} As in camera calibration, we calibrate the projector by relying on 2D-3D correspondences, yet for projector calibration we obtain them by \textit{projecting} a calibration pattern that we detect using the calibrated downwards-facing camera.
+\paragraph{2D-3D correspondences for projector calibration} As in camera calibration, we calibrate the projector by relying on 2D-3D correspondences.
+For projector calibration, however, obtaining those correspondences involves \textit{projecting} a calibration pattern to the ground plane.
+The 2D pixel positions of the projected pattern are obtained by
+
+ are known \textit{a priori}      that we detect using the calibrated downwards-facing camera.
+
  The pattern we project is one of circles (cf.\ Figure~\ref{fig:2d}(a)), and we rely on an algorithm to detect the circle center points in the asymmetrical circle pattern in the image plane of the projector \cite{bradski2000opencv}, giving the 2D positions of our 2D-3D correspondences for calibrating the projector.
@@
 \begin{figure}
-    \centerline{\includegraphics[scale=.35]{images/2d3d.png}}
-    \caption{Scene plane (gray) recovered via spatial resection with respect to 2D-3D correspondences obtained using a chessboard pattern (blue); 3D circle center points of the asymmetrical circle pattern---i.e., the 3D positions of the 2D-3D correspondences to be used for calibrating the projector---obtained by intersection with the scene plane of back-projections (likewise gray) of the 2D circle center points of the asymmetrical circle pattern detected in the image plane (red).
-Note that as in the figures of this section that follow, the rendering in the figure corresponds to the projection calibration image in Figure~\ref{fig:2d}(b), acquired by the downwards-facing camera (frustum of the camera in black, with up vector in red).}
+    \centerline{\includegraphics[scale=.375]{images/2d3d.png}}
+    \caption{Obtaining 2D-3D correspondences for projector calibration, using a projector calibration image (cf.\ Figure~\ref{fig:2d}(b)).
+First, the scene plane (gray) is recovered via spatial resection with respect to 2D-3D correspondences obtained using a chessboard pattern (blue); 3D circle center points of the asymmetrical circle pattern---i.e., the 3D positions of the 2D-3D correspondences to be used for calibrating the projector (red)---obtained by intersection with the scene plane of back-projections (likewise gray) of the 2D circle center points of the asymmetrical circle pattern detected in the image plane.
+The downwards-facing camera is shown in terms of its camera frustum (black, with up vector in red).}
     \label{fig:3d}
@@
     \centering
-    \subfloat[\centering Oblique view.]{{\includegraphics[height=4.2cm]{images/1_big.pdf}}}
+    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=4.2cm]{images/3_cam.png}}}
     \qquad
-    \subfloat[\centering Downwards view.]{{\includegraphics[height=4.2cm]{images/3_above3.png}}}
-    \caption{The virtual camera is obtained by (i) rotating the camera (top left, black) about its center of projection such that its optical axis be made parallel with the normal vector of the scene plane.
-The virtual projector is obtained by (ii) rotating the projector (bottom left, black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (iii) rotating the $X$- and $Y$-axes to align them with those of the virtual camera, and (iv) translating along the normal direction to achieve the desired metric projected image dimensions.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
+    \subfloat[\centering Downwards view.]{{\includegraphics[height=4.2cm]{images/3_above3_cam.png}}}
+    \caption{Obtaining the virtual camera.
+A physical camera placed to face downwards is almost certain to not face downwards precisely.
+We correct for this possibility by rotating the recovered downwards-facing camera (black) about its center of projection such that its optical axis be made parallel with the plane's normal.
+Since it is in accordance with our downwards-facing camera the we wish to orient our virtual projector and since our virtual projector is required to face directly downwards to the scene plane to compute our corrective image warp, it is the orientiation of the virtual camera that we use to subsequently orient our virtual projector.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
+This rotated camera is our virtual camera, and it is the orientation of the virtual camera that we subsequently use to orient our virtual projector.
+    \label{fig:virtualcam}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=4.2cm]{images/1_big.pdf}}}
+    \qquad
+    \subfloat[\centering Downwards view.]{{\includegraphics[height=4.2cm]{images/3_above3_proj.png}}}
+    \caption{Obtaining the virtual projector.
+The virtual projector is then obtained by (i) rotating the recovered projector (black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (ii) rotating the $X$- and $Y$-axes to align them with those of the virtual camera, and (iii) translating along the normal direction to achieve the desired metric projected image dimensions.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
     \label{fig:virtualproj}
@@
 a form of `plane-induced' 2D homography \cite{Hartley2004}.
-For convenience, we enable optional rotation of the image to be projected \textit{before} applying $\mathtt{H}$, about the image center; that rotation, parameterized in degrees, is thus in effect carried out---likewise intuitively---with respect to the ground plane.
+For convenience, we enable optional rotation of the image to be projected \textit{before} applying $\mathtt{H}$, about the image center; that rotation, parameterized in degrees, is thus in effect carried out intuitively with respect to the ground plane.
 

### 2021-10-09 src/apps/calibrateProj/calibrateProj.cpp

@@ -9,6 +9,7 @@
  *   calibrateProj 0.0565 4 6 4 11 C:\Users\micha\Desktop\spatial-ar\in_out\calibrateProj\out 11 C:\Users\micha\Desktop\spatial-ar\in_out\splitZed\projCalib\outLeft C:\Users\micha\Desktop\spatial-ar\in_out\splitZed\projCalib\outRight C:\Users\micha\Desktop\spatial-ar\in_out\calibrateCam\out\cam_0.yml  C:\Users\micha\Desktop\spatial-ar\in_out\calibrateCam\out\cam_1.yml C:\Users\micha\Desktop\spatial-ar\in_out\calibrateProj\acircles_pattern_960x600.png 1.0 3 C:\Users\micha\Desktop\spatial-ar\in_out\applyHomography\holodeck.png
  */
 
+// CMake projects should use: "-DCMAKE_TOOLCHAIN_FILE=C:/Users/micha/vcpkg/scripts/buildsystems/vcpkg.cmake"
 
 #include <iostream>
 #include <fstream>
@@ -59,9 +60,9 @@ vector<vector<Mat>> capturedPattern;
 cv::Vec3d centroidChessboardObjectPts;
 
 PointCloud * pointCloud;
-PointCloud * pointCloudCircles;
+std::vector<PointCloud*> pointCloudCirclesVec;
 PointCloud* pointCloud2Circles;
-PointCloud* triangulatedPointCloudCircles;
+std::vector<PointCloud*> triangulatedPointCloudCirclesVec;
 PointCloud* pointCloudImVis;
 PointCloud* pointCloud2ImVis;
 
@@ -194,10 +195,15 @@ void display()
                 pointCloud2Circles->display(pointSize, 0, 1, 0);
             else
             {
-                pointCloudCircles->display(pointSize, 1, 0, 0);
-                triangulatedPointCloudCircles->display(pointSize, 0, 1, 0);
+                
 
-                std::vector<cv::Point3f> points = pointCloudCircles->getPoints();
+                for (int i = 0; i < triangulatedPointCloudCirclesVec.size(); i++)
+                {
+                    pointCloudCirclesVec[i]->display(pointSize, 1, 0, 0);
+                    triangulatedPointCloudCirclesVec[i]->display(pointSize, 0, 1, 0);
+                }
+
+                std::vector<cv::Point3f> points = pointCloudCirclesVec[viewIdx]->getPoints();
[…]

### 2021-10-09 src/cpslib/libCalib/CalibPattern.cpp

@@ -83,9 +83,7 @@ void CalibPattern::findCirclesImPts(cv::Mat& im, cv::Size circlesPatternSize,
     cv::Mat imUndistort, imGrayUndistort, imGrayUndistortInvert;
 
     if (applyIntrinsics)
-    {
         cv::undistort(im, imUndistort, K, distCoeffs);
-    }
     else
         im.copyTo(imUndistort);
 

### 2021-10-09 src/cpslib/libGeom/Camera.cpp

@@ -49,7 +49,6 @@ Camera::Camera(const Camera& cam)
 	cam.M33d_.copyTo(M33d_);
 	cam.MInv33d_.copyTo(MInv33d_);
 	cam.Rt34d_.copyTo(Rt34d_);
-	cam.P34d_.copyTo(P34d_);
 	cam.Rt44d_.copyTo(Rt44d_);
 	cam.RtInv44d_.copyTo(RtInv44d_);
 

### 2021-10-09 src/cpslib/libGeom/Plane.cpp

@@ -1,5 +1,12 @@
 #include "Plane.h"
 
+#include <pcl/ModelCoefficients.h>
+#include <pcl/io/pcd_io.h>
+#include <pcl/point_types.h>
+#include <pcl/sample_consensus/method_types.h>
+#include <pcl/sample_consensus/model_types.h>
+#include <pcl/segmentation/sac_segmentation.h>
+
 Plane::Plane(const Plane & plane)
 {
 	normal_ = cv::Vec3d(plane.normal_);
@@ -50,32 +57,67 @@ Plane::Plane(cv::Vec3d normal, float distance)
 		rigid_.at<double>(i, 3) = ptEigen[i];
 }
 
-// from https://stackoverflow.com/questions/40589802/eigen-best-fit-of-a-plane-to-n-points
-Plane::Plane(std::vector<cv::Point3f> points)
+Plane::Plane(std::vector<cv::Point3f> points, bool ransac)
 {
-	// copy coordinates to  matrix in Eigen format
-	size_t numPts = points.size();
-	Eigen::MatrixXd pointsEigen(3, numPts);
-	for (size_t i = 0; i < numPts; ++i)
-		pointsEigen.col(i) = Eigen::Vector3d(points[i].x, points[i].y, points[i].z);
+	if (ransac) // todo
+	{
+		std::cout << "carrying out RANSAC plane fit" << std::endl;
 
-	// calculate centroid
-	Eigen::Vector3d centroidEigen(pointsEigen.row(0).mean(), pointsEigen.row(1).mean(), pointsEigen.row(2).mean());
+		pcl::PointCloud<pcl::PointXYZ>::Ptr cloud(new pcl::PointCloud<pcl::PointXYZ>);
 
-	// subtract centroid
-	pointsEigen.row(0).array() -= centroidEigen(0);
-	pointsEigen.row(1).array() -= centroidEigen(1);
-	pointsEigen.row(2).array() -= centroidEigen(2);
+		cloud->width = points.size();
+		cloud->height = 1;
+		cloud->points.resize(cloud->width * cloud->height);
 
-	// we only need the left-singular matrix here
-	//  http://math.stackexchange.com/questions/99299/best-fitting-plane-given-a-set-of-points
+		for (int i = 0; i < points.size(); i++)
+		{
+			cloud->at(i).x = points[i].x;
+			cloud->at(i).y = points[i].y;
[…]

### 2021-10-09 src/cpslib/libGeom/Plane.h

@@ -9,7 +9,7 @@ public:
 
 	Plane(cv::Mat & rigid);
 	Plane(cv::Vec3d normal, float distance);
-	Plane(std::vector<cv::Point3f> points);
+	Plane(std::vector<cv::Point3f> points, bool ransac = false);
 
 	void rigidTransform(cv::Mat & rigid44d);
 

### 2021-10-22 paper/elsarticle/elsarticle-template.tex
paper

@@
 Reliance on head-mounted displays has two adverse consequences of its own: a head-mounted display must be worn by each individual wishing to partake in the augmentation, and such a head-mounted display---in some cases taking the form of a helmet in order to house multiple sensors in support of accurately tracking the viewpoint of the viewer relative to the scene---can be obtrusive.
-A HoloLens 2, for instance, is such a head-mounted display; at a weight of 566~grams \cite{hololens}, it places an additional ergonomic load on the worker, in particular if worn over a longer duration or in bent positions.
+A HoloLens 2 (Microsoft Corporation, USA), for instance, is such a head-mounted display; at a weight of 566~grams \cite{hololens}, it places an additional ergonomic load on the worker, in particular if worn over a longer duration or in bent positions.
 
@@
 
-Calibrating a projector \cite{moreno2012simple,zhang2009projector,zhang2006novel} can be carried out in precisely the same manner as calibrating a camera insofar as step (ii) is concerned; the major difference in projector calibration relative to the calibrating a camera concerns the manner in which 2D-3D correspondences are identified, i.e., between pixels in the image plane of the projector and the corresponding points in the scene.
-Since our approach to projector calibration relies on a calibrated camera, what remains of this section is concerned primarily with the recovery of 2D-3D correspondences first in support of calibrating cameras, then of projectors.
-A consequence of the approach we take to identifying the 3D points of the 2D-3D correspondences we use for projector calibration is, for each target location, recovery of the local scene plane.
+Calibrating a projector \cite{moreno2012simple,zhang2009projector,zhang2006novel} can be carried out in precisely the same manner as calibrating a camera insofar as step (ii) is concerned; if proceeding accordingly, the major difference in projector calibration relative to the calibrating a camera concerns the manner in which 2D-3D correspondences are identified, i.e., between pixels in the image plane of the \textit{projector} and the corresponding points in the scene.
+What remains of this section is concerned primarily with the recovery of 2D-3D correspondences, first in support of calibrating the downwards-facing camera (used to obtain the camera calibration matrix~$\mathtt{K}_\text{cam}$), then---in a manner that makes use of $\mathtt{K}_\text{cam}$---for calibrating the projector (used for obtaining the projector calibration matrix~$\mathtt{K}_\text{proj}$ and the projector poses~($\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j) \in SE(3)$, for each projector calibration image).
+
+\paragraph{2D-3D correspondences for camera calibration} We recover 2D-3D correspondences in support of calibrating the downwards-facing camera by relying on a planar calibration surface to automatically identify correspondences between the 3D points on the calibration surface and their 2D correspondences in the image plane.
+The classical calibration surface is a chessboard pattern, and is such a pattern that we use as well.
+The 3D corner points of the chessboard are obtained \textit{a priori} in a coordinate system defined in the plane of the chessboard\footnote{E.g., $(0,0,0), (1.5,0,0), (3,0,0), \dots, (9,7.5,0)$ for a chessboard with $7~\times{}~6$ corners ($8~\times{}~7$ squares), with each square of length and width of 1.5 unit, respectively.
+Note that the units of the chessboard's 3D points give the units of the camera calibration, and---since our projector calibration relies on the camera calibration---the units of the projector calibration as well.}, requiring knowledge only of the dimensions of the chessboard pattern and of the length of a side of a chessboard square.
+The corresponding 2D points are obtained, in the same order, using a specialized algorithm \cite{bradski2000opencv}.
+A set of calibration images is acquired, each with the calibration pattern visible in a different part of the image plane, and such that the center and all corners and edges of the image plane are covered, the camera's autofocus setting be off, and the camera's zoom factor remain fixed.
+2D-3D correspondences are then recovered for each calibration image, respectively, and the resulting list of correspondences is used to recover the camera calibration matrix~$\mathtt{K}_\text{cam}$.
+
+\begin{figure}[ht]
+    \centering
+    \subfloat[\centering Projector calibration image (detections overlain).]{{\includegraphics[height=3cm]{images/detections.png} }}
+    \qquad
+    \subfloat[\centering 3D positions for projector calibration (red, in scene plane).]{{\includegraphics[height=3cm]{images/2d3d.png}}}
+    \caption{Obtaining the 3D positions of the 2D-3D correspondences for projector calibration.
+(a) For each target location, a projector calibration image is acquired from the viewpoint of the downwards-facing camera, containing a projected asymmetric circle pattern at the target location and a chessboard pattern placed on the floor nearby.
+Automatically detected circle pattern center points and chessboard corners overlain for illustration.
+(b) Given a projector calibration image, the scene plane (gray) is recovered in the coordinate frame of the camera (black, up vector red) via spatial resection with respect to 2D-3D correspondences obtained using the chessboard pattern (blue); the 3D positions of the 2D-3D correspondences to be used for calibrating the projector (red, in scene plane) are obtained by intersection with the scene plane of back-projections (rays, likewise gray) of the 2D center points of the circle pattern detected in the image plane of the camera.} %The corresponding 2D positions are detected in the image of the asymmetric circles pattern image projected by the projector (not shown).
+The downwards-facing camera is shown relative to the scene plane in terms of its camera frustum (black, with up vector in red).
+    \label{fig:2d}
+\end{figure}
+
+\paragraph{2D-3D correspondences for projector calibration} For projector calibration, we obtain the 2D-3D correspondences by \textit{projecting} a calibration pattern to the scene plane, once for each target location.
+Since we know that the 3D positions of the projected pattern must lie in the scene plane, we can recover the 3D correspondence of a pattern point detected in the 2D image plane of the downwards-facing camera by intersecting the back-projection of the 2D position with the scene plane.
+We reuse the chessboard pattern---used above for camera calibration---to recover the scene plane locally to each target location via spatial resection \cite{Hartley2004}, by placing the chessboard pattern on the floor near the projected pattern and using the resulting 2D-3D chessboard correspondences as input to a PnP algorithm \cite{collins2014infinitesimal}.
+While a single image of such a chessboard pattern placed on the floor could be sufficient if the floor is even, we recover a scene plane per projector calibration image to account for the possibility of an uneven floor.
+Since our projector calibration images thus already contain a chessboard, we choose an alternative calibration pattern as the pattern to project, namely an asymmetric circle pattern.
+We again use a specialized algorithm \cite{bradski2000opencv} to obtain the circle centers of the asymmetric circle pattern, capable of recovering circle centers even under projective distortions.
+
+Given a projector calibration image acquired using the downwards-facing camera, we detect the circle centers of the \textit{projected} asymmetrical circle pattern (cf.\ Figure~\ref{fig:2d}(a)); given the scene plane and such a 2D circle center $\mathbf{x}$, we obtain its 3D correspondence by intersecting the back-projection~$\mathtt{K}_\text{cam}^{-1}(\mathbf{x}^\top, 1)^\top \in \mathbb{P}^2$ of $\mathbf{x}$ with the recovered scene plane (cf.\ Figure~\ref{fig:2d}(b)).
+We obtain the 2D positions of the 2D-3D correspondences by simply running the algorithm for detecting circle centers on the original asymmetric circles pattern image (i.e., the one we feed to the projector).
+Since---as in the case of recovering chessboard pattern corners---the algorithm that yields 2D circle centers does so in a consistent ordering, we thus obtain respective 2D-3D correspondences for each projector calibration image, which we use to recover the projector calibration matrix~$\mathtt{K}_\text{proj}$.
+
+Recall that the pose~($\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j) \in SE(3)$ of the projector, for each projector calibration image~$j$, is provided alongside $\mathtt{K}_\text{proj}$, relative to the 3D points of the pattern and thus in the coordinate frame of the downwards-facing camera.
+Note that for a fixed projector with steerable mirror, given a projector calibration image, the recovered projector's pose is the pose the projector would have to have had to project to the given target location \textit{in the absence of the mirror} (i.e., facing directly towards the target location).
+This is sufficient for our needs in Section~\ref{sec:approach:homography}, and it is in this sense that our system is able to handle a projector equipped with a steerable mirror, without need for modeling the steerable mirror explicitly.
+
+\subsection{Correcting for Projective Distortion}\label{sec:approach:homography}
 
@@
     \centering
-    \subfloat[\centering Asymmetrical circle pattern image, in image plane of projector (detections overlaid).]{{\includegraphics[height=3.25cm]{images/circles.png} }}
-    \qquad
-    \subfloat[\centering Projector calibration image (one for each target location), in image plane of camera (detections overlaid).]{{\includegraphics[height=3.25cm]{images/detections.png}}}
-    \caption{Recovering 2D positions in support of projector calibration.
-(a) The asymmetrical circle pattern that we project to each of the target locations in the scene plane.
-2D positions of the 2D-3D correspondences to be used for calibrating the projector are detected automatically, as the respective circle centers in the asymmetrical circle pattern image and given in a fixed ordering.
-(b) For each such target location, a projector calibration image is acquired from the viewpoint of the camera and the circle centers of the projected asymmetrical circle pattern are detected using the same algorithm as in (a).
-A chessboard pattern to be used for recovering the local scene plane is placed near the projected pattern, whose corners are likewise detected.
-Detected 2D projected asymmetrical circle pattern circle center points and chessboard corners overlain for illustration.}
-    \label{fig:2d}
-\end{figure}
-
-\paragraph{2D-3D correspondences for camera calibration} We recover 2D-3D correspondences in support of calibrating the downwards-facing camera by relying on a planar calibration surface to automatically identify correspondences between the 3D points on the calibration surface and their 2D correspondences in the image plane.
-The classical calibration surface is a chessboard pattern.
-The 3D corner points of the chessboard are obtained \textit{a priori} in a coordinate system defined in the plane of the chessboard\footnote{E.g., $(0,0,0), (1.5,0,0), (3,0,0), \dots, (9,7.5,0)$ for a chessboard with $7~\times{}~6$ corners ($8~\times{}~7$ squares), with each square of length and width of 1.5 unit, respectively.
-Note that the units of the chessboard's 3D points give the units in terms of which the camera calibration is carried out, and---owing to how our projector calibration relies on the camera calibration---the units of the projector calibration as well.}, requiring knowledge only of the dimensions of the chessboard pattern and of the length of a side of a chessboard square.
-The corresponding 2D points are obtained, in the same order, using a specialized algorithm \cite{bradski2000opencv}.
-A set of calibration images is acquired, each with the calibration pattern visible in a different part of the image plane, and such that the center and all corners and edges of the image plane are covered, the camera's autofocus setting be off, and the camera's zoom factor remain fixed.
-2D-3D correspondences are then recovered for each calibration image, and the resulting list is passed on as input to an optimization procedure that relies on bundle adjustment to yield the camera calibration matrix~$\mathtt{K}_\text{cam}$.
-
-\paragraph{2D-3D correspondences for projector calibration} As in camera calibration, we calibrate the projector by relying on 2D-3D correspondences.
-For projector calibration, however, obtaining those correspondences involves \textit{projecting} a calibration pattern to the ground plane.
-The 2D pixel positions of the projected pattern are obtained by
-
- are known \textit{a priori}      that we detect using the calibrated downwards-facing camera.
-
- The pattern we project is one of circles (cf.\ Figure~\ref{fig:2d}(a)), and we rely on an algorithm to detect the circle center points in the asymmetrical circle pattern in the image plane of the projector \cite{bradski2000opencv}, giving the 2D positions of our 2D-3D correspondences for calibrating the projector.
-We project the asymmetrical circle pattern to each of the target locations, and use the the calibrated downwards-facing camera to acquire a projector calibration image for each.
-Given a projector calibration image acquired using camera, we detect the circle centers of the \textit{projected} asymmetrical circle pattern (cf.\ Figure~\ref{fig:2d}(b)); given the scene plane (the recovery of which we shall return to in the paragraph that follows) and such a 2D circle center $\mathbf{x}$, we obtain its 3D correspondence by intersecting the back-projection~$\mathtt{K}_\text{cam}^{-1}(\mathbf{x}^\top, 1)^\top \in \mathbb{P}^2$ of $\mathbf{x}$ with the recovered scene plane (cf.\ Figure~\ref{fig:3d}).
-Since the algorithm that yields 2D circle centers does so in a consistent ordering, we thus obtain the 2D-3D correspondences between the projector's image plane and the scene required for projector calibration, yielding the projector calibration matrix~$\mathtt{K}_\text{proj}$.
-
-\begin{figure}
-    \centerline{\includegraphics[scale=.375]{images/2d3d.png}}
-    \caption{Obtaining 2D-3D correspondences for projector calibration, using a projector calibration image (cf.\ Figure~\ref{fig:2d}(b)).
-First, the scene plane (gray) is recovered via spatial resection with respect to 2D-3D correspondences obtained using a chessboard pattern (blue); 3D circle center points of the asymmetrical circle pattern---i.e., the 3D positions of the 2D-3D correspondences to be used for calibrating the projector (red)---obtained by intersection with the scene plane of back-projections (likewise gray) of the 2D circle center points of the asymmetrical circle pattern detected in the image plane.
-The downwards-facing camera is shown in terms of its camera frustum (black, with up vector in red).}
-    \label{fig:3d}
-\end{figure}
-
-We recover the scene plane via spatial resection by applying a PnP algorithm \cite{collins2014infinitesimal} to the 2D-3D correspondences obtained using a chessboard pattern.
-Note that this step is separate from camera calibration, yet could well be carried out using the same calibration pattern used in the camera calibration step.\footnote{The critical point is that the pattern should ideally be coplanar with the local scene plane, meaning its height above the scene plane should not exceed a few millimeters.} While a single image of such a chessboard pattern placed on the floor could be sufficient if the floor is even, we account for the possibility of an uneven floor by placing a chessboard pattern in close proximity to the projected asymmetrical circle pattern in each projector calibration image in order to recover the scene plane locally to each target location (cf.\ Figure~\ref{fig:2d}(b)).
-Note that in principle, we could project a chessboard pattern instead of an asymmetrical circle pattern to obtain the 2D-3D correspondences needed for projector calibration; it is, however, because we rely on detecting the corners of a chessboard pattern to recover the scene plane that we opt instead for a alternative pattern.
-
-The pose of the projector, for each projector calibration image, is provided relative to the 3D points of the pattern---and thus in the coordinate frame of the camera---alongside $\mathtt{K}_\text{proj}$ by the aforementioned optimization procedure.
-Note that for a fixed projector with steerable mirror, given a projector calibration image, the recovered projector's pose is the pose the projector would have to have had to project to the given target location \textit{in the absence of the mirror} (i.e., facing directly towards the target location).
-This is sufficient for our needs in Section~\ref{sec:approach:homography}, and it is in this sense that our system is able to handle a projector equipped with a steerable mirror, without need for modeling the steerable mirror explicitly.
-
-\begin{figure}
-    \centering
-    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=4.2cm]{images/3_cam.png}}}
+    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=4.2cm]{images/3_cam.pdf}}}
     \qquad
@@
     \caption{Obtaining the virtual camera.
-A physical camera placed to face downwards is almost certain to not face downwards precisely.
-We correct for this possibility by rotating the recovered downwards-facing camera (black) about its center of projection such that its optical axis be made parallel with the plane's normal.
-Since it is in accordance with our downwards-facing camera the we wish to orient our virtual projector and since our virtual projector is required to face directly downwards to the scene plane to compute our corrective image warp, it is the orientiation of the virtual camera that we use to subsequently orient our virtual projector.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
+It is in accordance with the $X$- and $Y$-axes of the downwards-facing camera the we wish to orient our virtual projector.
+To avoid projective distortions, the virtual projector is required to face directly downwards to the scene plane; a physical camera placed to face downwards, however, is almost certain to not face downwards 	precisely.
+Accordingly, we rotate the recovered downwards-facing camera (black) about its center of projection such that its optical axis be made parallel with the scene plane's normal with respect to the minimum arc-length rotation.
+It is then the $X$- and $Y$-axes of the \textit{virtual} camera (gray) that we subsequently use to orient our virtual projector.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
 This rotated camera is our virtual camera, and it is the orientation of the virtual camera that we subsequently use to orient our virtual projector.
@@
 
-\begin{figure}
-    \centering
-    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=4.2cm]{images/1_big.pdf}}}
-    \qquad
-    \subfloat[\centering Downwards view.]{{\includegraphics[height=4.2cm]{images/3_above3_proj.png}}}
-    \caption{Obtaining the virtual projector.
-The virtual projector is then obtained by (i) rotating the recovered projector (black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (ii) rotating the $X$- and $Y$-axes to align them with those of the virtual camera, and (iii) translating along the normal direction to achieve the desired metric projected image dimensions.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
-    \label{fig:virtualproj}
-\end{figure}
-
-\subsection{Correcting for Projective Distortion}\label{sec:approach:homography}
-
 If the projector is calibrated and its pose relative to the scene plane is known, a `virtual' projector (with the same calibration $\mathtt{K}$ and lens distortion model coefficients) can be placed elsewhere relative to the scene plane.
-If we for a moment imagine that the projector---at its recovered pose---functions as a camera,\footnote{Recall that the calibration matrix~$\mathtt{K}$ enables computing both (i) the projection of a scene point to the image plane (the function of a camera), or (ii) the back-projection of a pixel in the image plane, giving a ray into the scene (along which a projector illuminates the scene with the given pixel).} then (i) projecting an image to the scene plane \textit{from the viewpoint of the virtual projector} and (ii) acquiring the resulting projected image from the viewpoint of the recovered projector gives the desired corrective warp.
+If we for a moment imagine that the projector---at its recovered pose---functions as a camera,\footnote{Recall that the calibration matrix~$\mathtt{K}$ enables computing (i) the projection of a scene point to the image plane (the function of a camera), or (ii) the back-projection of a pixel in the image plane, giving a ray into the scene (along which a projector illuminates the scene with the given pixel).} then (i) projecting an image to the scene plane \textit{from the viewpoint of the virtual projector} and (ii) acquiring the resulting projected image from the viewpoint of the recovered projector gives the desired corrective warp.
 Projecting an image warped in this manner to the scene plane \textit{from the viewpoint of the recovered projector} then has the same effect as projecting the original image to the scene plane from the viewpoint of the virtual projector.
@@
     \centering
-    \subfloat[\centering Projecting original image.]{{\includegraphics[height=4.2cm]{images/proj_projim.png}}}
+    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=4.2cm]{images/3_proj.pdf}}}
     \qquad
-    \subfloat[\centering Projecting our warped image.]{{\includegraphics[height=4.2cm]{images/proj_projim_waped.png}}}
-    \caption{Projection to the scene plane from the recovered projector viewpoint (bottom left, black; virtual projector in center, gray) of the original image and of the warped image.
-(a) Projecting the original image to the scene plane.
-(b) After warping the original image according to our plane-induced homography for the given target location, the image is projected in a manner that appears free of projective distortions, aligned with the axes of the virtual camera (via the virtual projector), and to have the desired dimensions in the scene plane, expressed in metric units.
-Note that unprojected background is shown set to black.}
+    \subfloat[\centering Downwards view.]{{\includegraphics[height=4.2cm]{images/3_above3_proj.png}}}
+    \caption{Obtaining the virtual projector.
+The virtual projector is then obtained by (i) rotating the recovered projector (black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (ii) rotating the $X$- and $Y$-axes to align them with those of the virtual camera (cf.\ Figure \ref{fig:virtualcam}), and (iii) translating along the normal direction to achieve the desired metric projected image dimensions.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
+    \label{fig:virtualproj}
+\end{figure}
+
+\paragraph{Virtual projector} The placement of the virtual projector determines from which pose the image to be projected is to \textit{appear} to have been projected.
+We carry out this placement according to a small handful of steps.
+First, we rotate the camera about its center of projection to align its optical axis\footnote{Strictly speaking, the optical axis is the back-projection of the principal point; in our usage, we understand it to refer to the ray through center of the image plane.} with the normal vector of the scene plane by applying the minimum arc-length rotation that relates the two, giving a \textit{virtual camera} facing directly\footnote{A physical camera placed to face downwards is almost certain to not face downwards precisely; in contrast, the virtual camera's optical axis is aligned exactly with the scene plane's normal vector, rendering it genuinely fronto-parallel with respect to the scene plane.} downwards to the scene plane (cf.\ Figure~\ref{fig:virtualcam}).
+It is to the virtual camera's $X$- and $Y$-axes that we subsequently set the $X$- and $Y$-axes of the virtual projector.
+To obtain the virtual projector, we (i) intersect the scene plane with the optical axis and rotate the projector's placement about that point of intersection, aligning the optical axis with the scene plane's normal vector and (ii) align the $X$- and $Y$-axes with those of the virtual camera.
+Finally, we (iii) translate along the scene plane's normal, in order to satisfy the desired projected image dimensions provided in metric units.
+The resulting virtual projector is thus rendered fronto-parallel with the scene plane---enabling projection to the scene plane absent of projective distortions---and projects to the scene plane with the desired scale.
+
+It is, strictly speaking, in the sense that we set the $X$- and $Y$-axes of the virtual projector to the $X$- and $Y$-axes of the virtual camera that we place the horizontal and vertical axes of the desired augmentations in accordance with the horizontal and vertical image axes of the camera; the placement of the camera thereby intuitively determines the principal axes according to which augmentations are to be placed.
+Note further that a consequence of placing the virtual projector by rotating about the point of intersection of the projector's optical axis with the scene plane is that \textit{the center of the projector's image plane remains invariant} to the placement of the virtual projector.
+A steerable mirror can be aimed with respect to a point projected from the center of the projector's image plane, thereby facilitating placement of target locations.
+
+\begin{figure}
+    \centering
+    \subfloat[\centering Original image.]{{\includegraphics[height=3.4cm]{images/projim.png}}}
+    \qquad
+    \subfloat[\centering Projecting the original image (downwards view).]{{\includegraphics[height=3.4cm]{images/proj_projim.png}}}
+    \caption{Projection to the scene plane from the viewpoint of the recovered projector (bottom left, black; virtual projector in center, gray).
+(a) Original image.
+(b) Projecting the original image to the scene plane introduces projective distortions if the projector is not fronto-parallel with the scene plane, as is the case here.}
+    \label{fig:proj}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \subfloat[\centering Our warped image.]{{\includegraphics[height=3.4cm]{images/projim_warped_scaled.png}}}
+    \qquad
+    \subfloat[\centering Projecting our warped image (downwards view).]{{\includegraphics[height=3.4cm]{images/proj_projim_waped.png}}}
+    \caption{Projection to the scene plane from the viewpoint of the recovered projector (bottom left, black; virtual projector in center, gray).
+(a) Original image warped using our approach.
+(b) After warping the original image according to our approach for the given target location, the image is projected with the effect of having the original image from the viewpoint of the virtual projector.
+This renders the projection free of projective distortions, aligned with the axes of the virtual camera, and with the desired dimensions in the scene plane, expressed in metric units.
+Note that pixels shown in black will not be projected to the scene.}
     \label{fig:warp}
@@
     \caption{Evaluation scenario.
-(a) Our hardware setup, comprised of a Panasonic PT-RZ660BE projector, a steerable mirror system manufactured by Dynamic Projection Institute, and a Stereolabs Zed 2 stereo camera, of which we used only the left view.
+(a) Our hardware setup, comprised of a Panasonic PT-RZ660BE projector, a steerable mirror system manufactured by Dynamic Projection Institute, and a Stereolabs Zed 2 stereo camera (extreme front, black), of which we used only the left view.
 (b) Example augmentation produced by projecting---to one of our 15 target locations---an image warped using our approach, as seen from the downwards-facing camera.
@@
 \begin{table}[ht]
-\caption{Deviations in measured lengths from 50~cm~$\times$~31.25~cm across the 15 augmentations produced using our approach, provided in cm and in pixel units of the downwards-facing camera (pixel size 3.5 mm).
-Note that in no instance do mean and standard deviation exceed 1 pixel from the viewpoint of the camera.}
+\caption{Means and standard deviations (in units of centimeters and pixels, respectively) of deviations in measured lengths from 50 cm for top and bottom and from 21.25 cm for left and right, with respect to the 15 augmentations produced using our approach (intended augmentation dimensions 50~cm~$\times$~31.25~cm).
+Pixel size given the downward-facing camera's height above the ground plane was 3.5 mm; note that in no instance do mean and standard deviation exceed 1 pixel from the viewpoint of the camera.}
 \label{table:length}
@@
 \begin{table}[ht]
-\caption{Deviations in measured angles from $90^\circ{}$ across the 15 augmentations produced using our approach.}
+\caption{Means and standard deviations of deviations in measured angles from $90^\circ{}$ with respect to the 15 augmentations produced using our approach.}
 \label{table:angle}

### 2021-10-24 paper/elsarticle/elsarticle-template.tex
paper

@@
 \end{equation}
-where $\hat{\mathtt{R}}_j\mathbf{X}_{i,j} + \hat{\mathbf{t}}_j \in \mathbb{R}^3$ expresses the point~$\mathbf{X}_{i,j}$ in the local coordinate frame of the camera corresponding to the $j^{\text{th}}$~calibration image and $(\hat{\mathbf{x}}_{i,j}'^\top, 1)^\top \sim \hat{\mathtt{K}}(\hat{\mathtt{R}}_j\mathbf{X}_{i,j} + \hat{\mathbf{t}}_j) \in \mathbb{P}^2$ gives the projection~$\hat{\mathbf{x}}_{i,j}' \in \mathbb{R}^2$ in pixels of $\mathbf{X}_{i,j}$ to that image.
-The function~$d_{\hat{\omega}}$ is a distance function that computes a distance with respect to two pixel positions that it first corrects for lens distortions, in accordance with the lens distortion model coefficients~$\hat{\omega}$ \cite{duane1971close,weng1992camera}.
-Inverting the rigid body transformation~$(\mathtt{R}_j, \mathbf{t}_j)$ gives the pose~$(\hat{\mathtt{R}}_j^{-1}, -\hat{\mathtt{R}}_j^{-1}\hat{\mathbf{t}}_j^{}) \in SE(3)$ of the $j^{\text{th}}$~camera relative to the 3D scene points of the calibration pattern, with $-\hat{\mathtt{R}}_j^{-1}\hat{\mathbf{t}}_j^{} \in \mathbb{R}^3$ the corresponding center of projection.
+where $\hat{\mathbf{x}}_{i,j}' \in \mathbb{R}^2$ gives the projection, in units of pixels, of $\mathbf{X}_{i,j}$ to the image plane of the camera underlying the $j^{\text{th}}$~calibration image, subject to the maximum likelihood estimates $\hat{\mathtt{K}}, (\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j)$.
+The function~$d_{\hat{\omega}}$ is a distance function that computes a distance with respect to two pixel positions that it first corrects for lens distortions, in accordance with the maximum likelihood estimate of the lens distortion model coefficients~$\hat{\omega}$ \cite{duane1971close,weng1992camera}.
+More generally, the projection $\mathbf{x}' \in \mathbb{R}^2$ of $\mathbf{X}$ to the image plane of a camera given calibration matrix~$\mathtt{K}$ and pose~$(\mathtt{R}, \mathbf{t})$ is given by
+\begin{equation}
+(\mathbf{x}'^\top, 1)^\top \sim \mathtt{K}(\mathtt{R}\mathbf{X} + \mathbf{t}) \in \mathbb{P}^2,
+\end{equation}
+where $\mathtt{R}\mathbf{X} + \mathbf{t} \in \mathbb{R}^3$ expresses the point~$\mathbf{X}$ in the coordinate frame of the camera.
+Inverting the rigid body transformation~$(\mathtt{R}_j, \mathbf{t}_j)$ gives the pose~$(\mathtt{R}_j^{-1}, -\mathtt{R}_j^{-1}\mathbf{t}_j^{}) \in SE(3)$ of the $j^{\text{th}}$~camera relative to the 3D scene points of the calibration pattern, with $-\mathtt{R}_j^{-1}\mathbf{t}_j^{} \in \mathbb{R}^3$ the corresponding center of projection.
 
@@
 
-Given a projector calibration image acquired using the downwards-facing camera, we detect the circle centers of the \textit{projected} asymmetrical circle pattern (cf.\ Figure~\ref{fig:2d}(a)); given the scene plane and such a 2D circle center $\mathbf{x}$, we obtain its 3D correspondence by intersecting the back-projection~$\mathtt{K}_\text{cam}^{-1}(\mathbf{x}^\top, 1)^\top \in \mathbb{P}^2$ of $\mathbf{x}$ with the recovered scene plane (cf.\ Figure~\ref{fig:2d}(b)).
+Given a projector calibration image acquired using the downwards-facing camera, we detect the circle centers of the \textit{projected} asymmetrical circle pattern (cf.\ Figure~\ref{fig:2d}(a)); given the scene plane and such a 2D circle center $\mathbf{x}$, we obtain its 3D correspondence by intersecting the back-projection
+\begin{equation}
+\mathtt{K}_\text{cam}^{-1}(\mathbf{x}^\top, 1)^\top \in \mathbb{P}^2
+\end{equation}
+of $\mathbf{x}$ with the recovered scene plane (cf.\ Figure~\ref{fig:2d}(b)).
 We obtain the 2D positions of the 2D-3D correspondences by simply running the algorithm for detecting circle centers on the original asymmetric circles pattern image (i.e., the one we feed to the projector).
@@
     \caption{Obtaining the virtual camera.
-It is in accordance with the $X$- and $Y$-axes of the downwards-facing camera the we wish to orient our virtual projector.
-To avoid projective distortions, the virtual projector is required to face directly downwards to the scene plane; a physical camera placed to face downwards, however, is almost certain to not face downwards 	precisely.
+Our physical downwards-facing camera---in terms of which we wish to orient our augmentations---is almost certain to not face downwards precisely.
 Accordingly, we rotate the recovered downwards-facing camera (black) about its center of projection such that its optical axis be made parallel with the scene plane's normal with respect to the minimum arc-length rotation.
-It is then the $X$- and $Y$-axes of the \textit{virtual} camera (gray) that we subsequently use to orient our virtual projector.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
+It is then, for each respective target position, the $X$- and $Y$-axes of the resulting \textit{virtual} camera (gray) that we subsequently use to orient our virtual projector.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
 This rotated camera is our virtual camera, and it is the orientation of the virtual camera that we subsequently use to orient our virtual projector.
@@
     \caption{Obtaining the virtual projector.
-The virtual projector is then obtained by (i) rotating the recovered projector (black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (ii) rotating the $X$- and $Y$-axes to align them with those of the virtual camera (cf.\ Figure \ref{fig:virtualcam}), and (iii) translating along the normal direction to achieve the desired metric projected image dimensions.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
+For a given target location, the virtual projector is obtained by (i) rotating the recovered projector (black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (ii) rotating the $X$- and $Y$-axes to align them with those of the virtual camera (cf.\ Figure \ref{fig:virtualcam}), and (iii) translating along the normal direction to achieve the desired metric projected image dimensions.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
     \label{fig:virtualproj}
@@
     \subfloat[\centering Projecting the original image (downwards view).]{{\includegraphics[height=3.4cm]{images/proj_projim.png}}}
-    \caption{Projection to the scene plane from the viewpoint of the recovered projector (bottom left, black; virtual projector in center, gray).
+    \caption{Projection to the scene plane from the viewpoint of the recovered projector (bottom left, black; virtual projector in center, gray) for a given target location.
 (a) Original image.
@@
     \subfloat[\centering Projecting our warped image (downwards view).]{{\includegraphics[height=3.4cm]{images/proj_projim_waped.png}}}
-    \caption{Projection to the scene plane from the viewpoint of the recovered projector (bottom left, black; virtual projector in center, gray).
+    \caption{Projection to the scene plane from the viewpoint of the recovered projector (bottom left, black; virtual projector in center, gray) for a given target location.
 (a) Original image warped using our approach.

### 2021-10-24 paper/elsarticle/elsarticle-template.tex
paper

@@
 \end{equation}
-where $\hat{\mathbf{x}}_{i,j}' \in \mathbb{R}^2$ gives the projection, in units of pixels, of $\mathbf{X}_{i,j}$ to the image plane of the camera underlying the $j^{\text{th}}$~calibration image, subject to the maximum likelihood estimates $\hat{\mathtt{K}}, (\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j)$.
-The function~$d_{\hat{\omega}}$ is a distance function that computes a distance with respect to two pixel positions that it first corrects for lens distortions, in accordance with the maximum likelihood estimate of the lens distortion model coefficients~$\hat{\omega}$ \cite{duane1971close,weng1992camera}.
-More generally, the projection $\mathbf{x}' \in \mathbb{R}^2$ of $\mathbf{X}$ to the image plane of a camera given calibration matrix~$\mathtt{K}$ and pose~$(\mathtt{R}, \mathbf{t})$ is given by
+where $\hat{\mathbf{x}}_{i,j}' \in \mathbb{R}^2$ gives the projection of the point~$\mathbf{X}_{i,j} \in \mathbb{R}^3$ to the image plane of the camera underlying the $j^{\text{th}}$~calibration image, subject to the maximum likelihood estimates $\hat{\mathtt{K}}, (\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j)$.
+The function~$d_{\hat{\omega}}$, in turn, is a function that first (i) corrects $\mathbf{x}_{i,j}, \hat{\mathbf{x}}_{i,j}'$ for lens distortions in accordance with the maximum likelihood estimate of the lens distortion model coefficients~$\hat{\omega}$ \cite{duane1971close,weng1992camera}, and then (ii) computes a distance in pixels with respect to two resulting corrected positions.
+More generally, the projection~$\mathbf{x}' \in \mathbb{R}^2$ of a point~$\mathbf{X} \in \mathbb{R}^3$ to the image plane of a camera given calibration matrix~$\mathtt{K}$ and pose~$(\mathtt{R}, \mathbf{t})$ is given by
 \begin{equation}
@@
 \end{equation}
-where $\mathtt{R}\mathbf{X} + \mathbf{t} \in \mathbb{R}^3$ expresses the point~$\mathbf{X}$ in the coordinate frame of the camera.
+where $\mathtt{R}\mathbf{X} + \mathbf{t} \in \mathbb{R}^3$ expresses $\mathbf{X}$ in the coordinate frame of the camera.
 Inverting the rigid body transformation~$(\mathtt{R}_j, \mathbf{t}_j)$ gives the pose~$(\mathtt{R}_j^{-1}, -\mathtt{R}_j^{-1}\mathbf{t}_j^{}) \in SE(3)$ of the $j^{\text{th}}$~camera relative to the 3D scene points of the calibration pattern, with $-\mathtt{R}_j^{-1}\mathbf{t}_j^{} \in \mathbb{R}^3$ the corresponding center of projection.
@@
 (b) After warping the original image according to our approach for the given target location, the image is projected with the effect of having the original image from the viewpoint of the virtual projector.
-This renders the projection free of projective distortions, aligned with the axes of the virtual camera, and with the desired dimensions in the scene plane, expressed in metric units.
-Note that pixels shown in black will not be projected to the scene.}
+This renders the projection free of projective distortions, aligned with the axes of the virtual camera, and with the desired metric dimensions (i.e., desired scale) in the scene plane.
+Note that pixels shown in black will in effect be ignored by the projector.}
     \label{fig:warp}
@@
 \paragraph{Plane-induced homography} Let $\mathtt{K}_\text{proj}$ express the calibration matrix of the recovered projector and $(\mathtt{R}, \mathbf{t}) \in SE(3)$ the rigid body transformation that transforms points from the coordinate frame of the recovered projector to that of the virtual projector, for a given target location.
-Moreover, let $(\mathbf{n}^\top, -d)^\top$ give the scene plane, expressed in the coordinate frame of the recovered projector, where $\mathbf{n} \in \mathbb{R}^3$ is the scene plane's normal vector and $d = \mathbf{n}^\top\mathbf{X}$ for any point~$\mathbf{X}$ in the plane, so that $(\mathbf{n}^\top, -d) (\mathbf{X}^\top, 1)^\top = 0$.
-The transformation that warps the image to be projected to the scene plane by the recovered projector such that it appear as if were projected to the scene plane by the virtual projector (cf.\ Figure~\ref{fig:warp}(b)) is given the by the $3~\times~3$ matrix
+Moreover, let $(\mathbf{n}^\top, -d)^\top \in \mathbb{P}^3$ give the scene plane, expressed in the coordinate frame of the recovered projector, where $\mathbf{n} \in \mathbb{R}^3$ is the scene plane's normal vector and $d = \mathbf{n}^\top\mathbf{X}$ for any point~$\mathbf{X} \in \mathbb{R}^3$ in the plane, so that $(\mathbf{n}^\top, -d) (\mathbf{X}^\top, 1)^\top = 0$.
+The transformation that warps the image to be projected---by the recovered projector to the scene plane---such that it appear as if were projected to the scene plane by the virtual projector (cf.\ Figure~\ref{fig:warp}(b)) is given the by the $3~\times~3$ matrix
 \begin{equation}

### 2021-10-29 paper/elsarticle/elsarticle-template.tex
paper

@@
 If the projector is calibrated and its pose relative to the scene plane is known, a `virtual' projector (with the same calibration $\mathtt{K}$ and lens distortion model coefficients) can be placed elsewhere relative to the scene plane.
-If we for a moment imagine that the projector---at its recovered pose---functions as a camera,\footnote{Recall that the calibration matrix~$\mathtt{K}$ enables computing (i) the projection of a scene point to the image plane (the function of a camera), or (ii) the back-projection of a pixel in the image plane, giving a ray into the scene (along which a projector illuminates the scene with the given pixel).} then (i) projecting an image to the scene plane \textit{from the viewpoint of the virtual projector} and (ii) acquiring the resulting projected image from the viewpoint of the recovered projector gives the desired corrective warp.
-Projecting an image warped in this manner to the scene plane \textit{from the viewpoint of the recovered projector} then has the same effect as projecting the original image to the scene plane from the viewpoint of the virtual projector.
+If we for a moment imagine that the projector---at its recovered pose---functions as a camera,\footnote{Recall that the calibration matrix~$\mathtt{K}$ enables computing (i) the projection of a scene point to the image plane (the function of a camera), or (ii) the back-projection of a pixel in the image plane, giving a ray into the scene (along which a projector illuminates the scene with the given pixel).} then (i) projecting an image to the scene plane \textit{from the viewpoint of the virtual projector} and (ii) acquiring the resulting projected image \textit{from the viewpoint of the recovered projector} gives the desired corrective warp.
+Projecting an image warped in this manner to the scene plane from the viewpoint of the recovered projector then has the same effect as projecting the original image to the scene plane from the viewpoint of the virtual projector.
 This warp can be effected using a plane-induced homography, computed analytically as a function of the scene plane, the projector, and the virtual projector.
@@
     \caption{Obtaining the virtual projector.
-For a given target location, the virtual projector is obtained by (i) rotating the recovered projector (black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (ii) rotating the $X$- and $Y$-axes to align them with those of the virtual camera (cf.\ Figure \ref{fig:virtualcam}), and (iii) translating along the normal direction to achieve the desired metric projected image dimensions.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
+For a given target location, the virtual projector is obtained by (i) rotating the recovered projector (bottom left, black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (ii) rotating the $X$- and $Y$-axes to align them with those of the virtual camera (cf.\ Figure \ref{fig:virtualcam}), and (iii) translating along the normal direction to achieve the desired metric projected image dimensions.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
     \label{fig:virtualproj}
@@
 (a) Original image warped using our approach.
-(b) After warping the original image according to our approach for the given target location, the image is projected with the effect of having the original image from the viewpoint of the virtual projector.
-This renders the projection free of projective distortions, aligned with the axes of the virtual camera, and with the desired metric dimensions (i.e., desired scale) in the scene plane.
+(b) After warping the original image according to our approach for the given target location, the image is projected with the effect of projecting the original image from the viewpoint of the virtual projector.
+This renders the projection free of projective distortions, aligned with the axes of the virtual camera, and at the desired metric scale.
 Note that pixels shown in black will in effect be ignored by the projector.}
@@
     \centering
-    \subfloat[\centering Hardware setup (projector, steerable mirror system, camera).]{{\includegraphics[height=3.3cm]{images/system2.png}}}
+    \subfloat[\centering Hardware setup (projector, steerable mirror system, camera).]{{\includegraphics[height=3.3cm]{images/system2.pdf}}}
     \quad
-    \subfloat[\centering Example projection of image warped for target location using our approach.]{{\includegraphics[height=3.3cm]{images/hans_undist_inset.png}}}
+    \subfloat[\centering Example projection of image warped for target location using our approach.]{{\includegraphics[height=3.3cm]{images/hans_undist.pdf}}}
     \caption{Evaluation scenario.
 (a) Our hardware setup, comprised of a Panasonic PT-RZ660BE projector, a steerable mirror system manufactured by Dynamic Projection Institute, and a Stereolabs Zed 2 stereo camera (extreme front, black), of which we used only the left view.
-(b) Example augmentation produced by projecting---to one of our 15 target locations---an image warped using our approach, as seen from the downwards-facing camera.
+(b) Example augmentation produced by projecting---to one of our 15 target locations across the floorspace---an image warped using our approach, as seen from the downwards-facing camera.
 Note that the warp places the axes of the augmentation in accordance with the axes of the camera, and that the dimensions of the augmentation are in line with the desired target dimensions (50~cm~$\times$~31.25~cm).
@@
 \begin{table}[ht]
-\caption{Means and standard deviations (in units of centimeters and pixels, respectively) of deviations in measured lengths from 50 cm for top and bottom and from 21.25 cm for left and right, with respect to the 15 augmentations produced using our approach (intended augmentation dimensions 50~cm~$\times$~31.25~cm).
-Pixel size given the downward-facing camera's height above the ground plane was 3.5 mm; note that in no instance do mean and standard deviation exceed 1 pixel from the viewpoint of the camera.}
+\caption{Means and standard deviations (in units of centimeters and pixels, respectively) of deviations in measured lengths from 50 cm for top and bottom and from 21.25 cm for left and right, with respect to the 15 augmentations produced using our approach  for 15 respective target locations (intended augmentation dimensions 50~cm~$\times$~31.25~cm).
+Pixel size given the downward-facing camera's height above the ground plane was 3.5 mm; note that in no instance did mean and standard deviation exceed 1 pixel from the viewpoint of the camera.}
 \label{table:length}
@@
 \begin{table}[ht]
-\caption{Means and standard deviations of deviations in measured angles from $90^\circ{}$ with respect to the 15 augmentations produced using our approach.}
+\caption{Means and standard deviations of deviations in measured angles from $90^\circ{}$ with respect to the 15 augmentations produced using our approach for 15 respective target locations.}
 \label{table:angle}

### 2021-10-29 paper/elsarticle/elsarticle-template.tex
paper

@@
     \quad
-    \subfloat[\centering Example projection of image warped for target location using our approach.]{{\includegraphics[height=3.3cm]{images/hans_undist.pdf}}}
+    \subfloat[\centering Projection of warped image from Figure~\ref{fig:warp}(a) to corresponding target location.]{{\includegraphics[height=3.3cm]{images/hans_undist.pdf}}}
     \caption{Evaluation scenario.
-(a) Our hardware setup, comprised of a Panasonic PT-RZ660BE projector, a steerable mirror system manufactured by Dynamic Projection Institute, and a Stereolabs Zed 2 stereo camera (extreme front, black), of which we used only the left view.
+(a) Our hardware setup, comprised of a Panasonic PT-RZ660BE projector, a steerable mirror system manufactured by Dynamic Projection Institute, and a Stereolabs Zed 2 stereo camera, of which we used only the left view.
 (b) Example augmentation produced by projecting---to one of our 15 target locations across the floorspace---an image warped using our approach, as seen from the downwards-facing camera.
 Note that the warp places the axes of the augmentation in accordance with the axes of the camera, and that the dimensions of the augmentation are in line with the desired target dimensions (50~cm~$\times$~31.25~cm).
-Corners of the full projected image extent projected to the floor in light green.}
+Corners of the full projected image extent projected to the floor in light green (emphasized with overlain dashed circles).}
     \label{fig:eval}

### 2021-10-29 paper/elsarticle/elsarticle-template.tex
paper

@@
 We evaluate our approach by using a 960 pixel~$\times$~600 pixel image of the TU Wien logo (cf.\ Figure \ref{fig:warp}) to augment 15 locations across the floorspace at the Pilotfabrik\footnote{\url{https://www.pilotfabrik.at/}} of TU Wien, a collaborative space for research on Industry 4.0 topics situated in Vienna, Austria.
-We contrast our approach with a baseline approach involving manual keystone correction, by aiming with both approaches to place the same image for each location aligned with the principal axes of the floorspace, absent of projective distortions, and with the same metric dimensions of 50~cm~$\times$~31.25~cm\footnote{The dimensions in pixels of the image we project are 960~$\times$~600; we chose for our experiments to set the projected metric length of the horizontal axis of the image to 50~cm, which implies 31.25~cm for the vertical axis if aspect ratio is to be preserved.}.
+We contrast our approach with a baseline approach involving manual keystone correction, by aiming with both approaches to place the same image for each location aligned with the principal axes of the floorspace (i.e., the yellow and red lines in Figure~\ref{fig:eval}(b)), absent of projective distortions, and with the same metric dimensions of 50~cm~$\times$~31.25~cm\footnote{The dimensions in pixels of the image we project are 960~$\times$~600; we chose for our experiments to set the projected metric length of the horizontal axis of the image to 50~cm, which implies 31.25~cm for the vertical axis if aspect ratio is to be preserved.}.
 All experiments were carried out by the same technician, experienced in both approaches.

### 2021-10-29 paper/elsarticle/elsarticle-template.tex
paper

@@
 
+\begin{figure}
+    \centering
+    \subfloat[\centering Original image.]{{\includegraphics[height=3.4cm]{images/projim.png}}}
+    \qquad
+    \subfloat[\centering Projecting the original image (downwards view).]{{\includegraphics[height=3.4cm]{images/proj_projim.png}}}
+    \caption{Projection to the scene plane from the viewpoint of the recovered projector, facing a given target location.
+(a) Original image.
+(b) Projecting the original image from the viewpoint of the recovered projector (bottom left, black) to the scene plane introduces projective distortions if the projector is not fronto-parallel with the scene plane, as is the case here.}
+    \label{fig:proj}
+\end{figure}
+
 While interest in spatial AR in industrial settings has been growing, its deployment in industrial settings remains well behind that of wearable AR systems such as handhelds or head-mounted displays.
@@
 Yet the increase in the area to be covered comes with noteworthy challenges in terms of ensuring that the images to be projected appear free of distortions to the users.
-Considering a planar surface to be augmented (e.g., a floor, wall, or table), unless the projector faces the surface frontally, the bounds of a projected rectangular image will not appear rectangular, but will instead be subject to projective distortions.
+Considering a planar surface to be augmented (e.g., a floor, wall, or table), unless the projector faces the surface frontally, the bounds of a projected rectangular image will not appear rectangular, but will instead be subject to projective distortions (cf.\ Figure~\ref{fig:proj}).
 Such distortions can be corrected for by carrying out a cumbersome manual process referred to as keystone correction---involving warping an image by perturbing its corners until the desired effect is produced when projected to the scene---to appropriately warp the image to be projected, often using software bundled with the projector.
@@
 
+\begin{figure}
+    \centering
+    \subfloat[\centering Our warped image.]{{\includegraphics[height=3.4cm]{images/projim_warped_scaled.png}}}
+    \qquad
+    \subfloat[\centering Projecting our warped image (downwards view).]{{\includegraphics[height=3.4cm]{images/proj_projim_waped.png}}}
+    \caption{Projection to the scene plane from the viewpoint of the recovered projector for a given target location.
+(a) Original image warped using our approach.
+(b) After warping the original image according to our approach for the given target location, the image is projected from the viewpoint of the recovered projector (bottom left, black) with the effect of projecting the original image from the viewpoint of the virtual projector (center, gray).
+This renders the projection free of projective distortions, aligned with the axes of the virtual camera, and at the desired metric scale.
+Note that black pixels will in effect be ignored by the projector.}
+    \label{fig:warp}
+\end{figure}
+
 Our contribution is to propose a wide-area spatial AR system for planar scenes that produces the effect of keystone correction analytically, free of any manual interaction.
@@
 This facilitates placement of augmentations, since establishing the principal axes of the augmentation thus reduces to placing the downwards-facing camera in a manner such that the $X$- and $Y$-axes of the image align with the intended $X$- and $Y$-axes of the augmentation.
-We achieve this by warping the image to be projected using a plane-induced homography computed to produce the effect of projecting the image not from the actual projector viewpoint, but in accordance with the viewpoint of a \textit{virtual} projector.
+We achieve this by warping the image to be projected using a plane-induced homography computed to produce the effect of projecting the image not from the actual projector viewpoint, but in accordance with the viewpoint of a \textit{virtual} projector (cf.\ Figure~\ref{fig:warp}).
 We place this virtual projector to (i) face directly downwards to the target location in the scene plane and (ii) rotate its axes in accordance with those of the camera.
@@
 If the projector is calibrated and its pose relative to the scene plane is known, a `virtual' projector (with the same calibration $\mathtt{K}$ and lens distortion model coefficients) can be placed elsewhere relative to the scene plane.
-If we for a moment imagine that the projector---at its recovered pose---functions as a camera,\footnote{Recall that the calibration matrix~$\mathtt{K}$ enables computing (i) the projection of a scene point to the image plane (the function of a camera), or (ii) the back-projection of a pixel in the image plane, giving a ray into the scene (along which a projector illuminates the scene with the given pixel).} then (i) projecting an image to the scene plane \textit{from the viewpoint of the virtual projector} and (ii) acquiring the resulting projected image \textit{from the viewpoint of the recovered projector} gives the desired corrective warp.
+If we for a moment imagine that the projector---at its recovered pose---functions as a camera,\footnote{Recall that the calibration matrix~$\mathtt{K}$ enables computing (i) the projection of a scene point to the image plane (the function of a camera), or (ii) the back-projection of a pixel in the image plane, giving a ray into the scene (along which a projector illuminates the scene with the given pixel).} then (i) projecting an image to the scene plane \textit{from the viewpoint of the virtual projector} and (ii) acquiring the resulting projected image \textit{from the viewpoint of the recovered projector} gives the desired corrective warp (cf.\ Figure~\ref{fig:warp}).
 Projecting an image warped in this manner to the scene plane from the viewpoint of the recovered projector then has the same effect as projecting the original image to the scene plane from the viewpoint of the virtual projector.
@@
 
-\begin{figure}
-    \centering
-    \subfloat[\centering Original image.]{{\includegraphics[height=3.4cm]{images/projim.png}}}
-    \qquad
-    \subfloat[\centering Projecting the original image (downwards view).]{{\includegraphics[height=3.4cm]{images/proj_projim.png}}}
-    \caption{Projection to the scene plane from the viewpoint of the recovered projector (bottom left, black; virtual projector in center, gray) for a given target location.
-(a) Original image.
-(b) Projecting the original image to the scene plane introduces projective distortions if the projector is not fronto-parallel with the scene plane, as is the case here.}
-    \label{fig:proj}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \subfloat[\centering Our warped image.]{{\includegraphics[height=3.4cm]{images/projim_warped_scaled.png}}}
-    \qquad
-    \subfloat[\centering Projecting our warped image (downwards view).]{{\includegraphics[height=3.4cm]{images/proj_projim_waped.png}}}
-    \caption{Projection to the scene plane from the viewpoint of the recovered projector (bottom left, black; virtual projector in center, gray) for a given target location.
-(a) Original image warped using our approach.
-(b) After warping the original image according to our approach for the given target location, the image is projected with the effect of projecting the original image from the viewpoint of the virtual projector.
-This renders the projection free of projective distortions, aligned with the axes of the virtual camera, and at the desired metric scale.
-Note that pixels shown in black will in effect be ignored by the projector.}
-    \label{fig:warp}
-\end{figure}
-
 \paragraph{Plane-induced homography} Let $\mathtt{K}_\text{proj}$ express the calibration matrix of the recovered projector and $(\mathtt{R}, \mathbf{t}) \in SE(3)$ the rigid body transformation that transforms points from the coordinate frame of the recovered projector to that of the virtual projector, for a given target location.
@@
     \centering
-    \subfloat[\centering Hardware setup (projector, steerable mirror system, camera).]{{\includegraphics[height=3.3cm]{images/system2.pdf}}}
+    \subfloat[\centering Hardware setup (projector, steerable mirror system, camera).]{{\includegraphics[height=3.35cm]{images/system2.pdf}}}
     \quad
-    \subfloat[\centering Projection of warped image from Figure~\ref{fig:warp}(a) to corresponding target location.]{{\includegraphics[height=3.3cm]{images/hans_undist.pdf}}}
+    \subfloat[\centering Projection of warped image from Figure~\ref{fig:warp}(a) to corresponding target location.]{{\includegraphics[height=3.35cm]{images/hans_undist.pdf}}}
     \caption{Evaluation scenario.
-(a) Our hardware setup, comprised of a Panasonic PT-RZ660BE projector, a steerable mirror system manufactured by Dynamic Projection Institute, and a Stereolabs Zed 2 stereo camera, of which we used only the left view.
+(a) Our hardware setup, comprised of a projector, a steerable mirror system, and a stereo camera, of which we used only the left view.
 (b) Example augmentation produced by projecting---to one of our 15 target locations across the floorspace---an image warped using our approach, as seen from the downwards-facing camera.

### 2021-11-02 paper/elsarticle/elsarticle-template.tex
paper

@@
 \textit{Spatial} AR, in turn, is a form of AR whereby the augmentation of the scene is carried out using a projector, with the advantage of rendering the augmentation visible to all onlookers simultaneously without calling for any to hold a handheld device such as a tablet or for each to wear some form of head-mounted display.
-In carrying out spatial AR, however, care must be taken to appropriately warp an image to be projected in a manner that it, when projected, appear free of distortions to the viewer.
-For planar scene geometry (such as a floor, wall, or table), a manual process called keystone correction can be used to carry out an appropriate corrective image warp, a process that is cumbersome and time consuming.
-Another drawback of conventional projector-based spatial AR is that it is capable of augmenting only the portion of the scene within the projector's own static field of view, thereby hindering its applicability to use cases calling for augmentation of wide areas such as a factory floorspace.
+In carrying out spatial AR, however, care must be taken to appropriately warp the image to be projected such that it, when projected, appear free of projective distortions to the viewer.
+For planar scene geometry (such as a floor, wall, or table), a manual process called keystone correction can be used to carry out an appropriate corrective image warp, a process that can be cumbersome.
+Another drawback of conventional spatial AR relying only on a single projector is that it is capable of augmenting only the portion of the scene within the projector's static field of view, thereby hindering its applicability to use cases calling for augmentation of wide areas such as a factory floorspace.
 
@@
 Our system renders the placement of augmentations in the scene more intuitive than manual keystone correction in two ways.
-First, (i) the horizontal and vertical axes of the desired augmentations is carried out in accordance with the horizontal and vertical image axes of the camera, thereby making setting those axes as simple as appropriately rotating the camera relative to the scene.
-Second, (ii) the desired dimensions of the projected augmentations are specified in metric terms, thereby allowing for consistent scaling across all target locations.
+First, (i) the horizontal and vertical axes of the desired augmentations is carried out in accordance with the horizontal and vertical image axes of the camera, thereby making setting those axes as simple as appropriately rotating the camera relative to the scene plane, and thereby providing a consistent axial reference frame across all target locations.
+Second, (ii) the desired dimensions of the projected augmentations are specified in metric terms, thereby providing for consistent scaling, likewise across all target locations.
 %Our system produces such augmentations accurately, with compelling time savings relative manual keystone correction.
@@
 Augmented reality (AR) \cite{van2010survey,zhou2008trends} promises to enable use cases in industrial settings that include the embedding of assembly instructions directly into the scene~\cite{schlund2018moglichkeiten,uva2018evaluating,masood2019augmented,gattullo2019towards,aschenbrenner2019comparing,mayrhofer2019one,rupprecht2020information,Rupprecht2021}, potentially reducing or altogether doing away with the need for workers to refer to such instructions in paper form or on a statically situated screen.
-Typically, AR works by embedding the augmentation in an image of the scene acquired from the viewpoint of a single individual, with the resulting augmented image in turn displayed typically using some form of handheld or head-mounted display.
+Typically, AR works by embedding the augmentation in imagery of the scene acquired from the viewpoint of a single individual, with the resulting augmented image in turn displayed typically using some form of handheld or head-mounted display.
 Handheld devices such as tablets do not enable the user to work hands-free while using the device, or force the user to wear additional gear.
@@
 In turn, a form of augmented reality referred to as \textit{spatial} AR\cite{bimber2019spatial} is carried out not by embedding the augmentation in an image of the scene as with a head-mounted or handheld display, but by projection to the scene itself, thus eliminating the aforementioned problems.
-Spatial AR thus turns surfaces such as the product being manufactured, a desk, the floor, a wall, or even the ceiling (considering overhead work) into displays.
+Spatial AR thus turns surfaces such as the object being manufactured, a desk, the floor, a wall, or even the ceiling (considering overhead work) into displays.
 Information provision---of the work plan or of quality-related information---can be carried out by projection to the given workspace, while directly referring to exact positions within that workspace.
-Moreover, the fact that this information can be viewed by multiple users simultaneously serves to promote collaboration \cite{aschenbrenner2019comparing}.
+Moreover, the fact that this information can be viewed by multiple users simultaneously serves to facilitate collaboration \cite{aschenbrenner2019comparing}.
 
@@
     \qquad
-    \subfloat[\centering Projecting the original image (downwards view).]{{\includegraphics[height=3.4cm]{images/proj_projim.png}}}
-    \caption{Projection to the scene plane from the viewpoint of the recovered projector, facing a given target location.
+    \subfloat[\centering Projection to target location in scene plane (downwards view).]{{\includegraphics[height=3.4cm]{images/proj_projim.png}}}
+    \caption{Projection to a given target location in the scene plane (e.g., the floor, a wall, or table).
 (a) Original image.
-(b) Projecting the original image from the viewpoint of the recovered projector (bottom left, black) to the scene plane introduces projective distortions if the projector is not fronto-parallel with the scene plane, as is the case here.}
+(b) Projecting the original image from the viewpoint of the projector (bottom left, black) to the scene plane introduces projective distortions if the projector is not fronto-parallel with the scene plane, as is the case here.}
     \label{fig:proj}
@@
 While interest in spatial AR in industrial settings has been growing, its deployment in industrial settings remains well behind that of wearable AR systems such as handhelds or head-mounted displays.
-One of the main shortcomings of typical projector-based spatial AR relative to AR relying on wearables is the applicability of spatial AR only to augmenting scenes within the projector's own static field of view.
-Here, dynamic projection systems---using mirror heads to deflect the projection across wider areas---has the potential to bridge the gap to wider industrial penetration.
-Yet the increase in the area to be covered comes with noteworthy challenges in terms of ensuring that the images to be projected appear free of distortions to the users.
+One of the main shortcomings of typical projector-based spatial AR relative to AR relying on wearables is the ability to project augmentations only within the projector's own static field of view.
+Here, dynamic projection systems---using steerable mirror heads to direct the projection across wider areas---has the potential to bridge the gap to wider industrial penetration.
+Yet the increase in the area to be covered comes with noteworthy challenges in terms of preparing the augmentations, in particular as the number of target locations in the scene becomes larger than a small handful.
 Considering a planar surface to be augmented (e.g., a floor, wall, or table), unless the projector faces the surface frontally, the bounds of a projected rectangular image will not appear rectangular, but will instead be subject to projective distortions (cf.\ Figure~\ref{fig:proj}).
@@
     \qquad
-    \subfloat[\centering Projecting our warped image (downwards view).]{{\includegraphics[height=3.4cm]{images/proj_projim_waped.png}}}
-    \caption{Projection to the scene plane from the viewpoint of the recovered projector for a given target location.
-(a) Original image warped using our approach.
-(b) After warping the original image according to our approach for the given target location, the image is projected from the viewpoint of the recovered projector (bottom left, black) with the effect of projecting the original image from the viewpoint of the virtual projector (center, gray).
+    \subfloat[\centering Projection to target location in scene plane (downwards view).]{{\includegraphics[height=3.4cm]{images/proj_projim_waped.png}}}
+    \caption{Projection to a given target location in the scene plane.
+(a) Original image (cf.\ Figure~\ref{fig:proj}(a)) warped with respect to the given target location using our approach.
+(b) Projecting the warped image from the viewpoint of the projector (bottom left, black) has the effect of projecting the \textit{original} image from the viewpoint of the virtual projector (center, gray).
 This renders the projection free of projective distortions, aligned with the axes of the virtual camera, and at the desired metric scale.
@@
 Our contribution is to propose a wide-area spatial AR system for planar scenes that produces the effect of keystone correction analytically, free of any manual interaction.
-Most importantly, we do this in a manner placing the axes of the augmentation in accordance with the axes of a camera placed to face directly towards the scene plane (a camera we shall in what follows refer to as `downwards facing', for brevity).
-This facilitates placement of augmentations, since establishing the principal axes of the augmentation thus reduces to placing the downwards-facing camera in a manner such that the $X$- and $Y$-axes of the image align with the intended $X$- and $Y$-axes of the augmentation.
+Notably, we do this in a manner placing the axes of the augmentation in accordance with the axes of a camera placed to face directly towards the scene plane (a camera we shall in what follows refer to as `downwards facing', for brevity).
+This facilitates placement of augmentations, since establishing the principal axes of the augmentation thus reduces to appropriately rotating the downwards-facing camera relative to the scene plane.
 We achieve this by warping the image to be projected using a plane-induced homography computed to produce the effect of projecting the image not from the actual projector viewpoint, but in accordance with the viewpoint of a \textit{virtual} projector (cf.\ Figure~\ref{fig:warp}).
@@
 Moreover, our system enables specifying the dimensions of augmentations in metric terms, which we achieve by placing the virtual projector at the appropriate height above the scene plane.
-Finally, our system set up to handle a projector equipped with a steerable mirror (without need for explicitly modeling the action of the steerable mirror on the projector), thereby enabling wide-area applications exceeding the immediate field of view of the projector without needing to rely on multiple projectors.
+Finally, our system is set up to handle a projector equipped with a steerable mirror in a manner that does not call for explicitly modeling the action of the steerable mirror on the projector, thereby enabling wide-area applications exceeding the immediate field of view of the projector without needing to rely on multiple projectors.
 
@@
 For the purposes of the prototype, a projector mounted with a steerable mirror was set up with a camera to demonstrate a variety of use cases, including collaborative assembly encompassing the projection of assembly instructions and an interactive projected user interface \cite{kjeldsen2002interacting,pinhanez2003applications}.
-The authors, however, carry out keystone correction manually, by interactively adjusting for the position and orientation of a virtual scene plane and a scaling factor and computing a 2D homography accordingly.
+The authors, however, carry out keystone correction manually, by interactively adjusting for the position and orientation of a virtual scene plane and a scaling factor and computing a corresponding 2D homography accordingly.
 %Rather than rely on a steerable mirror to support spatial AR to multiple locations using a single projector, some works mount the projector and a camera on a rigid rig and subject the rig to motion \cite{ehnes2004projected,borkowski2004spatial,butz2006applying}.
@@
 More generally, keystone correction for planar scenes can be said to reduce to computing a 2D homography \cite{Hartley2004}, an invertible transformation that preserves colinearity and is thus fittingly alternatively referred to as a `colineation'; this is the case whether computing the homography relies on manual interaction as in the case of Pinharez, or is obtained free of manual interaction as in ours.
-The intuition for why it is that a transformation that preserves colinearity---i.e., maps lines to lines---can serve to model an appropriate corrective image warp can be drawn from considering a planar chessboard pattern: looking at an image of a chessboard acquired from an oblique angle, one observes that lines parallel in the chessboard appear to meet in respective vanishing points; looking at an image of the same chessboard acquired frontally with respect to the plane of the chessboard, lines parallel in the chessboard appear parallel in the image (i.e., they are said to meet `at infinity').
-To warp the former image (where lines parallel in the scene meet in respective vanishing points) such that the lines of the chessboard appear as in the latter image (where lines parallel in the scene meet at infinity) calls for a transformation that maps lines to lines.
+The intuition for why it is that a transformation that preserves colinearity---i.e., maps lines to lines---can serve to model an appropriate corrective image warp can be drawn from considering a planar chessboard pattern: looking at a chessboard from an oblique angle, one observes that lines parallel in the chessboard appear to meet in respective vanishing points, yet remain lines; looking at the same chessboard frontally with respect to the plane of the chessboard, lines parallel in the chessboard appear parallel in the image (i.e., they are said to meet `at infinity').
+To warp an image of the chessboard acquired at one angle such that the lines of the chessboard appear as if the chessboard were acquired at another angle accordingly calls for a transformation that maps lines to lines; it therefore follows that there exists a homography that models that transformation.
 
@@
 Such a homography is said to be `induced' by a plane \cite{Hartley2004}.
-We exploit the fact that cameras and projectors alike can be modeled using a pinhole camera model \cite{bimber2019spatial}, proceed to recover the relative geometry relating the scene plane and projector, and compute a particular form of plane-induced homography to effect the intended corrective image warp, taking into account both desired in-plane orientation and metric scale.
+We exploit the fact that cameras and projectors alike can be modeled using a pinhole camera model \cite{bimber2019spatial}, proceed to recover the relative geometry relating the scene plane and projector, and compute a particular form of plane-induced homography to effect the intended corrective image warp, yet additionally taking into account both desired in-plane orientation and metric scale.
 
@@
 
-Correcting for projective distortions of the sort outlined in Section~\ref{sec:intro} can be achieved by modeling the manner in which the respective rays through the pixels of the projector's image plane fan out into the scene (i.e., by `calibrating' the projector) and the geometry of the scene itself (i.e., by recovering the scene plane relative to the projector) within at least the projector's field of view.
-This is because the scene point `illuminated' by a pixel in the projector's image plane can be recovered by intersecting its corresponding ray with the geometry of the scene surface; modeling the geometry that relates projector and scene plane thus enables one to reason about the problem of correcting for distortions in terms of geometry.
+Correcting for projective distortions of the sort outlined in Section~\ref{sec:intro}---and exemplified in Figure~\ref{fig:proj}---can be achieved by means of an appropriate plane-induced homography \cite{Hartley2004}, computed by modeling the manner in which the respective rays through the pixels of the projector's image plane fan out into the scene (i.e., by `calibrating' the projector) and the geometry of the scene itself (i.e., by recovering the scene plane relative to the projector) within at least the projector's field of view.
+This is because the scene point `illuminated' by a pixel in the projector's image plane can be recovered by intersecting its corresponding ray with the geometry of the scene plane; modeling the geometry that relates projector and scene plane thus enables one to reason about the problem of correcting for distortions in terms of geometry.
 To model this interaction, we (i) carry out a one-time projector calibration, which in our approach calls for additionally calibrating a camera facing downwards to the scene plane and includes recovery of the scene plane as a convenient side effect.
-Next, we use the relative camera-projector-scene plane geometry to (ii) compute a plane-induced homography that warps the image to be projected in a manner that it appear undistorted to the viewer, and placed in alignment with the axes of a proxy image of the scene.
+Next, we use the relative camera-projector-scene plane geometry to (ii) compute a plane-induced homography that warps the image to be projected in a manner that it appear undistorted to the viewer, placed in alignment with the axes of a proxy image of the scene, and at the desired metric scale.
 These two points are treated in Sections~\ref{sec:approach:geometry} and \ref{sec:approach:homography}, respectively.
@@
 \end{equation}
-where $f$ and $(x_0, y_0)^\top$ are both expressed in units of pixels.
-The calibration matrix~$\mathtt{K}$ models a so-called pinhole camera; such a model is applicable to an image acquired using a real-world camera assuming the image has been corrected for lens distortion effects, by having applied the appropriate lens distortion model coefficients.
+where $f$, $x_0$, and $y_0$ are each expressed in units of pixels.
+The calibration matrix~$\mathtt{K}$ models a so-called pinhole camera \cite{Hartley2004}; such a model is applicable to an image acquired using a real-world camera assuming the image has in a prior step been corrected for lens distortion effects.
+%, by having applied the appropriate lens distortion model coefficients.
 
@@
 
-Calibrating a projector \cite{moreno2012simple,zhang2009projector,zhang2006novel} can be carried out in precisely the same manner as calibrating a camera insofar as step (ii) is concerned; if proceeding accordingly, the major difference in projector calibration relative to the calibrating a camera concerns the manner in which 2D-3D correspondences are identified, i.e., between pixels in the image plane of the \textit{projector} and the corresponding points in the scene.
-What remains of this section is concerned primarily with the recovery of 2D-3D correspondences, first in support of calibrating the downwards-facing camera (used to obtain the camera calibration matrix~$\mathtt{K}_\text{cam}$), then---in a manner that makes use of $\mathtt{K}_\text{cam}$---for calibrating the projector (used for obtaining the projector calibration matrix~$\mathtt{K}_\text{proj}$ and the projector poses~($\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j) \in SE(3)$, for each projector calibration image).
+Calibrating a projector \cite{moreno2012simple,zhang2009projector,zhang2006novel} can be carried out in precisely the same manner as calibrating a camera insofar as step (ii) is concerned; the major difference in projector calibration relative to the calibrating a camera then concerns the manner in which 2D-3D correspondences are identified, i.e., between pixels in the image plane of the \textit{projector} and the corresponding points in the scene.
+What remains of this section is concerned primarily with the recovery of 2D-3D correspondences, first in support of calibrating the downwards-facing camera (giving the camera calibration matrix~$\mathtt{K}_\text{cam}$), then---in a manner that makes use of $\mathtt{K}_\text{cam}$---for calibrating the projector (giving the projector calibration matrix~$\mathtt{K}_\text{proj}$ and the projector poses~($\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j) \in SE(3)$, for each projector calibration image).
 
@@
     \qquad
-    \subfloat[\centering 3D positions for projector calibration (red, in scene plane).]{{\includegraphics[height=3cm]{images/2d3d.png}}}
-    \caption{Obtaining the 3D positions of the 2D-3D correspondences for projector calibration.
+    \subfloat[\centering 3D positions for use in projector calibration (red, in scene plane).]{{\includegraphics[height=3cm]{images/2d3d.png}}}
+    \caption{Obtaining the 3D positions of the 2D-3D correspondences for use in projector calibration.
 (a) For each target location, a projector calibration image is acquired from the viewpoint of the downwards-facing camera, containing a projected asymmetric circle pattern at the target location and a chessboard pattern placed on the floor nearby.
-Automatically detected circle pattern center points and chessboard corners overlain for illustration.
-(b) Given a projector calibration image, the scene plane (gray) is recovered in the coordinate frame of the camera (black, up vector red) via spatial resection with respect to 2D-3D correspondences obtained using the chessboard pattern (blue); the 3D positions of the 2D-3D correspondences to be used for calibrating the projector (red, in scene plane) are obtained by intersection with the scene plane of back-projections (rays, likewise gray) of the 2D center points of the circle pattern detected in the image plane of the camera.} %The corresponding 2D positions are detected in the image of the asymmetric circles pattern image projected by the projector (not shown).
+Circle pattern center points and chessboard corners detected automatically (overlain for illustration).
+(b) Given a projector calibration image, the scene plane is recovered in the coordinate frame of the camera (black) via spatial resection with respect to 2D-3D correspondences obtained using the chessboard pattern (blue); the 3D positions of the 2D-3D correspondences to be used in calibrating the projector (red, in scene plane) are obtained by intersection with the scene plane of back-projections (rays, likewise gray) of the 2D center points of the circle pattern detected in the image plane of the camera.} %The corresponding 2D positions are detected in the image of the asymmetric circles pattern image projected by the projector (not shown).
 The downwards-facing camera is shown relative to the scene plane in terms of its camera frustum (black, with up vector in red).
@@
 Note further that a consequence of placing the virtual projector by rotating about the point of intersection of the projector's optical axis with the scene plane is that \textit{the center of the projector's image plane remains invariant} to the placement of the virtual projector.
-A steerable mirror can be aimed with respect to a point projected from the center of the projector's image plane, thereby facilitating placement of target locations.
+A steerable mirror can be aimed with respect to a point projected from the center of the projector's image plane, thereby facilitating placement.
 
@@
 \section{Evaluation}
-We evaluate our approach by using a 960 pixel~$\times$~600 pixel image of the TU Wien logo (cf.\ Figure \ref{fig:warp}) to augment 15 locations across the floorspace at the Pilotfabrik\footnote{\url{https://www.pilotfabrik.at/}} of TU Wien, a collaborative space for research on Industry 4.0 topics situated in Vienna, Austria.
-We contrast our approach with a baseline approach involving manual keystone correction, by aiming with both approaches to place the same image for each location aligned with the principal axes of the floorspace (i.e., the yellow and red lines in Figure~\ref{fig:eval}(b)), absent of projective distortions, and with the same metric dimensions of 50~cm~$\times$~31.25~cm\footnote{The dimensions in pixels of the image we project are 960~$\times$~600; we chose for our experiments to set the projected metric length of the horizontal axis of the image to 50~cm, which implies 31.25~cm for the vertical axis if aspect ratio is to be preserved.}.
+We evaluate our approach by using a 960 pixel~$\times$~600 pixel image of the TU Wien logo (cf.\ Figure \ref{fig:warp}(a)) to augment 15 locations across the floorspace at the Pilotfabrik\footnote{\url{https://www.pilotfabrik.at/}} of TU Wien, a collaborative space for research on Industry 4.0 topics situated in Vienna, Austria.
+We contrast our approach with a baseline approach involving manual keystone correction, by aiming with both approaches to place the same image for each location aligned with the principal axes of the floorspace (i.e., the yellow and red lines in Figure~\ref{fig:2d}(b)), absent of projective distortions, and with the same metric dimensions of 50~cm~$\times$~31.25~cm\footnote{The dimensions in pixels of the image we project are 960~$\times$~600; we chose for our experiments to set the projected metric length of the horizontal axis of the image to 50~cm, which implies 31.25~cm for the vertical axis if aspect ratio is to be preserved.}.
 All experiments were carried out by the same technician, experienced in both approaches.
 
-The hardware setup employed in the evaluation (cf.\ Figure \ref{fig:eval}(a)) comprised a Panasonic PT-RZ660BE projector with a steerable mirror system---used in our experiments to point the projection to each of the 15 locations---manufactured by Dynamic Projection Institute \cite{rupprecht2020information,Rupprecht2021}.
+The hardware setup employed in the evaluation (cf.\ Figure \ref{fig:eval}(a)) comprised a Panasonic PT-RZ660BE projector (Panasonic Corporation, Japan) with a steerable mirror system---used in our experiments to point the projection to each of the 15 locations---manufactured by Dynamic Projection Institute (Dynamic Projection Institute, Austria) \cite{rupprecht2020information,Rupprecht2021}.
 The steerable mirror system was bundled with the MDC-X software for steering the mirror, loading imagery, and optionally carrying out manual keystone correction, such that each position and (warped) image can be registered as a preset.
-In addition, we used a downwards-facing Zed 2 stereo camera manufactured by Stereolabs, yet relied only on the left view.
+In addition, we used a downwards-facing Stereolabs Zed 2 stereo camera (Stereolabs Inc., France), yet relied only on the left view.
 The floorspace used for our experiments measured dimensions of ca.\ 6~m~$\times$~4~m; the projector was mounted at approximately the center of this space, at a height of ca.\ 3.5 m.

### 2021-11-07 paper/elsarticle/elsarticle-template.tex
paper

@@
     \qquad
-    \subfloat[\centering Projection to target location in scene plane (downwards view).]{{\includegraphics[height=3.4cm]{images/proj_projim.png}}}
-    \caption{Projection to a given target location in the scene plane (e.g., the floor, a wall, or table).
-(a) Original image.
-(b) Projecting the original image from the viewpoint of the projector (bottom left, black) to the scene plane introduces projective distortions if the projector is not fronto-parallel with the scene plane, as is the case here.}
+    \subfloat[\centering Projection to target location in scene plane (view downwards to scene plane).]{{\includegraphics[height=3.4cm]{images/proj_projim.png}}}
+    \caption{Na\"ive projection to a given target location in the scene plane (e.g., the floor, a wall, or table).
+(a) Original image, which we wish to project to the scene plane.
+(b) Projecting the original image to the scene plane from the viewpoint of the projector directed to the target location (bottom left, black with up vector red) introduces projective distortions if the projector is not fronto-parallel with the scene plane (as is the case here), with orientation and scale varying as a function of target location.}
     \label{fig:proj}
@@
     \qquad
-    \subfloat[\centering Projection to target location in scene plane (downwards view).]{{\includegraphics[height=3.4cm]{images/proj_projim_waped.png}}}
-    \caption{Projection to a given target location in the scene plane.
-(a) Original image (cf.\ Figure~\ref{fig:proj}(a)) warped with respect to the given target location using our approach.
-(b) Projecting the warped image from the viewpoint of the projector (bottom left, black) has the effect of projecting the \textit{original} image from the viewpoint of the virtual projector (center, gray).
+    \subfloat[\centering Projection to target location in scene plane (view downwards to scene plane).]{{\includegraphics[height=3.4cm]{images/proj_projim_waped.png}}}
+    \caption{Projection to a given target location in the scene plane, correcting for projective distortions using our approach.
+(a) Original image (cf.\ Figure~\ref{fig:proj}(a)) warped using our approach with respect to the given target location.
+(b) Projecting the warped image to the scene plane from the viewpoint of the projector directed to the target location (bottom left, black with up vector red; same as in Figure~\ref{fig:proj}(b)) has the effect of projecting the \textit{original} image from the viewpoint of the \textit{virtual} projector (center, gray with up vector red).
 This renders the projection free of projective distortions, aligned with the axes of the virtual camera, and at the desired metric scale.
@@
 Our contribution is to propose a wide-area spatial AR system for planar scenes that produces the effect of keystone correction analytically, free of any manual interaction.
-Notably, we do this in a manner placing the axes of the augmentation in accordance with the axes of a camera placed to face directly towards the scene plane (a camera we shall in what follows refer to as `downwards facing', for brevity).
-This facilitates placement of augmentations, since establishing the principal axes of the augmentation thus reduces to appropriately rotating the downwards-facing camera relative to the scene plane.
-We achieve this by warping the image to be projected using a plane-induced homography computed to produce the effect of projecting the image not from the actual projector viewpoint, but in accordance with the viewpoint of a \textit{virtual} projector (cf.\ Figure~\ref{fig:warp}).
-We place this virtual projector to (i) face directly downwards to the target location in the scene plane and (ii) rotate its axes in accordance with those of the camera.
+Notably, we do this in a manner placing the axes of the augmentation in accordance with the axes of a camera placed to face towards the scene plane, a camera we shall in what follows refer to as `downwards facing', for brevity.
+To compensate for the likely possibility that the camera not face downwards to the scene plane precisely, we rotate the recovered `downwards-facing' camera to face downwards directly, yielding what we refer to as the \textit{virtual} camera.
+Relying on the placement of the `downwards-facing' camera thus facilitates placement of augmentations, since setting the desired principal axes of augmentations with respect to the scene plane thus reduces to appropriately rotating the `downwards-facing' camera relative to the scene plane.
+We compute our corrective image warp for a given target location by warping the image to be projected using a plane-induced homography computed to produce the effect of projecting the image not from the actual projector viewpoint, but in accordance with the viewpoint of a \textit{virtual} projector (cf.\ Figure~\ref{fig:warp}).
+We place this virtual projector to (i) face directly downwards to the target location in the scene plane and (ii) rotate its axes to match those of the virtual camera.
 Moreover, our system enables specifying the dimensions of augmentations in metric terms, which we achieve by placing the virtual projector at the appropriate height above the scene plane.
-Finally, our system is set up to handle a projector equipped with a steerable mirror in a manner that does not call for explicitly modeling the action of the steerable mirror on the projector, thereby enabling wide-area applications exceeding the immediate field of view of the projector without needing to rely on multiple projectors.
+Finally, our system is set up for use with a projector equipped with a steerable mirror in a manner that does not call for explicitly modeling the action of the steerable mirror on the projector, thereby enabling wide-area applications exceeding the immediate field of view of the projector without needing to rely on multiple projectors.
 
@@
 This is because the scene point `illuminated' by a pixel in the projector's image plane can be recovered by intersecting its corresponding ray with the geometry of the scene plane; modeling the geometry that relates projector and scene plane thus enables one to reason about the problem of correcting for distortions in terms of geometry.
-To model this interaction, we (i) carry out a one-time projector calibration, which in our approach calls for additionally calibrating a camera facing downwards to the scene plane and includes recovery of the scene plane as a convenient side effect.
+To model this interaction, we (i) carry out a one-time projector calibration, which in our approach calls for additionally calibrating a camera facing downwards to the scene plane and includes recovery of the scene plane.
 Next, we use the relative camera-projector-scene plane geometry to (ii) compute a plane-induced homography that warps the image to be projected in a manner that it appear undistorted to the viewer, placed in alignment with the axes of a proxy image of the scene, and at the desired metric scale.
@@
 %, by having applied the appropriate lens distortion model coefficients.
+
+\begin{figure}[ht]
+    \centering
+    \subfloat[\centering Projector calibration image (2D pattern position detections overlain).]{{\includegraphics[height=3.1cm]{images/detections.png} }}
+    \qquad
+    \subfloat[\centering Corresponding 3D positions for use in projector calibration (red, in scene plane).]{{\includegraphics[height=3.1cm]{images/2d3d.png}}}
+    \caption{Obtaining the 3D positions of the 2D-3D correspondences for use in projector calibration.
+(a) For each target location, a projector calibration image is acquired from the viewpoint of the `downwards-facing' camera, containing a projected asymmetric circle pattern at the target location and a chessboard pattern placed on the floor nearby.
+Circle pattern center points and chessboard corners detected automatically (overlain for illustration).
+(b) Given a projector calibration image, the scene plane is recovered in the coordinate frame of the camera (black with up vector red) via spatial resection with respect to 2D-3D correspondences obtained using the chessboard pattern (blue); the 3D positions of the 2D-3D correspondences to be used in calibrating the projector (red, in scene plane) are obtained by intersection with the scene plane of the back-projections (rays, likewise gray) of the 2D center points of the circle pattern detected in the image plane of the camera.} %The corresponding 2D positions are detected in the image of the asymmetric circles pattern image projected by the projector (not shown).
+The `downwards-facing' camera is shown relative to the scene plane in terms of its camera frustum (black, with up vector in red).
+    \label{fig:2d}
+\end{figure}
 
@@
 Calibrating a projector \cite{moreno2012simple,zhang2009projector,zhang2006novel} can be carried out in precisely the same manner as calibrating a camera insofar as step (ii) is concerned; the major difference in projector calibration relative to the calibrating a camera then concerns the manner in which 2D-3D correspondences are identified, i.e., between pixels in the image plane of the \textit{projector} and the corresponding points in the scene.
-What remains of this section is concerned primarily with the recovery of 2D-3D correspondences, first in support of calibrating the downwards-facing camera (giving the camera calibration matrix~$\mathtt{K}_\text{cam}$), then---in a manner that makes use of $\mathtt{K}_\text{cam}$---for calibrating the projector (giving the projector calibration matrix~$\mathtt{K}_\text{proj}$ and the projector poses~($\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j) \in SE(3)$, for each projector calibration image).
+What remains of this section is concerned primarily with the recovery of 2D-3D correspondences, first in support of calibrating the `downwards-facing' camera (giving the camera calibration matrix~$\mathtt{K}_\text{cam}$), then---in a manner that makes use of $\mathtt{K}_\text{cam}$---for calibrating the projector (giving the projector calibration matrix~$\mathtt{K}_\text{proj}$ and the projector poses~($\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j) \in SE(3)$, for each projector calibration image).
 
-\paragraph{2D-3D correspondences for camera calibration} We recover 2D-3D correspondences in support of calibrating the downwards-facing camera by relying on a planar calibration surface to automatically identify correspondences between the 3D points on the calibration surface and their 2D correspondences in the image plane.
+\paragraph{2D-3D correspondences for camera calibration} We recover 2D-3D correspondences in support of calibrating the `downwards-facing' camera by relying on a planar calibration surface to automatically identify correspondences between the 3D points on the calibration surface and their 2D correspondences in the image plane.
 The classical calibration surface is a chessboard pattern, and is such a pattern that we use as well.
@@
 
-\begin{figure}[ht]
-    \centering
-    \subfloat[\centering Projector calibration image (detections overlain).]{{\includegraphics[height=3cm]{images/detections.png} }}
-    \qquad
-    \subfloat[\centering 3D positions for use in projector calibration (red, in scene plane).]{{\includegraphics[height=3cm]{images/2d3d.png}}}
-    \caption{Obtaining the 3D positions of the 2D-3D correspondences for use in projector calibration.
-(a) For each target location, a projector calibration image is acquired from the viewpoint of the downwards-facing camera, containing a projected asymmetric circle pattern at the target location and a chessboard pattern placed on the floor nearby.
-Circle pattern center points and chessboard corners detected automatically (overlain for illustration).
-(b) Given a projector calibration image, the scene plane is recovered in the coordinate frame of the camera (black) via spatial resection with respect to 2D-3D correspondences obtained using the chessboard pattern (blue); the 3D positions of the 2D-3D correspondences to be used in calibrating the projector (red, in scene plane) are obtained by intersection with the scene plane of back-projections (rays, likewise gray) of the 2D center points of the circle pattern detected in the image plane of the camera.} %The corresponding 2D positions are detected in the image of the asymmetric circles pattern image projected by the projector (not shown).
-The downwards-facing camera is shown relative to the scene plane in terms of its camera frustum (black, with up vector in red).
-    \label{fig:2d}
-\end{figure}
-
 \paragraph{2D-3D correspondences for projector calibration} For projector calibration, we obtain the 2D-3D correspondences by \textit{projecting} a calibration pattern to the scene plane, once for each target location.
-Since we know that the 3D positions of the projected pattern must lie in the scene plane, we can recover the 3D correspondence of a pattern point detected in the 2D image plane of the downwards-facing camera by intersecting the back-projection of the 2D position with the scene plane.
+Since we know that the 3D positions of the projected pattern must lie in the scene plane, we can recover the 3D correspondence of a pattern point detected in the 2D image plane of the `downwards-facing' camera by intersecting the back-projection of the 2D position with the scene plane.
 We reuse the chessboard pattern---used above for camera calibration---to recover the scene plane locally to each target location via spatial resection \cite{Hartley2004}, by placing the chessboard pattern on the floor near the projected pattern and using the resulting 2D-3D chessboard correspondences as input to a PnP algorithm \cite{collins2014infinitesimal}.
@@
 
-Given a projector calibration image acquired using the downwards-facing camera, we detect the circle centers of the \textit{projected} asymmetrical circle pattern (cf.\ Figure~\ref{fig:2d}(a)); given the scene plane and such a 2D circle center $\mathbf{x}$, we obtain its 3D correspondence by intersecting the back-projection
+Given a projector calibration image acquired using the `downwards-facing' camera, we detect the circle centers of the \textit{projected} asymmetrical circle pattern (cf.\ Figure~\ref{fig:2d}(a)); given the scene plane and such a 2D circle center $\mathbf{x}$, we obtain its 3D correspondence by intersecting the back-projection
 \begin{equation}
@@
 
-Recall that the pose~($\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j) \in SE(3)$ of the projector, for each projector calibration image~$j$, is provided alongside $\mathtt{K}_\text{proj}$, relative to the 3D points of the pattern and thus in the coordinate frame of the downwards-facing camera.
+Recall that the pose~($\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j) \in SE(3)$ of the projector, for each projector calibration image~$j$, is provided alongside $\mathtt{K}_\text{proj}$, relative to the 3D points of the pattern and thus in the coordinate frame of the `downwards-facing' camera.
 Note that for a fixed projector with steerable mirror, given a projector calibration image, the recovered projector's pose is the pose the projector would have to have had to project to the given target location \textit{in the absence of the mirror} (i.e., facing directly towards the target location).
@@
     \qquad
-    \subfloat[\centering Downwards view.]{{\includegraphics[height=4.2cm]{images/3_above3_cam.png}}}
+    \subfloat[\centering View downwards to scene plane.]{{\includegraphics[height=4.2cm]{images/3_above3_cam.png}}}
     \caption{Obtaining the virtual camera.
-Our physical downwards-facing camera---in terms of which we wish to orient our augmentations---is almost certain to not face downwards precisely.
-Accordingly, we rotate the recovered downwards-facing camera (black) about its center of projection such that its optical axis be made parallel with the scene plane's normal with respect to the minimum arc-length rotation.
+Our physical `downwards-facing' camera---in terms of whose rotation relative to the scene plane we wish to orient our augmentations---is almost certain to not face downwards precisely.
+Accordingly, we rotate the recovered `downwards-facing' camera (black) about its center of projection such that its optical axis be made parallel with the scene plane's normal, with respect to the minimum arc-length rotation.
 It is then, for each respective target position, the $X$- and $Y$-axes of the resulting \textit{virtual} camera (gray) that we subsequently use to orient our virtual projector.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
@@
     \qquad
-    \subfloat[\centering Downwards view.]{{\includegraphics[height=4.2cm]{images/3_above3_proj.png}}}
+    \subfloat[\centering View downwards to scene plane.]{{\includegraphics[height=4.2cm]{images/3_above3_proj.png}}}
     \caption{Obtaining the virtual projector.
-For a given target location, the virtual projector is obtained by (i) rotating the recovered projector (bottom left, black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (ii) rotating the $X$- and $Y$-axes to align them with those of the virtual camera (cf.\ Figure \ref{fig:virtualcam}), and (iii) translating along the normal direction to achieve the desired metric projected image dimensions.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
+For a given target location, the virtual projector is obtained by (i) rotating the recovered projector (bottom left, black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (ii) rotating the $X$- and $Y$-axes to align them with those of the virtual camera guaranteed---in contrast to the `downwards-facing' camera---to face \textit{directly} downwards to the scene plane (cf.\ Figure \ref{fig:virtualcam}), and (iii) translating along the normal direction to achieve the desired metric projected image dimensions.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
     \label{fig:virtualproj}
@@
 We carry out this placement according to a small handful of steps.
-First, we rotate the camera about its center of projection to align its optical axis\footnote{Strictly speaking, the optical axis is the back-projection of the principal point; in our usage, we understand it to refer to the ray through center of the image plane.} with the normal vector of the scene plane by applying the minimum arc-length rotation that relates the two, giving a \textit{virtual camera} facing directly\footnote{A physical camera placed to face downwards is almost certain to not face downwards precisely; in contrast, the virtual camera's optical axis is aligned exactly with the scene plane's normal vector, rendering it genuinely fronto-parallel with respect to the scene plane.} downwards to the scene plane (cf.\ Figure~\ref{fig:virtualcam}).
-It is to the virtual camera's $X$- and $Y$-axes that we subsequently set the $X$- and $Y$-axes of the virtual projector.
-To obtain the virtual projector, we (i) intersect the scene plane with the optical axis and rotate the projector's placement about that point of intersection, aligning the optical axis with the scene plane's normal vector and (ii) align the $X$- and $Y$-axes with those of the virtual camera.
+It is in accordance with the $X$- and $Y$-axes of the `downwards-facing' camera that we wish to place the $X$- and $Y$-axes of the augmentation; recognizing that a physical camera placed to face downwards is almost certain to not face downwards precisely, we first rotate the camera about its center of projection to align its optical axis\footnote{Strictly speaking, the optical axis is the back-projection of the principal point; in our usage, we understand it to refer to the ray through center of the image plane.} with the normal vector of the scene plane by applying the minimum arc-length rotation that relates the two, giving a \textit{virtual} camera facing directly downwards to the scene plane (cf.\ Figure~\ref{fig:virtualcam}).
+To obtain the virtual projector, we then (i) intersect the scene plane with the optical axis and rotate the projector's placement about that point of intersection, aligning the optical axis with the scene plane's normal vector and (ii) align the $X$- and $Y$-axes with those of the virtual camera.
 Finally, we (iii) translate along the scene plane's normal, in order to satisfy the desired projected image dimensions provided in metric units.
-The resulting virtual projector is thus rendered fronto-parallel with the scene plane---enabling projection to the scene plane absent of projective distortions---and projects to the scene plane with the desired scale.
+These steps are illustrated in Figure~\ref{fig:virtualproj}.
+The resulting virtual projector is thus rendered fronto-parallel with the scene plane---enabling projection to the scene plane absent of projective distortions---and projects to the scene plane in the desired scale and with the the $X$- and $Y$-axes of the augmentation set in accordance with those of the `downwards-facing' camera.
+%\footnote{A physical camera placed to face downwards is almost certain to not face downwards precisely; in contrast, we align the virtual camera's optical axis exactly with the scene plane's normal vector, rendering it genuinely fronto-parallel with respect to the scene plane.}
 
-It is, strictly speaking, in the sense that we set the $X$- and $Y$-axes of the virtual projector to the $X$- and $Y$-axes of the virtual camera that we place the horizontal and vertical axes of the desired augmentations in accordance with the horizontal and vertical image axes of the camera; the placement of the camera thereby intuitively determines the principal axes according to which augmentations are to be placed.
+It is thus in the sense that we set the $X$- and $Y$-axes of the virtual projector to the $X$- and $Y$-axes of the virtual camera that we place the horizontal and vertical axes of the desired augmentations `in accordance' with the horizontal and vertical image axes of the `downwards-facing' camera; the placement of the camera thereby intuitively determines the principal axes according to which augmentations are to be placed.
 Note further that a consequence of placing the virtual projector by rotating about the point of intersection of the projector's optical axis with the scene plane is that \textit{the center of the projector's image plane remains invariant} to the placement of the virtual projector.
@@
 Moreover, let $(\mathbf{n}^\top, -d)^\top \in \mathbb{P}^3$ give the scene plane, expressed in the coordinate frame of the recovered projector, where $\mathbf{n} \in \mathbb{R}^3$ is the scene plane's normal vector and $d = \mathbf{n}^\top\mathbf{X}$ for any point~$\mathbf{X} \in \mathbb{R}^3$ in the plane, so that $(\mathbf{n}^\top, -d) (\mathbf{X}^\top, 1)^\top = 0$.
-The transformation that warps the image to be projected---by the recovered projector to the scene plane---such that it appear as if were projected to the scene plane by the virtual projector (cf.\ Figure~\ref{fig:warp}(b)) is given the by the $3~\times~3$ matrix
+The transformation that warps the image to be projected---by the recovered projector to the scene plane---such that it appear as if were projected to the scene plane by the virtual projector (cf.\ Figure~\ref{fig:warp}) is given the by the $3~\times~3$ matrix
 \begin{equation}
@@
 (a) Our hardware setup, comprised of a projector, a steerable mirror system, and a stereo camera, of which we used only the left view.
-(b) Example augmentation produced by projecting---to one of our 15 target locations across the floorspace---an image warped using our approach, as seen from the downwards-facing camera.
+(b) Example augmentation produced by projecting---to one of our 15 target locations across the floorspace---an image warped using our approach, as seen from the `downwards-facing' camera.
 Note that the warp places the axes of the augmentation in accordance with the axes of the camera, and that the dimensions of the augmentation are in line with the desired target dimensions (50~cm~$\times$~31.25~cm).
@@
 \section{Evaluation}
-We evaluate our approach by using a 960 pixel~$\times$~600 pixel image of the TU Wien logo (cf.\ Figure \ref{fig:warp}(a)) to augment 15 locations across the floorspace at the Pilotfabrik\footnote{\url{https://www.pilotfabrik.at/}} of TU Wien, a collaborative space for research on Industry 4.0 topics situated in Vienna, Austria.
-We contrast our approach with a baseline approach involving manual keystone correction, by aiming with both approaches to place the same image for each location aligned with the principal axes of the floorspace (i.e., the yellow and red lines in Figure~\ref{fig:2d}(b)), absent of projective distortions, and with the same metric dimensions of 50~cm~$\times$~31.25~cm\footnote{The dimensions in pixels of the image we project are 960~$\times$~600; we chose for our experiments to set the projected metric length of the horizontal axis of the image to 50~cm, which implies 31.25~cm for the vertical axis if aspect ratio is to be preserved.}.
+We evaluate our approach by using a 960 pixel~$\times$~600 pixel image of the TU Wien logo (cf.\ Figure \ref{fig:proj}(a)) to augment 15 locations across the floorspace at the Pilotfabrik\footnote{\url{https://www.pilotfabrik.at/}} of TU Wien, a collaborative space for research on Industry 4.0 topics situated in Vienna, Austria.
+We contrast our approach with a baseline approach involving manual keystone correction, by aiming with both approaches to place the same image of the TU Wien logo for each of the 15 locations, aligned with the principal axes of the floorspace (i.e., the yellow and red lines in Figure~\ref{fig:2d}(b)), absent of projective distortions, and with the same metric dimensions of 50~cm~$\times$~31.25~cm\footnote{The dimensions in pixels of the image we project are 960~$\times$~600; we chose for our experiments to set the projected metric length of the horizontal axis of the image to 50~cm, which implies 31.25~cm for the vertical axis if aspect ratio is to be preserved.}.
 All experiments were carried out by the same technician, experienced in both approaches.
@@
 The steerable mirror system was bundled with the MDC-X software for steering the mirror, loading imagery, and optionally carrying out manual keystone correction, such that each position and (warped) image can be registered as a preset.
-In addition, we used a downwards-facing Stereolabs Zed 2 stereo camera (Stereolabs Inc., France), yet relied only on the left view.
+In addition, we used a Stereolabs Zed 2 stereo camera (Stereolabs Inc., France) as our `downwards-facing' camera, making use of only its left view.
 The floorspace used for our experiments measured dimensions of ca.\ 6~m~$\times$~4~m; the projector was mounted at approximately the center of this space, at a height of ca.\ 3.5 m.
@@
 In order to provide a template for manually carrying out keystone correction, we prepared a rectangular 50~cm~$\times$~31.25~cm cutout of cardboard.
-For each target location, we placed this piece of cardboard in a manner that its axes were aligned with the principal axes of the scene (cf.\ the yellow and red lines in Figure \ref{fig:eval}(b)); we then manually warped the image for the given target location so that any corner of the augmentation deviated by at most 1 cm from the corresponding corner of the template.
+For each target location, we placed this piece of cardboard in a manner that its axes were aligned with the principal axes of the floorspace; we then manually warped the image of the TU Wien logo for the given target location so that any corner of the augmentation deviated by at most 1 cm from the corresponding corner of the template.
 End-to-end, the process of setting the presets in the MDC-X software having carried out keystone correction manually took ca.\ 32 min for the 15 target locations.
@@
 \paragraph{Our approach} We began by carrying out a calibration of the camera, acquiring 10 camera calibration images (cf.\ Section~\ref{sec:approach:geometry}) of a chessboard calibration pattern with $6~\times~4$ corners ($7~\times~5$ squares) and feeding the images as input to our camera calibration module.
-Separately, for each of the 15 target locations, we produced a projector calibration image (cf.\ again Section~\ref{sec:approach:geometry}) by projecting a $11~\times~4$ asymmetrical circle pattern image to the location in question using the steerable mirror, placing a chessboard pattern beside the projected pattern, and acquiring the image using the downwards-facing camera.
+Separately, for each of the 15 target locations, we produced a projector calibration image (cf.\ again Section~\ref{sec:approach:geometry}) by projecting a $11~\times~4$ asymmetrical circle pattern image to the target location in question by directing the projection using the steerable mirror, placing a chessboard pattern beside the projected pattern, and acquiring the image using the `downwards-facing' camera.
 We then fed these images alongside the output of the camera calibration module to our projector calibration module.
-For each of the target locations, the steerable mirror was made to point to that location, the asymmetrical circle pattern image was projected to the scene plane, an image was acquired using the camera, and the location was registered in the MDC-X software as a preset.
+For each of the target locations, the target location was registered in the MDC-X software as a preset.
 The output of the projector calibration module is a homography per input projector calibration image (cf.\ Section~\ref{sec:approach:homography}).
-Next, we warped the images to be projected to the respective locations using their corresponding homography, using a third dedicated custom module.
-These warped images were finally imported into the MDC-X software and associated with their respective location presets (cf.\ Figure~\ref{fig:eval}(b)).
+Next, we warped the image of the TU Wien logo to the respective target locations using the corresponding homography, using a third dedicated custom module responsible for warping an image according to a homography.
+These warped images were finally imported into the MDC-X software and associated with their respective target location presets (cf.\ Figure~\ref{fig:eval}(b)).
 
@@
 Lengths were measured using a measuring tape for the top, bottom, left, and right sides of the augmentations; angles were measured using an angle gauge at the augmentations' top-left and bottom-right corners (the measuring tape and angle gauge are both visible in Figure~\ref{fig:eval}(b)).
-Note that if deviations in measured lengths are expressed in units of pixels of the downwards-facing camera, mean and standard deviation in no instance exceed 1 pixel.
+Note that if deviations in measured lengths are expressed in units of pixels of the `downwards-facing' camera, mean and standard deviation in no instance exceed 1 pixel.
 The total amount of time expended for carrying out all the above steps amounted to ca.\ 20 min, with ca.\ 2 min going to acquisition of the camera calibration images, and ca.\ 5 min going to that of projector calibration images.
@@
 Moreover, our method allows for specifying the desired dimensions of augmentations in metric terms, thus enabling consistent scaling.
-Our evaluation demonstrated our approach to produce compelling results---with mean and standard deviation of errors in measured lengths within the downwards-facing camera's pixel size---at less time than a more cumbersome manual approach to keystone correction.
+Our evaluation demonstrated our approach to produce compelling results---with mean and standard deviation of errors in measured lengths within the `downwards-facing' camera's pixel size---at less time than a more cumbersome manual approach to keystone correction.
 

### 2021-11-09 paper/elsarticle/elsarticle-template.tex
paper

@@
 To obtain the virtual projector, we then (i) intersect the scene plane with the optical axis and rotate the projector's placement about that point of intersection, aligning the optical axis with the scene plane's normal vector and (ii) align the $X$- and $Y$-axes with those of the virtual camera.
-Finally, we (iii) translate along the scene plane's normal, in order to satisfy the desired projected image dimensions provided in metric units.
-These steps are illustrated in Figure~\ref{fig:virtualproj}.
+Finally, we (iii) translate along the scene plane's normal, in order to satisfy the desired projected image dimensions provided in metric units; the height $h_\text{proj}$ [m] above the ground plane at which the virtual projector is to be placed if the image---with original image width of $w_\text{im}$ [px]---is to be projected to the scene plane with a metric width of $w$ [m] is given by
+\begin{equation}
+h_\text{proj} = w \times f_\text{proj} / w_\text{im},
+\end{equation}
+where $f_\text{proj}$ [px] is the focal length of the projector.
+These three steps are illustrated in Figure~\ref{fig:virtualproj}.
 The resulting virtual projector is thus rendered fronto-parallel with the scene plane---enabling projection to the scene plane absent of projective distortions---and projects to the scene plane in the desired scale and with the the $X$- and $Y$-axes of the augmentation set in accordance with those of the `downwards-facing' camera.
@@
 Note further that a consequence of placing the virtual projector by rotating about the point of intersection of the projector's optical axis with the scene plane is that \textit{the center of the projector's image plane remains invariant} to the placement of the virtual projector.
-A steerable mirror can be aimed with respect to a point projected from the center of the projector's image plane, thereby facilitating placement.
+A steerable mirror can be aimed with respect to a point projected from the center of the projector's image plane, thereby further facilitating placement.
 

### 2021-11-09 paper/elsarticle/elsarticle-template.tex
paper

@@
 To obtain the virtual projector, we then (i) intersect the scene plane with the optical axis and rotate the projector's placement about that point of intersection, aligning the optical axis with the scene plane's normal vector and (ii) align the $X$- and $Y$-axes with those of the virtual camera.
-Finally, we (iii) translate along the scene plane's normal, in order to satisfy the desired projected image dimensions provided in metric units; the height $h_\text{proj}$ [m] above the ground plane at which the virtual projector is to be placed if the image---with original image width of $w_\text{im}$ [px]---is to be projected to the scene plane with a metric width of $w$ [m] is given by
+Finally, we (iii) translate along the scene plane's normal, in order to satisfy the desired metric projected image dimensions; the height $h_\text{proj}$ [m] above the ground plane at which the virtual projector is to be placed if the image---with original image width of $w_\text{im}$ [px]---is to be projected to the scene plane with a metric width of $w$ [m] is given by
 \begin{equation}
@@
 (a) Our hardware setup, comprised of a projector, a steerable mirror system, and a stereo camera, of which we used only the left view.
-(b) Example augmentation produced by projecting---to one of our 15 target locations across the floorspace---an image warped using our approach, as seen from the `downwards-facing' camera.
+(b) Example augmentation produced by projecting---to one of the 15 target locations across the floorspace considered in our evaluation---an image warped using our approach, as seen from the `downwards-facing' camera.
 Note that the warp places the axes of the augmentation in accordance with the axes of the camera, and that the dimensions of the augmentation are in line with the desired target dimensions (50~cm~$\times$~31.25~cm).

### 2021-11-10 paper/elsarticle/elsarticle-template.tex
paper

@@
 In addition, we used a Stereolabs Zed 2 stereo camera (Stereolabs Inc., France) as our `downwards-facing' camera, making use of only its left view.
-The floorspace used for our experiments measured dimensions of ca.\ 6~m~$\times$~4~m; the projector was mounted at approximately the center of this space, at a height of ca.\ 3.5 m.
+The floorspace used for our experiments measured dimensions of ca.\ 6~m~$\times$~4~m; the projector and camera were mounted at approximately the center of this space, at a height of ca.\ 3.5 m.
 At this height, the pixel size (i.e., length or width of a pixel if projected to the scene plane) of the camera was ca.\ 3.5 mm.
 
-\paragraph{Manual approach} Having pointed the steerable mirror to the desired location, manual keystone correction using the MDC-X software involves warping the image to be projected by manipulating the corners of the image until the desired effect is produced on the projection surface.
-In order to provide a template for manually carrying out keystone correction, we prepared a rectangular 50~cm~$\times$~31.25~cm cutout of cardboard.
-For each target location, we placed this piece of cardboard in a manner that its axes were aligned with the principal axes of the floorspace; we then manually warped the image of the TU Wien logo for the given target location so that any corner of the augmentation deviated by at most 1 cm from the corresponding corner of the template.
-End-to-end, the process of setting the presets in the MDC-X software having carried out keystone correction manually took ca.\ 32 min for the 15 target locations.
-
-\begin{table}[ht]
-\caption{Means and standard deviations (in units of centimeters and pixels, respectively) of deviations in measured lengths from 50 cm for top and bottom and from 21.25 cm for left and right, with respect to the 15 augmentations produced using our approach  for 15 respective target locations (intended augmentation dimensions 50~cm~$\times$~31.25~cm).
-Pixel size given the downward-facing camera's height above the ground plane was 3.5 mm; note that in no instance did mean and standard deviation exceed 1 pixel from the viewpoint of the camera.}
+\begin{table}[ht!]
+\caption{Mean absolute error (MAE), minimum absolute error, and maximum absolute error in units of centimeters, with respect to the four sides (top, bottom, left, right) of the 15~augmentations produced using our approach for 15 target locations spread across the floorspace, respectively.
+Errors were computed against intended augmentation dimensions of 50~cm~$\times$~31.25~cm (i.e., 50~cm for top and bottom, 31.25~cm---rounded to 31.3~cm for the purposes of the computing the errors---for left and right).} %Note that absolute errors in no instance exceeded 1~cm for any side of any of the 15~augmentations.
 \label{table:length}
@@
 \midrule
-Mean {\tiny[cm]} & 0.03 & 0.07 & 0.09 & 0.06 \\
-Stdev {\tiny[cm]} & 0.33 & 0.17 & 0.17 & 0.17 \\
-\midrule
-Mean {\tiny[px]} & 0.1 & 0.19 & 0.26 & 0.18 \\
-Stdev {\tiny[px]} & 0.95 & 0.48 & 0.49 & 0.49 \\
+MAE {\tiny[cm]} & 0.23 & 0.17 & 0.22 & 0.2 \\
+Min AE {\tiny[cm]} & 0.0 & 0.0 & 0.0 & 0.0 \\
+Max AE {\tiny[cm]} & 0.8 & 0.6 & 0.5 & 0.5 \\
 \bottomrule
@@
 
-\begin{table}[ht]
-\caption{Means and standard deviations of deviations in measured angles from $90^\circ{}$ with respect to the 15 augmentations produced using our approach for 15 respective target locations.}
+\paragraph{Manual approach} Having pointed the steerable mirror to the desired location, manual keystone correction using the MDC-X software involves warping the image to be projected by manipulating the corners of the image until the desired effect is produced on the projection surface.
+In order to provide a template for manually carrying out keystone correction, we prepared a rectangular 50~cm~$\times$~31.25~cm cutout of cardboard.
+For each target location, we placed this piece of cardboard in a manner that its axes were aligned with the principal axes of the floorspace; we then manually warped the image of the TU Wien logo for the given target location so that any corner of the augmentation deviated by at most 1 cm from the corresponding corner of the template.
+End-to-end, the process of setting the presets in the MDC-X software having carried out keystone correction manually took ca.\ 32 min for the 15 target locations.
+
+\begin{table}[ht!]
+\caption{Mean absolute error (MAE), minimum absolute error, and maximum absolute error in units of degrees, with respect to the top-left and bottom-right corners of the 15~augmentations produced using our approach for 15 target locations spread across the floorspace, respectively.
+Errors computed in terms of deviation from $90^\circ{}$.}
 \label{table:angle}
@@
 \midrule
-Mean {\tiny[deg]} & -0.03 & 0.0 \\
-Stdev {\tiny[deg]} & 0.4 & 0.34 \\
+MAE {\tiny[deg]} & 0.17 & 0.32 \\
+Min AE {\tiny[deg]} & 0.0 & 0.0 \\
+Max AE {\tiny[deg]} & 0.9 & 0.5 \\
 \bottomrule
@@
 
-\paragraph{Our approach} We began by carrying out a calibration of the camera, acquiring 10 camera calibration images (cf.\ Section~\ref{sec:approach:geometry}) of a chessboard calibration pattern with $6~\times~4$ corners ($7~\times~5$ squares) and feeding the images as input to our camera calibration module.
-Separately, for each of the 15 target locations, we produced a projector calibration image (cf.\ again Section~\ref{sec:approach:geometry}) by projecting a $11~\times~4$ asymmetrical circle pattern image to the target location in question by directing the projection using the steerable mirror, placing a chessboard pattern beside the projected pattern, and acquiring the image using the `downwards-facing' camera.
+\paragraph{Our approach} We began by carrying out a calibration of the camera, acquiring 10~camera calibration images (cf.\ Section~\ref{sec:approach:geometry}) of a chessboard calibration pattern with $6~\times~4$ corners ($7~\times~5$ squares) and feeding the images as input to our camera calibration module.
+Separately, for each of the 15~target locations, we produced a projector calibration image (cf.\ again Section~\ref{sec:approach:geometry}) by projecting a $11~\times~4$ asymmetrical circle pattern image to the target location in question by directing the projection using the steerable mirror, placing a chessboard pattern beside the projected pattern, and acquiring the image using the `downwards-facing' camera.
 We then fed these images alongside the output of the camera calibration module to our projector calibration module.
@@
 
-Tables~\ref{table:length} and \ref{table:angle} give the deviations in measured lengths and angles, respectively, from the intended quantities.
-Lengths were measured using a measuring tape for the top, bottom, left, and right sides of the augmentations; angles were measured using an angle gauge at the augmentations' top-left and bottom-right corners (the measuring tape and angle gauge are both visible in Figure~\ref{fig:eval}(b)).
-Note that if deviations in measured lengths are expressed in units of pixels of the `downwards-facing' camera, mean and standard deviation in no instance exceed 1 pixel.
-The total amount of time expended for carrying out all the above steps amounted to ca.\ 20 min, with ca.\ 2 min going to acquisition of the camera calibration images, and ca.\ 5 min going to that of projector calibration images.
+Tables~\ref{table:length} and \ref{table:angle} give the errors in measured lengths and angles, respectively, with respect to the intended quantities.
+Lengths were measured using a measuring tape for the top, bottom, left, and right sides of the 15~augmentations; angles were measured using an angle gauge at the 15~augmentations' top-left and bottom-right corners.
+Note that absolute errors in lengths in no instance exceeded 1~cm for any side of any of the 15~augmentations, or $1^\circ{}$ for any of the measured corners.
+The total amount of time expended for carrying out all the above steps amounted to ca.\ 20~min, with ca.\ 2 min going to acquisition of the camera calibration images, and ca.\ 5 min going to that of projector calibration images.
 The remainder of the time was spent running our modules or working with the MDC-X software.

### 2021-11-12 paper/elsarticle/elsarticle-template.tex
paper

@@
 
+\begin{figure}
+    \centering
+    \includegraphics[height=3.75cm]{images/hans_undist_.pdf}
+    \caption{Projection of warped image from Figure \ref{fig:warp}(a) to corresponding target location, acquired by the `downwards-facing' camera.
+Note that besides correcting for projective distortions, the warp places the axes of the augmentation in accordance with the axes of the `downwards-facing' camera, and that the dimensions of the augmentation are in line with the desired target dimensions (50~cm~$\times$~31.25~cm).
+To provide context, the corners of the full projected image extent are projected to the floor in light green (emphasized with overlain dashed circles).}
+    \label{fig:final_proj}
+\end{figure}
+
 \paragraph{Plane-induced homography} Let $\mathtt{K}_\text{proj}$ express the calibration matrix of the recovered projector and $(\mathtt{R}, \mathbf{t}) \in SE(3)$ the rigid body transformation that transforms points from the coordinate frame of the recovered projector to that of the virtual projector, for a given target location.
@@
 a form of `plane-induced' 2D homography \cite{Hartley2004}.
-For convenience, we enable optional rotation of the image to be projected \textit{before} applying $\mathtt{H}$, about the image center; that rotation, parameterized in degrees, is thus in effect carried out intuitively with respect to the ground plane.
+For convenience, we enable optional rotation of the image to be projected \textit{before} applying $\mathtt{H}$, about the image center; that rotation, parameterized in degrees, is thus in effect carried out---again intuitively---with respect to the ground plane.
 
@@
     \centering
-    \subfloat[\centering Hardware setup (projector, steerable mirror system, camera).]{{\includegraphics[height=3.35cm]{images/system2.pdf}}}
+    \subfloat[\centering Hardware setup (projector, steerable mirror system, camera).]{{\includegraphics[height=3.6cm]{images/system3_.pdf}}}
     \quad
-    \subfloat[\centering Projection of warped image from Figure~\ref{fig:warp}(a) to corresponding target location.]{{\includegraphics[height=3.35cm]{images/hans_undist.pdf}}}
+    \subfloat[\centering Measurement for example target location, using a steel rule and a digital angle gauge.]{{\includegraphics[height=3.6cm]{images/meas2_.pdf}}}
     \caption{Evaluation scenario.
 (a) Our hardware setup, comprised of a projector, a steerable mirror system, and a stereo camera, of which we used only the left view.
-(b) Example augmentation produced by projecting---to one of the 15 target locations across the floorspace considered in our evaluation---an image warped using our approach, as seen from the `downwards-facing' camera.
-Note that the warp places the axes of the augmentation in accordance with the axes of the camera, and that the dimensions of the augmentation are in line with the desired target dimensions (50~cm~$\times$~31.25~cm).
-Corners of the full projected image extent projected to the floor in light green (emphasized with overlain dashed circles).}
+(b) Example augmentation produced by projecting---to one of the 15 target locations across the floorspace considered in our evaluation---an image warped using our approach.
+The angle gauge reads $89.6^\circ{}$, and the steel rule 49.9 cm.
+Corners of the full projected image extent projected to the floor in light green.}
     \label{fig:eval}
@@
 \section{Evaluation}
+
 We evaluate our approach by using a 960 pixel~$\times$~600 pixel image of the TU Wien logo (cf.\ Figure \ref{fig:proj}(a)) to augment 15 locations across the floorspace at the Pilotfabrik\footnote{\url{https://www.pilotfabrik.at/}} of TU Wien, a collaborative space for research on Industry 4.0 topics situated in Vienna, Austria.
@@
 Tables~\ref{table:length} and \ref{table:angle} give the errors in measured lengths and angles, respectively, with respect to the intended quantities.
-Lengths were measured using a measuring tape for the top, bottom, left, and right sides of the 15~augmentations; angles were measured using an angle gauge at the 15~augmentations' top-left and bottom-right corners.
+Lengths were measured using a steel rule for the top, bottom, left, and right sides of the 15~augmentations; angles were measured using a digital angle gauge at the 15~augmentations' top-left and bottom-right corners.
 Note that absolute errors in lengths in no instance exceeded 1~cm for any side of any of the 15~augmentations, or $1^\circ{}$ for any of the measured corners.

### 2021-11-12 paper/elsarticle/elsarticle-template.tex
paper

@@
 Moreover, let $(\mathbf{n}^\top, -d)^\top \in \mathbb{P}^3$ give the scene plane, expressed in the coordinate frame of the recovered projector, where $\mathbf{n} \in \mathbb{R}^3$ is the scene plane's normal vector and $d = \mathbf{n}^\top\mathbf{X}$ for any point~$\mathbf{X} \in \mathbb{R}^3$ in the plane, so that $(\mathbf{n}^\top, -d) (\mathbf{X}^\top, 1)^\top = 0$.
-The transformation that warps the image to be projected---by the recovered projector to the scene plane---such that it appear as if were projected to the scene plane by the virtual projector (cf.\ Figure~\ref{fig:warp}) is given the by the $3~\times~3$ matrix
+The transformation that warps the image to be projected---by the recovered projector to the scene plane---such that it appear as if were projected to the scene plane by the virtual projector (cf.\ Figures~\ref{fig:warp} and \ref{fig:final_proj}) is given the by the $3~\times~3$ matrix
 \begin{equation}
@@
     \quad
-    \subfloat[\centering Measurement for example target location, using a steel rule and a digital angle gauge.]{{\includegraphics[height=3.6cm]{images/meas2_.pdf}}}
+    \subfloat[\centering Measurement for augmentation, using a steel rule and a digital angle gauge.]{{\includegraphics[height=3.6cm]{images/meas2_.pdf}}}
     \caption{Evaluation scenario.
@@
 (b) Example augmentation produced by projecting---to one of the 15 target locations across the floorspace considered in our evaluation---an image warped using our approach.
-The angle gauge reads $89.6^\circ{}$, and the steel rule 49.9 cm.
-Corners of the full projected image extent projected to the floor in light green.}
+The angle gauge reads $89.6^\circ{}$ for the top-left corner, and the steel rule 49.9 cm for the bottom edge.
+Corners of the full projected image extent projected to the floor in light green to provide context.}
     \label{fig:eval}

### 2021-11-15 paper/elsarticle/elsarticle-template.tex
paper

@@
 To obtain the virtual projector, we then (i) intersect the scene plane with the optical axis and rotate the projector's placement about that point of intersection, aligning the optical axis with the scene plane's normal vector and (ii) align the $X$- and $Y$-axes with those of the virtual camera.
-Finally, we (iii) translate along the scene plane's normal, in order to satisfy the desired metric projected image dimensions; the height $h_\text{proj}$ [m] above the ground plane at which the virtual projector is to be placed if the image---with original image width of $w_\text{im}$ [px]---is to be projected to the scene plane with a metric width of $w$ [m] is given by
+Finally, we (iii) translate along the scene plane's normal, in order to satisfy the desired metric projected image dimensions; the height $h$ [m] above the ground plane at which the virtual projector is to be placed if the image---with original image width of $w_\text{im}$ [px]---is to be projected to the scene plane with a metric width of $w$ [m] is given by
 \begin{equation}
-h_\text{proj} = w \times f_\text{proj} / w_\text{im},
+h = w \times f_\text{proj} / w_\text{im},
 \end{equation}
@@
     \includegraphics[height=3.75cm]{images/hans_undist_.pdf}
-    \caption{Projection of warped image from Figure \ref{fig:warp}(a) to corresponding target location, acquired by the `downwards-facing' camera.
+    \caption{Projection of the warped image from Figure \ref{fig:warp}(a) to corresponding target location, acquired by the `downwards-facing' camera.
 Note that besides correcting for projective distortions, the warp places the axes of the augmentation in accordance with the axes of the `downwards-facing' camera, and that the dimensions of the augmentation are in line with the desired target dimensions (50~cm~$\times$~31.25~cm).
@@
 The remainder of the time was spent running our modules or working with the MDC-X software.
-Note that once the camera is calibrated, that calibration can be reused if the camera's intrinsics remain fixed, in particular if no change is made to the zoom factor or focus settings of the camera.
+Note that once the camera is calibrated, that calibration can be reused if the camera's optics remain fixed, notably if no change is made to the zoom factor or focus settings of the camera.
 

### 2021-11-17 paper/elsarticle/elsarticle-template.tex
paper

@@
     \centering
-    \includegraphics[height=3.75cm]{images/hans_undist_.pdf}
-    \caption{Projection of the warped image from Figure \ref{fig:warp}(a) to corresponding target location, acquired by the `downwards-facing' camera.
+    \includegraphics[height=3.85cm]{images/hans_undist_.pdf}
+    \caption{Projection of the warped image from Figure \ref{fig:warp}(a) to corresponding target location at the upper-right corner of the floorspace, acquired by the `downwards-facing' camera.
 Note that besides correcting for projective distortions, the warp places the axes of the augmentation in accordance with the axes of the `downwards-facing' camera, and that the dimensions of the augmentation are in line with the desired target dimensions (50~cm~$\times$~31.25~cm).
@@
 (a) Our hardware setup, comprised of a projector, a steerable mirror system, and a stereo camera, of which we used only the left view.
-(b) Example augmentation produced by projecting---to one of the 15 target locations across the floorspace considered in our evaluation---an image warped using our approach.
-The angle gauge reads $89.6^\circ{}$ for the top-left corner, and the steel rule 49.9 cm for the bottom edge.
-Corners of the full projected image extent projected to the floor in light green to provide context.}
+(b) Example augmentation produced by projecting---to one of the 15 target locations across the floorspace considered in our evaluation, to the upper-left corner of our floorspace---an image warped using our approach.
+The digital angle gauge reads $89.6^\circ{}$ for the top-left corner, and the steel rule 49.9 cm for the bottom edge.
+Corners of the full projected image extent projected to the floor in light green to provide context.
+Note that blur has a minimal impact on the quality of the augmentation, and hence on the accuracy of the measurements.}
     \label{fig:eval}
@@
 
-A natural extension of this work would be to consider the impact of varying calibration patterns, their dimensions, and their scale relative to the camera.
+A natural extension of this work would be to consider the impact of varying calibration patterns, their dimensions, and their scale relative to the camera, or to consider the impact of the `downwards-facing' camera's pixel size.
 Still another would be to address the augmentation of non-planar scenes.

### 2021-12-03 paper/elsarticle/elsarticle-template.tex
click to point doc intro

@@
     \qquad
-    \subfloat[\centering Corresponding 3D positions for use in projector calibration (red, in scene plane).]{{\includegraphics[height=3.1cm]{images/2d3d.png}}}
+    \subfloat[\centering Corresponding 3D positions for use in projector calibration (red, in scene plane).]{{\includegraphics[height=3.26cm]{images/2d3d.png}}}
     \caption{Obtaining the 3D positions of the 2D-3D correspondences for use in projector calibration.
@@
 Accordingly, we rotate the recovered `downwards-facing' camera (black) about its center of projection such that its optical axis be made parallel with the scene plane's normal, with respect to the minimum arc-length rotation.
-It is then, for each respective target position, the $X$- and $Y$-axes of the resulting \textit{virtual} camera (gray) that we subsequently use to orient our virtual projector.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
+It is then, for each respective target location, the $X$- and $Y$-axes of the resulting \textit{virtual} camera (gray) that we subsequently use to orient the virtual projector (cf.\ Figure~\ref{fig:virtualproj}) in terms of which the augmentation is to appear to be projected, which is itself placed to face downwards towards the scene plane precisely.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
 This rotated camera is our virtual camera, and it is the orientation of the virtual camera that we subsequently use to orient our virtual projector.
@@
     \centering
-    \includegraphics[height=3.85cm]{images/hans_undist_.pdf}
+    \includegraphics[height=4cm]{images/hans_undist_.pdf}
     \caption{Projection of the warped image from Figure \ref{fig:warp}(a) to corresponding target location at the upper-right corner of the floorspace, acquired by the `downwards-facing' camera.

### 2022-01-07 paper/elsarticle/elsarticle-template.tex
paper

@@
     \centering
-    \subfloat[\centering Projector calibration image (2D pattern position detections overlain).]{{\includegraphics[height=3.1cm]{images/detections.png} }}
+    \subfloat[\centering Projector calibration image (2D pattern position detections overlain).]{{\includegraphics[width=5.15cm]{images/detections_.png}}}
     \qquad
-    \subfloat[\centering Corresponding 3D positions for use in projector calibration (red, in scene plane).]{{\includegraphics[height=3.26cm]{images/2d3d.png}}}
+    \subfloat[\centering Corresponding 3D positions for use in projector calibration (red, in scene plane).]{{\includegraphics[height=3.5cm]{images/2d3d.png}}}
     \caption{Obtaining the 3D positions of the 2D-3D correspondences for use in projector calibration.
@@
 
-\paragraph{Virtual projector} The placement of the virtual projector determines from which pose the image to be projected is to \textit{appear} to have been projected.
+\paragraph{Virtual projector} The placement of the virtual projector---for the recovered projector given a target location---determines from which pose the image to be projected is to \textit{appear} to have been projected.
 We carry out this placement according to a small handful of steps.
-It is in accordance with the $X$- and $Y$-axes of the `downwards-facing' camera that we wish to place the $X$- and $Y$-axes of the augmentation; recognizing that a physical camera placed to face downwards is almost certain to not face downwards precisely, we first rotate the camera about its center of projection to align its optical axis\footnote{Strictly speaking, the optical axis is the back-projection of the principal point; in our usage, we understand it to refer to the ray through center of the image plane.} with the normal vector of the scene plane by applying the minimum arc-length rotation that relates the two, giving a \textit{virtual} camera facing directly downwards to the scene plane (cf.\ Figure~\ref{fig:virtualcam}).
-To obtain the virtual projector, we then (i) intersect the scene plane with the optical axis and rotate the projector's placement about that point of intersection, aligning the optical axis with the scene plane's normal vector and (ii) align the $X$- and $Y$-axes with those of the virtual camera.
+It is in accordance with the $X$- and $Y$-axes of the `downwards-facing' camera that we wish to place the $X$- and $Y$-axes of the augmentation, since we wish to intuitively be able to use the rotation of the camera relative to the scene plane to set those axes.
+It is, however, only if the virtual projector be placed to face the scene plane directly that the projection will appear free of distortions; this is noteworthy, since a physical camera placed to face downwards is almost certain to not face downwards to the scene plane precisely.
+Accordingly, we rotate the recovered camera about its center of projection to align its optical axis\footnote{Strictly speaking, the optical axis is the back-projection of the principal point; in our usage, we understand it to refer to the ray through center of the image plane.} with the normal vector of the scene plane by applying the minimum arc-length rotation that relates the two, giving a \textit{virtual} camera facing directly downwards to the scene plane (cf.\ Figure~\ref{fig:virtualcam}).
+To obtain the virtual projector, we then (i) intersect the scene plane with the optical axis of the projector and rotate the projector's placement about that point of intersection, aligning the optical axis with the scene plane's normal vector and (ii) align the $X$- and $Y$-axes with those of the virtual camera.
 Finally, we (iii) translate along the scene plane's normal, in order to satisfy the desired metric projected image dimensions; the height $h$ [m] above the ground plane at which the virtual projector is to be placed if the image---with original image width of $w_\text{im}$ [px]---is to be projected to the scene plane with a metric width of $w$ [m] is given by
@@
     \centering
-    \includegraphics[height=4cm]{images/hans_undist_.pdf}
+    \includegraphics[height=4.5cm]{images/hans_undist_.pdf}
     \caption{Projection of the warped image from Figure \ref{fig:warp}(a) to corresponding target location at the upper-right corner of the floorspace, acquired by the `downwards-facing' camera.
@@
     \centering
-    \subfloat[\centering Hardware setup (projector, steerable mirror system, camera).]{{\includegraphics[height=3.6cm]{images/system3_.pdf}}}
+    \subfloat[\centering Hardware setup (projector, steerable mirror system, camera).]{{\includegraphics[height=4.75cm]{images/system3_.pdf}}}  %3.6
     \quad
-    \subfloat[\centering Measurement for augmentation, using a steel rule and a digital angle gauge.]{{\includegraphics[height=3.6cm]{images/meas2_.pdf}}}
+    \subfloat[\centering Measurement for augmentation, using a steel rule and a digital angle gauge (enlargements on the right).]{{\includegraphics[height=4.75cm]{images/meas2_.pdf}}}
     \caption{Evaluation scenario.
@@
 \begin{table}[ht!]
-\caption{Mean absolute error (MAE), minimum absolute error, and maximum absolute error in units of centimeters, with respect to the four sides (top, bottom, left, right) of the 15~augmentations produced using our approach for 15 target locations spread across the floorspace, respectively.
+\caption{Mean absolute error (MAE) and maximum absolute error in units of centimeters, with respect to the four sides (top, bottom, left, right) of the 15~augmentations produced using our approach for 15 target locations spread across the floorspace, respectively.
 Errors were computed against intended augmentation dimensions of 50~cm~$\times$~31.25~cm (i.e., 50~cm for top and bottom, 31.25~cm---rounded to 31.3~cm for the purposes of the computing the errors---for left and right).} %Note that absolute errors in no instance exceeded 1~cm for any side of any of the 15~augmentations.
@@
 MAE {\tiny[cm]} & 0.23 & 0.17 & 0.22 & 0.2 \\
-Min AE {\tiny[cm]} & 0.0 & 0.0 & 0.0 & 0.0 \\
+%Min AE {\tiny[cm]} & 0.0 & 0.0 & 0.0 & 0.0 \\
 Max AE {\tiny[cm]} & 0.8 & 0.6 & 0.5 & 0.5 \\
@@
 \begin{table}[ht!]
-\caption{Mean absolute error (MAE), minimum absolute error, and maximum absolute error in units of degrees, with respect to the top-left and bottom-right corners of the 15~augmentations produced using our approach for 15 target locations spread across the floorspace, respectively.
+\caption{Mean absolute error (MAE) and maximum absolute error in units of degrees, with respect to the top-left and bottom-right corners of the 15~augmentations produced using our approach for 15 target locations spread across the floorspace, respectively.
 Errors computed in terms of deviation from $90^\circ{}$.}
@@
 MAE {\tiny[deg]} & 0.17 & 0.32 \\
-Min AE {\tiny[deg]} & 0.0 & 0.0 \\
+%Min AE {\tiny[deg]} & 0.0 & 0.0 \\
 Max AE {\tiny[deg]} & 0.9 & 0.5 \\

### 2022-01-07 paper/elsarticle/elsarticle-template.tex
paper

@@
 
+\begin{figure}[t]
+    \centering
+    \subfloat[\centering Original image.]{{\includegraphics[height=3.65cm]{images/projim.png}}}
+    \quad
+    \subfloat[\centering Projection of original image to target location in scene plane (view downwards to scene plane).]{{\includegraphics[height=3.7cm]{images/proj_projim.png}}}
+    \caption{Na\"ive projection to a given target location in the scene plane (e.g., the floor, a wall, or table).
+(a) Original image, which we wish to project to the scene plane.
+(b) Projecting the original image to the scene plane from the viewpoint of the projector (shown as a frustum in the bottom left, in black with up vector red) directed to the target location introduces projective distortions if the projector is not fronto-parallel with the scene plane (as is the case here), with orientation and scale varying as a function of target location.}
+    \label{fig:proj}
+\end{figure}
+
 In turn, a form of augmented reality referred to as \textit{spatial} AR\cite{bimber2019spatial} is carried out not by embedding the augmentation in an image of the scene as with a head-mounted or handheld display, but by projection to the scene itself, thus eliminating the aforementioned problems.
@@
 
-\begin{figure}
+\begin{figure}[t]
     \centering
-    \subfloat[\centering Original image.]{{\includegraphics[height=3.4cm]{images/projim.png}}}
-    \qquad
-    \subfloat[\centering Projection to target location in scene plane (view downwards to scene plane).]{{\includegraphics[height=3.4cm]{images/proj_projim.png}}}
-    \caption{Na\"ive projection to a given target location in the scene plane (e.g., the floor, a wall, or table).
-(a) Original image, which we wish to project to the scene plane.
-(b) Projecting the original image to the scene plane from the viewpoint of the projector directed to the target location (bottom left, black with up vector red) introduces projective distortions if the projector is not fronto-parallel with the scene plane (as is the case here), with orientation and scale varying as a function of target location.}
-    \label{fig:proj}
+    \subfloat[\centering Our warped image.]{{\includegraphics[height=3.65cm]{images/projim_warped_scaled.png}}}
+    \quad
+    \subfloat[\centering Projection of our warped image to target location in scene plane (view downwards to scene plane).]{{\includegraphics[height=3.7cm]{images/proj_projim_waped.png}}}
+    \caption{Projection to a given target location in the scene plane, correcting for projective distortions in terms of a \textit{virtual} projector placed and utilized using our approach.
+(a) Original image (cf.\ Figure~\ref{fig:proj}(a)) warped using our approach.
+(b) Projecting the warped image to the scene plane from the viewpoint of the projector (bottom left, frustum in black with up vector red; same as in Figure~\ref{fig:proj}(b)) directed to the target location has the effect of projecting the \textit{original} image from the viewpoint of a virtual projector (center, frustum in gray with up vector red).
+This renders the projection free of projective distortions, aligned with the axes of the virtual camera, and at the desired metric scale.
+Note that black pixels will in effect be ignored by the projector.}
+    \label{fig:warp}
 \end{figure}
@@
 Especially manufacturing processes of large products that remain at a static mounting position during the assembly process---i.e., industrial site assembly---offer a significant application potential for the use of spatial AR as part of worker assistance systems \cite{mayrhofer2019one}.
-
-\begin{figure}
-    \centering
-    \subfloat[\centering Our warped image.]{{\includegraphics[height=3.4cm]{images/projim_warped_scaled.png}}}
-    \qquad
-    \subfloat[\centering Projection to target location in scene plane (view downwards to scene plane).]{{\includegraphics[height=3.4cm]{images/proj_projim_waped.png}}}
-    \caption{Projection to a given target location in the scene plane, correcting for projective distortions using our approach.
-(a) Original image (cf.\ Figure~\ref{fig:proj}(a)) warped using our approach with respect to the given target location.
-(b) Projecting the warped image to the scene plane from the viewpoint of the projector directed to the target location (bottom left, black with up vector red; same as in Figure~\ref{fig:proj}(b)) has the effect of projecting the \textit{original} image from the viewpoint of the \textit{virtual} projector (center, gray with up vector red).
-This renders the projection free of projective distortions, aligned with the axes of the virtual camera, and at the desired metric scale.
-Note that black pixels will in effect be ignored by the projector.}
-    \label{fig:warp}
-\end{figure}
 
@@
 The keystone correction approach of Sukthankar \textit{et al.} \cite{sukthankar2001smarter} reduces to using a segmentation approach to identify the four corners of projection screen in an image acquired by a camera, and computing a homography that maps the resulting four corners of the projection screen to the four corners of the projector's image plane.
-Additional sensors can be used to inform the computation of a homography; Raskar and Beardsley \cite{raskar2001self} use a tilt sensor to recover the projector's gravity vector, which they use to carry out a rotational correction in projecting to a wall.
+Additional sensors can be used to inform the computation of a homography; Raskar and Beardsley \cite{raskar2001self} use a tilt sensor to recover the projector's gravity (i.e., down) vector, which they use to carry out a rotational correction in projecting to a wall.
 Still another way to compute a homography is as a function of a plane and a pair of pinhole cameras; the effect produced by the homography is one of projecting an image to the plane from the viewpoint of the one camera, and acquiring the projected image from the viewpoint of the other.
@@
 
+\begin{figure}[t]
+    \centering
+    \subfloat[\centering Projector calibration image (2D pattern position detections overlain).]{{\includegraphics[width=5.15cm]{images/detections_.png}}}
+    \quad
+    \subfloat[\centering Recovered 3D positions corresponding to projector calibration image for use in projector calibration (red, in scene plane).]{{\includegraphics[height=3.7cm]{images/2d3d.png}}}
+    \caption{Obtaining the 3D positions of the 2D-3D correspondences for use in projector calibration.
+(a) For each target location, a projector calibration image is acquired from the viewpoint of the `downwards-facing' camera, containing a projected asymmetric circle pattern at the target location and a chessboard pattern placed on the floor nearby.
+Circle pattern center points and chessboard corners detected automatically (overlain for illustration).
+(b) Given a projector calibration image, the scene plane is recovered in the coordinate frame of the camera (black with up vector red) via spatial resection with respect to 2D-3D correspondences obtained using the chessboard pattern (blue); the 3D positions of the 2D-3D correspondences to be used in calibrating the projector (red, in scene plane) are obtained by intersection with the scene plane of the back-projections (rays, likewise gray) of the 2D center points of the circle pattern detected in the image plane of the camera.} %The corresponding 2D positions are detected in the image of the asymmetric circles pattern image projected by the projector (not shown).
+The `downwards-facing' camera is shown relative to the scene plane in terms of its camera frustum (black, with up vector in red).
+    \label{fig:2d}
+\end{figure}
+
 Calibration of a projector (or camera) in the sense we employ the term here\footnote{We are referring to a geometric calibration; not, e.g., to a color calibration.} renders one able to project a scene point~$\mathbf{X} \in \mathbb{R}^3$ to its corresponding pixel~$\mathbf{x} \in \mathbb{R}^2$ in the projector's (or camera's) image plane, or to compute the `back-projection' of $\mathbf{x}$, i.e., the ray from the projector's (or camera's) center of projection through $\mathbf{x}$ along which a projecting point~$\mathbf{X}$ must lie.
@@
 
-\begin{figure}[ht]
-    \centering
-    \subfloat[\centering Projector calibration image (2D pattern position detections overlain).]{{\includegraphics[width=5.15cm]{images/detections_.png}}}
-    \qquad
-    \subfloat[\centering Corresponding 3D positions for use in projector calibration (red, in scene plane).]{{\includegraphics[height=3.5cm]{images/2d3d.png}}}
-    \caption{Obtaining the 3D positions of the 2D-3D correspondences for use in projector calibration.
-(a) For each target location, a projector calibration image is acquired from the viewpoint of the `downwards-facing' camera, containing a projected asymmetric circle pattern at the target location and a chessboard pattern placed on the floor nearby.
-Circle pattern center points and chessboard corners detected automatically (overlain for illustration).
-(b) Given a projector calibration image, the scene plane is recovered in the coordinate frame of the camera (black with up vector red) via spatial resection with respect to 2D-3D correspondences obtained using the chessboard pattern (blue); the 3D positions of the 2D-3D correspondences to be used in calibrating the projector (red, in scene plane) are obtained by intersection with the scene plane of the back-projections (rays, likewise gray) of the 2D center points of the circle pattern detected in the image plane of the camera.} %The corresponding 2D positions are detected in the image of the asymmetric circles pattern image projected by the projector (not shown).
-The `downwards-facing' camera is shown relative to the scene plane in terms of its camera frustum (black, with up vector in red).
-    \label{fig:2d}
-\end{figure}
-
 A camera can be calibrated by (i) establishing 2D-3D correspondences between pixels in the camera's image plane and the corresponding points in the scene, and on (ii) using those correspondences as input to an optimization procedure that relies on bundle adjustment \cite{triggs1999bundle} to output maximum likelihood estimates of the calibration matrix, the associated lens distortion model coefficients, and, for each calibration image, the pose (i.e., position and orientation) of the camera relative to the 3D scene points from among the 2D-3D correspondences \cite{Hartley2004,zhang2000flexible}.
 Calibration images containing a calibration surface such as a chessboard pattern of known dimensions and scale are acquired from varying viewpoints using the camera, to be used to establish the 2D-3D correspondences~$\{\mathbf{x}_{i,j} \leftrightarrow \mathbf{X}_{i,j}\}$, $i \in \{ 1, \dots, n_\text{pt} \}, j \in \{ 1, \dots, n_\text{im} \}$, where $n_\text{pt}$ gives the number of correspondences obtained from one calibration image of the pattern, $n_\text{im}$ the number of such images, and $\mathbf{X}_{i,j} = \mathbf{X}_{i,j'}$, for $j, j' \in \{ 1, \dots, n_\text{im} \}$.
-Bundle adjustment is then used to obtain the maximum likelihood estimates~$(\hat{\omega}, \hat{\mathtt{K}}, \{(\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j)\}), j \in \{ 1, \dots, n_\text{im} \}$ of the lens distortion model coefficients~$\omega$, calibration matrix~$\mathtt{K}$, and rigid body transformations~$(\mathtt{R}_j, \mathbf{t}_j) \in SE(3), j \in \{ 1, \dots, n_\text{im} \}$.
+Bundle adjustment is then used to obtain the maximum likelihood estimates~$(\hat{\omega}, \hat{\mathtt{K}}, \{(\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j)\}), j \in \{ 1, \dots, n_\text{im} \}$ of the lens distortion model coefficients~$\omega$, calibration matrix~$\mathtt{K}$, and rigid body transformations~$(\mathtt{R}_j, \mathbf{t}_j), j \in \{ 1, \dots, n_\text{im} \}$.
 These are obtained using bundle adjustment by minimizing the cumulative reprojection error
@@
 where $\mathtt{R}\mathbf{X} + \mathbf{t} \in \mathbb{R}^3$ expresses $\mathbf{X}$ in the coordinate frame of the camera.
-Inverting the rigid body transformation~$(\mathtt{R}_j, \mathbf{t}_j)$ gives the pose~$(\mathtt{R}_j^{-1}, -\mathtt{R}_j^{-1}\mathbf{t}_j^{}) \in SE(3)$ of the $j^{\text{th}}$~camera relative to the 3D scene points of the calibration pattern, with $-\mathtt{R}_j^{-1}\mathbf{t}_j^{} \in \mathbb{R}^3$ the corresponding center of projection.
+Inverting the rigid body transformation~$(\mathtt{R}_j, \mathbf{t}_j)$ gives the pose~$(\mathtt{R}_j^{-1}, -\mathtt{R}_j^{-1}\mathbf{t}_j^{})$ of the $j^{\text{th}}$~camera relative to the 3D scene points of the calibration pattern, with $-\mathtt{R}_j^{-1}\mathbf{t}_j^{} \in \mathbb{R}^3$ the corresponding center of projection.
 
 Calibrating a projector \cite{moreno2012simple,zhang2009projector,zhang2006novel} can be carried out in precisely the same manner as calibrating a camera insofar as step (ii) is concerned; the major difference in projector calibration relative to the calibrating a camera then concerns the manner in which 2D-3D correspondences are identified, i.e., between pixels in the image plane of the \textit{projector} and the corresponding points in the scene.
-What remains of this section is concerned primarily with the recovery of 2D-3D correspondences, first in support of calibrating the `downwards-facing' camera (giving the camera calibration matrix~$\mathtt{K}_\text{cam}$), then---in a manner that makes use of $\mathtt{K}_\text{cam}$---for calibrating the projector (giving the projector calibration matrix~$\mathtt{K}_\text{proj}$ and the projector poses~($\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j) \in SE(3)$, for each projector calibration image).
+What remains of this section is concerned primarily with the recovery of 2D-3D correspondences, first in support of calibrating the `downwards-facing' camera (giving the camera calibration matrix~$\mathtt{K}_\text{cam}$), then---in a manner that makes use of $\mathtt{K}_\text{cam}$---for calibrating the projector (giving the projector calibration matrix~$\mathtt{K}_\text{proj}$ and the projector poses~($\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j)$, for each projector calibration image).
 
@@
 
-Recall that the pose~($\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j) \in SE(3)$ of the projector, for each projector calibration image~$j$, is provided alongside $\mathtt{K}_\text{proj}$, relative to the 3D points of the pattern and thus in the coordinate frame of the `downwards-facing' camera.
+Recall that the pose~($\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j)$ of the projector, for each projector calibration image~$j$, is provided alongside $\mathtt{K}_\text{proj}$, relative to the 3D points of the pattern and thus in the coordinate frame of the `downwards-facing' camera.
 Note that for a fixed projector with steerable mirror, given a projector calibration image, the recovered projector's pose is the pose the projector would have to have had to project to the given target location \textit{in the absence of the mirror} (i.e., facing directly towards the target location).
-This is sufficient for our needs in Section~\ref{sec:approach:homography}, and it is in this sense that our system is able to handle a projector equipped with a steerable mirror, without need for modeling the steerable mirror explicitly.
+It is in this sense that our system is able to handle a projector equipped with a steerable mirror, without need for modeling the steerable mirror explicitly.
 
@@
     \centering
-    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=4.2cm]{images/3_cam.pdf}}}
+    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=5.25cm]{images/3_cam.pdf}}}
     \qquad
-    \subfloat[\centering View downwards to scene plane.]{{\includegraphics[height=4.2cm]{images/3_above3_cam.png}}}
+    \subfloat[\centering View downwards to scene plane.]{{\includegraphics[height=5.25cm]{images/3_above3_cam.png}}}
     \caption{Obtaining the virtual camera.
-Our physical `downwards-facing' camera---in terms of whose rotation relative to the scene plane we wish to orient our augmentations---is almost certain to not face downwards precisely.
-Accordingly, we rotate the recovered `downwards-facing' camera (black) about its center of projection such that its optical axis be made parallel with the scene plane's normal, with respect to the minimum arc-length rotation.
-It is then, for each respective target location, the $X$- and $Y$-axes of the resulting \textit{virtual} camera (gray) that we subsequently use to orient the virtual projector (cf.\ Figure~\ref{fig:virtualproj}) in terms of which the augmentation is to appear to be projected, which is itself placed to face downwards towards the scene plane precisely.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
+Our physical `downwards-facing' camera---placed to face downwards towards the scene plane, and in terms of whose rotation relative to the normal vector of the scene plane we wish to orient our augmentations---is almost certain to not face downwards towards the scene plane precisely.
+Accordingly, we rotate the recovered `downwards-facing' camera (black) about its center of projection such that its optical axis be made parallel with the scene plane's normal, with respect to the minimum arc-length rotation relating the two.} %It is then, for each respective target location, the $X$- and $Y$-axes of the resulting \textit{virtual} camera (gray) that we subsequently use to orient the virtual projector (cf.\ Figure~\ref{fig:virtualproj}) in terms of which the augmentation is to appear to be projected, which is itself placed to face downwards towards the scene plane precisely.
+The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
 This rotated camera is our virtual camera, and it is the orientation of the virtual camera that we subsequently use to orient our virtual projector.
@@
     \centering
-    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=4.2cm]{images/3_proj.pdf}}}
+    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=5.25cm]{images/3_proj.pdf}}}
     \qquad
-    \subfloat[\centering View downwards to scene plane.]{{\includegraphics[height=4.2cm]{images/3_above3_proj.png}}}
+    \subfloat[\centering View downwards to scene plane.]{{\includegraphics[height=5.25cm]{images/3_above3_proj.png}}}
     \caption{Obtaining the virtual projector.
@@
 
-\paragraph{Plane-induced homography} Let $\mathtt{K}_\text{proj}$ express the calibration matrix of the recovered projector and $(\mathtt{R}, \mathbf{t}) \in SE(3)$ the rigid body transformation that transforms points from the coordinate frame of the recovered projector to that of the virtual projector, for a given target location.
+\paragraph{Plane-induced homography} Let $\mathtt{K}_\text{proj}$ express the calibration matrix of the recovered projector and $(\mathtt{R}, \mathbf{t})$ the rigid body transformation that transforms points from the coordinate frame of the recovered projector to that of the virtual projector, for a given target location.
 Moreover, let $(\mathbf{n}^\top, -d)^\top \in \mathbb{P}^3$ give the scene plane, expressed in the coordinate frame of the recovered projector, where $\mathbf{n} \in \mathbb{R}^3$ is the scene plane's normal vector and $d = \mathbf{n}^\top\mathbf{X}$ for any point~$\mathbf{X} \in \mathbb{R}^3$ in the plane, so that $(\mathbf{n}^\top, -d) (\mathbf{X}^\top, 1)^\top = 0$.
@@
     \centering
-    \subfloat[\centering Hardware setup (projector, steerable mirror system, camera).]{{\includegraphics[height=4.75cm]{images/system3_.pdf}}}  %3.6
+    \subfloat[\centering Hardware setup consisting of projector, steerable mirror system, `downwards-facing' camera (i.e., camera placed to face downards towards the scene plane).]{{\includegraphics[height=5cm]{images/system3_.pdf}}}  %3.6
     \quad
-    \subfloat[\centering Measurement for augmentation, using a steel rule and a digital angle gauge (enlargements on the right).]{{\includegraphics[height=4.75cm]{images/meas2_.pdf}}}
+    \subfloat[\centering Measurement for augmentation, using a steel rule and a digital angle gauge (enlargements on the right).]{{\includegraphics[height=5cm]{images/meas2_.pdf}}}
     \caption{Evaluation scenario.

### 2022-01-07 paper/elsarticle/elsarticle-template.tex
paper

@@
     \centering
-    \subfloat[\centering Original image.]{{\includegraphics[height=3.65cm]{images/projim.png}}}
-    \quad
-    \subfloat[\centering Projection of original image to target location in scene plane (view downwards to scene plane).]{{\includegraphics[height=3.7cm]{images/proj_projim.png}}}
-    \caption{Na\"ive projection to a given target location in the scene plane (e.g., the floor, a wall, or table).
+    \subfloat[\centering Original image.]{\raisebox{10mm}{\includegraphics[width=5cm]{images/projim.png}}}
+    \qquad
+    \subfloat[\centering Projection of original image to target location in scene plane (view directly downwards to scene plane).]{{\includegraphics[width=6cm]{images/proj_projim_full_.png}}}
+    \caption{Na\"ive projection to a given target location in the scene plane (e.g., the floor, a wall, or table) if not facing the scene plane directly.
 (a) Original image, which we wish to project to the scene plane.
-(b) Projecting the original image to the scene plane from the viewpoint of the projector (shown as a frustum in the bottom left, in black with up vector red) directed to the target location introduces projective distortions if the projector is not fronto-parallel with the scene plane (as is the case here), with orientation and scale varying as a function of target location.}
+(b) Projecting the original image to the scene plane from the viewpoint of the projector (shown as a frustum in the bottom left, in black with up vector red) directed to the target location introduces projective distortions if the projector does not face the scene plane directly (as is the case here), with orientation and scale of the augmentation varying in the scene plane as a function of target location.}
     \label{fig:proj}
@@
     \centering
-    \subfloat[\centering Our warped image.]{{\includegraphics[height=3.65cm]{images/projim_warped_scaled.png}}}
-    \quad
-    \subfloat[\centering Projection of our warped image to target location in scene plane (view downwards to scene plane).]{{\includegraphics[height=3.7cm]{images/proj_projim_waped.png}}}
+    \subfloat[\centering Our warped image.]{\raisebox{10mm}{\includegraphics[width=5cm]{images/projim_warped_scaled.png}}}
+    \qquad
+    \subfloat[\centering Projection of our warped image to target location in scene plane (view directly downwards to scene plane).]{{\includegraphics[width=6cm]{images/proj_projim_waped_full_.png}}}
     \caption{Projection to a given target location in the scene plane, correcting for projective distortions in terms of a \textit{virtual} projector placed and utilized using our approach.
@@
 Our contribution is to propose a wide-area spatial AR system for planar scenes that produces the effect of keystone correction analytically, free of any manual interaction.
-Notably, we do this in a manner placing the axes of the augmentation in accordance with the axes of a camera placed to face towards the scene plane, a camera we shall in what follows refer to as `downwards facing', for brevity.
-To compensate for the likely possibility that the camera not face downwards to the scene plane precisely, we rotate the recovered `downwards-facing' camera to face downwards directly, yielding what we refer to as the \textit{virtual} camera.
-Relying on the placement of the `downwards-facing' camera thus facilitates placement of augmentations, since setting the desired principal axes of augmentations with respect to the scene plane thus reduces to appropriately rotating the `downwards-facing' camera relative to the scene plane.
+Notably, we do this in a manner placing the $X$- and $Y$-axes of the augmentation in accordance with the $X$- and $Y$-axes of a camera placed to face downwards towards the scene plane (cf.\ Figure~\ref{fig:eval}(a)).
+In what follows,  \textit{we shall refer to this camera as `downwards facing'} for brevity, in quotes since the pose of the camera is likely to deviate at least gently from facing downwards to the scene plane directly (an effect for which we shall carry out a correction).
+Relying on the placement of the `downwards-facing' camera thus facilitates placement of augmentations, since setting the $X$- and $Y$-axes of augmentations with respect to the scene plane thus reduces to appropriately rotating the `downwards-facing' camera relative to the scene plane.
 We compute our corrective image warp for a given target location by warping the image to be projected using a plane-induced homography computed to produce the effect of projecting the image not from the actual projector viewpoint, but in accordance with the viewpoint of a \textit{virtual} projector (cf.\ Figure~\ref{fig:warp}).
@@
 Finally, our system is set up for use with a projector equipped with a steerable mirror in a manner that does not call for explicitly modeling the action of the steerable mirror on the projector, thereby enabling wide-area applications exceeding the immediate field of view of the projector without needing to rely on multiple projectors.
+
+%This makes setting those axes particularly easy from the user viewpoint, since it reduces to appropriately rotating the `downwards-facing' camera relative to the scene plane.
+Because we wish to set the principal axes of our augmentations in terms of a rotation about the normal vector of the scene plane,  To compensate for the likely possibility that the camera not face downwards to the scene plane precisely, we rotate the recovered `downwards-facing' camera to face downwards directly, yielding what we refer to as the \textit{virtual} camera.
 
@@
 
-\begin{figure}
+\begin{figure}[t]
     \centering
@@
     \qquad
-    \subfloat[\centering View downwards to scene plane.]{{\includegraphics[height=5.25cm]{images/3_above3_cam.png}}}
+    \subfloat[\centering View directly downwards to scene~plane.]{{\includegraphics[height=5.25cm]{images/3_above3_cam.png}}}
     \caption{Obtaining the virtual camera.
 Our physical `downwards-facing' camera---placed to face downwards towards the scene plane, and in terms of whose rotation relative to the normal vector of the scene plane we wish to orient our augmentations---is almost certain to not face downwards towards the scene plane precisely.
-Accordingly, we rotate the recovered `downwards-facing' camera (black) about its center of projection such that its optical axis be made parallel with the scene plane's normal, with respect to the minimum arc-length rotation relating the two.} %It is then, for each respective target location, the $X$- and $Y$-axes of the resulting \textit{virtual} camera (gray) that we subsequently use to orient the virtual projector (cf.\ Figure~\ref{fig:virtualproj}) in terms of which the augmentation is to appear to be projected, which is itself placed to face downwards towards the scene plane precisely.
+Accordingly, we rotate the recovered `downwards-facing' camera (frustum in black) about its center of projection such that its optical axis be made parallel with the scene plane's normal, with respect to the minimum arc-length rotation relating the two.} %It is then, for each respective target location, the $X$- and $Y$-axes of the resulting \textit{virtual} camera (gray) that we subsequently use to orient the virtual projector (cf.\ Figure~\ref{fig:virtualproj}) in terms of which the augmentation is to appear to be projected, which is itself placed to face downwards towards the scene plane precisely.
 The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
@@
     \qquad
-    \subfloat[\centering View downwards to scene plane.]{{\includegraphics[height=5.25cm]{images/3_above3_proj.png}}}
+    \subfloat[\centering View directly downwards to scene~plane.]{{\includegraphics[height=5.25cm]{images/3_above3_proj.png}}}
     \caption{Obtaining the virtual projector.
-For a given target location, the virtual projector is obtained by (i) rotating the recovered projector (bottom left, black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (ii) rotating the $X$- and $Y$-axes to align them with those of the virtual camera guaranteed---in contrast to the `downwards-facing' camera---to face \textit{directly} downwards to the scene plane (cf.\ Figure \ref{fig:virtualcam}), and (iii) translating along the normal direction to achieve the desired metric projected image dimensions.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
+For a given target location, the virtual projector is obtained by (i) rotating the recovered projector (bottom left, frustum in black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (ii) rotating the $X$- and $Y$-axes to align them with those of the virtual camera guaranteed---in contrast to the `downwards-facing' camera---to face \textit{directly} downwards to the scene plane (cf.\ Figure \ref{fig:virtualcam}), and (iii) translating along the normal direction to achieve the desired metric projected image dimensions.
+Note in (b) that the virtual projector's principal axes are aigned with those of the virtual camera in Figure~\ref{fig:virtualcam}(b).} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
     \label{fig:virtualproj}
@@
 
-\paragraph{Virtual projector} The placement of the virtual projector---for the recovered projector given a target location---determines from which pose the image to be projected is to \textit{appear} to have been projected.
+\paragraph{Virtual projector} The placement of the virtual projector---for a given a target location---determines from which pose the image to be projected is to \textit{appear} to have been projected, with the effect correcting for projective distortions, placing the axes of the augmentation in accordance with the axes of the `downwards-facing' camera, and setting the dimensions of the augmentation in line with the desired target dimensions.
 We carry out this placement according to a small handful of steps.
@@
 Accordingly, we rotate the recovered camera about its center of projection to align its optical axis\footnote{Strictly speaking, the optical axis is the back-projection of the principal point; in our usage, we understand it to refer to the ray through center of the image plane.} with the normal vector of the scene plane by applying the minimum arc-length rotation that relates the two, giving a \textit{virtual} camera facing directly downwards to the scene plane (cf.\ Figure~\ref{fig:virtualcam}).
-To obtain the virtual projector, we then (i) intersect the scene plane with the optical axis of the projector and rotate the projector's placement about that point of intersection, aligning the optical axis with the scene plane's normal vector and (ii) align the $X$- and $Y$-axes with those of the virtual camera.
+To obtain the virtual projector, we then (i) intersect the scene plane with the optical axis of the projector and rotate the projector's placement about that point of intersection, aligning the optical axis with the scene plane's normal vector, and (ii) align the $X$- and $Y$-axes with those of the virtual camera.
 Finally, we (iii) translate along the scene plane's normal, in order to satisfy the desired metric projected image dimensions; the height $h$ [m] above the ground plane at which the virtual projector is to be placed if the image---with original image width of $w_\text{im}$ [px]---is to be projected to the scene plane with a metric width of $w$ [m] is given by
@@
     \centering
-    \subfloat[\centering Hardware setup consisting of projector, steerable mirror system, `downwards-facing' camera (i.e., camera placed to face downards towards the scene plane).]{{\includegraphics[height=5cm]{images/system3_.pdf}}}  %3.6
+    \subfloat[\centering Hardware setup consisting of projector, steerable mirror system, `downwards-facing' camera (i.e., camera placed to face downards towards the scene plane, here the left view of a stereo camera).]{{\includegraphics[height=5cm]{images/system3_.pdf}}}  %3.6
     \quad

### 2022-01-07 paper/elsarticle/elsarticle-template.tex
paper

@@
     \centering
-    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=5.25cm]{images/3_cam.pdf}}}
+    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=5.25cm]{images/3_cam_.pdf}}}
     \qquad
-    \subfloat[\centering View directly downwards to scene~plane.]{{\includegraphics[height=5.25cm]{images/3_above3_cam.png}}}
+    \subfloat[\centering View directly downwards to scene~plane.]{{\includegraphics[height=5.25cm]{images/3_above3_cam.pdf}}}
     \caption{Obtaining the virtual camera.
@@
     \centering
-    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=5.25cm]{images/3_proj.pdf}}}
+    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=5.25cm]{images/3_proj_.pdf}}}
     \qquad
-    \subfloat[\centering View directly downwards to scene~plane.]{{\includegraphics[height=5.25cm]{images/3_above3_proj.png}}}
+    \subfloat[\centering View directly downwards to scene~plane.]{{\includegraphics[height=5.25cm]{images/3_above3_proj.pdf}}}
     \caption{Obtaining the virtual projector.

### 2022-01-07 paper/elsarticle/elsarticle-template.tex
paper

--- 2022-01-07 paper/elsarticle/elsarticle-template.tex
@@
     \centering
-    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=5.25cm]{images/3_cam.pdf}}}
+    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=5.25cm]{images/3_cam_.pdf}}}
     \qquad
@@
     \centering
-    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=5.25cm]{images/3_proj.pdf}}}
+    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=5.25cm]{images/3_proj_.pdf}}}
     \qquad

### 2022-01-07 paper/elsarticle/elsarticle-template.tex
paper

@@
 Our system renders the placement of augmentations in the scene more intuitive than manual keystone correction in two ways.
-First, (i) the horizontal and vertical axes of the desired augmentations is carried out in accordance with the horizontal and vertical image axes of the camera, thereby making setting those axes as simple as appropriately rotating the camera relative to the scene plane, and thereby providing a consistent axial reference frame across all target locations.
+First, (i) the horizontal and vertical axes of the desired augmentations are set in accordance with the horizontal and vertical image axes of the camera, thereby making setting those axes as simple as appropriately rotating the camera relative to the scene plane, and thereby providing a consistent axial reference frame across all target locations.
 Second, (ii) the desired dimensions of the projected augmentations are specified in metric terms, thereby providing for consistent scaling, likewise across all target locations.
@@
 
-While interest in spatial AR in industrial settings has been growing, its deployment in industrial settings remains well behind that of wearable AR systems such as handhelds or head-mounted displays.
+While interest in spatial AR in industrial settings has been growing, its deployment in industrial settings lags well behind that of wearable AR systems such as handhelds or head-mounted displays.
 One of the main shortcomings of typical projector-based spatial AR relative to AR relying on wearables is the ability to project augmentations only within the projector's own static field of view.
-Here, dynamic projection systems---using steerable mirror heads to direct the projection across wider areas---has the potential to bridge the gap to wider industrial penetration.
+Here, dynamic projection systems---using steerable mirror heads to direct the projection across wider areas---have the potential to bridge the gap to wider industrial penetration.
 Yet the increase in the area to be covered comes with noteworthy challenges in terms of preparing the augmentations, in particular as the number of target locations in the scene becomes larger than a small handful.
@@
 Notably, we do this in a manner placing the $X$- and $Y$-axes of the augmentation in accordance with the $X$- and $Y$-axes of a camera placed to face downwards towards the scene plane (cf.\ Figure~\ref{fig:eval}(a)).
-In what follows,  \textit{we shall refer to this camera as `downwards facing'} for brevity, in quotes since the pose of the camera is likely to deviate at least gently from facing downwards to the scene plane directly (an effect for which we shall carry out a correction).
+In what follows,  \textit{we shall refer to this camera as `downwards facing'} for brevity, in quotes since the pose of the camera is likely to deviate at least gently from facing downwards towards the scene plane directly (an effect for which we shall carry out a correction).
 Relying on the placement of the `downwards-facing' camera thus facilitates placement of augmentations, since setting the $X$- and $Y$-axes of augmentations with respect to the scene plane thus reduces to appropriately rotating the `downwards-facing' camera relative to the scene plane.
@@
 More generally, keystone correction for planar scenes can be said to reduce to computing a 2D homography \cite{Hartley2004}, an invertible transformation that preserves colinearity and is thus fittingly alternatively referred to as a `colineation'; this is the case whether computing the homography relies on manual interaction as in the case of Pinharez, or is obtained free of manual interaction as in ours.
-The intuition for why it is that a transformation that preserves colinearity---i.e., maps lines to lines---can serve to model an appropriate corrective image warp can be drawn from considering a planar chessboard pattern: looking at a chessboard from an oblique angle, one observes that lines parallel in the chessboard appear to meet in respective vanishing points, yet remain lines; looking at the same chessboard frontally with respect to the plane of the chessboard, lines parallel in the chessboard appear parallel in the image (i.e., they are said to meet `at infinity').
-To warp an image of the chessboard acquired at one angle such that the lines of the chessboard appear as if the chessboard were acquired at another angle accordingly calls for a transformation that maps lines to lines; it therefore follows that there exists a homography that models that transformation.
+The intuition for why it is that a transformation that preserves colinearity---i.e., maps lines to lines---can serve to model an appropriate corrective image warp can be drawn from considering a planar chessboard pattern: looking at a chessboard from one angle, one observes that lines parallel in the chessboard may appear to meet in respective vanishing points, yet always remain lines; looking at the same chessboard from another angle, one observes those same lines remain lines.
+Accordingly, to warp an image of the chessboard acquired at one angle such that the lines of the chessboard appear as if the chessboard were acquired at another angle accordingly calls for a transformation that maps lines to lines.
+Since this holds true for any set of lines in a plane, it follows that there exists a line-to-line mapping (i.e., a homography) that models the perspective transformation relating a pair of views of a plane.
 
-One way to compute a homography is by identifying at least four correspondences between pixel positions in two images of a planar surface \cite{Hartley2004}.
+One way to compute such a homography is by identifying at least four correspondences between pixel positions in two images of a planar surface \cite{Hartley2004}.
 The keystone correction approach of Sukthankar \textit{et al.} \cite{sukthankar2001smarter} reduces to using a segmentation approach to identify the four corners of projection screen in an image acquired by a camera, and computing a homography that maps the resulting four corners of the projection screen to the four corners of the projector's image plane.
-Additional sensors can be used to inform the computation of a homography; Raskar and Beardsley \cite{raskar2001self} use a tilt sensor to recover the projector's gravity (i.e., down) vector, which they use to carry out a rotational correction in projecting to a wall.
+Additional sensors can be used to inform the computation of a homography; Raskar and Beardsley \cite{raskar2001self} use a tilt sensor to recover the projector's gravity vector relative to the projector, which they use to carry out a rotational correction in projecting to a wall.
 Still another way to compute a homography is as a function of a plane and a pair of pinhole cameras; the effect produced by the homography is one of projecting an image to the plane from the viewpoint of the one camera, and acquiring the projected image from the viewpoint of the other.
@@
     \centering
-    \subfloat[\centering Hardware setup consisting of projector, steerable mirror system, `downwards-facing' camera (i.e., camera placed to face downards towards the scene plane, here the left view of a stereo camera).]{{\includegraphics[height=5cm]{images/system3_.pdf}}}  %3.6
+    \subfloat[\centering Hardware setup consisting of a Panasonic PT-RZ660BE projector, a Dynamic Projection Institute steerable mirror system, and the left view of a Stereolabs Zed 2 stereo camera to serve as the `downwards-facing' camera.]{{\includegraphics[height=5cm]{images/system3_.pdf}}}  %3.6
     \quad

### 2022-01-11 paper/elsarticle/elsarticle-template.tex
paper

@@
 
-\begin{figure}[t]
-    \centering
-    \subfloat[\centering Original image.]{\raisebox{10mm}{\includegraphics[width=5cm]{images/projim.png}}}
-    \qquad
-    \subfloat[\centering Projection of original image to target location in scene plane (view directly downwards to scene plane).]{{\includegraphics[width=6cm]{images/proj_projim_full_.png}}}
-    \caption{Na\"ive projection to a given target location in the scene plane (e.g., the floor, a wall, or table) if not facing the scene plane directly.
-(a) Original image, which we wish to project to the scene plane.
-(b) Projecting the original image to the scene plane from the viewpoint of the projector (shown as a frustum in the bottom left, in black with up vector red) directed to the target location introduces projective distortions if the projector does not face the scene plane directly (as is the case here), with orientation and scale of the augmentation varying in the scene plane as a function of target location.}
-    \label{fig:proj}
-\end{figure}
-
 In turn, a form of augmented reality referred to as \textit{spatial} AR\cite{bimber2019spatial} is carried out not by embedding the augmentation in an image of the scene as with a head-mounted or handheld display, but by projection to the scene itself, thus eliminating the aforementioned problems.
@@
     \centering
-    \subfloat[\centering Our warped image.]{\raisebox{10mm}{\includegraphics[width=5cm]{images/projim_warped_scaled.png}}}
+    \subfloat[\centering Original image (to be projected).]{\raisebox{10mm}{\includegraphics[width=5cm]{images/projim.png}}}
     \qquad
-    \subfloat[\centering Projection of our warped image to target location in scene plane (view directly downwards to scene plane).]{{\includegraphics[width=6cm]{images/proj_projim_waped_full_.png}}}
-    \caption{Projection to a given target location in the scene plane, correcting for projective distortions in terms of a \textit{virtual} projector placed and utilized using our approach.
-(a) Original image (cf.\ Figure~\ref{fig:proj}(a)) warped using our approach.
-(b) Projecting the warped image to the scene plane from the viewpoint of the projector (bottom left, frustum in black with up vector red; same as in Figure~\ref{fig:proj}(b)) directed to the target location has the effect of projecting the \textit{original} image from the viewpoint of a virtual projector (center, frustum in gray with up vector red).
-This renders the projection free of projective distortions, aligned with the axes of the virtual camera, and at the desired metric scale.
-Note that black pixels will in effect be ignored by the projector.}
-    \label{fig:warp}
+    \subfloat[\centering Projection from side to scene plane (viewed directly downwards to scene plane).]{{\includegraphics[width=6cm]{images/proj_projim_full_.png}}}
+    \caption{Na\"ive projection to a given target location in the scene plane (e.g., the floor, a wall, or table) if not facing the scene plane directly.
+(a) Original image, which we wish to project to the scene plane.
+(b) Projecting the original image to the scene plane from the viewpoint of the projector (shown as a frustum in the bottom left, in black with up vector red) directed to the target location introduces projective distortions if the projector does not face the scene plane directly (as is the case here), with orientation and scale of the augmentation varying in the scene plane as a function of target location.}
+    \label{fig:proj}
 \end{figure}
@@
 Especially manufacturing processes of large products that remain at a static mounting position during the assembly process---i.e., industrial site assembly---offer a significant application potential for the use of spatial AR as part of worker assistance systems \cite{mayrhofer2019one}.
+
+\begin{figure}[t]
+    \centering
+    \subfloat[\centering Our warped image (to be projected).]{\raisebox{10mm}{\includegraphics[width=5cm]{images/projim_warped_scaled.png}}}
+    \qquad
+    \subfloat[\centering Projection from side to scene plane (viewed directly downwards to scene plane).]{{\includegraphics[width=6cm]{images/proj_projim_waped_full_.png}}}
+    \caption{Projection to a given target location in the scene plane, correcting for projective distortions in terms of a \textit{virtual} projector placed and utilized according to our approach.
+(a) Original image (cf.\ Figure~\ref{fig:proj}(a)) warped using our approach.
+(b) Projecting the warped image to the scene plane from the viewpoint of the projector (bottom left, frustum in black with up vector red; same as in Figure~\ref{fig:proj}(b)) directed to the target location has the effect of projecting the \textit{original} image from the viewpoint of a virtual projector (center, frustum in gray with up vector red).
+This renders the projection free of projective distortions, aligned with the axes of the virtual camera, and at the desired metric scale.
+Note that black pixels will in effect be ignored by the projector.}
+    \label{fig:warp}
+\end{figure}
 
@@
 Circle pattern center points and chessboard corners detected automatically (overlain for illustration).
-(b) Given a projector calibration image, the scene plane is recovered in the coordinate frame of the camera (black with up vector red) via spatial resection with respect to 2D-3D correspondences obtained using the chessboard pattern (blue); the 3D positions of the 2D-3D correspondences to be used in calibrating the projector (red, in scene plane) are obtained by intersection with the scene plane of the back-projections (rays, likewise gray) of the 2D center points of the circle pattern detected in the image plane of the camera.} %The corresponding 2D positions are detected in the image of the asymmetric circles pattern image projected by the projector (not shown).
+(b) Given a projector calibration image, the scene plane is recovered in the coordinate frame of the `downwards-facing' camera (black with up vector red) via spatial resection with respect to 2D-3D correspondences obtained using the chessboard pattern (blue); the 3D positions of the 2D-3D correspondences to be used in calibrating the projector (red, in scene plane) are obtained by intersection with the scene plane of the back-projections (rays, likewise gray) of the 2D center points of the circle pattern detected in the image plane of the camera.} %The corresponding 2D positions are detected in the image of the asymmetric circles pattern image projected by the projector (not shown).
 The `downwards-facing' camera is shown relative to the scene plane in terms of its camera frustum (black, with up vector in red).
@@
 
-\begin{figure}[t]
-    \centering
-    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=5.25cm]{images/3_proj_.pdf}}}
-    \qquad
-    \subfloat[\centering View directly downwards to scene~plane.]{{\includegraphics[height=5.25cm]{images/3_above3_proj.png}}}
-    \caption{Obtaining the virtual projector.
-For a given target location, the virtual projector is obtained by (i) rotating the recovered projector (bottom left, frustum in black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (ii) rotating the $X$- and $Y$-axes to align them with those of the virtual camera guaranteed---in contrast to the `downwards-facing' camera---to face \textit{directly} downwards to the scene plane (cf.\ Figure \ref{fig:virtualcam}), and (iii) translating along the normal direction to achieve the desired metric projected image dimensions.
-Note in (b) that the virtual projector's principal axes are aigned with those of the virtual camera in Figure~\ref{fig:virtualcam}(b).} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
-    \label{fig:virtualproj}
-\end{figure}
-
-\paragraph{Virtual projector} The placement of the virtual projector---for a given a target location---determines from which pose the image to be projected is to \textit{appear} to have been projected, with the effect correcting for projective distortions, placing the axes of the augmentation in accordance with the axes of the `downwards-facing' camera, and setting the dimensions of the augmentation in line with the desired target dimensions.
+\paragraph{Virtual camera and projector} The placement of the virtual projector---for a given a target location---determines from which pose the image to be projected is to \textit{appear} to have been projected, with the effect of (i) correcting for projective distortions, (ii) placing the axes of the augmentation in accordance with the axes of the `downwards-facing' camera, and (iii) setting the dimensions of the augmentation in line with the desired target dimensions.
 We carry out this placement according to a small handful of steps.
 It is in accordance with the $X$- and $Y$-axes of the `downwards-facing' camera that we wish to place the $X$- and $Y$-axes of the augmentation, since we wish to intuitively be able to use the rotation of the camera relative to the scene plane to set those axes.
-It is, however, only if the virtual projector be placed to face the scene plane directly that the projection will appear free of distortions; this is noteworthy, since a physical camera placed to face downwards is almost certain to not face downwards to the scene plane precisely.
+It is, however, only if the virtual projector be placed to face downwards to the scene plane precisely that the projection will appear free of distortions; this is noteworthy, since a physical camera placed to face downwards is almost certain to not face precisely downwards to the scene plane.
 Accordingly, we rotate the recovered camera about its center of projection to align its optical axis\footnote{Strictly speaking, the optical axis is the back-projection of the principal point; in our usage, we understand it to refer to the ray through center of the image plane.} with the normal vector of the scene plane by applying the minimum arc-length rotation that relates the two, giving a \textit{virtual} camera facing directly downwards to the scene plane (cf.\ Figure~\ref{fig:virtualcam}).
@@
 %\footnote{A physical camera placed to face downwards is almost certain to not face downwards precisely; in contrast, we align the virtual camera's optical axis exactly with the scene plane's normal vector, rendering it genuinely fronto-parallel with respect to the scene plane.}
+
+\begin{figure}[t]
+    \centering
+    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=5.25cm]{images/3_proj_.pdf}}}
+    \qquad
+    \subfloat[\centering View directly downwards to scene~plane.]{{\includegraphics[height=5.25cm]{images/3_above3_proj.png}}}
+    \caption{Obtaining the virtual projector, with $X$- and $Y$-axes aligned with those of the virtual camera (cf.\ Figure~\ref{fig:virtualcam}).
+For a given target location, the virtual projector is obtained by (i) rotating the recovered projector (bottom left, frustum in black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (ii) rotating the $X$- and $Y$-axes to align them with those of the virtual camera facing \textit{directly} downwards to the scene plane (cf.\ Figure \ref{fig:virtualcam}), and (iii) translating along the normal direction to achieve the desired metric projected image dimensions.
+Projection from the viewpoint of the virtual projector thus produces an augmentation absent of projective distortions, aligns the augmentation \texit{in accordance with} the horizontal and vertical image axes of the `downwards-facing' camera, and ensures the augmentation appears in the scene plane with the desired metric dimensions.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
+    \label{fig:virtualproj}
+\end{figure}
 
@@
     \centering
-    \subfloat[\centering Hardware setup consisting of a Panasonic PT-RZ660BE projector, a Dynamic Projection Institute steerable mirror system, and the left view of a Stereolabs Zed 2 stereo camera to serve as the `downwards-facing' camera.]{{\includegraphics[height=5cm]{images/system3_.pdf}}}  %3.6
+    \subfloat[\centering Hardware setup consisting of a Panasonic PT-RZ660BE projector, a steerable mirror system manufactured by Dynamic Projection Institute, and the left view of a Stereolabs Zed 2 stereo camera---placed to face downwards towards the scene plane---to serve as the `downwards-facing' camera.]{{\includegraphics[height=5cm]{images/system3_.pdf}}}  %3.6
     \quad

### 2022-01-11 paper/elsarticle/elsarticle-template.tex
paper

@@
     \centering
-    \subfloat[\centering Projector calibration image (2D pattern position detections overlain).]{{\includegraphics[width=5.15cm]{images/detections_.png}}}
+    \subfloat[\centering Projector calibration image (2D pattern position detections overlain).]{{\includegraphics[width=5.85cm]{images/detections_.png}}}
     \quad
-    \subfloat[\centering Recovered 3D positions corresponding to projector calibration image for use in projector calibration (red, in scene plane).]{{\includegraphics[height=3.7cm]{images/2d3d.png}}}
+    \subfloat[\centering Recovered 3D positions corresponding to projector calibration image for use in projector calibration (red, in scene plane).]{{\includegraphics[height=4.75cm]{images/2d3d.pdf}}}
     \caption{Obtaining the 3D positions of the 2D-3D correspondences for use in projector calibration.

### 2022-01-12 paper/elsarticle/elsarticle-template.tex
paper

@@
     \qquad
-    \subfloat[\centering Projection from side to scene plane (viewed directly downwards to scene plane).]{{\includegraphics[width=6cm]{images/proj_projim_full_.png}}}
+    \subfloat[\centering Projection from side to scene plane (viewed directly downwards to scene plane).]{{\includegraphics[width=6cm]{images/proj_projim_full_.pdf}}}
     \caption{Na\"ive projection to a given target location in the scene plane (e.g., the floor, a wall, or table) if not facing the scene plane directly.
@@
     \qquad
-    \subfloat[\centering Projection from side to scene plane (viewed directly downwards to scene plane).]{{\includegraphics[width=6cm]{images/proj_projim_waped_full_.png}}}
+    \subfloat[\centering Projection from side to scene plane (viewed directly downwards to scene plane).]{{\includegraphics[width=6cm]{images/proj_projim_waped_full_.pdf}}}
     \caption{Projection to a given target location in the scene plane, correcting for projective distortions in terms of a \textit{virtual} projector placed and utilized according to our approach.

### 2022-01-12 paper/elsarticle/elsarticle-template.tex
paper

@@
     \centering
-    \subfloat[\centering Projector calibration image (2D pattern position detections overlain).]{{\includegraphics[width=5.85cm]{images/detections_.png}}}
+    \subfloat[\centering Projector calibration image (2D position detections of chessboard pattern and asymmetric circle pattern overlain).]{{\includegraphics[width=5.85cm]{images/detections_.png}}}
     \quad
-    \subfloat[\centering Recovered 3D positions corresponding to projector calibration image for use in projector calibration (red, in scene plane).]{{\includegraphics[height=4.75cm]{images/2d3d.pdf}}}
+    \subfloat[\centering Recovery of corresponding 3D positions of asymmetric circle pattern for use in projector calibration (red, in scene plane).]{{\includegraphics[height=4.75cm]{images/2d3d.pdf}}}
     \caption{Obtaining the 3D positions of the 2D-3D correspondences for use in projector calibration.
@@
 Circle pattern center points and chessboard corners detected automatically (overlain for illustration).
-(b) Given a projector calibration image, the scene plane is recovered in the coordinate frame of the `downwards-facing' camera (black with up vector red) via spatial resection with respect to 2D-3D correspondences obtained using the chessboard pattern (blue); the 3D positions of the 2D-3D correspondences to be used in calibrating the projector (red, in scene plane) are obtained by intersection with the scene plane of the back-projections (rays, likewise gray) of the 2D center points of the circle pattern detected in the image plane of the camera.} %The corresponding 2D positions are detected in the image of the asymmetric circles pattern image projected by the projector (not shown).
+(b) Given a projector calibration image, the scene plane is recovered in the coordinate frame of the `downwards-facing' camera (frustum in black with up vector red) via spatial resection with respect to 2D-3D correspondences obtained using the chessboard pattern (blue), with 3D positions known \textit{a priori}.
+With the scene plane available, the 3D positions of the projected asymmetric circles pattern to be used in calibrating the projector (red, in scene plane) are obtained by intersection with the scene plane of the back-projections (gray rays) of the detected 2D center points of the circle pattern (red, in image plane of camera).} %The corresponding 2D positions are detected in the image of the asymmetric circles pattern image projected by the projector (not shown).
 The `downwards-facing' camera is shown relative to the scene plane in terms of its camera frustum (black, with up vector in red).
@@
 The classical calibration surface is a chessboard pattern, and is such a pattern that we use as well.
-The 3D corner points of the chessboard are obtained \textit{a priori} in a coordinate system defined in the plane of the chessboard\footnote{E.g., $(0,0,0), (1.5,0,0), (3,0,0), \dots, (9,7.5,0)$ for a chessboard with $7~\times{}~6$ corners ($8~\times{}~7$ squares), with each square of length and width of 1.5 unit, respectively.
+The 3D corner points of the chessboard are known \textit{a priori} in a coordinate system defined in the plane of the chessboard\footnote{E.g., $(0,0,0), (1.5,0,0), (3,0,0), \dots, (9,7.5,0)$ for a chessboard with $7~\times{}~6$ corners ($8~\times{}~7$ squares), with each square of length and width of 1.5 unit, respectively.
 Note that the units of the chessboard's 3D points give the units of the camera calibration, and---since our projector calibration relies on the camera calibration---the units of the projector calibration as well.}, requiring knowledge only of the dimensions of the chessboard pattern and of the length of a side of a chessboard square.
@@
     \centering
-    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=5.25cm]{images/3_cam_.pdf}}}
+    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=5.25cm]{images/3_cam.pdf}}}
     \qquad

### 2022-01-12 paper/elsarticle/elsarticle-template.tex
paper

@@
 Yet the increase in the area to be covered comes with noteworthy challenges in terms of preparing the augmentations, in particular as the number of target locations in the scene becomes larger than a small handful.
+
 Considering a planar surface to be augmented (e.g., a floor, wall, or table), unless the projector faces the surface frontally, the bounds of a projected rectangular image will not appear rectangular, but will instead be subject to projective distortions (cf.\ Figure~\ref{fig:proj}).
-Such distortions can be corrected for by carrying out a cumbersome manual process referred to as keystone correction---involving warping an image by perturbing its corners until the desired effect is produced when projected to the scene---to appropriately warp the image to be projected, often using software bundled with the projector.
+More generally, na\"ively projecting an image to a planar surface causes not only projective distortions as a function of target location, but also variation in scale and orientation.
+Such effects can be corrected for by carrying out a cumbersome manual process referred to as keystone correction---involving warping an image by perturbing its corners until the desired effect is produced when projected to the scene---to appropriately warp the image to be projected, often using software bundled with the projector.
 Automating this manual, time-consuming, and error-prone process significantly reduces the time needed for the industrial engineering of a spatial AR application and therefore has potential to promote adoption by industry.
@@
 (b) Projecting the warped image to the scene plane from the viewpoint of the projector (bottom left, frustum in black with up vector red; same as in Figure~\ref{fig:proj}(b)) directed to the target location has the effect of projecting the \textit{original} image from the viewpoint of a virtual projector (center, frustum in gray with up vector red).
-This renders the projection free of projective distortions, aligned with the axes of the virtual camera, and at the desired metric scale.
+This renders the projection free of projective distortions, oriented in accordance with the axes of the `downwards-facing' camera, and at the desired metric scale.
 Note that black pixels will in effect be ignored by the projector.}
@@
 
-Our contribution is to propose a wide-area spatial AR system for planar scenes that produces the effect of keystone correction analytically, free of any manual interaction.
-Notably, we do this in a manner placing the $X$- and $Y$-axes of the augmentation in accordance with the $X$- and $Y$-axes of a camera placed to face downwards towards the scene plane (cf.\ Figure~\ref{fig:eval}(a)).
-In what follows,  \textit{we shall refer to this camera as `downwards facing'} for brevity, in quotes since the pose of the camera is likely to deviate at least gently from facing downwards towards the scene plane directly (an effect for which we shall carry out a correction).
-Relying on the placement of the `downwards-facing' camera thus facilitates placement of augmentations, since setting the $X$- and $Y$-axes of augmentations with respect to the scene plane thus reduces to appropriately rotating the `downwards-facing' camera relative to the scene plane.
-We compute our corrective image warp for a given target location by warping the image to be projected using a plane-induced homography computed to produce the effect of projecting the image not from the actual projector viewpoint, but in accordance with the viewpoint of a \textit{virtual} projector (cf.\ Figure~\ref{fig:warp}).
-We place this virtual projector to (i) face directly downwards to the target location in the scene plane and (ii) rotate its axes to match those of the virtual camera.
-Moreover, our system enables specifying the dimensions of augmentations in metric terms, which we achieve by placing the virtual projector at the appropriate height above the scene plane.
-Finally, our system is set up for use with a projector equipped with a steerable mirror in a manner that does not call for explicitly modeling the action of the steerable mirror on the projector, thereby enabling wide-area applications exceeding the immediate field of view of the projector without needing to rely on multiple projectors.
+Our contribution is to propose a spatial AR system for planar scenes that produces the effect of manual keystone correction analytically in a manner for use with a projector equipped with a steerable mirror, thereby enabling wide-area applications exceeding the immediate field of view of the projector without needing to rely on multiple projectors.
+Besides correcting for projective distortions, our system enables placement of augmentations at a desired metric scale.
+Notably, the system moreover enables setting the $X$- and $Y$-axes of the augmentation in accordance with the $X$- and $Y$-axes of a camera placed to face downwards towards the scene plane (cf.\ Figure~\ref{fig:eval}(a)), which in our case is a factory floor (but could well be generalized to other planar surfaces).
+We refer to this camera as `downwards facing' for brevity, in quotes since the camera likely to deviate at least gently from facing downwards towards the scene plane directly.\footnote{More precisely, we understand a camera to face downwards towards the scene plane \textit{directly} if the camera's optical axis is parallel with the scene plane's normal vector, and correct for possible deviation.} Relying on the placement of the `downwards-facing' camera to orient augmentations substantially facilitates placement relative to manual keystone correction, since setting the $X$- and $Y$-axes of augmentations thus reduces to appropriately rotating the `downwards-facing' camera relative to the scene plane.
+
+We compute our corrective image warp for a given target location by warping the image to be projected using a plane-induced homography computed to produce the effect of projecting the image not from the actual projector viewpoint, but from the viewpoint of a \textit{virtual} projector (cf.\ Figure~\ref{fig:warp}).
+We place this virtual projector to (i) face directly downwards to the target location in the scene plane and (ii) rotate its axes in accordance with those of the `downwards-facing' camera.
+Finally, we (iii) place the virtual projector at the appropriate height above the scene plane to produce the augmentation with the desired metric dimensions in the scene plane.
 
@@
 Additional sensors can be used to inform the computation of a homography; Raskar and Beardsley \cite{raskar2001self} use a tilt sensor to recover the projector's gravity vector relative to the projector, which they use to carry out a rotational correction in projecting to a wall.
+
 Still another way to compute a homography is as a function of a plane and a pair of pinhole cameras; the effect produced by the homography is one of projecting an image to the plane from the viewpoint of the one camera, and acquiring the projected image from the viewpoint of the other.
 Such a homography is said to be `induced' by a plane \cite{Hartley2004}.
-We exploit the fact that cameras and projectors alike can be modeled using a pinhole camera model \cite{bimber2019spatial}, proceed to recover the relative geometry relating the scene plane and projector, and compute a particular form of plane-induced homography to effect the intended corrective image warp, yet additionally taking into account both desired in-plane orientation and metric scale.
+We exploit the fact that cameras and projectors alike can be modeled using a pinhole camera model \cite{bimber2019spatial}, proceed to recover the relative geometry relating the scene plane and projector, and compute a particular form of plane-induced homography to effect the intended corrective image warp, additionally taking into account both desired in-plane orientation and metric scale.
 
@@
 A camera can be calibrated by (i) establishing 2D-3D correspondences between pixels in the camera's image plane and the corresponding points in the scene, and on (ii) using those correspondences as input to an optimization procedure that relies on bundle adjustment \cite{triggs1999bundle} to output maximum likelihood estimates of the calibration matrix, the associated lens distortion model coefficients, and, for each calibration image, the pose (i.e., position and orientation) of the camera relative to the 3D scene points from among the 2D-3D correspondences \cite{Hartley2004,zhang2000flexible}.
+
 Calibration images containing a calibration surface such as a chessboard pattern of known dimensions and scale are acquired from varying viewpoints using the camera, to be used to establish the 2D-3D correspondences~$\{\mathbf{x}_{i,j} \leftrightarrow \mathbf{X}_{i,j}\}$, $i \in \{ 1, \dots, n_\text{pt} \}, j \in \{ 1, \dots, n_\text{im} \}$, where $n_\text{pt}$ gives the number of correspondences obtained from one calibration image of the pattern, $n_\text{im}$ the number of such images, and $\mathbf{X}_{i,j} = \mathbf{X}_{i,j'}$, for $j, j' \in \{ 1, \dots, n_\text{im} \}$.
@@
     \caption{Obtaining the virtual camera.
-Our physical `downwards-facing' camera---placed to face downwards towards the scene plane, and in terms of whose rotation relative to the normal vector of the scene plane we wish to orient our augmentations---is almost certain to not face downwards towards the scene plane precisely.
-Accordingly, we rotate the recovered `downwards-facing' camera (frustum in black) about its center of projection such that its optical axis be made parallel with the scene plane's normal, with respect to the minimum arc-length rotation relating the two.} %It is then, for each respective target location, the $X$- and $Y$-axes of the resulting \textit{virtual} camera (gray) that we subsequently use to orient the virtual projector (cf.\ Figure~\ref{fig:virtualproj}) in terms of which the augmentation is to appear to be projected, which is itself placed to face downwards towards the scene plane precisely.
+Our physical `downwards-facing' camera---placed to face downwards towards the scene plane, and in terms of whose rotation relative to the normal vector of the scene plane we wish to orient our augmentations---is almost certain to not face downwards towards the scene plane directly.
+Accordingly, we rotate the recovered `downwards-facing' camera (frustum in black) about its center of projection such that its optical axis be made parallel with the scene plane's normal, with respect to the minimum arc-length rotation relating the two.
+It is the $X$- and $Y$-axes of the virtual camera that we then use to set the $X$- and $Y$-axes of the virtual projector.} %It is then, for each respective target location, the $X$- and $Y$-axes of the resulting \textit{virtual} camera (gray) that we subsequently use to orient the virtual projector (cf.\ Figure~\ref{fig:virtualproj}) in terms of which the augmentation is to appear to be projected, which is itself placed to face downwards towards the scene plane precisely.
 The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
@@
 We carry out this placement according to a small handful of steps.
+
 It is in accordance with the $X$- and $Y$-axes of the `downwards-facing' camera that we wish to place the $X$- and $Y$-axes of the augmentation, since we wish to intuitively be able to use the rotation of the camera relative to the scene plane to set those axes.
-It is, however, only if the virtual projector be placed to face downwards to the scene plane precisely that the projection will appear free of distortions; this is noteworthy, since a physical camera placed to face downwards is almost certain to not face precisely downwards to the scene plane.
-Accordingly, we rotate the recovered camera about its center of projection to align its optical axis\footnote{Strictly speaking, the optical axis is the back-projection of the principal point; in our usage, we understand it to refer to the ray through center of the image plane.} with the normal vector of the scene plane by applying the minimum arc-length rotation that relates the two, giving a \textit{virtual} camera facing directly downwards to the scene plane (cf.\ Figure~\ref{fig:virtualcam}).
-To obtain the virtual projector, we then (i) intersect the scene plane with the optical axis of the projector and rotate the projector's placement about that point of intersection, aligning the optical axis with the scene plane's normal vector, and (ii) align the $X$- and $Y$-axes with those of the virtual camera.
+It is, however, only if the virtual projector is placed to face downwards to the scene plane directly that the projection will appear free of distortions; this is noteworthy, since a physical camera placed to face downwards is almost certain to not face directly downwards to the scene plane.
+Accordingly, we rotate the recovered camera about its center of projection to align its optical axis with the normal vector of the scene plane by computing and applying the minimum arc-length rotation that relates the two, giving a \textit{virtual} camera facing directly downwards to the scene plane (cf.\ Figure~\ref{fig:virtualcam}).
+
+To obtain the virtual projector, we (i) intersect the scene plane with the optical axis of the projector and rotate the projector's placement about that point of intersection, aligning the optical axis with the scene plane's normal vector, and (ii) align the $X$- and $Y$-axes with those of the virtual camera.
 Finally, we (iii) translate along the scene plane's normal, in order to satisfy the desired metric projected image dimensions; the height $h$ [m] above the ground plane at which the virtual projector is to be placed if the image---with original image width of $w_\text{im}$ [px]---is to be projected to the scene plane with a metric width of $w$ [m] is given by

### 2022-01-14 paper/elsarticle/back/reviewer_feedback_and_responses.txt
paper submission + adding ceres solver

The authors wish to thank the reviewers for valuable feedback and for the chance to address that feedback in a revised version of the manuscript. Please find below our response to each of the points raised by the reviewers (our responses in each instance begin with an arrow '->').

Reviewers' comments:

Reviewer #1: The manuscript describes an interesting approach to registering a scene to a projector by way of a separate camera, and using the relative position and orientation of scene and projector to apply warp correction and scaling to the projected image. My biggest concern is that the measurements being used to validate the scaling are not sufficiently described. Given that this journal has a strong focus on measurement science, the authors should strengthen this part of the manuscript.

Specific feedback:

Minor English mistakes throughout, though nothing catastrophic.

-> We have gone through the entire paper and paid attention to ensuring those mistakes have been removed

Line 16: After mentioning a product name, it is typical to put in parentheses the name of the company and its country, i.e. "(Microsoft Corporation, USA)"

-> Done (in all instances)

Line 138: Shouldn't (px,py)T actually be (xo,yo)T?

-> Corrected (the reviewer was correct)

Pages 10-12: The section on 'Projector calibration' is difficult to understand. Please, consider rewording.

-> Addressed (we believe the section should now be substantially more readable)

Lines 246-248: The sentence "Finally, we (iii)... (cf. again Figure 3)" feels like it's not complete. Please, consider rewording.

-> Addressed

Figure 5b: Please accentuate that the inset image is a magnified view of a portion of the normal image, e.g. using borders around both the magnified view and the portion that is being magnified, with lines connecting the corners of both.

-> We decided to remove the inset and instead provide additonal figures to the paper

Footnote 8: Please note that this footnote carries forward from page 15 into page 16.

-> Addressed (no footnote carries over across multiple pages any more)

Table 1: It is difficult to understand how sub-mm measurements can be made using a tape measure. Can you elaborate on how these measurements were made, and perhaps provide some information about the tape measure?

-> In the earlier reviewer copy, measurements were not sub-mm, but the computed means and standard deviations were provided at sub-mm. Reflecting feedback of reviewer 2 we have switched to providing errors in terms of absolute values; we give min and max abs val error with respect to actual measurements (not sub-mm), and mean absolute error (sub-mm, to one decimal point)

Table 1: More information is needed on how the standard deviation is calculated. What do the multiple measurements in this calculation correspond to? Repeated measurements? How many?

-> Reflecting feedback of reviewer 2 we have switched to providing errors in terms of absolute values; we give min and max abs val error with respect to actual measurements (not sub-mm), and mean absolute error (sub-mm, to one decimal point). The numbers are computed with respect to 15 augmentations produced at varying target locations across the floorspace (we have tried in the latest version to further emphasize this point, both in the captions to the tables and by providing a new Figure 7(b) for illustration)

Reviewer #2: In this paper, the authors presented a keystone correction method for image projection on a planar scene surface. The keystone correction is produced as a function of the relative geometry of the projector and scene plane, and a steerable mirror is used to direct the projection across varying target locations.  A camera facing the scene plane is used to calibrate the system. It is demonstrated that the system can be adjusted more efficiently than manual keystone correction, and with an average dimension error of less than 1 mm for 50 cm x 31.25 cm projected images over 15 locations of the projection floor.

The work is interesting as the method has practical application for industry such as displaying user manual on the factory floor. However, the reviewer found the paper a little bit to follow, hence suggest the authors put a flow chart for their procedures. Also, some figures are of poor quality, that needs to be redrawn and properly labeled. Most important, the reviewer has concern about how the measurement in Table 1 and Table 2 was made. The detailed suggestions/questions are listed below. They need to be addressed before the paper can be considered for publication.

1.      The projector is calibrated by circles projected at different target locations (Figure 1b). These circles will be distorted due to the projection angle. Since the camera and the projector are not at the same physical location, the camera calibration (by chessboard pattern) cannot fully compensate for this distortion. Will this affect the center point detection for the circles?

-> Correct, the circles undergo projective distortion when not projected in a fronto-parallel manner to the ground plane. However, the OpenCV circle pattern center detection algorithm we rely on is - like the OpenCV algorithm for detecting chessboard corners - intended to be invariant to such projective effects (i.e., it does *not* proceed by looking for circles in the image, but more generally is intended to handle ellipses). We agree with the reviewer that the question of the impact of the choice of calibration pattern on the results is not an uninteresting one; in both the reivewer copy and the latest version, we mentioned in the conclusion that an open question for future work would be to consider the impact of the choice of pattern on accuracy. In a followup paper, it is our intention to present a variant of our approach that would no longer require the presence of two separate patterns in each projector calibration image (as is the case now); it is as part of that paper that we wish to evaluate the impact of the choice of calibration pattern (chessboard vs. circles)

2.      Figure 2 does not show very well under normal size. The author should consider redrawing it.  The authors mentioned "scene plane (gray) recovered", "scene plane of back-projections", and "image plane (red)". Please label these planes in the figure, as the gray color does not show very well and there is no plane in red.

-> We have reworded the caption (now Figure 3(b)); we hope that doing so makes the figure as is - which is not merely an illustration, but is the actual rendering of our recovered camera relative to the scene plane) more clear

3.      Figure 3 is very difficult to follow. For example, the figure caption says: "The virtual projector is obtained by (ii) rotating the projector (bottom left, black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector". 3a shows an arrow labeled as (ii), which indicates the projector is moved to a new position, which is different from the camera. However, 3b shows the camera and projector optical axis is aligned (after step (ii)?).  It seems the four steps are (1) Align the camera plane with the scene plane (so they are parallel to each other). (2) Align the projector plane with the scene plane. (3) Align the projector plane x-y coordinate with the camera plane. (4) Adjust the height of the projector. It is better to put each step into a separate figure.

-> The reviewer's understanding of the steps is correct (minor detail: in the reviewer's step 3, a more accurate phrasing would be "Align the projector plane x-y axes with those of the virtual camera"). To increase clarity, we split Figure 3 from the reviewer copy into two new figures (in the latest version): Figure 4 for the virtual camera, and Figure 5 for the virtual projector. 
4.      The virtual projector is shown for both 4a and 4b. It is understood the warped image in 4b is 'projected' by the virtual camera, but what is the role of the virtual camera in projecting the original image in 4a?

-> The warped image is projected by the projector; the effect of the warp is the appearance of projecting the unwarped original image from the virtual projector. The role of the virtual camera is to supply the x-y axes of the virtual projector. Ideally, we would get them directly from the 'downwards-facing' physical camera (since it is one of our claims that we orient augmentations 'in accordance' with the 'downwards-facing' camera), but this only works if the 'downwards-facing' physical camera accurately (i.e., precisely) faces downwards to the scene plane. This is because the virtual projector itself is precisely downwards facing. Since it is almost certain to not be the case that the 'downwards-facing' camera faces downwards to the scene plane exactly (at least by some small deviation; cf. Figure 4 for the real-world example of the deviation present in the setup used in the evaluation), we (i) apply the min arc length rotation to the recovered 'downwards-facing' physical camera that rotates that camera to make it face precisely downwards to the scene plane (by making its view direction colinear with the normal vector of the scene plane) and then (ii) obtain the x-y axes we use for orienting the virtual projector are obtained by applying. It is our hope that the relevance of the virtual camera is more clear in the latest version (cf. e.g. the caption of Figure 4 in the latet version). In the current version, we now refer to the physical placed to face downwards in single quotes as the 'downwards-facing' camera to further emphasize the distinction between the physical 'downwards-facing' camera and the virtual camera.

5.      Figure 5. It is not clear where is the camera in 5a, please mark it with an arrow. The projected image insert at lower-left corner of 5b is at very low resolution. The warped image is the most important result from this study, please show it with a high-resolution image.

-> Figure 5 from the reviewer copy is now Figure 7a in the latest version; in the latest version, we have marked the three components of the hardware setup textually

6.      The authors mentioned, "the measuring tape and angle gauge are both visible in Figure 5(b)" (line 46, page 18). The image is not clear but appears it is a regular tape and angle gauge after enlarged the image. Table 1 shows, for example, the mean measure error for Top is 0.03 cm with a standard deviation of 0.33. Similarly, for Table 2, the mean error for the bottom right is 0.0 degree, with stdev 0.34. The reviewer's questions are:
(a)     What is the mean error if the absolute value is taken for each measurement before the average?
(b)     What is the measurement uncertainty due to the resolution of the tape ruler and the angle gauge?
(c)     What is the error caused by the edge blur, especially when the image was projected at an angle on the floor, some regions will be out of the focus?
(d)     The pixel size is 3.5mm and the distortion adjustment has to be based on an integer number of pixels, so theoretically speaking one can only correct the image within half pixel size. Is this understanding correct? If it is, reporting an error below this limit may not be meaningful?

-> For the current version of the paper, we redid the measurements by newly projecting our 15 augmentations to the respective 15 target locations across the floorspace and using a metal rule (visible in Figure 7(b) of the current version) to measure error in lengths. We did this because the placement of the metal rule can be shown more clearly in a figure than the previously used measuring tape (in Figure 7(b), where both the metal rule and digital angle gauge are now visible; the original measuring tape remains visible in remains visible in Figure 6, but no mention is made of it anywhere in the paper). (a) In Tables 1 and 2 we now report MEA (to one sub-mm decimal place), min error, and max error. (b) Our main reported claim concerning accuracy is now that errors in measured lengths now in no instance exceed 1 cm, and errors in angles at the measured corners in no instance exceed 1 deg. The metal rule and digital angle gauge are with certainty accurate enough to be able to make this claim. (c) Given the 6 m x 4 m extent of the floorspace considered in the evaluation, blur was not observed increase as the projection was directed further and further from directly downwards.  Figure 7(b) shows an augmentation carried out at one of the corners of the floorspace and shows that blur has a minimal impact on the quality of the augmentation, and hence on the accuracy of the measurements (as now mentioned in the figure caption). (d) The pixel size certainly has an impact on the accuracy our approach is able to provide, but is not necessarily to be seen so strictly (e.g., the algorithms we rely on to detect circle pattern center points or chessboard corners provide outputs at subpixel accuracy); we leave the question of the impact of the pixel size of the camera as an interesting question for potential future work, and make mention of this in the conclusion in the current version.

### 2022-01-14 paper/elsarticle/reviewer_feedback_and_responses_20220112.txt
paper submission + adding ceres solver

The authors would like to thank the reviewers for their valuable feedback to the manuscript. Please find below our responses to the reviewer feedback.

Reviewer #2:

- The reviewer appreciates the authors' effort, the revised version has addressed most of the reviewer's comments. However, there is still a major concern for the errors reported in Table 1. The Min AE (minimum absolute error) shows 0.0 cm, but the minimum division on the steel ruler is 0.1 cm. Not sure if that is proper to claim that ruler can measure a projected image edge on a shop floor down to 0.05 cm resolution.  More important, the impact of the pixel resolution of the dimension accuracy of the projected image can not be ignored. Question 6D in the original review is about the size of the projected image having to be an integer number of the pixel, not about the measurement accuracy of the image. For example, if the correction algorithm calculates that the corrected dimension is equivalent to 100.5 pixels, but the size can only be set either to 100 or 101 pixels. That difference is half of the pixel resolution,  or 1.75 mm. This is a limitation of digital projection. Therefore, reporting an error of 0.0 cm in Table 1 is not convincing. It could only happen when the corrected dimension matches the integer multiple of the pixel resolution very well. Therefore, a very special case.  It is recommended to take the Min AE lines out from Table 1.

Response: The reviewer is of course correct; the quality of the augmentation is bounded by the resolution of the projector, even if a perfect corrective warp is computed and applied. We accordingly acknowledge the reviewer's reservations about reporting errors of 0 mm. In line with the reviewer's recommendation, we have removed the Min AE lines from the error tables.

How big an issue this is depends on the use case; , the corrective image warp is applied in a manner that bilinear interpolation (rather than nearest neighbour interpolation) is used to produce a smooth warped image (cf. Figure 2(a) - note that the edges appear smooth). This means that corrected dimensions of 100.5 pixels may arguably appear like 100.5 pixels to the viewer, even if in fact 101 pixels are used for the effect.

Reviewer #3:

- The diagrams in the figurers are hard to read, they should be enlarged with better image quality

Response: We have enlarged most of the figures and added additional annotations to improve clarity. Additionally, we have refined several figure captions.

- The terms "up vector" and gravity vector" are used in the manuscript.  Are they describing the same vector?  If so, harmonization of terms are needed, otherwise better explanation of these terms should be provided.

Response: The up vector of a camera (or projector) gives the vertical direction of the image plane in the camera's (or projector's) local coordinate frame (cf. the red vector in e.g. Figure 3(b)). A gravity vector points in the direction of the force of gravity.

- The jargons used such as "fronto-parallel", "frontally", "face towards", or "facing downwards" are not precise descriptions in metrology domain. Better descriptions such as "projector (or camera) optical axis normal (or orthogonal) to the scene plane" are needed.

Response: We now clarify this point in Footnote 1.

- The acronym SE(3) associated with (R_j, t_j) is not defined.

Response: SE(3) is the set of 3D rigid body transformations (cf. https://www.seas.upenn.edu/~meam620/slides/kinematicsI.pdf). We agree that the notation is a bit obscure and does not add to the readability of the paper, so we removed it.

### 2022-01-14 paper/elsarticle/elsarticle-template.tex
paper submission + adding ceres solver

@@
 Considering a planar surface to be augmented (e.g., a floor, wall, or table), unless the projector faces the surface frontally, the bounds of a projected rectangular image will not appear rectangular, but will instead be subject to projective distortions (cf.\ Figure~\ref{fig:proj}).
-More generally, na\"ively projecting an image to a planar surface causes not only projective distortions as a function of target location, but also variation in scale and orientation.
+More generally, na\"ively projecting an image to a planar surface causes not only projective distortions as a function of target location on the surface, but also variation in scale and orientation.
 Such effects can be corrected for by carrying out a cumbersome manual process referred to as keystone correction---involving warping an image by perturbing its corners until the desired effect is produced when projected to the scene---to appropriately warp the image to be projected, often using software bundled with the projector.
@@
 Moreover, our method allows for specifying the desired dimensions of augmentations in metric terms, thus enabling consistent scaling.
-Our evaluation demonstrated our approach to produce compelling results---with mean and standard deviation of errors in measured lengths within the `downwards-facing' camera's pixel size---at less time than a more cumbersome manual approach to keystone correction.
+Our evaluation demonstrated our approach to produce compelling results at less time than a more cumbersome manual approach to keystone correction.
 

### 2022-02-02 paper/elsarticle/elsarticle-template_proof.tex
work on pan tilt

--- 2022-01-14 paper/elsarticle/elsarticle-template.tex
@@
-\documentclass[review]{elsarticle}
+\documentclass[proof]{elsarticle}
 
@@
     \centering
-    \subfloat[\centering Original image (to be projected).]{\raisebox{10mm}{\includegraphics[width=5cm]{images/projim.png}}}
-    \qquad
-    \subfloat[\centering Projection from side to scene plane (viewed directly downwards to scene plane).]{{\includegraphics[width=6cm]{images/proj_projim_full_.pdf}}}
+    \subfloat[\centering ]{\raisebox{10mm}{\includegraphics[width=5cm]{images/projim.png}}}
+    \qquad\quad
+    \subfloat[\centering ]{{\includegraphics[width=6cm]{images/proj_projim_full_.pdf}}}
     \caption{Na\"ive projection to a given target location in the scene plane (e.g., the floor, a wall, or table) if not facing the scene plane directly.
@@
     \centering
-    \subfloat[\centering Our warped image (to be projected).]{\raisebox{10mm}{\includegraphics[width=5cm]{images/projim_warped_scaled.png}}}
-    \qquad
-    \subfloat[\centering Projection from side to scene plane (viewed directly downwards to scene plane).]{{\includegraphics[width=6cm]{images/proj_projim_waped_full_.pdf}}}
+    \subfloat[\centering ]{\raisebox{10mm}{\includegraphics[width=5cm]{images/projim_warped_scaled.png}}}
+    \qquad\quad
+    \subfloat[\centering ]{{\includegraphics[width=6cm]{images/proj_projim_waped_full_.pdf}}}
     \caption{Projection to a given target location in the scene plane, correcting for projective distortions in terms of a \textit{virtual} projector placed and utilized according to our approach.
@@
     \centering
-    \subfloat[\centering Projector calibration image (2D position detections of chessboard pattern and asymmetric circle pattern overlain).]{{\includegraphics[width=5.85cm]{images/detections_.png}}}
+    \subfloat[\centering ]{{\includegraphics[width=5.85cm]{images/detections_.png}}}
     \quad
-    \subfloat[\centering Recovery of corresponding 3D positions of asymmetric circle pattern for use in projector calibration (red, in scene plane).]{{\includegraphics[height=4.75cm]{images/2d3d.pdf}}}
+    \subfloat[\centering ]{{\includegraphics[height=4.75cm]{images/2d3d.pdf}}}
     \caption{Obtaining the 3D positions of the 2D-3D correspondences for use in projector calibration.
@@
     \centering
-    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=5.25cm]{images/3_cam.pdf}}}
+    \subfloat[\centering ]{{\includegraphics[height=5.25cm]{images/3_cam.pdf}}}
     \qquad
-    \subfloat[\centering View directly downwards to scene~plane.]{{\includegraphics[height=5.25cm]{images/3_above3_cam.png}}}
+    \subfloat[\centering ]{{\includegraphics[height=5.25cm]{images/3_above3_cam.png}}}
     \caption{Obtaining the virtual camera.
@@
     \centering
-    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=5.25cm]{images/3_proj_.pdf}}}
+    \subfloat[\centering ]{{\includegraphics[height=5.25cm]{images/3_proj_.pdf}}}
     \qquad
-    \subfloat[\centering View directly downwards to scene~plane.]{{\includegraphics[height=5.25cm]{images/3_above3_proj.png}}}
+    \subfloat[\centering ]{{\includegraphics[height=5.25cm]{images/3_above3_proj.png}}}
     \caption{Obtaining the virtual projector, with $X$- and $Y$-axes aligned with those of the virtual camera (cf.\ Figure~\ref{fig:virtualcam}).
@@
     \centering
-    \subfloat[\centering Hardware setup consisting of a Panasonic PT-RZ660BE projector, a steerable mirror system manufactured by Dynamic Projection Institute, and the left view of a Stereolabs Zed 2 stereo camera---placed to face downwards towards the scene plane---to serve as the `downwards-facing' camera.]{{\includegraphics[height=5cm]{images/system3_.pdf}}}  %3.6
+    \subfloat[\centering ]{{\includegraphics[height=3.6cm]{images/system3_.pdf}}}  %3.6
     \quad
-    \subfloat[\centering Measurement for augmentation, using a steel rule and a digital angle gauge (enlargements on the right).]{{\includegraphics[height=5cm]{images/meas2_.pdf}}}
+    \subfloat[\centering ]{{\includegraphics[height=3.6cm]{images/meas2_.pdf}}}
     \caption{Evaluation scenario.

### 2022-02-02 src/apps/calibrateProj/calibrateProj.cpp

@@ -9,7 +9,6 @@
  *   calibrateProj 0.0565 4 6 4 11 C:\Users\micha\Desktop\spatial-ar\in_out\calibrateProj\out 11 C:\Users\micha\Desktop\spatial-ar\in_out\splitZed\projCalib\outLeft C:\Users\micha\Desktop\spatial-ar\in_out\splitZed\projCalib\outRight C:\Users\micha\Desktop\spatial-ar\in_out\calibrateCam\out\cam_0.yml  C:\Users\micha\Desktop\spatial-ar\in_out\calibrateCam\out\cam_1.yml C:\Users\micha\Desktop\spatial-ar\in_out\calibrateProj\acircles_pattern_960x600.png 1.0 3 C:\Users\micha\Desktop\spatial-ar\in_out\applyHomography\holodeck.png
  */
 
-// CMake projects should use: "-DCMAKE_TOOLCHAIN_FILE=C:/Users/micha/vcpkg/scripts/buildsystems/vcpkg.cmake"
 
 #include <iostream>
 #include <fstream>
@@ -35,7 +34,7 @@ using namespace cv;
 GLint viewportWidth;
 GLint viewportHeight;
 
-GLFWwindow * window;
+GLFWwindow* window;
 
 bool showDistances;
 bool showVirtual;
@@ -44,7 +43,7 @@ bool showImVis;
 /* Camera variables */
 GLdouble defaultCameraDist, defaultCameraYaw, defaultCameraRoll;
 GLdouble cameraDist, cameraYaw, cameraRoll;
-GLdouble * cameraPos;
+GLdouble* cameraPos;
 
 int camIdx;
 int viewIdx;
@@ -59,18 +58,14 @@ vector<vector<Mat>> capturedPattern;
 
 cv::Vec3d centroidChessboardObjectPts;
 
-PointCloud * pointCloud;
-std::vector<PointCloud*> pointCloudCirclesVec;
+PointCloud* pointCloud;
+PointCloud* pointCloudCircles;
 PointCloud* pointCloud2Circles;
-std::vector<PointCloud*> triangulatedPointCloudCirclesVec;
 PointCloud* pointCloudImVis;
 PointCloud* pointCloud2ImVis;
-
-Plane * planeVis;
-Plane* fittedPlaneVis;
+Plane* planeVis;
 
 std::vector<Plane> planes;
-std::vector<Plane> fittedPlanes;
 
 cv::Vec3d projCamIntersection;
 cv::Vec3d projPlaneIntersection;
[…]

### 2022-02-02 src/apps/calibrateProjMirror/calibrateProjMirror.cpp

@@ -1,0 +1,1343 @@
+/*
+ *   2022 Michael Hornacek
+ *   michael.hornacek@gmail.com
+ *   IMW-CPS TU Vienna, Austria
+ *
+ *   calibrateProj <boardSqSize> <boardDimsX> <boardDimsY> <circlesDimsX> <circlesDimsY> <outDir> <numIms> <cam0ImDir> <cam1ImDir> <cam0Path> <cam1Path> <circlesImPath> <targetWidth> [<visImIdx>] [<visIm>]
+ *
+ *   Example invocation:
+ *   calibrateProjMirror 0.0565 4 6 4 11 C:\Users\micha\Desktop\spatial-ar\in_out\calibrateProj\out 11 C:\Users\micha\Desktop\spatial-ar\in_out\splitZed\projCalib\outLeft C:\Users\micha\Desktop\spatial-ar\in_out\splitZed\projCalib\outRight C:\Users\micha\Desktop\spatial-ar\in_out\calibrateCam\out\cam_0.yml  C:\Users\micha\Desktop\spatial-ar\in_out\calibrateCam\out\cam_1.yml C:\Users\micha\Desktop\spatial-ar\in_out\calibrateProj\acircles_pattern_960x600.png 1.0 3 C:\Users\micha\Desktop\spatial-ar\in_out\applyHomography\holodeck.png
+ */
+
+// CMake projects should use: "-DCMAKE_TOOLCHAIN_FILE=C:/Users/micha/vcpkg/scripts/buildsystems/vcpkg.cmake"
+
+#include <iostream>
+#include <fstream>
+#include <vector>
+#include <sstream>
+
+#include "headers.h"
+#include "Camera.h"
+#include "PointCloud.h"
+#include "Plane.h"
+#include "CalibPattern.h"
+#include "Homography.h"
+
+#include <opencv2/core.hpp>
+#include <opencv2/highgui.hpp>
+#include <opencv2/calib3d.hpp>
+#include <opencv2/imgproc.hpp>
+#include <opencv2/opencv_modules.hpp>
+
+#include "ceres/ceres.h"
+#include "ceres/rotation.h"
+
+using namespace std;
+using namespace cv;
+
+GLint viewportWidth;
+GLint viewportHeight;
+
+GLFWwindow * window;
+
+bool showDistances;
+bool showVirtual;
+bool showImVis;
+
+/* Camera variables */
[…]

### 2022-02-02 src/apps/visPanTilt/visPanTilt.cpp

@@ -1,0 +1,1339 @@
+/*
+ *   2022 Michael Hornacek
+ *   michael.hornacek@gmail.com
+ *   IMW-CPS TU Vienna, Austria
+ *
+ */
+
+// CMake projects should use: "-DCMAKE_TOOLCHAIN_FILE=C:/Users/micha/vcpkg/scripts/buildsystems/vcpkg.cmake"
+
+#include <iostream>
+#include <fstream>
+#include <vector>
+#include <sstream>
+
+#include "headers.h"
+#include "Camera.h"
+#include "PointCloud.h"
+#include "Plane.h"
+#include "CalibPattern.h"
+#include "Homography.h"
+
+#include <opencv2/core.hpp>
+#include <opencv2/highgui.hpp>
+#include <opencv2/calib3d.hpp>
+#include <opencv2/imgproc.hpp>
+#include <opencv2/opencv_modules.hpp>
+
+#include "ceres/ceres.h"
+#include "ceres/rotation.h"
+
+using namespace std;
+using namespace cv;
+
+GLint viewportWidth;
+GLint viewportHeight;
+
+GLFWwindow * window;
+
+bool showDistances;
+bool showVirtual;
+bool showImVis;
+
+/* Camera variables */
+GLdouble defaultCameraDist, defaultCameraYaw, defaultCameraRoll;
+GLdouble cameraDist, cameraYaw, cameraRoll;
+GLdouble * cameraPos;
+
+int camIdx;
+int viewIdx;
+vector<Camera> cams;
+int imWidth, imHeight;
+
+Camera projCam;
+int projWidth, projHeight;
+
+int numPatterns;
+vector<vector<Mat>> capturedPattern;
+
+cv::Vec3d centroidChessboardObjectPts;
+
+PointCloud * pointCloud;
+std::vector<PointCloud*> pointCloudCirclesVec;
+PointCloud* pointCloud2Circles;
+std::vector<PointCloud*> triangulatedPointCloudCirclesVec;
+PointCloud* pointCloudImVis;
+PointCloud* pointCloud2ImVis;
+
+Plane * planeVis;
+Plane* fittedPlaneVis;
+
+std::vector<Plane> planes;
+std::vector<Plane> fittedPlanes;
+
+cv::Vec3d projCamIntersection;
+cv::Vec3d projPlaneIntersection;
+cv::Vec3d camPlaneIntersection;
+cv::Vec3d virtualCamPlaneIntersection;
+
[…]
